### 2021-03-19 Paper/Paper.Rnw
added an outline for the paper, largely working on the theory section for the sup mat now

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xr}
\usepackage{authblk}

%new commands and so on
\providecommand{\keywords}[1]
{
  \small	
  \textbf{\textit{Keywords---}} #1
}
\newcommand{\mcP}{\mathcal{P}}
\newcommand{\mcB}{\mathcal{P}}

%external documents
\externaldocument[SM-]{SupMat}

%header material for paper
\title{How environmental drivers of spatial synchrony interact}
\date{}

%***Tentative author order
\author[a,b]{Daniel C. Reuman}
\author[c]{Max Castorani}
\author[d,e]{Tom Bell}
\author[f]{Kyle Cavanaugh}
\author[c]{Jonathan Walter}
\author[g]{Lawrence W. Sheppard}

\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
\affil[b]{Laboratory of Populations, Rockefeller University}
\affil[c]{University of Virginia}
\affil[d]{WHOI}
\affil[e]{UCSB}
\affil[f]{UCLA}
\affil[g]{MBA}

%***Need to indicate corresponding author Reuman

\begin{document}

%The following is where you load in the numeric results that will be embedded in the text
<<eval=T,echo=F,message=F,warning=F>>=
#place R code here for loading in necessary variables from Results
@

\maketitle

\begin{abstract}
Place abstract here
\end{abstract}

%***Add to these
\keywords{kelp, synchrony, Moran effects}

\section{Introduction}

\section{Theory}

\section{Methods}

\section{Results}

\section{Discussion}

\section{References}

\end{document}

### 2021-03-19 Paper/SupMat.Rnw
added an outline for the paper, largely working on the theory section for the sup mat now

\documentclass[letterpaper,11pt]{article}

%packages
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
\usepackage{url}
\usepackage{multirow}
\usepackage{longtable}
\usepackage{subfig}
\usepackage{float}
\usepackage{setspace}
\usepackage{lineno}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xr}
\usepackage{authblk}

\setcounter{figure}{0}
\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{section}{0}
\renewcommand{\thesection}{S\arabic{section}}

%new commands and so on
\newcommand{\mcP}{\mathcal{P}}
\newcommand{\mcB}{\mathcal{B}}
\newcommand{\Rp}{\text{Re}}
\newcommand{\cov}{\text{cov}}
\let\conj\overline

%external documents
\externaldocument[MT-]{Paper}

%header material for paper
\title{How environmental drivers of spatial synchrony interact}
\date{}

%***Tentative author order
\author[a,b]{Daniel C. Reuman}
\author[c]{Max Castorani}
\author[d,e]{Tom Bell}
\author[f]{Kyle Cavanaugh}
\author[c]{Jonathan Walter}
\author[g]{Lawrence W. Sheppard}

\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
\affil[b]{Laboratory of Populations, Rockefeller University}
\affil[c]{University of Virginia}
\affil[d]{WHOI}
\affil[e]{UCSB}
\affil[f]{UCLA}
\affil[g]{MBA}

\begin{document}

%The following is where you load in the numeric results that will be embedded in the text
<<eval=T,echo=F,message=F,warning=F>>=
#place R code here for loading in necessary variables from Results
@

\maketitle

\tableofcontents
\listoftables
\listoffigures

\section{Measuring synchrony} \label{Sect:measuring_sync}

If $w=(w_1,\ldots,w_N)$ is a stationary stochastic process, where the index $i=1,\ldots,N$ corresponds
to sampling locations, then we measure the synchrony of $w$ as 
\begin{equation}
\rho_{ww} = \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{w_i w_j},
\end{equation}
where $S_{ww}$ is the spectral matrix of the process $w$. This is a function of frequency or timescale.
To see that $\rho_{ww}$ is a real-valued function, note that $\conj{S_{ww}}=S_{ww}^\tau$, where the superscript 
$\tau$ denotes matrix transpose (REF to Brillinger theorem 2.5.1 on p. 24).
%***DAN: Don't forget to fill in this reference!
Therefore 
\begin{align}
\conj{\rho_{ww}} &= \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} \conj{S_{w_i w_j}} \\
&= \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{w_j w_i} \\
&= \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{w_i w_j} \\
&= \rho_{ww}.
\end{align} 
The quantity $\rho_{ww}$ makes sense as a measure of synchrony in part because 
\begin{align}
\rho_{ww} &= \Rp(\rho_{ww}) \\
&= \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} \Rp(S_{w_i w_j}).
\end{align}
The quantity $\Rp(S_{w_i w_j})$ is the cospectrum of the component processes $w_i$ and $w_j$, 
and an appropriate integral across frequencies of the cospectrum equals the covariance, 
$\cov(w_i,w_j)$ (REF to some specific page/theorem in Brillinger). 
%***DAN: Don't forget to add
Thus an integral of $\rho_{ww}$ across frequencies equals 
\begin{equation}
\left( \frac{1}{N^2-N} \right) \sum_{i \neq j} \cov(w_i,w_j),
\end{equation}
which is a standard quantity that can be used to measure non-frequency-specific synchrony.
Thus $\rho_{ww}$ is a frequency-specific generalization of a classic, non-frequency-specific 
measure of synchrony.

If $\epsilon^{(1)}=(\epsilon^{(1)}_1,\ldots,\epsilon^{(1)}_N)$ and 
$\epsilon^{(2)}=(\epsilon^{(2)}_1,\ldots,\epsilon^{(2)}_N)$
are two stationary stochastic processes, we define the \emph{cross-variable spatial synchrony}
(or, simply, the \emph{cross synchrony}) between $\epsilon^{(1)}$ and $\epsilon^{(2)}$ as
\begin{equation}
\rho_{\epsilon^{(1)} \epsilon^{(2)}} = \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon^{(1)}_i \epsilon^{(2)}_j}.
\end{equation}
This is a complex-valued function of frequency or timescale. Why is this interpretable as cross-variable 
spatial synchrony? A quantity representing spatial synchrony should compare values of the processes 
$\epsilon^{(1)}$ and $\epsilon^{(2)}$ in one location to values in another location, since spatial synchrony
is about relationships between distinct locations; hence our sum is over $i \neq j$. 
A quantity representing cross-variable synchrony should 
make comparisons between two variables, as opposed to comparisons between the same variable in different locations;
hence we use the cross spectrum between variable $1$ in location $i$ ($\epsilon^{(1)}_i$) and variable 
$2$ in location $j$ ($\epsilon^{(2)}_j$). 
Finally, a quantity representing cross-variable synchrony needs to
include the possibility of time delays in the relationship between the two variables. For instance, 
if $\epsilon^{(2)}_j$ is a time-lagged version of $\epsilon^{(1)}_i$, but the effects of 
$\epsilon^{(2)}_j$ on a biotic process $w_j$ are immediate, whereas the effects of $\epsilon^{(1)}_i$
on $w_i$ are delayed by the same lag, then the effects of $\epsilon^{(1)}_i$ on $w_i$ will be correlated
with the effects of $\epsilon^{(2)}_j$ on $w_j$, i.e., this combination of properties of $\epsilon^{(1)}$
and $\epsilon^{(2)}$ will produce synchrony, in the classic sense of correlation, between the biotic
process $w_i$ and $w_j$.
Thus we use the cross spectrum, $S_{\epsilon^{(1)}_i \epsilon^{(2)}_j}$, for which the phase angle at 
frequency $f$ measures the time delay between frequency-$f$ oscillations of $\epsilon^{(1)}_i$ and 
frequency-$f$ oscillations of $\epsilon^{(2)}_j$. The following examples should help clarify these features of
the definition of cross synchrony.

%***DAN: See page 51 of your hand-written notes for descriptions of the examples, which you need to code up and 
%make some figure or something and describe and report here.

\section{Model specification, full details}

The population model is specified starting with the equation
\begin{align}
w_i(t) &= b_1 w_i(t-1)+ \cdots + b_n w_i(t-n) \label{main_model_1} \\ 
&+ p_0^{(1)} \epsilon_i^{(1)}(t)+ \cdots + p_{m_1}^{(1)} \epsilon_i^{(1)}(t-m_1) \label{main_model_2} \\
&+ p_0^{(2)} \epsilon_i^{(2)}(t)+ \cdots + p_{m_2}^{(2)} \epsilon_i^{(2)}(t-m_2) \label{main_model_3} \\
&+ \delta_i(t), \label{main_model_4}
\end{align}
where $i=1,\ldots,N$ indexes habitat patches or sampling locations, $w_i(t)$ is a measure of the 
population in location $i$ at time $t$, and $\epsilon^{(1)}=(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)})$,
$\epsilon^{(2)}=(\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)})$ and $\delta=(\delta_1,\ldots,\delta_N)$ are
environmental processes influencing the populations. Note that this model assumes from the outset 
a homogeneous influence of the environmental variables on the populations in the different
sampling locations, i.e., the parameters $p$ are spatially homogeneous. We also assume that the 
combined process $(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ is a second-order stationary stochastic process (REF)
%***DAN: REFS here to Brillinger
with the expected values of its component processes equal to $0$. Finally, we assume the complex roots of the
polynomial $1-b_1 z - \cdots - b_n z^n = 0$ have modulus greater than $1$, an assumption needed for invertability of
the filter $\mcB$ which we will define below. See REF
%***DAN: add refs to Brilliger, Shumway and Stoffer, Rinsel
for background on the statistical theory of stationary stochastic processes.

Letting $I_N$ denote the $N \times N$ identity matrix, we define the linear filters 
\begin{align}
\mcP^{(1)} &= p_0^{(1)} I_N + p_1^{(1)} I_N B + \cdots + p_{m_1}^{(1)} I_N B^{m_1}, \\
\mcP^{(2)} &= p_0^{(2)} I_N + p_1^{(2)} I_N B + \cdots + p_{m_2}^{(2)} I_N B^{m_2}, \\
\mcB &= I_N - b_1 I_N B - b_2 I_N B^2 - \cdots b_n I_N B^n. \\
\end{align}
Using the language of filters and stochastic processes, the original model 
(\ref{main_model_1}-\ref{main_model_4}) can then be written as
\begin{equation}
\mcB w = \mcP^{(1)} \epsilon^{(1)} +\mcP^{(2)} \epsilon^{(2)} +\delta.\label{main_model_stochproc}
\end{equation}
Writing the model in the language of stochastic processes and linear filters makes it straightforward
to calculate the spectral matrix, $S_{ww}$, of the stochastic process, $w$, which we do in section \ref{sect:derivation}.

%***DAN: Do you need to have some sort of section gtiving background info on spectral stuff, or is that all assumed?
%Probably no, we just assume it here, and give some info in the main text on how to interpret these quantities, but
%we do not have to give any definitions and so on because we are not going to give anything like a proper into
%to stochastic processes and spectral approaches either in the main text or here, so better just to assume it 
%(here) and take the attitude in the main text of "you won't understand where this comes from (see sup mat for that)
%but we can tell you how to interpret it.

\section{Derivation of the main equation}\label{sect:derivation}

Computing the spectral matrix of both sides of the model (\ref{main_model_stochproc}) gives
\begin{align}
T(\mcB)S_{ww} T(\mcB)^* &= T(\mcP^{(1)}) S_{\epsilon^{(1)} \epsilon^{(1)}} T(\mcP^{(1)})^* \label{spectmat_eq1_t1}\\
  &+ T(\mcP^{(2)}) S_{\epsilon^{(2)} \epsilon^{(2)}} T(\mcP^{(2)})^* \\
  &+ S_{\delta \delta} \\
  &+ T(\mcP^{(1)}) S_{\epsilon^{(1)} \epsilon^{(2)}} T(\mcP^{(2)})^*+T(\mcP^{(2)}) S_{\epsilon^{(2)} \epsilon^{(1)}} T(\mcP^{(1)})^* \\
  &+ T(\mcP^{(1)}) S_{\epsilon^{(1)}\delta} + S_{\delta \epsilon^{(1)}} T(\mcP^{(1)})^* \\
  &+ T(\mcP^{(2)}) S_{\epsilon^{(2)}\delta} + S_{\delta \epsilon^{(2)}} T(\mcP^{(2)})^*, \label{spectmat_eq1_t6}
\end{align}
where $T(\mcB)$, $T(\mcP^{(1)})$ and $T(\mcP^{(2)})$ are the tranfer functions of the filters
$\mcB$, $\mcP^{(1)}$ and $\mcP^{(2)}$, respectively, $S_{ww}$ is the spectral matrix of the process $w$,
$S_{\epsilon^{(i)} \epsilon^{(i)}}$ is the spectral matrix of the process $\epsilon^{(i)}$, $S_{\delta \delta}$
is the spectral matrix of the process $\delta$,  $S_{\epsilon^{(i)} \epsilon^{(j)}}$ is the cross spectral
matrix of the processes $\epsilon^{(i)}$ and $\epsilon^{(j)}$, $S_{\epsilon^{(i)} \delta}$ is the cross 
spectral matrix of the processes $\epsilon^{(i)}$ and $\delta$, $S_{\delta \epsilon^{(i)}}$ is the cross
spectral matrix of the processes $\delta$ and $\epsilon^{(i)}$, and the superscript $*$ denotes
the conjugate transpose of a matrix.

Letting $\mu=\exp(-2\pi i f)$ where $f$ is frequency in units of cycles per time steps, we have
\begin{align}
T(\mcB) &= (1-b_1 \mu - b_2 \mu^2 - \cdots - b_n \mu^n) I_N \\
T(\mcP^{(1)}) &= (p_0^{(1)} + p_1^{(1)} \mu + \cdots + p_{m_1}^{(1)} \mu^{m_1}) I_N \\
T(\mcP^{(2)}) &= (p_0^{(2)} + p_1^{(2)} \mu + \cdots + p_{m_2}^{(2)} \mu^{m_2}) I_N.
\end{align}
\noindent These are all scalar matrices, so (\ref{spectmat_eq1_t1}-\ref{spectmat_eq1_t6}) simplifies to
\begin{align}
S_{ww} &= \frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} S_{\epsilon^{(1)} \epsilon^{(1)}} \label{Sww_T1}\\
&+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} S_{\epsilon^{(2)} \epsilon^{(2)}} \\
&+ \frac{1}{|f_{\mcB}|^2} S_{\delta \delta}\\
&+ \frac{1}{|f_{\mcB}|^2} \left[  f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}  S_{\epsilon^{(1)} \epsilon^{(2)}} + 
f_{\mcP^{(2)}} \conj{f_{\mcP^{(1)}}} S_{\epsilon^{(2)} \epsilon^{(1)}}\right] \label{Sww_T4}\\
&+ \frac{1}{|f_{\mcB}|^2} \left[f_{\mcP^{(1)}} S_{\epsilon^{(1)} \delta} + \conj{f_{\mcP^{(1)}}}  S_{\delta \epsilon^{(1)}}\right]\\
&+ \frac{1}{|f_{\mcB}|^2} \left[ f_{\mcP^{(2)}} S_{\epsilon^{(2)} \delta} + \conj{f_{\mcP^{(2)}}}  S_{\delta \epsilon^{(2)}} \right], \label{Sww_T6}
\end{align}
where 
\begin{align}
f_{\mcB} &= 1-b_1 \mu - b_2 \mu^2 - \cdots - b_n \mu^n \\
f_{\mcP^{(1)}} &= p_0^{(1)} + p_1^{(1)} \mu + \cdots + p_{m_1}^{(1)} \mu^{m_1} \\
f_{\mcP^{(2)}} &= p_0^{(2)} + p_1^{(2)} \mu + \cdots + p_{m_2}^{(2)} \mu^{m_2}.
\end{align}
We here made use of the assumption that the complex roots of the
polynomial $1-b_1 z - \cdots - b_n z^n = 0$ have modulus greater than $1$, so that $f_{\mcB}$ is non-zero.

Now, averaging the off-diagonal entries of both sides of (\ref{Sww_T1}-\ref{Sww_T6}) gives
\begin{align}
\left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{w_i w_j} &= 
\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\epsilon_i^{(1)} \epsilon_j^{(1)}} \\
 &+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\epsilon_i^{(2)} \epsilon_j^{(2)}} \\
 &+ \frac{1}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} S_{\delta_i \delta_j} \\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} 
  \Rp \left( f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} S_{\epsilon_i^{(1)} \epsilon_j^{(2)}} \right) \label{SumSww_T4}\\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right)  \sum_{i \neq j} \Rp \left( f_{\mcP^{(1)}}  
 S_{\epsilon_i^{(1)} \delta_j}\right)  \label{SumSww_T5}\\
 &+ \frac{2}{|f_{\mcB}|^2} \left(\frac{1}{N^2-N}\right) \sum_{i \neq j} \Rp \left( f_{\mcP^{(2)}}    
 S_{\epsilon_i^{(2)} \delta_j}\right), \label{SumSww_T6}
\end{align}
\noindent where $\Rp(\cdot)$ denotes the real part, and where (\ref{SumSww_T4}-\ref{SumSww_T6})
follows from two observations: that the two summands in the brackets of each of the terms 
(\ref{Sww_T4}-\ref{Sww_T6}) are conjugate-transpose pairs; and that, for any complex matrix, $A$,
$\sum_{i \neq j} (A_{ij} +\conj{A_{ji}}) = \sum_{i \neq j} (A_{ij} +\conj{A_{ij}}) = 2 \sum_{i \neq j} \Rp (A_{ij})$.
Now, making use of the definitions of section \ref{Sect:measuring_sync}, this becomes
\begin{align}
\rho_{ww} &= \frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}
+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}
+ \frac{1}{|f_{\mcB}|^2} \rho_{\delta \delta} \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right) \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(1)}} \rho_{\epsilon^{(1)} \delta} \right) \\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(2)}} \rho_{\epsilon^{(2)} \delta} \right).
\end{align}
If the processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ were measured and the process $\delta$ represents
the aggregate effects of unknown processes, then we can write
\begin{align}
\rho_{ww} &= \frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}} \label{rhoww_T1}\\
&+ \frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}} \label{rhoww_T2}\\
&+ \frac{2}{|f_{\mcB}|^2} \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right) \label{rhoww_Tint} \\
&+ \text{other contributions}.
\end{align}
The term (\ref{rhoww_T1}) is contributions to the synchrony of $w$ due to direct Moran effects of $\epsilon^{(1)}$.
The term (\ref{rhoww_T2}) is contributions to the synchrony of $w$ due to direct Moran effects of $\epsilon^{(2)}$.
The term (\ref{rhoww_Tint}) is contributions to the synchrony of $w$ due to interactions between the Moran effects
of $\epsilon^{(1)}$ and $\epsilon^{(2)}$. The ``other contributions'' may include direct Moran effects of other,
unmeasured factors, and interactions between these effects and Moran effects of the measured factors
$\epsilon^{(1)}$ and $\epsilon^{(2)}$.

\end{document}

### 2021-03-22 Code/AnalysisStep_SpectralDecomp.R
changes mostly to do with moving toward making the figs for a paper

@@ -376,8 +376,8 @@ do_analysis<-function(Args)
                fnpre=paste0(fnpre,"_cov"),plottype,PlotForMSArgs=h)
 
     #make the plot that explains the main effects
-    explain_direct_Moran_ForMS(freq=freq,frg=frg,fP1=fP1,fP2=fP2,Se1e1_avg=Se1e1_avg_cov,Se2e2_avg=Se2e2_avg_cov,
-                               fnpre=paste0(fnpre,"_cov"),plottype=plottype)
+    explain_direct_Moran_ForMS(freq=freq,frg=frg,fP1=fP1,fP2=fP2,fB=fB,Se1e1_avg=Se1e1_avg_cov,Se2e2_avg=Se2e2_avg_cov,
+                               fnpre=paste0(fnpre,"_cov"),plottype=plottype,PlotForMSArgs=PlotForMSArgs)
     
     
   }
@@ -387,13 +387,32 @@ do_analysis<-function(Args)
 
 #A function for making the manuscript plots explaining direct Moran effects
 #
-explain_direct_Moran_ForMS<-function(freq,frg,fP1,fP2,Se1e1_avg,Se2e2_avg,fnpre,plottype)
+explain_direct_Moran_ForMS<-function(freq,frg,fP1,fP2,fB,Se1e1_avg,Se2e2_avg,fnpre,plottype,PlotForMSArgs)
 {
+  #unpack plotting params
+  fP1_short_ylims<-PlotForMSArgs$ExplainDirectMoran_fP1_short_ylims
+  fP1_mid_ylims<-PlotForMSArgs$ExplainDirectMoran_fP1_mid_ylims
+  fP1_long_ylims<-PlotForMSArgs$ExplainDirectMoran_fP1_long_ylims
+  Se1e1_short_ylims<-PlotForMSArgs$ExplainDirectMoran_Se1e1_short_ylims
+  Se1e1_mid_ylims<-PlotForMSArgs$ExplainDirectMoran_Se1e1_mid_ylims
+  Se1e1_long_ylims<-PlotForMSArgs$ExplainDirectMoran_Se1e1_long_ylims
+  fP2_short_ylims<-PlotForMSArgs$ExplainDirectMoran_fP2_short_ylims
+  fP2_mid_ylims<-PlotForMSArgs$ExplainDirectMoran_fP2_mid_ylims
[…]

### 2021-03-22 Paper/SupMat.Rnw
initial write up of theory for the sup mat, prior to writing a summary for the main text. Not done, need to write some examples.

@@
 of $\epsilon^{(1)}$ and $\epsilon^{(2)}$. The ``other contributions'' may include direct Moran effects of other,
-unmeasured factors, and interactions between these effects and Moran effects of the measured factors
-$\epsilon^{(1)}$ and $\epsilon^{(2)}$.
+unmeasured environmental factors, as well as interactions between these effects and Moran effects of the 
+measured factors $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
 

### 2021-03-22 Paper/Paper.Rnw
rapid try at paragraph outline for Intro added

@@
 
+Tentative paragraph outline written on my phone on 3/20/2021
+1) Importance of synchrony
+2) Classic problem of identifying causes of synchrony, but recently largely resolved due to new methods
+  -elaborate briefly in a couple methods
+3) in fact, it has recently been discovered that environmental causes of synchrony can interact, so that ...
+  -related to, but distinct from, interaction between environmental and dispersal causes of synchrony discussed, for instance by Kendal et al
+  -make sure you describe the two cases
+4) the simplest form of interacting moran effects can be demonstrated when a simple model, though elaboration of the true nature of this mechanism will take the rest of this manuscript
+  -use Moran model
+  -but with two drivers
+  -moran theorem says correlation of pops equals correlation of sum of drivers, so work it out
+5) general idea of two noises lagged or time delayed relative to each other and their effects oppositely lagged
+6) paragraph on why we would expect interacting moran effects to be common
+7) final para on goals, ma structure, etc.
+
 \section{Theory}

### 2021-07-29 Code/AnalysisStep_LinearModels.R
small changes only

@@ -157,16 +157,14 @@ saveRDS(allfsgtd,file=paste0(resloc2,"allfsgtd"))
 #Now do the same linear regressions, but for only selected geographic regions
 #***
 
-#***The following code copied from AnalysisStep_LinearModels.R
-
 #from google maps
-MontBLatLon<-c(36.62393626182818, -121.8511283749223)
-CarBLatLon<-c(36.52316357182972, -121.95535513873469)
-MorBLatLon<-c(35.427553615598676, -120.89133343717805)
-PtConcLatLon<-c(34.44830424836278,-120.47125509005721)
-SBLatLon<-c(34.407876790966135, -119.68428533917637)
-OxLatLon<-c(34.2310543912163, -119.2726684255394)
-LALatLon<-c(34.00907538429378, -118.50422514984244)
+MontBLatLon<-c(36.62393626182818, -121.8511283749223) #Monterey Bay, closer to the southern end
+CarBLatLon<-c(36.52316357182972, -121.95535513873469) #The point to the south of Carmel Bay
+MorBLatLon<-c(35.427553615598676, -120.89133343717805) #Morro Bay, closer to the northern end
+PtConcLatLon<-c(34.44830424836278,-120.47125509005721) #Pt Conception
+SBLatLon<-c(34.407876790966135, -119.68428533917637) #Santa Barbara pier
+OxLatLon<-c(34.2310543912163, -119.2726684255394) #Oxnard, where the Santa Clara River empties
+LALatLon<-c(34.00907538429378, -118.50422514984244) #LA, near the Santa Monica pier
 
 dists<-ncf::gcdist(x=c(MontBLatLon[2],locs$Lon),y=c(MontBLatLon[1],locs$Lat))
 dists<-dists[1,2:(dim(dists)[2])]
@@ -220,8 +218,6 @@ h<-c(CC2locstouse[1],CC2locstouse[length(CC1locstouse)])
 ncf::gcdist(locs$Lon[h],locs$Lat[h]) #78.46km
 
 #So my three roughly comparable comparable regions are SBlocstouse, CC1locstouse, CC2locstouse
-
-#***The above code copied from AnalysisStep_LinearModels.R
 
 #***Now do central cal, region 1
 
@@ -313,7 +309,7 @@ for (counter in 1:5)
 names(allfsgtd_SB)<-c("Rsq","NO3_l0","NO3_l1","waves_l0","waves_l1")
[…]

### 2021-07-29 Paper/SupMat.Rnw
mostly small changes

@@
 
-If $w=(w_1,\ldots,w_N)$ is a stationary stochastic process, where the index $i=1,\ldots,N$ corresponds
+If $w=(w_1,\ldots,w_N)$ is an $N$-dimensional stationary stochastic process, where the index $i=1,\ldots,N$ corresponds
 to sampling locations, then we measure the synchrony of $w$ as 
@@
 \end{equation}
-where $S_{ww}$ is the spectral matrix of the process $w$. This is a function of frequency or timescale.
+where $S_{ww}$ is the spectral matrix of the process $w$. This is a function of frequency or timescale because
+$S_{w_i w_j}$ is a function of frequency or timescale.
 To see that $\rho_{ww}$ is a real-valued function, note that $\conj{S_{ww}}=S_{ww}^\tau$, where the superscript 
@@
 $\epsilon^{(2)}=(\epsilon^{(2)}_1,\ldots,\epsilon^{(2)}_N)$
-are two stationary stochastic processes, we define the \emph{cross-variable spatial synchrony}
+are two $N$-dimensional stationary stochastic processes, we define the \emph{cross-variable spatial synchrony}
 (or, simply, the \emph{cross synchrony}) between $\epsilon^{(1)}$ and $\epsilon^{(2)}$ as
@@
 and $\epsilon^{(2)}$ will produce synchrony, in the classic sense of correlation, between the biotic
-process $w_i$ and $w_j$.
+processes $w_i$ and $w_j$.
 Thus we use the cross spectrum, $S_{\epsilon^{(1)}_i \epsilon^{(2)}_j}$, for which the phase angle at 
@@
 a homogeneous influence of the environmental variables on the populations in the different
-sampling locations, i.e., the parameters $p$ are spatially homogeneous. We also assume that the 
+sampling locations, i.e., the parameters $p$ are spatially homogeneous. 
+%We also assume spatial homogeneity of the AR coefficients, need to add a mention of that
+We also assume that the 
 combined process $(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ is a second-order stationary stochastic process (REF)
@@
 with the expected values of its component processes equal to $0$. Finally, we assume the complex roots of the
-polynomial $1-b_1 z - \cdots - b_n z^n = 0$ have modulus greater than $1$, an assumption needed for invertability of
+polynomial $1-b_1 z - \cdots - b_n z^n = 0$ have modulus greater than $1$, an assumption needed for the 
+existence of a stationary stochastic process solution for the model and for invertability of
 the filter $\mcB$ which we will define below. See REF
@@
 
-Letting $I_N$ denote the $N \times N$ identity matrix, we define the linear filters 
+Letting $I_N$ denote the $N \times N$ identity matrix and letting $B$ denote the backshift operator, 
+we define the linear filters 
 \begin{align}
@@
 \mcP^{(2)} &= p_0^{(2)} I_N + p_1^{(2)} I_N B + \cdots + p_{m_2}^{(2)} I_N B^{m_2}, \\
-\mcB &= I_N - b_1 I_N B - b_2 I_N B^2 - \cdots b_n I_N B^n. \\
+\mcB &= I_N - b_1 I_N B - b_2 I_N B^2 - \cdots b_n I_N B^n.
 \end{align}
@@
 &+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(1)}} \rho_{\epsilon^{(1)} \delta} \right) \\
-&+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(2)}} \rho_{\epsilon^{(2)} \delta} \right).
+&+ \frac{2}{|f_{\mcB}|^2} \Rp \left( f_{\mcP^{(2)}} \rho_{\epsilon^{(2)} \delta} \right),
 \end{align}
+which expresses population synchrony in terms of noise synchrony and cross synchrony.
 If the processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ were measured and the process $\delta$ represents
@@
 
+\section{Details of data and data cleaning}
+
+%***DAN: This should include specification of the regions CC1, CC2, SB.
+%You have to get some info from Tom about the nature of the spatial sampling, and the process leading up to what he gave me.
+%You have to then describe the stages of cleaning I did.
+
+\section{Correlation analyses to help select autoregressive model lags}
+
+%***DAN: Describe the surrogating process
+%Probably you only need to include the results in allcorfsgtd_RE, in the Results section. 
+%See line 480 and above in AnalysisStep_LinearModels.R
+
 \end{document}

### 2021-07-29 Code/WhichBreak.R
added a new utility function

@@ -1,0 +1,34 @@
+#Takes a set of breaks as in "hist" and a set of values and returns a set of indices of 
+#the same length as vals indicating between which two breaks each value is.
+#
+#Args
+#breaks         A set of breaks, as for the function hist - evenly spaced increasing values
+#vals           A set of values, assumed to be >breaks[1] and <=breaks[length(breaks)]
+#
+#Output
+#A vector of the same length as vals with numbers between 1 and length(breaks)-1 indicating
+#which of the breaks the corresponding value fell between.
+#
+WhichBreak<-function(breaks,vals)
+{
+  if (!isTRUE(all.equal(diff(breaks),rep(breaks[2]-breaks[1],length(breaks)-1))))
+  {
+    stop("Error in WhichBreak: breaks must have evenly spaced values")
+  }
+  if (breaks[2]-breaks[1]<=0)
+  {
+    stop("Error in WhichBreak: breaks must have increasing values")
+  }
+  if (any(vals<=breaks[1]) || any(vals>breaks[length(breaks)]))
+  {
+    stop("Error in WhichBreak: some entries of vals fall outside the range of breaks")
+  }
+  
+  res<-NA*numeric(length(vals))
+  for (counter in 1:length(vals))
+  {
+    res[counter]<-max(which(breaks<vals[counter]))
+  }
+  
+  return(res)
+}

### 2021-07-29 Code/tests_WhichBreak.R

@@ -1,0 +1,13 @@
+context("WhichBreak.R")
+
+source("WhichBreak.R")
+
+test_that("test it",{
+  #test the error catching
+  testthat::expect_error(WhichBreak(c(1,2,4),rep(.5,5)),"Error in WhichBreak: breaks must have evenly spaced values")
+  testthat::expect_error(WhichBreak(c(3,2,1),rep(.5,5)),"Error in WhichBreak: breaks must have increasing values")
+  testthat::expect_error(WhichBreak(c(1,2,3),c(-1,2,2.5)),"Error in WhichBreak: some entries of vals fall outside the range of breaks")
+  
+  #test a simple case
+  testthat::expect_equal(WhichBreak(c(0,1,2,3,4),c(.1,1.1,2,3,4)),c(1,2,2,3,4))
+})

### 2021-07-29 Code/plotColLine.R
added another utility function

@@ -1,0 +1,29 @@
+#Plots a line, but with shifting colors.
+#
+#Args
+#x, y       Vectors of the same length to plot
+#cols       Same length as x and y, colors to use
+#lnwid      Line width to use      
+#...        Passed to "plot" command, used for controlling axis limits and the like
+#
+#Notes
+#Makes a line plot using the colors on the corresponding parts of the plot.
+#
+plotColLine<-function(x,y,cols,lnwid,...)
+{
+  lx<-length(x)
+  
+  plot(x,y,type="n",...)
+  rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4],col="grey90")
+  
+  lines(c(x[1],mean(x[1:2])),c(y[1],mean(y[1:2])),lwd=lnwid,col=cols[1])
+  for (counter in 2:(length(x)-1))
+  {
+    lines(c(mean(c(x[counter-1],x[counter])),x[counter]),
+          c(mean(c(y[counter-1],y[counter])),y[counter]),lwd=lnwid,col=cols[counter])
+    lines(c(x[counter],mean(c(x[counter],x[counter+1]))),
+          c(y[counter],mean(c(y[counter],y[counter+1]))),lwd=lnwid,col=cols[counter])
+  }
+  lines(c(mean(x[(lx-1):lx]),x[lx]),c(mean(y[(lx-1):lx]),y[lx]),lwd=lnwid,col=cols[lx])
+}
+

### 2021-07-29 Code/plottests_plotColLine.R

@@ -1,0 +1,10 @@
+#This tests the function in plotColLine.R, but since that's a plotting function this script is
+#not covered by the automated testing through the testthat package, which is why the file name
+#does not start with "tests".
+
+source("plotColLine.R")
+
+x<-1:5
+y<-c(0,1,0,1,0)
+cols<-c("red","blue","green","yellow","purple")
+plotColLine(x,y,cols,lnwid=2)

### 2021-07-29 Code/AnalysisStep_SpectralDecomp.R
working on getting a complete set of the main plots for the paper

@@ -8,10 +8,12 @@ rm(list=ls())
 #External codes needed
 #***
 
-#packages needed (invoked with "::"): parallel, wsyn, ncf, graphics
+#packages needed (invoked with "::"): parallel, wsyn, ncf, graphics, latex2exp
 source("SpectralTools.R")
 source("AvgOffDiags.R")
 source("PlotPhaseFunc.R")
+source("WhichBreak.R")
+source("plotColLine.R")
 
 #***
 #Location for storing results and other prep
@@ -51,13 +53,13 @@ lents<-dim(kelp)[2]
 #***
 
 #from google maps
-MontBLatLon<-c(36.62393626182818, -121.8511283749223)
-CarBLatLon<-c(36.52316357182972, -121.95535513873469)
-MorBLatLon<-c(35.427553615598676, -120.89133343717805)
-PtConcLatLon<-c(34.44830424836278,-120.47125509005721)
-SBLatLon<-c(34.407876790966135, -119.68428533917637)
-OxLatLon<-c(34.2310543912163, -119.2726684255394)
-LALatLon<-c(34.00907538429378, -118.50422514984244)
+MontBLatLon<-c(36.62393626182818, -121.8511283749223) #Monterey Bay, closer to the southern end
+CarBLatLon<-c(36.52316357182972, -121.95535513873469) #The point to the south of Carmel Bay
+MorBLatLon<-c(35.427553615598676, -120.89133343717805) #Morro Bay, closer to the northern end
+PtConcLatLon<-c(34.44830424836278,-120.47125509005721) #Pt Conception
+SBLatLon<-c(34.407876790966135, -119.68428533917637) #Santa Barbara pier
+OxLatLon<-c(34.2310543912163, -119.2726684255394) #Oxnard, where the Santa Clara River empties
+LALatLon<-c(34.00907538429378, -118.50422514984244) #LA, near the Santa Monica pier
 
 dists<-ncf::gcdist(x=c(MontBLatLon[2],locs$Lon),y=c(MontBLatLon[1],locs$Lat))
 dists<-dists[1,2:(dim(dists)[2])]
[…]

### 2021-07-29 Code/plottests_PlotPhaseFunc.R

@@ -1,5 +1,5 @@
-#This tests the function in PLotPhaseFunc.R, but since that's a plotting function this script is
-#not covered by the automated testing through the testthat package, which is while the file name
+#This tests the function in PlotPhaseFunc.R, but since that's a plotting function this script is
+#not covered by the automated testing through the testthat package, which is why the file name
 #does not start with "tests".
 
 source("PlotPhaseFunc.R")

### 2021-07-29 Code/AnalysisStep_SpectralDecomp.R
finished the code for making the main results figures in the main text

@@ -2333,21 +2333,21 @@ PlotForMSArgs<-list(cov_MainPlot_PanLabs=c("A","B","C"),
                     ExplainDirectMoran_fB_mid_ylims=c(0.8,1.65),
                     ExplainDirectMoran_fB_long_ylims=c(1.2,3),
                     ExplainDirectMoran_PanLabs=LETTERS[1:15],
-                    ExplainInteractingMoran_fP1_short_ylims=c(NA,NA),
-                    ExplainInteractingMoran_fP1_mid_ylims=c(NA,NA),
-                    ExplainInteractingMoran_fP1_long_ylims=c(NA,NA),
-                    ExplainInteractingMoran_fP2_short_ylims=c(NA,NA),
-                    ExplainInteractingMoran_fP2_mid_ylims=c(NA,NA),
-                    ExplainInteractingMoran_fP2_long_ylims=c(NA,NA),
-                    ExplainInteractingMoran_fP1ConjfP2_short_ylims=c(NA,NA),
-                    ExplainInteractingMoran_fP1ConjfP2_mid_ylims=c(NA,NA),
-                    ExplainInteractingMoran_fP1ConjfP2_long_ylims=c(NA,NA),
-                    ExplainInteractingMoran_Se1e2_short_ylims=c(NA,NA),
-                    ExplainInteractingMoran_Se1e2_mid_ylims=c(NA,NA),
-                    ExplainInteractingMoran_Se1e2_long_ylims=c(NA,NA),
-                    ExplainInteractingMoran_Prod_short_ylims=c(NA,NA),
-                    ExplainInteractingMoran_Prod_mid_ylims=c(NA,NA),
-                    ExplainInteractingMoran_Prod_long_ylims=c(NA,NA),
+                    ExplainInteractingMoran_fP1_short_ylims=c(0.064,0.15),
+                    ExplainInteractingMoran_fP1_mid_ylims=c(0.0716,0.117),
+                    ExplainInteractingMoran_fP1_long_ylims=c(0.0725,0.1125),
+                    ExplainInteractingMoran_fP2_short_ylims=c(0.12,0.25),
+                    ExplainInteractingMoran_fP2_mid_ylims=c(0.12,0.25),
+                    ExplainInteractingMoran_fP2_long_ylims=c(0.12,0.25),
+                    ExplainInteractingMoran_fP1ConjfP2_short_ylims=c(0.014,0.032),
[…]

### 2021-07-29 Paper/Paper.Rnw
started writing theory section of the main text

--- 2021-03-19 Paper/Paper.Rnw
@@
 \newcommand{\mcB}{\mathcal{P}}
+\newcommand{\var}{\text{var}}
+\newcommand{\cov}{\text{cov}}
 
@@
 
-%***Tentative author order
+%Alternative titles
+%Interactions among Moran effects
+%Interacting Moran effects
+%Interactions between environmental drivers of spatial synchrony
+
+%***Tentative author order, who is last yet to be determined
 \author[a,b]{Daniel C. Reuman}
@@
 \author[c]{Jonathan Walter}
-\author[g]{Lawrence W. Sheppard}
+\author[a,g]{Lawrence W. Sheppard}
 
@@
 \affil[b]{Laboratory of Populations, Rockefeller University}
-\affil[c]{University of Virginia}
-\affil[d]{WHOI}
-\affil[e]{UCSB}
-\affil[f]{UCLA}
-\affil[g]{MBA}
+\affil[c]{Department of Environmental Sciences, University of Virginia}
+\affil[d]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
+\affil[e]{Earth Research Institute, University of California, Santa Barbara}
+\affil[f]{Department of Geography, University of California, Los Angeles}
+\affil[g]{Marine Biological Association of the United Kingdom}
 
@@
 
+%Tentative paragraph outline written on my phone on 3/20/2021
+%1) Importance of synchrony
+%2) Classic problem of identifying causes of synchrony, but recently largely resolved due to new methods
+%  -elaborate briefly on a couple methods
+%3) in fact, it has recently been discovered that environmental causes of synchrony can interact, so that ...
+%  -related to, but distinct from, interaction between environmental and dispersal causes of synchrony discussed, for instance by Kendal et al
+%  -make sure you describe the two cases, cite the earlier papers, mention a few words about the two systems
+%4) the simplest form of interacting moran effects can be demonstrated when a simple model, though elaboration of the true nature of this mechanism %will take the rest of this manuscript
+%  -use Moran model
+%  -but with two drivers
+%  -moran theorem says correlation of pops equals correlation of sum of drivers, so work it out
+%5) general idea of two noises lagged or time delayed relative to each other and their effects oppositely lagged - this may be where a pedagogical %fig comes in, or maybe it is just worsd here and then the pedagog fig comes in later, at end of theory section, in form of examples.
+%6) paragraph on why we would expect interacting moran effects to be common, return to this point in the discussion
+%7) final para on goals, manuscript structure, etc.
+%  -G1) Elaborate a mathematical/statistical theory of interacting Moran effects which both allows both a detailed understanding 
+%      of how the mechanism works and fecilitates statistical investigation of whether the effect is present in empirical systems.
+%  -G2) Apply the theory and associated stats to revisiting kelp, the intention being to further elaborate the mechanism itself
+%  -G3) Apply the theory and associated stats to plankton, the intention being to further elaborate the mechanism itself
+%
+%Notes:
+%1) You need to make it clear the earlier papers already identified interacting Moran effects, so what are we doing?
+%We are exploring theoreitically what the mechanism is, here, and how general it might be, and what different forms it 
+%might take, using the theory, and also using the theory to further illuminate the two examples. The emphasis here
+%is on illuminating the general mechanism of interacting Moran effects through the theory and its application, whereas
+%the earlier papers focussed on the specific cases. The Fourier theory here serves to illuminate the general mechanism,
+%whereas the earlier papers used wavelet statistics applied directly to data and did not really explore the general 
+%mechanism. Need to work some of these thoughts in without seeming defensive.
+
 \section{Theory}
+
+%Tentative broad outline:
+%DONE 1) Intro spectral quantities, the spectrum, the cospectrum, and the cross spectrum of two time series/processes. And the spectral
+%matrix for putting them all in one structure.
+%2) Define your measures of timescale/frequency-specific synchrony and crosss-variable synchrony, explain why these make sense as
+%ways to measure those things.
+%3) Present the model and the theory
+%4) Present "toy" examples and fig 1 (need to think a lot about these)
+%
+%Notes:
+%Compared to the deer paper, this is a much deeper presentation of the theory in the main text. I'm not sure that's avoidable.
+%For one thing, the theory is explicitly applied to get the results, later, and that was not the case for the deer paper.
+
+\noindent Our main theoretical work requires a conceptual understanding of the \emph{spectrum} of an environmental 
+or population time series, as well as of the \emph{cospectrum} and \emph{cross spectrum} of two time series.
+If $w_i(t)$ is a stochastic process or time series measured at location $i$, the spectrum $S_{w_i w_i}(f)$ is a function
+of frequency, $f$. For a periodic ocillation, $f$ can be defined as one over the timescale, or period, of the oscillation. 
+The spectrum, $S_{w_i w_i}(f)$ is larger for frequencies at which $w_i(t)$ oscillates more. So, for example, a 
+population $w_i(t)$ that shows strong oscillatory dynamics at $5$-year period will have a large value of 
+$S_{w_i w_i}(f)$ for $f=1/5 \text{y}^{-1}$. The spectrum separates the overall variance of a time series into
+contributions which occur on different timescales, in the sense that an integral of $S_{w_i w_i}(f)$ across all 
+frequencies equals $\var(w_i)$. In a similar way, the cospectrum of two time series, $w_i(t)$ and $w_j(t)$,
+is a function of $f$ that takes large values at frequencies for which oscillations in $w_i(t)$ and 
+$w_j(t)$ are strong, and are strongly correlated, i.e., they have the same or similar phase. 
+(Here, $j$ refers to another location where measurements were performed.) The cospectrum is 
+the real part of the cross spectrum, $S_{w_i w_j}(f)$, which is a complex-valued function of frequency.
+The cross-spectrum takes large-magnitude values at frequencies, $f$, for which the oscillatory components
+of $w_i$ and $w_j$ are strong; and the complex phase of the cross spectrum at $f$ then quantifies
+the phase lag between those oscillations. The notation $S_{ww}$ refers to the \emph{spectral matrix}, a 
+matrix for which the $ij^{\text{th}}$ entry is $S_{w_i w_j}$. 
+Spectral methods are standard tools in many fields of science, 
+including ecology \citep{Vasseur2007,Reuman2008,Defriez2017}, and many background references are available 
+[e.g., \cite{Brillinger2001}].
 
@@
 
+\bibliographystyle{ecology_letters2}
+\bibliography{refs}
+
 \end{document}

### 2021-07-30 Paper/Paper.Rnw
bit more progress on the theory section of main text

@@
 %Interactions between environmental drivers of spatial synchrony
+%Environmental drivers of spatia synchony can strongly interact
 
@@
 %The following is where you load in the numeric results that will be embedded in the text
-<<eval=T,echo=F,message=F,warning=F>>=
-#place R code here for loading in necessary variables from Results
-@
+%<<eval=T,echo=F,message=F,warning=F>>=
+%#place R code here for loading in necessary variables from Results
+%@
 
@@
 %2) Classic problem of identifying causes of synchrony, but recently largely resolved due to new methods
-%  -elaborate briefly on a couple methods
+%  -elaborate briefly on a couple methods, mention the importance of a timescale-resolved appproach
 %3) in fact, it has recently been discovered that environmental causes of synchrony can interact, so that ...
@@
 %matrix for putting them all in one structure.
-%2) Define your measures of timescale/frequency-specific synchrony and crosss-variable synchrony, explain why these make sense as
+%DONE 2) Define your measures of timescale/frequency-specific synchrony and crosss-variable synchrony, explain why these make sense as
 %ways to measure those things.
@@
 or population time series, as well as of the \emph{cospectrum} and \emph{cross spectrum} of two time series.
-If $w_i(t)$ is a stochastic process or time series measured at location $i$, the spectrum $S_{w_i w_i}(f)$ is a function
+If $w_i(t)$ is a stochastic process or time series measured at location $i$ (e.g., the population density time series
+of a species of interest in that location), the spectrum $S_{w_i w_i}(f)$ is a function
 of frequency, $f$. For a periodic ocillation, $f$ can be defined as one over the timescale, or period, of the oscillation. 
-The spectrum, $S_{w_i w_i}(f)$ is larger for frequencies at which $w_i(t)$ oscillates more. So, for example, a 
-population $w_i(t)$ that shows strong oscillatory dynamics at $5$-year period will have a large value of 
-$S_{w_i w_i}(f)$ for $f=1/5 \text{y}^{-1}$. The spectrum separates the overall variance of a time series into
-contributions which occur on different timescales, in the sense that an integral of $S_{w_i w_i}(f)$ across all 
+The spectrum $S_{w_i w_i}(f)$ is larger for frequencies at which $w_i(t)$ oscillates more. So, for example, a 
+population $w_i(t)$ that shows strong oscillatory dynamics of $5$-year period will have a large value of 
+$S_{w_i w_i}(f)$ for $f=1/5 \text{ y}^{-1}$. The spectrum separates the overall variance of a time series into
+contributions which occur on different frequencies, in the sense that an integral of $S_{w_i w_i}(f)$ across all 
 frequencies equals $\var(w_i)$. In a similar way, the cospectrum of two time series, $w_i(t)$ and $w_j(t)$,
 is a function of $f$ that takes large values at frequencies for which oscillations in $w_i(t)$ and 
-$w_j(t)$ are strong, and are strongly correlated, i.e., they have the same or similar phase. 
+$w_j(t)$ are both strong and strongly correlated, i.e., they have the same or similar phase. 
 (Here, $j$ refers to another location where measurements were performed.) The cospectrum is 
@@
 The cross-spectrum takes large-magnitude values at frequencies, $f$, for which the oscillatory components
-of $w_i$ and $w_j$ are strong; and the complex phase of the cross spectrum at $f$ then quantifies
-the phase lag between those oscillations. The notation $S_{ww}$ refers to the \emph{spectral matrix}, a 
-matrix for which the $ij^{\text{th}}$ entry is $S_{w_i w_j}$. 
-Spectral methods are standard tools in many fields of science, 
-including ecology \citep{Vasseur2007,Reuman2008,Defriez2017}, and many background references are available 
+of $w_i(t)$ and $w_j(t)$ are both strong; and the complex phase of the cross spectrum at $f$ then quantifies
+the phase lag between those components. The notation $S_{ww}(f)$ refers to the \emph{spectral matrix}, 
+the $ij^{\text{th}}$ entry of which is $S_{w_i w_j}(f)$. Spectral methods are standard tools in many fields, 
+including ecology [e.g., \cite{Vasseur2007,Reuman2008,Defriez2017}], and many background references are available 
 [e.g., \cite{Brillinger2001}].
+
+Much prior work demonstrates the importance of a frequency-specific approach to
+synchrony [e.g., \cite{Vasseur2007,Keitt2008,Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}], 
+and it will turn out (see below, and \emph{Discussion}) that a frequency-specific 
+approach is essential to the development of our new theory. 
+%***DAN: Good topic for a Discussion para, the fact that this could not really be done without taking frequency specificity into account, and why, though need to think a bit more before writing
+We therefore here define, using the spectral methods outlined above, a frequency-specific measure of spatial synchrony, 
+as well as a new concept of \emph{cross-variable spatial synchrony}. If time series data $w_i(t)$ for $t=1,\ldots,T$ 
+were gathered at locations $i=1,\ldots,N$, our spatial synchrony measure is simply
+$\rho_{ww}=\left( \sum_{i \neq j} S_{w_i w_j} \right)/(N^2-N)$, the average of the cross spectra for all pairs
+of distinct locations. This is a real-valued function of frequency, the integral (across frequencies) of which
+is the classic, non-frequency-specific measure $\left( \sum_{i \neq j} \cov(w_i,w_j) \right)/(N^2-N)$ (see SI section
+\ref{SM-Sect:measuring_sync} for details, here). If two time series $\epsilon_i^{(1)}(t)$ and $\epsilon_i^{(2)}(t)$
+($t=1,\ldots,T$ and $i,j=1,\ldots,N$) were measured at each sampling location (e.g., two environmental variables),
+cross-variable spatial synchrony (or, simply, \emph{cross synchrony}) between the variables
+is $\rho_{\epsilon^{(1)}\epsilon^{(2)}}=\left( \sum_{i \neq j} S_{\epsilon_i^{(1)}\epsilon_j^{(2)}} \right)/(N^2-N)$. 
+This is conceptually interpretable as pertaining to spatial synchrony because it makes comparisons across 
+distinct locations, i.e., $i \neq j$. 
+It is interpretable as cross-variable synchrony because comparisons are between the two variables, i.e., cross spectra 
+between time series components of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are used. The new index 
+takes into account possible time lags. For instance, if both $\epsilon^{(1)}$ and $\epsilon^{(2)}$ show strong $4$-year-period 
+oscillations, but peaks in the $\epsilon^{(2)}$ oscillations consistently lag peaks in the $\epsilon^{(1)}$ oscillations 
+by a year, then $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$, which is complex valued, will have high magnitude at
+frequency $f=1/4 \text{ y}^{-1}$, and will have  phase at that frequency equal to $\pi/2$, reflecting the one-year 
+lag between the variables.
+%***DAN: make sure I'm not making a sign error here when you create the additional examples for sup mat
+See SI section \ref{SM-Sect:measuring_sync} for further examples and explanations of the new cross synchrony index.
 

### 2021-07-30 Paper/SupMat.Rnw
started writing theory section of the main text
bit more progress on the theory section of main text

@@
 \newcommand{\Rp}{\text{Re}}
+\newcommand{\var}{\text{var}}
 \newcommand{\cov}{\text{cov}}

### 2021-08-10 Paper/Paper.Rnw
mostly outlining of Methods inthe main text

@@
 \newcommand{\mcP}{\mathcal{P}}
-\newcommand{\mcB}{\mathcal{P}}
+\newcommand{\mcB}{\mathcal{B}}
 \newcommand{\var}{\text{var}}
 \newcommand{\cov}{\text{cov}}
+\newcommand{\Rp}{\text{Re}}
+\let\conj\overline
 
@@
 %5) general idea of two noises lagged or time delayed relative to each other and their effects oppositely lagged - this may be where a pedagogical %fig comes in, or maybe it is just worsd here and then the pedagog fig comes in later, at end of theory section, in form of examples.
-%6) paragraph on why we would expect interacting moran effects to be common, return to this point in the discussion
+Yes, I
+%think I want that second approach.
+%6) brief paragraph on why we would expect interacting moran effects to be common, return to this point in the discussion
 %7) final para on goals, manuscript structure, etc.
@@
 %      of how the mechanism works and fecilitates statistical investigation of whether the effect is present in empirical systems.
-%  -G2) Apply the theory and associated stats to revisiting kelp, the intention being to further elaborate the mechanism itself
+%  -G2) Apply the theory and associated stats to kelp, the intention being to further elaborate the mechanism itself
 %  -G3) Apply the theory and associated stats to plankton, the intention being to further elaborate the mechanism itself
-%
+%   Final sentence about how this paper conceptually and theoretically illuminates a new and probably common mechanism of 
+%   synchrony, and provides statistics which facilitate investigation of the mechanism in empirical systems.
+
 %Notes:
@@
 %ways to measure those things.
-%3) Present the model and the theory
+%DONE 3) Present the model and the theory
 %4) Present "toy" examples and fig 1 (need to think a lot about these)
@@
 %***DAN: make sure I'm not making a sign error here when you create the additional examples for sup mat
-See SI section \ref{SM-Sect:measuring_sync} for further examples and explanations of the new cross synchrony index.
+See SI section \ref{SM-Sect:measuring_sync} for further examples of the new cross synchrony index.
+
+Our model is 
+\begin{align}
+w_i(t) &= b_1 w_i(t-1)+ \cdots + b_n w_i(t-n) \label{main_model_1} \\ 
+&+ p_0^{(1)} \epsilon_i^{(1)}(t)+ \cdots + p_{m_1}^{(1)} \epsilon_i^{(1)}(t-m_1) \label{main_model_2} \\
+&+ p_0^{(2)} \epsilon_i^{(2)}(t)+ \cdots + p_{m_2}^{(2)} \epsilon_i^{(2)}(t-m_2) \label{main_model_3} \\
+&+ \delta_i(t), \label{main_model_4}
+\end{align}
+where $w_i(t)$ is an index of the population in location $i=1,\ldots,N$ at time
+$t$, and $\epsilon^{(1)}=(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)})$,
+$\epsilon^{(2)}=(\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)})$ and $\delta=(\delta_1,\ldots,\delta_N)$
+are environmental processes. The processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
+taken to be measured environmental processes, whereas $\delta$ represents the 
+aggregate influence of unmeasured processes. We assume that the combined process 
+$(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ is an ergodic second-order stationary stochastic process \citep{Brillinger2001}
+with expected values of its component processes equal to $0$. We make additional mild
+regularity assumptions to ensure the model is stable 
+(SI section \ref{SM-sect:model_spec}). This model can straightforwardly be seen as 
+a linearization of a very general dynamical model, influenced by ``weak noise'' [see, e.g.,
+SI section S1.2 of \cite{Desharnais2018} and SI section S1 of \citep{Walter2017}].
+Linearization and ``weak noise'' assumptions are tactics which have been made very commonly to make
+theoretical progress in ecology (REFS), and it has been demonstrated that results based on a
+weak noise assumption often also hold for quite strong noise (REF).
+%***DAN: Add refs later
+See \cite{Brillinger2001} for background on stochastic processes.
+
+The outcome of our theory is an equation which expresses population spatial synchrony 
+in terms of synchrony of the processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$, and cross
+synchrony between these processes:
+\begin{align}
+\rho_{ww} &= \frac{1}{|f_{\mcB}|^2}\left[|f_{\mcP^{(1)}}|^2 \rho_{\epsilon^{(1)} \epsilon^{(1)}} 
++ |f_{\mcP^{(2)}}|^2 \rho_{\epsilon^{(2)} \epsilon^{(2)}} 
++ 2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)\right] \label{rhoww_T1to3} \\
+&+ \text{other contributions}.\label{rhoww_OC}
+\end{align}
+Here, $f_{\mcB} = 1-b_1 \mu - b_2 \mu^2 - \cdots - b_n \mu^n$, 
+$f_{\mcP^{(1)}} = p_0^{(1)} + p_1^{(1)} \mu + \cdots + p_{m_1}^{(1)} \mu^{m_1}$, and
+$f_{\mcP^{(2)}} = p_0^{(2)} + p_1^{(2)} \mu + \cdots + p_{m_2}^{(2)} \mu^{m_2}$, where 
+$\mu=\exp(-2 \pi \iota f)$, and where $\iota$, the Greek letter iota, represents the imaginary unit.
+The ``other contributions'' in this equation correspond to synchronizing influences of 
+$\delta$ and interactions between $\delta$ and the $\epsilon^{(i)}$. If data 
+$w_i(t)$, $\epsilon^{(1)}_i(t)$ and $\epsilon^{(2)}_i(t)$ are available for locations 
+$i=1,\ldots,N$ and times $t=1,\ldots,T$, then all the terms on the right side of (\ref{rhoww_T1to3})
+can be estimated from these data (see \emph{Methods}), providing an assessment of the relative 
+contributions of direct and interacting Moran effects to population synchrony.
+
+%***DAN: There is still one thing left to do here, see above outline. This is a very important thing, because it is 
+%supposed to use examples to demo how the interaction terms work, bringing in the lags in effects and the lags in
+%the env vars themselves, and the interactions between these lags. This is where we take the next step with the 
+%core of the concept, after the basic intro to it in the Intro.
+%***DAN: Consider the possibility that you just want to use the same model as Plos Comp Biol, but analyze it using
+%the new theory. On the other hand, don't want to give the impression that not much is new about this paper.
 
 \section{Methods}
+
+%Tentative broad outline
+%1) Para from Tom on how remote sensing data are gathered and coverted into what he handed to me. Brief, mostly 
+%citing other works where it is laid out in much more detail. 
+%2) Para from Tom on how he got to the wave and NO3 data
+%3) Para summarizing the data cleaning I did
+%Comment) Actually, might want to send the abive 3 paras to sup mat, and just have one para describing the final format of the
+%data, especially if you make this paper also be about plankton. But do that later if you do it.
+%4) Some paras about the plankton data, similar structure to how the kelp data were presented (and using sup mat as described)
+%5) Para about the linear modelling - this is how certain parts of the main theoretical equation were estimated
+%6) Para about spectral estimates. Brief, cites appropriate parts of Brillinger, uses sup mat if needed. This is how certain
+%other parts of the main theoretical equation were estimated. 
+%Comment) I could write these now, but then if I end up having to do something slightly different for plankton, I'll have to 
+%rewrite them. Maybe now is the time to decide if I want this to be a kelp-only or a kelp-an-plankton paper and then do the
+%plankton work.
+%Comment) Somehow need to get across what was done - different parts of main equation were examined to explore the different
+%aspects of how the mechanism works for the focal taxa. Elaborate as appropriate.
+
+%Para 1 from Tom
+A process was previously developed (REFS) using remote sensing to provide quarterly kelp biomass and kelp canopy area estimates,
+from quarter 1 of 1987 to quarter 4 of 2019, at $437$ locations along the California coast, from slightly north of Monterey Bay
+to San Diego. We very briefly describe the process, refering to our earlier published works for details.
+%Need to mention: The spacing and extent of each site; where the data can be downloaded (other than in our git repo, though I'll
+%a mention of that, too); lots of references as necessary; probably need slightly more formal statements of what "kelp biomass" 
+%"kelp canopy area" mean in these contextx; anything else important. Need bibliographic data for cited papers.
+
+%Para 2 from Tom
+Another process was previously developed (REFS) to provide mean nitrate concentration and maximum wave height data
+for the same quarters and locations. We again briefly describe the process, citing earlier works for details.
+
+%I asked Tom for these paragraphs by email on 2021 08 05
+
+%Need:
+%1) Some kind of map of the locations and display of some time series for the sup mat, to give a sense of the data. I think Max will
+%have made one of these for his paper with Lawrence, so maybe I could get him to make another one, same or appropriately modified?
 

### 2021-08-10 Paper/SupMat.Rnw
mostly outlining of Methods inthe main text

@@
 
-\section{Model specification, full details}
+\section{Model specification, full details}\label{sect:model_spec}
 
@@
 We also assume that the 
-combined process $(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ is a second-order stationary stochastic process (REF)
-%***DAN: REFS here to Brillinger
+combined process $(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ is a second-order stationary stochastic process \citep{Brillinger2001}
 with the expected values of its component processes equal to $0$. Finally, we assume the complex roots of the
@@
 existence of a stationary stochastic process solution for the model and for invertability of
-the filter $\mcB$ which we will define below. See REF
-%***DAN: add refs to Brilliger, Shumway and Stoffer, Rinsel
+the filter $\mcB$ which we will define below. See \cite{Reinsel1997,Shumway2000,Brillinger2001}
 for background on the statistical theory of stationary stochastic processes.
@@
 to calculate the spectral matrix, $S_{ww}$, of the stochastic process, $w$, which we do in section \ref{sect:derivation}.
-
-%***DAN: Do you need to have some sort of section gtiving background info on spectral stuff, or is that all assumed?
-%Probably no, we just assume it here, and give some info in the main text on how to interpret these quantities, but
-%we do not have to give any definitions and so on because we are not going to give anything like a proper into
-%to stochastic processes and spectral approaches either in the main text or here, so better just to assume it 
-%(here) and take the attitude in the main text of "you won't understand where this comes from (see sup mat for that)
-%but we can tell you how to interpret it.
 
@@
 
-Letting $\mu=\exp(-2\pi i f)$ where $f$ is frequency in units of cycles per time steps, we have
+Letting $\mu=\exp(-2\pi \iota f)$ where $\iota$ is the imaginary unit and 
+$f$ is frequency in units of cycles per time steps, we have
 \begin{align}
@@
 
+\section{References}
+
+\bibliographystyle{ecology_letters2}
+\bibliography{refs}
+
 \end{document}

### 2021-11-04 Paper/Paper.Rnw
various changes, catching up on forgetting to commit

@@
 %Environmental drivers of spatia synchony can strongly interact
+%Environmental drivers of spatial synchrony interact
 
@@
 \begin{document}
+\SweaveOpts{concordance=TRUE}
 
@@
 %7) final para on goals, manuscript structure, etc.
-%  -G1) Elaborate a mathematical/statistical theory of interacting Moran effects which both allows both a detailed understanding 
-%      of how the mechanism works and fecilitates statistical investigation of whether the effect is present in empirical systems.
-%  -G2) Apply the theory and associated stats to kelp, the intention being to further elaborate the mechanism itself
-%  -G3) Apply the theory and associated stats to plankton, the intention being to further elaborate the mechanism itself
+%  -G1) Elaborate a mathematical theory of interacting Moran effects which allows a detailed understanding 
+%      of how the mechanism works 
+%  -G2) Consider theoretical case studies, which emerge as important special cases of our general theory, and which illuminate the intuition behind how Moran effects may interact in real systems. 
+%  -G3) Apply the theory to populations giant kelp of the California coast, the intention being to further elaborate the mechanism itself
 %   Final sentence about how this paper conceptually and theoretically illuminates a new and probably common mechanism of 
-%   synchrony, and provides statistics which facilitate investigation of the mechanism in empirical systems.
+%   synchrony.
 
@@
 %DONE 3) Present the model and the theory
-%4) Present "toy" examples and fig 1 (need to think a lot about these)
+%4) Present "toy" examples and fig 1 (need to think a lot about these).
+DAN: perhaps these actually belong in Methods (describe the specific-case models we will consider) and results (apply the general theory to those specific cases)
 %
@@
 \noindent Our main theoretical work requires a conceptual understanding of the \emph{spectrum} of an environmental 
-or population time series, as well as of the \emph{cospectrum} and \emph{cross spectrum} of two time series.
+or population time series, as well as of the \emph{cospectrum} and \emph{cross spectrum} of two time series,
+so we briefly introduce these concepts.
 If $w_i(t)$ is a stochastic process or time series measured at location $i$ (e.g., the population density time series
@@
 population $w_i(t)$ that shows strong oscillatory dynamics of $5$-year period will have a large value of 
-$S_{w_i w_i}(f)$ for $f=1/5 \text{ y}^{-1}$. The spectrum separates the overall variance of a time series into
+$S_{w_i w_i}(f)$ for $f=\frac{1}{5} \text{ y}^{-1}$. The spectrum separates the overall variance of a time series into
 contributions which occur on different frequencies, in the sense that an integral of $S_{w_i w_i}(f)$ across all 
@@
 \ref{SM-Sect:measuring_sync} for details, here). If two time series $\epsilon_i^{(1)}(t)$ and $\epsilon_i^{(2)}(t)$
-($t=1,\ldots,T$ and $i,j=1,\ldots,N$) were measured at each sampling location (e.g., two environmental variables),
+($t=1,\ldots,T$ and $i=1,\ldots,N$) were measured at each sampling location (e.g., two environmental variables),
 cross-variable spatial synchrony (or, simply, \emph{cross synchrony}) between the variables
 is $\rho_{\epsilon^{(1)}\epsilon^{(2)}}=\left( \sum_{i \neq j} S_{\epsilon_i^{(1)}\epsilon_j^{(2)}} \right)/(N^2-N)$. 
-This is conceptually interpretable as pertaining to spatial synchrony because it makes comparisons across 
+This is interpretable as pertaining to spatial synchrony because it makes comparisons across 
 distinct locations, i.e., $i \neq j$. 
@@
 between time series components of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are used. The new index 
-takes into account possible time lags. For instance, if both $\epsilon^{(1)}$ and $\epsilon^{(2)}$ show strong $4$-year-period 
-oscillations, but peaks in the $\epsilon^{(2)}$ oscillations consistently lag peaks in the $\epsilon^{(1)}$ oscillations 
+takes into account possible time lags. For instance, if both $\epsilon^{(1)}$ and $\epsilon^{(2)}$ show strong,
+spatially synchronous, $4$-year-period 
+oscillations, but peaks in the $\epsilon^{(1)}$ oscillations consistently lag peaks in the $\epsilon^{(2)}$ oscillations 
 by a year, then $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$, which is complex valued, will have high magnitude at
-frequency $f=1/4 \text{ y}^{-1}$, and will have  phase at that frequency equal to $\pi/2$, reflecting the one-year 
+frequency $f=\frac{1}{4} \text{ y}^{-1}$, and will have  phase at that frequency equal to $\pi/2$, reflecting the one-year 
 lag between the variables.
-%***DAN: make sure I'm not making a sign error here when you create the additional examples for sup mat
-See SI section \ref{SM-Sect:measuring_sync} for further examples of the new cross synchrony index.
+See SI section \ref{SM-sect:examples_cross_sync} for further examples of the new cross synchrony index.
 
@@
 a linearization of a very general dynamical model, influenced by ``weak noise'' [see, e.g.,
-SI section S1.2 of \cite{Desharnais2018} and SI section S1 of \citep{Walter2017}].
-Linearization and ``weak noise'' assumptions are tactics which have been made very commonly to make
+SI section S1.2 of \cite{Desharnais2018} and SI section S1 of \cite{Walter2017}].
+Linearization and ``weak noise'' assumptions are tactics which have been used very commonly to make
 theoretical progress in ecology (REFS), and it has been demonstrated that results based on a
-weak noise assumption often also hold for quite strong noise (REF).
+weak noise assumption often also hold for noise which is quite strong (REF).
 %***DAN: Add refs later
@@
 
-The outcome of our theory is an equation which expresses population spatial synchrony 
-in terms of synchrony of the processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$, and cross
-synchrony between these processes:
+Fulfilling goal G1 of the Introduction (i.e., to elaborate theory of interacting Moran effects), 
+the outcome of our theory is an equation which expresses population spatial synchrony 
+in terms of synchrony of the environmenal processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$, and cross
+synchrony between those processes:
 \begin{align}
@@
 $\mu=\exp(-2 \pi \iota f)$, and where $\iota$, the Greek letter iota, represents the imaginary unit.
-The ``other contributions'' in this equation correspond to synchronizing influences of 
-$\delta$ and interactions between $\delta$ and the $\epsilon^{(i)}$. If data 
-$w_i(t)$, $\epsilon^{(1)}_i(t)$ and $\epsilon^{(2)}_i(t)$ are available for locations 
-$i=1,\ldots,N$ and times $t=1,\ldots,T$, then all the terms on the right side of (\ref{rhoww_T1to3})
-can be estimated from these data (see \emph{Methods}), providing an assessment of the relative 
-contributions of direct and interacting Moran effects to population synchrony.
-
-%***DAN: There is still one thing left to do here, see above outline. This is a very important thing, because it is 
-%supposed to use examples to demo how the interaction terms work, bringing in the lags in effects and the lags in
-%the env vars themselves, and the interactions between these lags. This is where we take the next step with the 
-%core of the concept, after the basic intro to it in the Intro.
-%***DAN: Consider the possibility that you just want to use the same model as Plos Comp Biol, but analyze it using
-%the new theory. On the other hand, don't want to give the impression that not much is new about this paper.
+The first term on the right of (\ref{rhoww_T1to3}) is the component of population spatial synchrony 
+due to the direct Moran effects of $\epsilon^{(1)}$, the second term is 
+the component of population synchrony due to the direct Moran effects of $\epsilon^{(2)}$, 
+and the third term is the component due to interactions between the Moran effects of the 
+two environmental drivers.
+The ``other contributions'' in the above equations correspond to synchronizing influences of 
+$\delta$ and of interactions between $\delta$ and the $\epsilon^{(i)}$. Such contributions would be
+difficult to assess because $\delta$ was unmeasured. Comparing the three terms on the right
+of (\ref{rhoww_T1to3}) gives the relative contributions of direct Moran effects and interactions 
+between Moran effects. The detailed derivation of the theory is in SI section \ref{SM-sect:derivation}.
 
 \section{Methods}
+
+\noindent We now describe how we pursue goal G2 of the Introduction, to develop 
+theoretical case studies that illuminate the intuition of interacting Moran effects. 
+We develop case studies A-C. For case study A,  the environmental variable 
+$\epsilon^{(1)}$ is assumed to have a simple positive effect on populations, but 
+lagged by one model time step, i.e.,
+$p_0^{(1)}=0$, $p_1^{(1)}>0$, and $p_i^{(1)}=0$ for $i>1$. 
+For case study B, which generalizes A, $\epsilon^{(1)}$ is again assumed to have a simple 
+positive effect on populations,
+but now lagged by $n$ time steps, so that $p_n^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n$.
+For both A and B, the effects of $\epsilon^{(2)}$
+are assumed to be un-lagged, i.e., $p_0^{(2)} \neq 0$ and $p_i^{(2)}=0$ for $i>0$.
+For case study C, the effects of $\epsilon^{(1)}$ are lagged by $n_1$ and the 
+effects of $\epsilon^{(2)}$ are lagged by $n_2$, i.e.,
+$p_{n_1}^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n_1$, and 
+$p_{n_2}^{(2)}>0$ while $p_i^{(2)}=0$ for $i \neq n_2$.
+The noise process
+$(\epsilon^{(1)},\epsilon^{(2)})$ is assumed to be a Gaussian white-noise process 
+for both A and B, i.e., 
+the random variables $(\epsilon^{(1)}(t),\epsilon^{(2)}(t))$ are independent and 
+identically distributed (iid) for distinct times, $t$. Noise may be correlated across space,
+and the components of $\epsilon^{(1)}(t)$ may be correlated with those of $\epsilon^{(2)}(t)$.
+For C, the noise processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
+each assumed to exhibit spatially synchronous periodic oscillations of the same period
+but potentially different phases. Such a scenario could be realized by seasonally 
+periodic environmental fluctuations if the model time step were quarterly or monthly,
+e.g., kelp wave action in central California peaks annually in the winter, whereas surface-water
+nitrate concentrations also fluctuate with period one year, but peak in the spring. 
+Details of the noise processes are specified in SI section \ref{SM-sect:examples_cross_sync}.
+For all case studies,
+the autoregressive order of population dynamics is assumed to be $1$, i.e, $b_1 \neq 0$ 
+but $b_i=0$ for $i>1$.
 
@@
 
+% What do we learn from case study A:
+% 
+% Warmup:
+% 1) interactions between Moran effects (ref to specific term of the theory) can easily be of comparable importance
+% to the main Moran effects
+% 2) interaction effects can be positive or negative, and can differ, and even differ in sign, by timescale
+% 
+% Actually breaking it down:
+% 1) At the highest frequencies (shortest timescales) of 0.5 cycles per sampling 
+% interval (period 2 sampling intervals), a lag of 1 
+% sampling interval in the population effects of \epsilon^{(1)} is half the period of oscialltion, so 
+% causes the effects of \epsilon^{(1)}
+% on those timescales to be exactly out of phase of the effects of \epsilon^{(2)}, and so we get 
+% cancellation, i.e., negative interaction effects. The effects of \epsilon1 are multiplied by 1.1,
+% those of \epsilon2 by 0.8, and the correlation between the two (which is made to have negative effects
+% by the lag) is 0.3. So the negative interaction effects are 1.1*0.8*0.3*2, where the last factor of
+% two is because there are actually two effects here, the one where \epsilon1 is operating, lagged, at i and 
+% \epsilon2 is operating at j; and the effect where \epsilon1 is operating, lagged, at j and \epsilon2 is 
+% operating at i. 
+% 2) As timescales get to 1 year = 4 quarters, a lag of 1 quarter in the population effects of 
+% \epsilon1 puts the population effects of \epsilon1 3/4 out of phase with those of \epsilon 2,
+% and the correlation of two sin waves of the same frequency that are 3/4 out of phase is 0, so 
+% interaction effects are 0 here.
+% 3) As timescales get long, a lag of 1 quarter in the population effects of \epsilon1 is a
+% negligible time compared to the timescale being considered, so the effects of 
+% \epsilon1 and \epsilon2 on populations are closer and closer to be in phase and correlated
+% as timescales get longer and longer. The effects of \epsilon1 are multiplied by 1.1, those of 
+% \epsilon2 by 0.8, and the correlation is 0.3, so the interaction effect is 1.1*.8*.3*2 (the 
+% reason for the 2 is explained above).
+% 
+% Need to put this into good writing
+% 
+% Need to carry out similar thoughts for case B and write them, too.
+
+%Thoughts on 2021 11 01
+% Topic sentence for first para on theory results:
+% Results confirm earlier empirical work (refs) that interaction effects between Moran drivers 1) can be comparable in strength to direct Moran effects; 2) can be synergistic or destructive; and 3) can depend strongly on timescale, even to the extent that the sign of interaction effects can differ by timescale.
+% 
+% Combine all three theory figs into one multi panel fig, change the order of panels again 
+% 
+% Subsequent paras will then break things down to understand how this stuff works
+% 
+% Possible topic sentence for subsequent para: Theory reveals, as illustrated by cases A and B, how interactions between Moran effects depend on the lags with which Moran drivers influence populations.
+Or theory reveals how interactions between Moran effects depend on the lags with which Moran drivers influence populations, and case study begins to illustrate how this works.
+Or something like that.
+% 
+%The also have a para for C
+%
+% Maybe get rid of the “confirm earlier work” phrase since it wasn’t necessarily explicitly listed in that way in earlier work
+
+%Para basically confirming earlier work, as a starter
+Theoretical results help demonstrate that interaction effects between Moran drivers: 1) can be 
+comparable in strength to direct Moran effects; 2) can be synergistic or destructive; and 3) 
+can depend strongly on timescale, even to the extent that the sign of interaction effects can 
+differ by timescale.
+
+%Current (2021 11 01) thoughts on figures:
+%1) Maybe a pedagog fig, not sure what this would look like
+%2) Combined theory fig, all 3 case studies
+%3) "main fig" for CC1 and SB systems
+%4-5) "explain direct effects" fig for CC1 and SB systems (2 figures, I think it'll have to be)
+%6-7) "explain indirect effects" figs for CC1 and SB systems (again, 2)
+%One too many figs here. What should be cut? Right now I don't know what the pedagog fig would be 
+%so that's the obvious one to cut. And then it will all fit, though the figs are really big and 
+%it seems like it is always good to have a pedagogical, warmup fig. If you do come up with a pedagog
+%fig idea, could maybe combine some figs to make it work, or put the pedagog content as the first
+%part of the theory fig? Don't worry about it for now.
+
+%***DAN: add case study C to this.
+\begin{figure}
+\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseA.pdf}\\
+\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseB.pdf}\\
+\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v01.pdf}\\
+\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v02.pdf}
+\caption{Theoretical case studies A (panels a-d), B (panels e-h) and C (i-p, two alternative sets of 
+parameters used for i-l and m-p, respectively). Left panels (a, e, i, m) show the terms on the right side
+of (\ref{rhoww_T1to3}). On those panels, the green line is 
+$\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}$, quantifying 
+population synchrony due to the direct Moran effects of $\epsilon^{(1)}$. The blue line
+is $\frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}$, quantifying 
+population synchrony due to the direct Moran effects of $\epsilon^{(2)}$.
+The dashed green-blue line is 
+$\frac{2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)}{|f_{\mcB}|^2}$, 
+which is population synchrony due to interacting Moran effects.
+%***DAN: More text here as we flesh out the presentation 
+Model parameters were: $N=5$ and $b_1=0.4$ for all case studies; 
+$p_1^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study A; 
+$p_3^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study B;
+$p_1^{(1)}=3$, $p_0^{(2)}=1.5$, $c_1=0$ and $c_2=-4/9$ for case study C;
+$v_{11}=v{22}=0.6$ and $v_{12}=0.3$ for case studies A and B;
+$w_{11}=w{22}=0.6$ and $w_{12}=0.3$ for case study C;
+$l_1=0$ and $l_2=1$ for case study C, panels i-l; and 
+$l_1=1$ and $l_2=0$ for case study C, panels m-p.
+See Methods and SI sections \ref{SM-sect:examples_cross_sync} and 
+\ref{SM-sect:details_case_studies} for the meaning of parameters and additional details.}\label{fig:TheoryCaseA}
+\end{figure}
+%#Case A
+%N<-5
+%ns<-c(.6,.6,.3)
+%p1<-c(0,1.1)
+%p2<-c(0.8)
+%b<-c(.4)
+% #Case B
+% N<-5
+% ns<-c(.6,.6,.3)
+% p1<-c(0,0,0,1.1)
+% p2<-c(0.8)
+% b<-c(.4)
+
 \section{Discussion}

### 2021-11-04 Paper/SupMat.Rnw
various changes, catching up on forgetting to commit

@@
 \usepackage{amsmath}
+\usepackage{amsthm}
 \usepackage{xr}
@@
 \let\conj\overline
+
+%set up the theorem environment
+\newtheorem{thm}{Theorem}
+\newtheorem{corl}[thm]{Corollary}
 
@@
 \tableofcontents
-\listoftables
+%\listoftables
 \listoffigures
@@
 To see that $\rho_{ww}$ is a real-valued function, note that $\conj{S_{ww}}=S_{ww}^\tau$, where the superscript 
-$\tau$ denotes matrix transpose (REF to Brillinger theorem 2.5.1 on p. 24).
-%***DAN: Don't forget to fill in this reference!
+$\tau$ denotes matrix transpose [\cite{Brillinger2001}, theorem 2.5.1 on p. 24].
 Therefore 
@@
 and an appropriate integral across frequencies of the cospectrum equals the covariance, 
-$\cov(w_i,w_j)$ (REF to some specific page/theorem in Brillinger). 
-%***DAN: Don't forget to add
+$\cov(w_i,w_j)$ \citep{Brillinger2001}.
 Thus an integral of $\rho_{ww}$ across frequencies equals 
@@
 frequency $f$ measures the time delay between frequency-$f$ oscillations of $\epsilon^{(1)}_i$ and 
-frequency-$f$ oscillations of $\epsilon^{(2)}_j$. The following examples should help clarify these features of
-the definition of cross synchrony.
+frequency-$f$ oscillations of $\epsilon^{(2)}_j$.
 
-%***DAN: See page 51 of your hand-written notes for descriptions of the examples, which you need to code up and 
-%make some figure or something and describe and report here.
+\section{Examples of cross synchrony}\label{sect:examples_cross_sync}
+
+The following examples should help clarify the properties of our definition of cross synchrony.
+For pedagogical clarity, we start with a white-noise example that does not illustrate the full value of the
+idea of cross synchrony, but which is simple; we subsequently move on to an example which is more
+complex but also more illustrative. These examples are also used in the theoretical case studies (main text).
+
+For the simple example, used in theoretical case studies A and B, we 
+assume $(\epsilon^{(1)}(t),\epsilon^{(2)}(t))$ are drawn, independently for each $t$, from a
+$2N$-dimensional multivariate normal distribution
+with mean $(0,...,0)$ and covariance matrix $\Sigma$, where the entries $\Sigma_{ij}$ of 
+$\Sigma$ are as follows. For $i=j$, $\Sigma_{ij}=1$. For $i \neq j$ and $1 \leq i,j \leq N$, 
+$\Sigma_{ij}$ equals a parameter $v_{11}$. For $i \neq j$ and $N+1 \leq i,j \leq 2N$, 
+$\Sigma_{ij}$ equals a second parameter $v_{22}$. The remaining entries of $\Sigma$ equal a 
+third parameter, $v_{12}$. The parameters $v_{11}$, $v_{22}$ and $v_{12}$ must be chosen to yield
+a positive definite matrix $\Sigma$.
+
+Now, having specified the noise process $(\epsilon^{(1)},\epsilon^{(2)})$, we derive the quantities 
+$\rho_{\epsilon^{(1)} \epsilon^{(1)}}$,
+$\rho_{\epsilon^{(2)} \epsilon^{(2)}}$, and 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$.
+For $a$ and $b$ equal to $1$ or $2$, the formula 
+$S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(f) = \sum_{u=-\infty}^\infty c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)
+\exp(-2 \pi \iota f u)$, where $c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)=\cov(\epsilon_i^{(a)}(t+u),\epsilon_j^{(b)}(t))$,
+follows from definitions of the spectrum and cross spectrum, given, for instance, in 2.5.2 on p. 23 of
+\cite{Brillinger2001}. The quantity 
+$c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(u)=\cov(\epsilon_i^{(a)}(t+u),\epsilon_j^{(b)}(t))$ is independent of 
+$t$ for any stationary process $(\epsilon^{(1)},\epsilon^{(2)})$, and in particular for the temporally iid process that 
+we assumed to operate here. Furthermore, because 
+$\epsilon_i^{(a)}(t+u)$ and $\epsilon_j^{(b)}(t)$ were assumed independent for $u \neq 0$, we have 
+\begin{equation}
+S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(f) = c_{\epsilon_i^{(a)}\epsilon_j^{(b)}}(0) =
+\begin{cases}
+1 & \text{if}\ a=b\ \text{and}\ i=j \\
+v_{11} & \text{if}\ a=b=1\ \text{and}\ i \neq j \\
+v_{22} & \text{if}\ a=b=2\ \text{and}\ i \neq j \\
+v_{12} & \text{if}\ a \neq b.
+\end{cases}
+\end{equation}
+Thus the spectra and cross spectra in this example are constant. It
+is then easy to see that $\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
+$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$, and 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$, based on the definitions of these 
+quantities given previously.
+
+For a more illustrative example, used in case study C, the noise process 
+$(\epsilon^{(1)},\epsilon^{(2)})$ is generated via a three-step process. 
+First, assume $(\tilde{\zeta}^{(1)}(t),\tilde{\zeta}^{(2)}(t))$ 
+are drawn, independently for each $t$, from a $2N$-dimensional multivariate normal distribution with mean $(0,\ldots,0)$
+and covariance matrix $\Psi$, where the entries $\Psi_{ij}$ of $\Psi$ are as follows. 
+For $i=j$, $\Psi_{ij}=1$. For $i \neq j$ and $1 \leq i,j \leq N$, 
+$\Psi_{ij}$ equals a parameter $w_{11}$. For $i \neq j$ and $N+1 \leq i,j \leq 2N$, 
+$\Psi_{ij}$ equals a second parameter $w_{22}$. The remaining entries of $\Psi$ equal a 
+third parameter, $w_{12}$. The parameters $w_{11}$, $w_{22}$ and $w_{12}$ must again be chosen to yield
+a positive definite matrix. Second, let $\zeta^{(1)}(t)=\tilde{\zeta}^{(1)}(t-l_1)$ and let
+$\zeta^{(2)}(t)=\tilde{\zeta}^{(2)}(t-l_2)$ for fixed integers $l_1$ and $l_2$. Finally,
+let $\epsilon_i^{(a)}(t)=c_1 \epsilon_i^{(a)}(t-1)+c_2 \epsilon_i^{(a)}(t-2)+\zeta_i^{(a)}(t)$
+for $i=1,\ldots,N$ and $a=1,2$. The second step above ensures that $\epsilon^{(1)}$ and $\epsilon^{(2)}$
+are related to each other in a time-delayed manner, and the third step ensures that 
+$\epsilon^{(1)}$ and $\epsilon^{(2)}$ can oscillate with a dominant periodic component, depending 
+on the values of $c_1$ and $c_2$ used.
+
+To compute $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ for this example, we first compute 
+$S_{\epsilon^{(1)}\epsilon^{(1)}}$, $S_{\epsilon^{(2)}\epsilon^{(2)}}$ and
+$S_{\epsilon^{(1)}\epsilon^{(2)}}$ using theorem \ref{thm:useful_thm} below.  
+Letting $\epsilon=(\epsilon^{(1)},\epsilon^{(2)})$ and $\zeta(t)=(\zeta^{(1)},\zeta^{(2)})$, we 
+have $\epsilon_i(t)=c_1 \epsilon_i(t-1)+c_2 \epsilon_i(t-2) + \zeta_i(t)$ for $i=1,\ldots,2N$. We then
+apply theorem \ref{thm:useful_thm} to get that $S_{\epsilon \epsilon}=\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 S_{\zeta \zeta}$.
+Considering upper-left $N \times N$ sub-matrices gives 
+$S_{\epsilon^{(1)} \epsilon^{(1)}}=\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 S_{\zeta^{(1)} \zeta^{(1)}}$.
+Considering lower-right $N \times N$ sub-matrices gives 
+$S_{\epsilon^{(2)} \epsilon^{(2)}}=\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 S_{\zeta^{(2)} \zeta^{(2)}}$.
+Considering upper-right $N \times N$ sub-matrices gives 
+$S_{\epsilon^{(1)} \epsilon^{(2)}}=\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 S_{\zeta^{(1)} \zeta^{(2)}}$.
+Because the $\zeta^{(1)}(t)$ are iid for distinct times $t$, we can again use the definitions of the spectrum and the
+cross spectrum and reasoning similar to that used above for the previous example to show that $S_{\zeta^{(1)} \zeta^{(1)}}$
+equals the upper-left $N \times N$ sub-matrix of $\Psi$, which is a matrix with $1$ in every diagonal entry and
+$w_{11}$ in all off-diagonal entries. Likewise, $S_{\zeta^{(2)} \zeta^{(2)}}$ equals the lower-right 
+$N \times N$ sub-matrix of $\Psi$, which has $1$ in all diagonal entries and $w_{22}$ in all off-diagonal entries.
+To get $S_{\zeta^{(1)} \zeta^{(2)}}$, we write 
+$S_{\zeta_i^{(1)}\zeta_j^{(2)}}(f)=\sum_{u=-\infty}^{\infty} c_{\zeta_i^{(1)}\zeta_j^{(2)}}(u) \exp(-2 \pi \iota f u)$.
+Here, 
+\begin{align}
+c_{\zeta_i^{(1)}\zeta_j^{(2)}}(u) &= \cov(\zeta_i^{(1)}(t+u),\zeta_j^{(2)}(t)) \\
+&= \cov(\tilde{\zeta}_i^{(1)}(t+u-l_1),\tilde{\zeta}_j^{(2)}(t-l_2)) \\
+&= 
+\begin{cases}
+w_{12} & \text{if}\ u-l_1=-l_2 \\
+0 & \text{otherwise}.
+\end{cases}
+\end{align}
+So $S_{\zeta_i^{(1)}\zeta_j^{(2)}}=w_{12} \exp(-2 \pi \iota f (l_1 - l_2))$. It then follows 
+straightforwardly that $\rho_{\epsilon^{(1)} \epsilon^{(1)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{11}$,
+$\rho_{\epsilon^{(2)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{22}$ and 
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{12} \exp(-2 \pi \iota f (l_1-l_2))$.
+The pre-factor $\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2$ in all these expressions reflects the autoregressive 
+manner in which the noises $\epsilon$ were defined. Autoregressive models of order $2$ are well known to be capable of
+producing oscillations for certain values of the parameters. For instance, we use $c_1=0$ and $c_2=-4/9$, values
+which cause the expression $\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2$ to have a single peak at
+frequency $0.25$ cycles per sampling interval (i.e., $1$ cycle per year). 
+The term $\exp(-2 \pi \iota f (l_1-l_2))$ in the expression for 
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ reflects that fact that cross variable synchrony for this example is lagged
+by an amount $l_1-l_2$. The lack of a similar term in the expressions for $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$
+or $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ reflects that fact that there are no lags built into within-variable
+synchrony in this example. Fig. \ref{fig:rhosForC} displays the quantities $\rho_{\epsilon^{(1)} \epsilon^{(1)}}$,
+$\rho_{\epsilon^{(2)} \epsilon^{(2)}}$, and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ for the two parameterizations
+of this noise that we use for theoretical case study C.
+
+\begin{figure}
+\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v01_rhos.pdf}\\
+\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v02_rhos.pdf}
+\caption[Spectral quantities for noise used in case study C]{The quantities $\rho_{\epsilon^{(1)} \epsilon^{(1)}}$,
+$\rho_{\epsilon^{(2)} \epsilon^{(2)}}$, and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ for the noise
+described in SI section \ref{sect:examples_cross_sync} and used for theoretical case study C. 
+Two alternative sets of parameters
+were used for panels a-c and d-f, respectively. See the caption of Fig. \ref{MT-fig:TheoryCaseA} in the main text
+for parameter values.}\label{fig:rhosForC}
+\end{figure}
+
+\begin{thm}\label{thm:useful_thm}
+If $\zeta(t)$ is an $N$-dimensional second-order stationary ergodic stochastic process with 
+component expected values $0$, if
+$\epsilon_i(t)=c_1 \epsilon_i(t-1)+\cdots+c_n \epsilon_i(t-n)+q_0 \zeta_i(t)+\cdots+q_m \zeta_i(t-m)$ for $i=1,\ldots,N$, 
+with $c_n \neq 0$ and $q_m \neq 0$, and if the complex roots of the polynomial
+$1-c_1 z - c_2 z^2 - \cdots c_n z^n = 0$ have modulus greater than $1$, then 
+$S_{\epsilon \epsilon}(f) = \left| \frac{\gamma}{1-\lambda} \right|^2 S_{\zeta \zeta}(f)$ where $\gamma=q_0+q_1 \mu+\cdots+q_m \mu^m$,
+$\lambda=c_1 \mu+c_2 \mu^2+\cdots+c_n \mu^n$, and $\mu=\exp(-2\pi \iota f)$, $\iota$ is the imaginary unit, and 
+$f$ is frequency in cycles per time step.
+\end{thm}
+\begin{proof}
+This theorem is identical to theorem 2 in SI section S1.5 of \cite{Anderson2021}, where a proof is also provided. 
+\end{proof}
 
@@
 Letting $\mu=\exp(-2\pi \iota f)$ where $\iota$ is the imaginary unit and 
-$f$ is frequency in units of cycles per time steps, we have
+$f$ is frequency in units of cycles per time step, we have
 \begin{align}
@@
 
+\section{Details of the theoretical case studies}\label{sect:details_case_studies}
+
+The noise processes $(\epsilon^{(1)},\epsilon^{(2)})$ used 
+in case studies A, B and C, which were only partly specified in the main text, are fully specified in 
+SI section \ref{sect:examples_cross_sync}. 
+We next derive, for our case studies, the values of the terms that appear in (\ref{MT-rhoww_T1to3}) of the main text.
+For case study A, parameter specifications presented in Methods of the main text straightforwardly imply that
+$f_{\mcB}=1-b_1 \mu$, $f_{\mcP^{(1)}}=p_1^{(1)} \mu$ and $f_{\mcP^{(2)}}=p_0^{(2)}$. In 
+SI section \ref{sect:examples_cross_sync}, we computed that 
+$\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
+$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$, and 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$.
+
+For case study B, details presented in Methods of the main text imply that 
+$f_{\mcB}=1-b_1 \mu$, $f_{\mcP^{(1)}}=p_n^{(1)} \mu^n$ and $f_{\mcP^{(2)}}=p_0^{(2)}$. 
+Because the noise process for B is taken to be the same as that for A, we again have 
+$\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
+$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$ and 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$.
+
+For case study C, again using details from Methods, $f_{\mcB}=1-b_1 \mu$, 
+$f_{\mcP^{(1)}}=p_{n_1}^{(1)} \mu^{n_1}$ and $f_{\mcP^{(2)}}=p_{n_2}^{(2)} \mu^{n_2}$. 
+In SI section \ref{sect:examples_cross_sync}, we computed that
+$\rho_{\epsilon^{(1)} \epsilon^{(1)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{11}$,
+$\rho_{\epsilon^{(2)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{22}$ and 
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{12} \exp(-2 \pi \iota f (l_1-l_2))$.
+
 \section{Details of data and data cleaning}
@@
 
-\section{Correlation analyses to help select autoregressive model lags}
-
+%\section{Correlation analyses to help select autoregressive model lags}
+%***DAN: Not sure to what extent I may or may not even have this section any more
 %***DAN: Describe the surrogating process

### 2021-11-04 Code/Kelp_AnalysisStep_LinearModelSelection.R

@@ -102,43 +102,60 @@ ncf::gcdist(locs$Lon[h],locs$Lat[h]) #73.35km
 h<-c(CC2locstouse[1],CC2locstouse[length(CC1locstouse)])
 ncf::gcdist(locs$Lon[h],locs$Lat[h]) #78.46km
 
-#So my three roughly comparable comparable regions are SBlocstouse, CC1locstouse, CC2locstouse
+#So my three roughly comparable regions are SBlocstouse, CC1locstouse, CC2locstouse
 
 #***
 #Pick lags for each region
 #***
 
-CC1_lag_res<-picklags(y=kelp[CC1locstouse,],x1=NO3[CC1locstouse,],x2=waves[CC1locstouse,],maxlags=c(28,8,1))
+CC1_lag_res<-picklags_mat(y=kelp[CC1locstouse,],x1=NO3[CC1locstouse,],x2=waves[CC1locstouse,],maxlags=c(28,8,1))
 CC1_lag_res_sortAIC<-CC1_lag_res[order(CC1_lag_res$AIC),]
 CC1_lag_res_sortBIC<-CC1_lag_res[order(CC1_lag_res$BIC),]
-head(CC1_lag_res_sortAIC)
-head(CC1_lag_res_sortBIC)
+head(CC1_lag_res_sortAIC,30)
+head(CC1_lag_res_sortBIC,20)
 indCC1<-which(CC1_lag_res$lagAR==4 & CC1_lag_res$lag1==1 & CC1_lag_res$lag2==0)
 CC1_lag_res[indCC1,]
-CC1lags<-CC1_lag_res_sortBIC[1,1:3]
+range(CC1_lag_res$lagAR[CC1_lag_res$DeltaAIC<=3])
+range(CC1_lag_res$lag1[CC1_lag_res$DeltaAIC<=3])
+range(CC1_lag_res$lag2[CC1_lag_res$DeltaAIC<=3])
+range(CC1_lag_res$lagAR[CC1_lag_res$DeltaBIC<=3])
+range(CC1_lag_res$lag1[CC1_lag_res$DeltaBIC<=3])
+range(CC1_lag_res$lag2[CC1_lag_res$DeltaBIC<=3])
 
[…]

### 2021-11-04 Code/Kelp_AnalysisStep_SpectralDecomp.R

@@ -105,7 +105,7 @@ LAInd<-which(dists==min(dists))
 #                   and 0 up to the value input for NO3, and the same for waves.
 #frg              The frequency range to use for display
 #fnpre            A prefix for file names of pdf plots exported by the function, with path but without the
-#                   extension "pdf" of "jpg"
+#                   extension "pdf" or "jpg"
 #plottype         Use "jpg" to make jpg plots, "pdf" to make pdf plots
 #PlotForMSArgs    Formatting arguments used for generating final, presentation-quality plots. If this is missing it 
 #                   means don't generate those plots.
@@ -2352,10 +2352,12 @@ PlotForMSArgs<-list(cov_MainPlot_PanLabs=c("A","B","C"),
 Args<-list(locstouse=CC1locstouse,lags=c(4,1,0),frg=c(0,.5),fnpre="RegionalAnalysis_CentCal1_KelpLag4",plottype="pdf",PlotForMSArgs=PlotForMSArgs)
 res_CentCal1_KelpLag4<-do_analysis(Args)
 saveRDS(res_CentCal1_KelpLag4,paste0(resloc,"res_CentCal1_KelpLag4.Rds"))
+#res_CentCal1_KelpLag4<-readRDS(paste0(resloc,"res_CentCal1_KelpLag4.Rds"))
 
 Args<-list(locstouse=CC1locstouse,lags=c(8,1,0),frg=c(0,.5),fnpre="RegionalAnalysis_CentCal1_KelpLag8",plottype="pdf",PlotForMSArgs=PlotForMSArgs)
 res_CentCal1_KelpLag8<-do_analysis(Args)
 saveRDS(res_CentCal1_KelpLag8,paste0(resloc,"res_CentCal1_KelpLag8.Rds"))
+#res_CentCal1_KelpLag8<-readRDS(paste0(resloc,"res_CentCal1_KelpLag8.Rds"))
 
 Args<-list(locstouse=CC1locstouse,lags=c(12,1,0),frg=c(0,.5),fnpre="RegionalAnalysis_CentCal1_KelpLag12",plottype="pdf",PlotForMSArgs=PlotForMSArgs)
 res_CentCal1_KelpLag12<-do_analysis(Args)
[…]

### 2021-11-04 Code/PickLags.R

@@ -1,7 +1,8 @@
 #This function picks lags in a linear regression setup relevant to the project. Read the specs and code for details.
+#Just picks lags for each location, one by one, using picklags_vect, below, and then aggregates across locations. 
 #
 #Args
-#y          An N by T matrix of time series, N=number of locations, T=time series length. The population variable.
+#y          An N by T matrix representing time series, N=number of locations, T=time series length. The population variable.
 #x1         An N by T matrix of time series. An environmental driver.
 #x2         Another such
 #maxlags    A vector of length 3. maxlags[1]>=0 is the max AR lag for the population variable, 0 meaning populations 
@@ -12,6 +13,47 @@
 #lagAR      The AR lag considered for the populations, >=0, meaning of 0 described above
 #lag1       The lag for x1 considered, >=-1, the meaning of -1 described above
 #lag2       Similar but for x2
+#Rsq        Rsq of the corresponding regression model, averaged across locations
+#AIC        AIC of the corresponding regression model, averaged across locations
+#BIC        BIC of the corresponding regression model, averaged across locations
+#
+#Notes
+#When a given lag is included in the linear regression model, all smaller lags are automatically included.
+#
+picklags_mat<-function(y,x1,x2,maxlags)
+{
+  #compute results for each location and compute average results
+  res_tot<-picklags_vect(y[1,],x1[1,],x2[1,],maxlags)
+  for (counter in 2:(dim(y)[1]))
+  {
+    res_loc<-picklags_vect(y[counter,],x1[counter,],x2[counter,],maxlags)
+    res_tot[,4:6]<-res_tot[,4:6]+res_loc[,4:6]
+  }
+  res<-res_tot
+  res[,4:6]<-res[,4:6]/(dim(y)[1])
+  
+  #put in the delta AIC and BIC columns
+  res$DeltaAIC<-res$AIC-min(res$AIC)
+  res$DeltaBIC<-res$BIC-min(res$BIC)
+  
+  return(res)
+}
+
+#This function picks lags in a linear regression setup relevant to the project. Read the specs and code for details.
+#This is for one location only, the above function iterates across locations.
+#
+#Args
+#y          A vector of length T representing a time series, T=time series length. The population variable.
[…]

### 2021-11-04 Code/Plankton_AnalysisStep_CleanData_Basic.R

@@ -13,7 +13,7 @@ rm(list=ls())
 #Location for storing results and other prep
 #***
 
-resloc<-"../Results/Plankton_DataAfterCleaning/"
+resloc<-"../Results/Plankton_DataAfterBasicCleaning/"
 if (!dir.exists(resloc))
 {
   dir.create(resloc,recursive=TRUE)
@@ -66,14 +66,36 @@ temp<-temp[2:(dim(temp)[1]),2:(dim(temp)[2])]
 testthat::expect_equal(dim(temp),dim(pci))
 rm(year_t,site_t)
 
+#***
+#save the results
+#***
+
+saveRDS(pci,file=paste0(resloc,"Pci_CleanedBasic.Rds"))
+write.table(pci,file=paste0(resloc,"Pci_CleanedBasic.csv"),row.names = FALSE,col.names = FALSE,sep=",")
+
+saveRDS(calfin,file=paste0(resloc,"Calfin_CleanedBasic.Rds"))
+write.table(calfin,file=paste0(resloc,"Calfin_CleanedBasic.csv"),row.names = FALSE,col.names = FALSE,sep=",")
+
+saveRDS(temp,file=paste0(resloc,"Temp_CleanedBasic.Rds"))
+write.table(temp,file=paste0(resloc,"Temp_CleanedBasic.csv"),row.names = FALSE,col.names = FALSE,sep=",")
+
+saveRDS(year,file=paste0(resloc,"Year_CleanedBasic.Rds"))
+write.table(year,file=paste0(resloc,"Year_CleanedBasic.csv"),row.names = FALSE,col.names = FALSE,sep=",")
+
+saveRDS(site,file=paste0(resloc,"Site_CleanedBasic.Rds"))
+write.table(site,file=paste0(resloc,"Site_CleanedBasic.csv"),row.names = FALSE,col.names = FALSE,sep=",")
+
 #These data just loaded have not been Box-Cox'd, or variance standardized, or detrended. They 
 #are just the "raw" annual values, with gaps filled as described in the documenation for the 
 #data and in the Plos Comp Biol paper.
 
 #At this point, the level of cleaning (i.e., none so far except what was done previously
-#for Plos Comp Biol) is similar to what was accomplished for kelp by 
-#Kelp_AnalysisStep_CleanData_SuperBasic. Next step is to look at the next level
-#of cleaning for kelp and do something appropriately similar for plankton
+#for the Plos Comp Biol Dryad-stored data) is similar to what was accomplished for kelp by 
[…]

### 2021-11-04 Code/main.R

@@ -6,6 +6,9 @@
 rm(list=ls())
 graphics.off()
 
+#Develop the theoretical case studies
+source("Theory_CaseStudyA.R")
+
 #Clean the kelp data
 source("Kelp_AnalysisStep_CleanData_SuperBasic.R")
 source("Kelp_AnalysisStep_CleanData_MoreInvolved.R")
@@ -16,8 +19,17 @@ source("Kelp_AnalysisStep_PlotTimeSeries.R")
 #Do linear models with kelp and related data to help demo the need for specific 
 #lags - this also performs and saves the final level of data cleaning
 source("Kelp_AnalysisStep_LinearModels.R")
+source("Kelp_AnalysisStep_LinearModelSelection.R")
 
 #Do the spectra decomposition analysis for kelp
 source("Kelp_AnalysisStep_SpectralDecomp.R")
 
 #Clean the plankton data
+source("Plankton_AnalysisStep_CleanData_Basic.R")
+
+#Plot plankton time series for visual inspection
+source("Plankton_AnalysisStep_PlotTimeSeries.R")
+
+#Do linear models with pci and related data to help demo the need for specific 
+#lags - this also performs and saves the final level of data cleaning
+source("Plankton_AnalysisStep_LinearModels.R")

### 2021-11-04 Code/plotColLine.R

@@ -27,3 +27,19 @@ plotColLine<-function(x,y,cols,lnwid,...)
   lines(c(mean(x[(lx-1):lx]),x[lx]),c(mean(y[(lx-1):lx]),y[lx]),lwd=lnwid,col=cols[lx])
 }
 
+#Same as the above, but plots on an existing plotting device.
+#
+linesColLine<-function(x,y,cols,lnwid)
+{
+  lx<-length(x)
+  
+  lines(c(x[1],mean(x[1:2])),c(y[1],mean(y[1:2])),lwd=lnwid,col=cols[1])
+  for (counter in 2:(length(x)-1))
+  {
+    lines(c(mean(c(x[counter-1],x[counter])),x[counter]),
+          c(mean(c(y[counter-1],y[counter])),y[counter]),lwd=lnwid,col=cols[counter])
+    lines(c(x[counter],mean(c(x[counter],x[counter+1]))),
+          c(y[counter],mean(c(y[counter],y[counter+1]))),lwd=lnwid,col=cols[counter])
+  }
+  lines(c(mean(x[(lx-1):lx]),x[lx]),c(mean(y[(lx-1):lx]),y[lx]),lwd=lnwid,col=cols[lx])
+}

### 2021-11-04 Code/tests_PickLags.R

@@ -9,35 +9,32 @@ test_that("test it",{
   #make up some data and call the function
   set.seed(101)
   T<-10
-  N<-3
-  y<-matrix(rnorm(T*N),T,N)
-  x1<-matrix(rnorm(T*N),T,N)
-  x2<-matrix(rnorm(T*N),T,N)
+  y<-rnorm(T)
+  x1<-rnorm(T)
+  x2<-rnorm(T)
   maxlags<-c(3,2,1)
   dat<-makedf(y,x1,x2,maxlags)
 
   #test the column names and dimensions of the result
   testthat::expect_equal(names(dat),c("y","y_l1","y_l2","y_l3","x1_l0","x1_l1","x1_l2","x2_l0","x2_l1"))
-  testthat::expect_equal(dim(dat)[2],9)
-  testthat::expect_equal(dim(dat)[1],(T-max(maxlags))*N)
-  
+  testthat::expect_equal(dim(dat)[1],7)
+
   #test the response variable column
-  mml<-max(maxlags)
-  testthat::expect_equal(dat$y,as.vector(y[1:(T-mml),]))
+  testthat::expect_equal(dat$y,y[4:10])
 
   #test the AR columns
-  testthat::expect_equal(dat$y_l1,as.vector(y[2:(T-mml+1),]))
-  testthat::expect_equal(dat$y_l2,as.vector(y[3:(T-mml+2),]))
-  testthat::expect_equal(dat$y_l3,as.vector(y[4:(T-mml+3),]))
+  testthat::expect_equal(dat$y_l1,as.vector(y[3:9]))
+  testthat::expect_equal(dat$y_l2,as.vector(y[2:8]))
+  testthat::expect_equal(dat$y_l3,as.vector(y[1:7]))
   
   #test the x1 columns
-  testthat::expect_equal(dat$x1_l0,as.vector(x1[1:(T-mml),]))
-  testthat::expect_equal(dat$x1_l1,as.vector(x1[2:(T-mml+1),]))
-  testthat::expect_equal(dat$x1_l2,as.vector(x1[3:(T-mml+2),]))
+  testthat::expect_equal(dat$x1_l0,x1[4:10])
+  testthat::expect_equal(dat$x1_l1,x1[3:9])
+  testthat::expect_equal(dat$x1_l2,x1[2:8])
   
   #test the x2 columns
[…]

### 2021-11-04 Code/Theory_CaseStudyAB.R
added code for theoretical case studies

@@ -1,0 +1,382 @@
+#This script implements theoretical case studies A and B
+
+rm(list=ls())
+
+#***
+#External codes needed
+#***
+
+#packages needed (invoked with "::"): matrixcalc, mvtnorm
+source("SpectralTools.R")
+source("WhichBreak.R")
+source("plotColLine.R")
+
+#***
+#Location for storing results and other prep
+#***
+
+resloc<-"../Results/Theory/"
+if (!dir.exists(resloc))
+{
+  dir.create(resloc,recursive=TRUE)
+}
+
+#***
+#The main function for doing both case studies
+#***
+
+#This function calculates the analytic results and plots them for case studies A and B.
+#The function is actually more general than the planned case studies that will be 
+#displayed in the paper.
+#
+#Args
+#b            The parameters (b_1,...,b_n) of the model
+#p1           The parameters (p_0^{(1)},...,p_{m_1}^{(1)}) of the model
+#p2           The parameters (p_0^{(2)},...,p_{m_1}^{(2)}) of the model
+#N            The number of habitat patches
+#ns           Parameters for the noise. A length-3 vector with some constraints.
+#               See the code or the paper for information on how the noise is
+#               constructed based on these parameters. Using notation from the paper,
+#               the entries are v11, v22, v12.
+#PlotName     A file name of the saved plots, including path, excluding file extension.
+#PanLabs      A vector of panel labels (typically c("A","B","C","D"), or similar)
+#XAxOn        Default TRUE means plot the x axis label and numbers, FALSE means don't,
+#               with the size of the plot correspondingly adjusted.
+#
+#Output - none, but plots saved.
+#
+#Notes
+#Very little error checking, so watch out!
+#
[…]

### 2021-11-04 Code/Theory_CaseStudyC.R

@@ -1,0 +1,563 @@
+#This script implements theoretical case study C
+
+rm(list=ls())
+
+#***
+#External codes needed
+#***
+
+#packages needed (invoked with "::"): matrixcalc, mvtnorm
+source("SpectralTools.R")
+source("WhichBreak.R")
+source("plotColLine.R")
+
+#***
+#Location for storing results and other prep
+#***
+
+resloc<-"../Results/Theory/"
+if (!dir.exists(resloc))
+{
+  dir.create(resloc,recursive=TRUE)
+}
+
+#***
+#The main function for doing both case studies
+#***
+
+#***
+#The main function for doing both case studies
+#***
+
+#This function calculates the analytic results and plots them for case study C.
+#The function is actually more general than the planned case study that will be 
+#displayed in the paper.
+#
+#Args
+#b            The parameters (b_1,...,b_n) of the model
+#p1           The parameters (p_0^{(1)},...,p_{m_1}^{(1)}) of the model
+#p2           The parameters (p_0^{(2)},...,p_{m_1}^{(2)}) of the model
+#N            The number of habitat patches
+#ns           Parameters for the noise. A length-7 vector with some constraints.
+#               See the code or the paper for information on how the noise is
+#               constructed based on these parameters. Using notation from the paper,
+#               the entries are c1, c2, l1, l2, w11, w22, w12.
+#PlotName     A file name of the saved plots, including path, excluding file extension.
+#PanLabs      A vector of panel labels (typically c("A","B","C","D"), or similar)
+#PanLabs2     Another similar, this one just 3 long, for the sup mat figure showing the 
+#               rhos
+#SyncYAx      A length 2 vector with the y axis extent for the synchrony plot. This is 
[…]

### 2021-11-04 Code/Plankton_AnalysisStep_LinearModelSelection.R
added plankton explorations

@@ -1,0 +1,73 @@
+#Does some linear models and information criterion analyses to select numbers of lags to use in 
+#ARMA-type models of plankton.
+
+rm(list=ls())
+
+#***
+#External codes needed
+#***
+
+#packages needed (invoked with "::"): ncf
+
+source("PickLags.R")
+
+#***
+#Locations for storing results, and other prep
+#***
+
+resloc<-"../Results/Plankton_LinearModelSelection/"
+if (!dir.exists(resloc))
+{
+  dir.create(resloc,recursive=TRUE)
+}
+
+#***
+#Load the data
+#***
+
+#load the data
+datloc<-"../Results/Plankton_DataAfterAllCleaning/"
+pci<-readRDS(paste0(datloc,"Pci_CleanedFinal.Rds")) 
+calfin<-readRDS(paste0(datloc,"Calfin_CleanedFinal.Rds")) 
+temp<-readRDS(paste0(datloc,"Temp_CleanedFinal.Rds")) 
+locs<-readRDS(paste0(datloc,"Site_CleanedFinal.Rds"))
+year<-readRDS(paste0(datloc,"Year_CleanedFinal.Rds"))
+times<-1:length(year)
+
+numts<-dim(pci)[1]
+lents<-dim(pci)[2]
+
+#***
+#Pick lags 
+#***
+
+lag_res<-picklags_mat(y=pci,x1=temp,x2=calfin,maxlags=c(1,1,2))
+
+lag_res$AICwt<-exp(-.5*lag_res$DeltaAIC)
+lag_res$AICwt<-lag_res$AICwt/sum(lag_res$AICwt)
+lag_res$BICwt<-exp(-.5*lag_res$DeltaBIC)
+lag_res$BICwt<-lag_res$BICwt/sum(lag_res$BICwt)
+
+lag_res_sortAIC<-lag_res[order(lag_res$AIC),]
+lag_res_sortBIC<-lag_res[order(lag_res$BIC),]
+lag_res_sortAIC$AICwt_cumsum<-cumsum(lag_res_sortAIC$AICwt)
+lag_res_sortBIC$BICwt_cumsum<-cumsum(lag_res_sortBIC$BICwt)
+
+head(lag_res_sortAIC,20)
+head(lag_res_sortBIC,20)
+
+range(lag_res$lagAR[lag_res$DeltaAIC<=3])
[…]

### 2021-11-04 Code/Plankton_AnalysisStep_LinearModels.R

@@ -1,0 +1,230 @@
+#Does some correlation analyses to confirm influences of temp and calfin on pci, and to get lags. 
+#Also finishes the data cleaning. The linear models done for kelp are skipped here, because we
+#ended up not needing them for kelp.
+
+rm(list=ls())
+
+#***
+#External codes needed
+#***
+
+#packages needed (invoked with "::"): wsyn, parallel
+
+source("MnCor_Tiny.R")
+
+#***
+#Locations for storing results, and other prep
+#***
+
+resloc1<-"../Results/Plankton_DataAfterAllCleaning/"
+if (!dir.exists(resloc1))
+{
+  dir.create(resloc1,recursive=TRUE)
+}
+
+resloc2<-"../Results/Plankton_LinearModelResults/"
+if (!dir.exists(resloc2))
+{
+  dir.create(resloc2,recursive=TRUE)
+}
+
+theseed<-101
+numsurrog<-1000
+saveRDS(theseed,paste0(resloc2,"theseed"))
+saveRDS(numsurrog,paste0(resloc2,"numsurrog"))
+
+#***
+#Load the data
+#***
+
+datloc<-"../Results/Plankton_DataAfterBasicCleaning/"
+pci<-readRDS(paste0(datloc,"Pci_CleanedBasic.Rds")) 
+calfin<-readRDS(paste0(datloc,"Calfin_CleanedBasic.Rds")) 
+temp<-readRDS(paste0(datloc,"Temp_CleanedBasic.Rds")) 
+year<-readRDS(paste0(datloc,"Year_CleanedBasic.Rds"))
+site<-readRDS(paste0(datloc,"Site_CleanedBasic.Rds"))
+
+numts<-dim(pci)[1]
+lents<-dim(pci)[2]
+
+# #***
+# #Get surrogates and create centered, appropriately normalized versions of everything - approach 1
+# #***
+# 
+# #centered (but not detrended) versions of the "environmental" variables (considered temp and calfin), 
+# #necessary to get surrog to work below
+# calfin_c<-wsyn::cleandat(calfin,year,1)$cdat
+# temp_c<-wsyn::cleandat(temp,year,1)$cdat
+# 
+# #Get surrogates for calfin and temp - use joint surrogates of both variables, synchrony-preserving, aaft.
+# #We do NOT detrend the data before surrogating because it makes the surrogates less good.
[…]

### 2021-11-04 Code/Plankton_AnalysisStep_PlotTimeSeries.R

@@ -1,0 +1,80 @@
+#Just plots time series after the "basic" cleaning (but prior to the final cleaning done in 
+#Plankton_AnalysisStep_LinearModels.R), for visual inspection.
+
+rm(list=ls())
+
+#***
+#Locations for storing results
+#***
+
+resloc<-"../Results/Plankton_TimeSeriesPlots/"
+if (!dir.exists(resloc))
+{
+  dir.create(resloc,recursive=TRUE)
+}
+
+#***
+#Load the data
+#***
+
+#load the data
+datloc<-"../Results/Plankton_DataAfterBasicCleaning/"
+pci<-readRDS(paste0(datloc,"Pci_CleanedBasic.Rds")) 
+calfin<-readRDS(paste0(datloc,"Calfin_CleanedBasic.Rds")) 
+temp<-readRDS(paste0(datloc,"Temp_CleanedBasic.Rds")) 
+year<-readRDS(paste0(datloc,"Year_CleanedBasic.Rds"))
+site<-readRDS(paste0(datloc,"Site_CleanedBasic.Rds"))
+
+numts<-dim(pci)[1]
+lents<-dim(pci)[2]
+
+#***
+#Make the plots
+#***
+
+pan.wd<-3
+pan.ht<-pan.wd
+gap<-0.2
+xax.ht<-.6
+yax.wd<-xax.ht
+tot.wd<-3*yax.wd+3*gap+3*pan.wd
+tot.ht<-xax.ht+pan.ht+gap
+
+for (counter in 1:(dim(pci)[1]))
+{
+  #plot the pci data
+  png(paste0(resloc,"SitePlot_",counter,".png"),res=300,units="in",width = tot.wd,height = tot.ht)
+  
+  #plot pci on the first set of axes
+  par(fig=c(yax.wd/tot.wd,
+            (yax.wd+pan.wd)/tot.wd,
+            (xax.ht)/tot.ht,
+            (xax.ht+pan.ht)/tot.ht),
+      mai=c(0,0,0,0),mgp=c(3,0.75,0))
+  plot(year,pci[counter,],type="l",col="green",xlab="year",ylab="pci") 
+  mtext("year",side=1,line=1.6)
+  mtext("pci",side=2,line=1.6)
+  
+  #temp on the second axes
+  par(fig=c((2*yax.wd+pan.wd+gap)/tot.wd,
+            (2*yax.wd+2*pan.wd+gap)/tot.wd,
[…]

### 2021-11-04 Code/Plankton_AnalysisStep_SpectralDecomp.R

@@ -1,0 +1,374 @@
+#Spectral analysis steps applied to plankton.
+
+rm(list=ls())
+
+#***
+#External codes needed
+#***
+
+#packages needed (invoked with "::"): parallel, wsyn, ncf, graphics, latex2exp
+source("SpectralTools.R")
+source("AvgOffDiags.R")
+source("PlotPhaseFunc.R")
+source("WhichBreak.R")
+source("plotColLine.R")
+
+#***
+#Location for storing results and other prep
+#***
+
+resloc<-"../Results/Plankton_SpectralDecompResults/"
+if (!dir.exists(resloc))
+{
+  dir.create(resloc,recursive=TRUE)
+}
+
+theseed<-101
+BiasVariance<-0.5
+
+saveRDS(theseed,file=paste0(resloc,"theseed.Rds"))
+saveRDS(BiasVariance,file=paste0(resloc,"BiasVariance.Rds"))
+
+#***
+#Load the data
+#***
+
+#load the data
+datloc<-"../Results/Plankton_DataAfterAllCleaning/"
+pci<-readRDS(paste0(datloc,"Pci_CleanedFinal.Rds")) 
+calfin<-readRDS(paste0(datloc,"Calfin_CleanedFinal.Rds")) 
+temp<-readRDS(paste0(datloc,"Temp_CleanedFinal.Rds")) 
+locs<-readRDS(paste0(datloc,"Site_CleanedFinal.Rds"))
+year<-readRDS(paste0(datloc,"Year_CleanedFinal.Rds"))
+times<-1:length(year)
+
+numts<-dim(pci)[1]
+lents<-dim(pci)[2]
+
+#***
+#The main function for doing the analysis 
+#***
+
+#Uses the spectral mathematics to get information about fractions of synchrony in pci explained by temperature, cal fin, 
+#and interactions. Assumes all the variables created above are in the workspace when run.
+#
+#Args - a list with these named elements
+#locstouse        Indices of locations to use. These refer to rows of the pci, temp and calfin data matrices above
+#lags             Max regression lags to use, pci then temp then calfin. In years. So if you want the model to
[…]

### 2021-11-04 Paper/SupMat.Rnw
edits

@@
 
-For the simple example, used in theoretical case studies A and B, we 
+For the simple example, to be used in theoretical case studies A and B (Methods), we 
 assume $(\epsilon^{(1)}(t),\epsilon^{(2)}(t))$ are drawn, independently for each $t$, from a
@@
 
-For a more illustrative example, used in case study C, the noise process 
+These results make intuitive sense, for a white-noise process, for the following reasons.
+First, $\rho_{\epsilon^{(1)} \epsilon^{(1)}}$ and $\rho_{\epsilon^{(2)} \epsilon^{(2)}}$,
+which are automatically real-valued (SI section \ref{Sect:measuring_sync}), should be flat for a white noise
+process (as we determined above that they are), i.e., independent of timescale, because spectral quantities are flat 
+for time series with no temporal dependencies.  For the same reason, $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$,
+our cross synchrony measure, should be flat (as we determined above that it is). It makes sense
+that cross synchrony in this example is real-valued because a complex-valued 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ would imply lagged or phase-delayed relationships between 
+$\epsilon^{(1)}$ and $\epsilon^{(2)}$, not possible with a white-noise process.
+
+For a more illustrative example, used in case study C (Methods), the noise process 
 $(\epsilon^{(1)},\epsilon^{(2)})$ is generated via a three-step process. 

### 2021-12-22 Paper/Paper.Rnw
write up of theoretical case studies

@@
 %Interacting Moran effects
+%How Moran effects interact
+%On interactions between Moran effects
 %Interactions between environmental drivers of spatial synchrony
-%Environmental drivers of spatia synchony can strongly interact
+%Environmental drivers of spatial synchony can strongly interact
 %Environmental drivers of spatial synchrony interact
 
-%***Tentative author order, who is last yet to be determined
-\author[a,b]{Daniel C. Reuman}
+%***Tentative author order, Bell last for his data contributions
+\author[a,b,*]{Daniel C. Reuman}
 \author[c]{Max Castorani}
-\author[d,e]{Tom Bell}
-\author[f]{Kyle Cavanaugh}
+\author[d]{Kyle Cavanaugh}
+\author[e]{Lawrence W. Sheppard}
 \author[c]{Jonathan Walter}
-\author[a,g]{Lawrence W. Sheppard}
+\author[f,g]{Tom Bell}
 
-\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
+\affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
 \affil[b]{Laboratory of Populations, Rockefeller University}
 \affil[c]{Department of Environmental Sciences, University of Virginia}
-\affil[d]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
-\affil[e]{Earth Research Institute, University of California, Santa Barbara}
-\affil[f]{Department of Geography, University of California, Los Angeles}
-\affil[g]{Marine Biological Association of the United Kingdom}
-
-%***Need to indicate corresponding author Reuman
+\affil[d]{Department of Geography, University of California, Los Angeles}
+\affil[e]{Marine Biological Association of the United Kingdom}
+\affil[f]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
+\affil[g]{Earth Research Institute, University of California, Santa Barbara}
+\affil[*]{Corresponding author, reuman@ku.edu}
 
@@
 theoretical case studies that illuminate the intuition of interacting Moran effects. 
-We develop case studies A-C. For case study A,  the environmental variable 
+We develop case studies A-C. For all cases, the model time step was assumed to be one quarter, i.e., 
+four model time steps per year. This makes no real mathematical difference, but was done to 
+facilitate later comparisons with results based on kelp data, which were available on a
+quarterly sampling frequency. 
+For case study A,  the environmental variable 
 $\epsilon^{(1)}$ is assumed to have a simple positive effect on populations, but 
@@
 For case study C, the effects of $\epsilon^{(1)}$ are lagged by $n_1$ and the 
-effects of $\epsilon^{(2)}$ are lagged by $n_2$, i.e.,
+effects of $\epsilon^{(2)}$ are lagged by $n_2$ time steps, i.e.,
 $p_{n_1}^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n_1$, and 
@@
 For C, the noise processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
-each assumed to exhibit spatially synchronous periodic oscillations of the same period
-but potentially different phases. Such a scenario could be realized by seasonally 
-periodic environmental fluctuations if the model time step were quarterly or monthly,
+each assumed to exhibit spatially synchronous periodic oscillations of period equal to one year, i.e.,
+four model time steps, but with different phases. Peaks in $\epsilon^{(1)}$ are assumed to 
+either lead or lag (we consider two model scenarios) peaks in $\epsilon^{(2)}$ by one quarter.
+Such a scenario could be realized by annually 
+periodic environmental fluctuations sampled quarterly,
 e.g., kelp wave action in central California peaks annually in the winter, whereas surface-water
-nitrate concentrations also fluctuate with period one year, but peak in the spring. 
+nitrate concentrations also fluctuate with period one year, but peak in the spring, a delay of one quarter
+compared to the wave peak. 
 Details of the noise processes are specified in SI section \ref{SM-sect:examples_cross_sync}.
 For all case studies,
-the autoregressive order of population dynamics is assumed to be $1$, i.e, $b_1 \neq 0$ 
+the autoregressive order of population dynamics is assumed, for simplicity, to be $1$, i.e, $b_1 \neq 0$ 
 but $b_i=0$ for $i>1$.
@@
 
-% What do we learn from case study A:
-% 
-% Warmup:
-% 1) interactions between Moran effects (ref to specific term of the theory) can easily be of comparable importance
-% to the main Moran effects
-% 2) interaction effects can be positive or negative, and can differ, and even differ in sign, by timescale
-% 
-% Actually breaking it down:
-% 1) At the highest frequencies (shortest timescales) of 0.5 cycles per sampling 
-% interval (period 2 sampling intervals), a lag of 1 
-% sampling interval in the population effects of \epsilon^{(1)} is half the period of oscialltion, so 
-% causes the effects of \epsilon^{(1)}
-% on those timescales to be exactly out of phase of the effects of \epsilon^{(2)}, and so we get 
-% cancellation, i.e., negative interaction effects. The effects of \epsilon1 are multiplied by 1.1,
-% those of \epsilon2 by 0.8, and the correlation between the two (which is made to have negative effects
-% by the lag) is 0.3. So the negative interaction effects are 1.1*0.8*0.3*2, where the last factor of
-% two is because there are actually two effects here, the one where \epsilon1 is operating, lagged, at i and 
-% \epsilon2 is operating at j; and the effect where \epsilon1 is operating, lagged, at j and \epsilon2 is 
-% operating at i. 
-% 2) As timescales get to 1 year = 4 quarters, a lag of 1 quarter in the population effects of 
-% \epsilon1 puts the population effects of \epsilon1 3/4 out of phase with those of \epsilon 2,
-% and the correlation of two sin waves of the same frequency that are 3/4 out of phase is 0, so 
-% interaction effects are 0 here.
-% 3) As timescales get long, a lag of 1 quarter in the population effects of \epsilon1 is a
-% negligible time compared to the timescale being considered, so the effects of 
-% \epsilon1 and \epsilon2 on populations are closer and closer to be in phase and correlated
-% as timescales get longer and longer. The effects of \epsilon1 are multiplied by 1.1, those of 
-% \epsilon2 by 0.8, and the correlation is 0.3, so the interaction effect is 1.1*.8*.3*2 (the 
-% reason for the 2 is explained above).
-% 
-% Need to put this into good writing
-% 
-% Need to carry out similar thoughts for case B and write them, too.
+%Para basically confirming earlier work, as a starter
+Our theoretical results demonstrate that interaction effects between Moran drivers: 1) can be 
+comparable in strength to direct Moran effects; 2) can be either synergistic or destructive, i.e., positive or negative; and 3) 
+can depend strongly on timescale, even having different signs on different timescales 
+(Fig. \ref{fig:TheoryCaseA}a,e,i,m). First, for all case studies, the magnitude of 
+interaction effects was comparable to that of direct Moran effects (compare the dashed and solid lines on 
+Fig. \ref{fig:TheoryCaseA}a,e,i,m). So interactions can contribute substantially to overall population 
+synchrony. Second, in contrast to the direct Moran effects of a single environmental variable, which provide non-negative 
+contributions to population synchrony at all timescales for which environmental synchrony 
+is positive, interactions between Moran effects can be positive or negative. For instance, case study A
+showed negative interaction effects at short timescales (Fig. \ref{fig:TheoryCaseA}a); and, for some
+parameters, case study C showed negative interactions at annual timescales (Fig. \ref{fig:TheoryCaseA}m).
+Thus interaction effects can either augment or reduce population synchrony from what its value would be 
+if only direct Moran effects operated. Finally, interactions between Moran effects depended strongly on timescale 
+for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and differed in sign across timescales for case studies
+A and B and for one of the C scenarios. This augments earlier studies demonstrating that direct Moran effects can 
+depend on timescale \citep{Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}. The points made in this paragraph
+also follow from the results of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not necessarily
+straightforwardly. Additionally, those are case studies, so our results also provide novel theoretical generality;
+and our theory can be interpreted to provide intuition about how interacting Moran effects work,
+as follows.
 
-%Thoughts on 2021 11 01
-% Topic sentence for first para on theory results:
-% Results confirm earlier empirical work (refs) that interaction effects between Moran drivers 1) can be comparable in strength to direct Moran effects; 2) can be synergistic or destructive; and 3) can depend strongly on timescale, even to the extent that the sign of interaction effects can differ by timescale.
-% 
-% Combine all three theory figs into one multi panel fig, change the order of panels again 
-% 
-% Subsequent paras will then break things down to understand how this stuff works
-% 
-% Possible topic sentence for subsequent para: Theory reveals, as illustrated by cases A and B, how interactions between Moran effects depend on the lags with which Moran drivers influence populations.
-Or theory reveals how interactions between Moran effects depend on the lags with which Moran drivers influence populations, and case study begins to illustrate how this works.
-Or something like that.
-% 
-%The also have a para for C
-%
-% Maybe get rid of the “confirm earlier work” phrase since it wasn’t necessarily explicitly listed in that way in earlier work
+%Case study A
+Case study A provides an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
+Moran effects on different timescales, and how our theory successfully captures that contrast.
+For case study A, for which $\epsilon^{(1)}$ effects on populations were lagged by 1 quarter but 
+$\epsilon^{(2)}$ effects were unlagged (Methods), interactions between Moran effects were negative on the shortest timescales 
+(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}a, left side of panel). This is because the 
+population influences of Moran effects of timescale 2
+quarters and lagged by 1 quarter are exactly out of phase with unlagged Moran effects of timescale
+2 quarters, producing destructive interference, i.e., negative interactions. Thus, on short timescales, total
+synchrony will be less, for this example, than it would be if only direct Moran effects operated. In contrast, 
+interaction effects on long timescales were synergistic, i.e., positive, for case study A 
+(Fig. \ref{fig:TheoryCaseA}a, right side of panel). 
+This is because, on long timescales, a lag of 1 quarter does not meaningfully alter the phase relationship between 
+Moran drivers, and drivers are therefore almost in phase, reinforcing each other.
+At an intermediate timescale of 1 year, a lag of 1 quarter causes the Moran effects
+of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be a quarter phase out of alignment (1 quarter is $1/4$ of a 1-year 
+timescale), resulting in 
+zero interaction effects, essentially because sine waves a quarter-phase out of alignment
+with each other are perfectly uncorrelated. Thus interaction effects between Moran drivers were zero 
+at timescale 1 year (Fig. \ref{fig:TheoryCaseA}a). Theory effectively captures these complexities because
+the size of the 1-quarter phase lag of this scenario, relative to a given timescale,
+is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$, which has phase $-\pi$ at short timescales
+and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b). Because 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ has phase $0$ for all timescales, the phase of 
+$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}c), which controls
+the sign of the real part of this quantity, is the same as the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$.
+Thus $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, which is the
+interaction term in our theory (\ref{rhoww_T1to3}), shifts 
+from negative, for short timescales, to positive for long timescales. Comparing
+Fig. \ref{fig:TheoryCaseA}a,d shows the additional influence of autoregressive population effects, which
+simply multiply all Moran influences by the same timescale-dependent non-negative 
+quantity without altering the relative importance or sign of those effects [see (\ref{rhoww_T1to3})].
 
-%Para basically confirming earlier work, as a starter
-Theoretical results help demonstrate that interaction effects between Moran drivers: 1) can be 
-comparable in strength to direct Moran effects; 2) can be synergistic or destructive; and 3) 
-can depend strongly on timescale, even to the extent that the sign of interaction effects can 
-differ by timescale.
+Case study B reveals how Moran effects can interact when lags are longer than 1 model time step. 
+For B, $\epsilon^{(1)}$ effects on populations were lagged by 3 quarters but 
+$\epsilon^{(2)}$ effects were unlagged (Methods). Thus interactions between Moran effects were 
+negative on the shortest timescales 
+(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, left side of panel). This is because the 
+population influences of Moran effects of timescale 2
+quarters and lagged by 3 quarters are exactly out of phase with unlagged Moran effects of timescale
+2 quarters, producing negative interactions. 
+Interaction effects on long timescales were instead synergistic, i.e., positive, for case study B 
+(Fig. \ref{fig:TheoryCaseA}e, right side of panel), for the same reason as case study A (see above).
+At timescale 3 quarters, Moran effects are, of course, reinforcing because the lag equals the timescale,
+producing Moran effects of the two drivers which are in phase 
+(Fig. \ref{fig:TheoryCaseA}e, peak at timescale 3 quarters = 0.75 years). 
+At a timescale of 1 year, a lag of 3 quarters causes the Moran effects of 
+$\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be $3/4$-phase out of alignment, resulting in zero interaction effects. 
+Likewise at a timescale of 3 years (12 quarters), a lag of 3 quarters 
+makes the Moran effects of 
+$\epsilon^{(1)}$ and $\epsilon^{(2)}$ be $1/4$-phase out of alignment, again producing zero interaction. 
+Finally, at timescale 1.5 years (6 quarters), the 3-quarter lag equals half the timescale, so 
+$\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects counteract each other, producing negative 
+interactions. Thus interactions between Moran drivers were zero 
+at timescales 1 and 3 years for B, were 
+negative for short timescales and for timescale 1.5 years, and were positive for 
+timescale 0.75 years and for long timescales (Fig. \ref{fig:TheoryCaseA}e).
+Theory captures these complexities because 
+the size of the 3-quarter phase lag of this scenario, relative to a given timescale,
+is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
+Because $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number for case study B,
+the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+is the same as that of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f); 
+that common phase then controls the sign of 
+$2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, which is the 
+interaction effect term in (\ref{rhoww_T1to3}).
+
+Cases A and B assumed white noise processes and real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$,
+but case study C illustrates what can happen when noise processes have periodic components, e.g., 
+both noise processes oscillate with annual periodicity according to the seasons, an important scenario, ecologically. 
+What matters then is whether and how the lag between the two seasonal noise processes aligns with the difference 
+between the lags of their effects on populations, as we now explain. 
+For the scenarios we considered under case study C (Fig. \ref{fig:TheoryCaseA}i-p), 
+$\epsilon^{(1)}$ effects on populations were lagged (i.e., delayed in their influence) by 1 quarter, but 
+$\epsilon^{(2)}$ effects were immediate (Methods and Fig. \ref{fig:TheoryCaseA} caption).
+For scenario 1 of case study C (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
+noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
+1 quarter. Thus lags cancelled and the annual peak population effects of $\epsilon^{(1)}$ and
+$\epsilon^{(2)}$ arrived in sync with each other each year, producing positive interactions 
+between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i), and more population
+synchrony than would have been expected if Moran effects operated independently. Contrastingly,
+for scenario 2 of case study C (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
+noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
+1 quarter. Thus lags compounded and the annual peak population effects of $\epsilon^{(1)}$
+arrived 2 quarters after those of $\epsilon^{(2)}$ every year, a half-phase shift for an 
+annually oscillating effect, which corresponds to negative interactions between Moran influences
+and less population synchrony than would have been expected if Moran effects operated independently 
+(Fig. \ref{fig:TheoryCaseA}m). The only difference between scenarios 1 and 2 of case study C
+is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
+Our theory captures these complexities because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
+quantifies the relative phase lag between the effects of $\epsilon^{(1)}$ and 
+$\epsilon^{(2)}$ relative to a given timescale, while 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the phase lag between 
+oscillations in the variables themselves, relative to the timescale.
+The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+therefore captures whether relative lags of effects and lags between 
+the variables themselves align or misalign, with the degree of alignment
+reflected in the real part of this quantity.
 
@@
 %fig idea, could maybe combine some figs to make it work, or put the pedagog content as the first
-%part of the theory fig? Don't worry about it for now.
+%part of the theory fig? Don't worry about it for now, but return to it later.
 
-%***DAN: add case study C to this.
 \begin{figure}
@@
 which is population synchrony due to interacting Moran effects.
-%***DAN: More text here as we flesh out the presentation 
+%***DAN: More text here as we flesh out the presentation - need to describe the middle two columns of panels 
+The functions plotted on d, h, l, p are those on a, e, i, m, respectively, times 
+$|f_{\mcB}|^2$, thereby showing the influence of autoregressive population effects on
+synchrony [see (\ref{rhoww_T1to3})-(\ref{rhoww_OC})].
 Model parameters were: $N=5$ and $b_1=0.4$ for all case studies; 
@@
 $p_3^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study B;
-$p_1^{(1)}=3$, $p_0^{(2)}=1.5$, $c_1=0$ and $c_2=-4/9$ for case study C;
-$v_{11}=v{22}=0.6$ and $v_{12}=0.3$ for case studies A and B;
-$w_{11}=w{22}=0.6$ and $w_{12}=0.3$ for case study C;
-$l_1=0$ and $l_2=1$ for case study C, panels i-l; and 
-$l_1=1$ and $l_2=0$ for case study C, panels m-p.
-See Methods and SI sections \ref{SM-sect:examples_cross_sync} and 
-\ref{SM-sect:details_case_studies} for the meaning of parameters and additional details.}\label{fig:TheoryCaseA}
+$p_1^{(1)}=3$, $p_0^{(2)}=1.5$ 
+%$c_1=0$ and $c_2=-4/9$ 
+for case study C. For C, peaks in the periodic noise process $\epsilon^{(2)}$
+lagged peaks in the periodic process $\epsilon^{(1)}$ by 1 quarter
+in panels i-l, and $\epsilon^{(1)}$ lagged $\epsilon^{(2)}$ by the same
+amount in m-p.
+See SI section \ref{SM-sect:examples_cross_sync} and Fig. \ref{SM-fig:rhosForC} for 
+parameters associated with the noise for each case study,
+and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.}\label{fig:TheoryCaseA}
 \end{figure}
-%#Case A
-%N<-5
-%ns<-c(.6,.6,.3)
-%p1<-c(0,1.1)
-%p2<-c(0.8)
-%b<-c(.4)
-% #Case B
-% N<-5
-% ns<-c(.6,.6,.3)
-% p1<-c(0,0,0,1.1)
-% p2<-c(0.8)
-% b<-c(.4)
 

### 2021-12-22 Paper/SupMat.Rnw
write up of theoretical case studies

@@
 
-%***Tentative author order
-\author[a,b]{Daniel C. Reuman}
+%***Tentative author order, Bell last for his data contributions
+\author[a,b,*]{Daniel C. Reuman}
 \author[c]{Max Castorani}
-\author[d,e]{Tom Bell}
-\author[f]{Kyle Cavanaugh}
+\author[d]{Kyle Cavanaugh}
+\author[e]{Lawrence W. Sheppard}
 \author[c]{Jonathan Walter}
-\author[a,g]{Lawrence W. Sheppard}
+\author[f,g]{Tom Bell}
 
-\affil[a]{Department of Ecology and Evolutionary Biology and Kansas Biological Survey, University of Kansas}
+\affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
 \affil[b]{Laboratory of Populations, Rockefeller University}
 \affil[c]{Department of Environmental Sciences, University of Virginia}
-\affil[d]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
-\affil[e]{Earth Research Institute, University of California, Santa Barbara}
-\affil[f]{Department of Geography, University of California, Los Angeles}
-\affil[g]{Marine Biological Association of the United Kingdom}
+\affil[d]{Department of Geography, University of California, Los Angeles}
+\affil[e]{Marine Biological Association of the United Kingdom}
+\affil[f]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
+\affil[g]{Earth Research Institute, University of California, Santa Barbara}
+\affil[*]{Corresponding author, reuman@ku.edu}
 
@@
 $\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$, based on the definitions of these 
-quantities given previously.
+quantities given previously. For theoretical case studies A and B in the main text, we used
+$N=5$, $v_{11}=v{22}=0.6$ and $v_{12}=0.3$.
 
@@
 $\rho_{\epsilon^{(1)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{12} \exp(-2 \pi \iota f (l_1-l_2))$.
+Fig. \ref{fig:rhosForC} displays these quantities for the two parameterizations
+of $(\epsilon^{(1)},\epsilon^{(2)})$ that we use for theoretical case study C.
+
+We now explore why the results above for $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ make sense for the noise processes we described, particularly for the sets of parameters 
+used for theoretical case study C in the main text (see Fig. \ref{fig:rhosForC}). 
 The pre-factor $\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2$ in all these expressions reflects the autoregressive 
 manner in which the noises $\epsilon$ were defined. Autoregressive models of order $2$ are well known to be capable of
-producing oscillations for certain values of the parameters. For instance, we use $c_1=0$ and $c_2=-4/9$, values
+producing oscillations for certain values of the parameters. For instance, for case study C, we use $c_1=0$ and $c_2=-4/9$, values
 which cause the expression $\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2$ to have a single peak at
-frequency $0.25$ cycles per sampling interval (i.e., $1$ cycle per year). 
+frequency $0.25$ cycles per sampling interval (i.e., $1$ cycle per year). It makes sense that the synchrony quantities
+$\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ for periodically oscillating quantities should also have a peak at the 
+timescale of oscillation. 
 The term $\exp(-2 \pi \iota f (l_1-l_2))$ in the expression for 
@@
 or $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ reflects that fact that there are no lags built into within-variable
-synchrony in this example. Fig. \ref{fig:rhosForC} displays the quantities $\rho_{\epsilon^{(1)} \epsilon^{(1)}}$,
-$\rho_{\epsilon^{(2)} \epsilon^{(2)}}$, and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ for the two parameterizations
-of this noise that we use for theoretical case study C.
+synchrony in this example.
 
@@
 described in SI section \ref{sect:examples_cross_sync} and used for theoretical case study C. 
-Two alternative sets of parameters
-were used for panels a-c and d-f, respectively. See the caption of Fig. \ref{MT-fig:TheoryCaseA} in the main text
-for parameter values.}\label{fig:rhosForC}
+These are complex-valued functions of timescale; their magnitude is displayed on vertical axes and
+their phase is displayed in color. Two alternative sets of parameters
+were used for panels a-c and d-f, respectively. Parameters were:
+$N=5$, $c_1=0$, $c_2=-4/9$, $w_{11}=w{22}=0.6$ and $w_{12}=0.3$ for both sets of panels;
+$l_1=0$ and $l_2=1$ for panels a-c; and 
+$l_1=1$ and $l_2=0$ for panels d-f.
+}\label{fig:rhosForC}
 \end{figure}

### 2021-12-22 Code/main.R
starting to change the overall control flow to drop the attempt to study the plankton

@@ -7,7 +7,8 @@ rm(list=ls())
 graphics.off()
 
 #Develop the theoretical case studies
-source("Theory_CaseStudyA.R")
+source("Theory_CaseStudyAB.R")
+source("Theory_CaseStudyC.R")
 
 #Clean the kelp data
 source("Kelp_AnalysisStep_CleanData_SuperBasic.R")
@@ -24,12 +25,3 @@ source("Kelp_AnalysisStep_LinearModelSelection.R")
 #Do the spectra decomposition analysis for kelp
 source("Kelp_AnalysisStep_SpectralDecomp.R")
 
-#Clean the plankton data
-source("Plankton_AnalysisStep_CleanData_Basic.R")
-
-#Plot plankton time series for visual inspection
-source("Plankton_AnalysisStep_PlotTimeSeries.R")
-
-#Do linear models with pci and related data to help demo the need for specific 
-#lags - this also performs and saves the final level of data cleaning
-source("Plankton_AnalysisStep_LinearModels.R")

### 2021-12-22 Code/Plankton_AnalysisStep_CleanData_Basic.R
removed plankton code and data

@@ -1,101 +1,0 @@
-#Clean the plankton data. Since data were already cleaned by our earlier efforts, 
-#published in Plos Comp Biol, not too much is needed here.
-
-rm(list=ls())
-
-#***
-#External codes needed
-#***
-
-#packages needed (invoked with "::"): 
-
-#***
-#Location for storing results and other prep
-#***
-
-resloc<-"../Results/Plankton_DataAfterBasicCleaning/"
-if (!dir.exists(resloc))
-{
-  dir.create(resloc,recursive=TRUE)
-}
-
-#***
-#Load the data
-#***
-
-datloc<-"../Data/PlanktonData/"
-
-#load the phytoplankton color index (pci) data
-pci<-read.csv(file=paste0(datloc,"PCIannual.csv"),header=FALSE)
-pci_mat<-as.matrix(pci)
-sum(pci_mat==pci)
-prod(dim(pci))
-pci<-pci_mat
-rm(pci_mat)
-year<-unname(pci[1,2:(dim(pci)[2])])
-site<-pci[2:(dim(pci)[1]),1]
-pci<-pci[2:(dim(pci)[1]),2:(dim(pci)[2])]
-
-#load the cal fin dat
-calfin<-read.csv(file=paste0(datloc,"Calanus_finmarchicus.csv"),header=FALSE)
-calfin_mat<-as.matrix(calfin)
-sum(calfin_mat==calfin)
-prod(dim(calfin))
-calfin<-calfin_mat
-rm(calfin_mat)
-year_cf<-unname(calfin[1,2:(dim(calfin)[2])])
-testthat::expect_equal(year,year_cf)
-site_cf<-calfin[2:(dim(calfin)[1]),1]
-testthat::expect_equal(site,site_cf)
-calfin<-calfin[2:(dim(calfin)[1]),2:(dim(calfin)[2])]
-testthat::expect_equal(dim(calfin),dim(pci))
-rm(year_cf,site_cf)
-
-#load the growing season temperature data
-temp<-read.csv(file=paste0(datloc,"temp3to9.csv"),header=FALSE)
-temp_mat<-as.matrix(temp)
-sum(temp_mat==temp)
-prod(dim(temp))
-temp<-temp_mat
-rm(temp_mat)
-year_t<-unname(temp[1,2:(dim(temp)[2])])
-testthat::expect_equal(year,year_t)
-site_t<-temp[2:(dim(temp)[1]),1]
[…]

### 2021-12-22 Code/Plankton_AnalysisStep_LinearModelSelection.R

@@ -1,73 +1,0 @@
-#Does some linear models and information criterion analyses to select numbers of lags to use in 
-#ARMA-type models of plankton.
-
-rm(list=ls())
-
-#***
-#External codes needed
-#***
-
-#packages needed (invoked with "::"): ncf
-
-source("PickLags.R")
-
-#***
-#Locations for storing results, and other prep
-#***
-
-resloc<-"../Results/Plankton_LinearModelSelection/"
-if (!dir.exists(resloc))
-{
-  dir.create(resloc,recursive=TRUE)
-}
-
-#***
-#Load the data
-#***
-
-#load the data
-datloc<-"../Results/Plankton_DataAfterAllCleaning/"
-pci<-readRDS(paste0(datloc,"Pci_CleanedFinal.Rds")) 
-calfin<-readRDS(paste0(datloc,"Calfin_CleanedFinal.Rds")) 
-temp<-readRDS(paste0(datloc,"Temp_CleanedFinal.Rds")) 
-locs<-readRDS(paste0(datloc,"Site_CleanedFinal.Rds"))
-year<-readRDS(paste0(datloc,"Year_CleanedFinal.Rds"))
-times<-1:length(year)
-
-numts<-dim(pci)[1]
-lents<-dim(pci)[2]
-
-#***
-#Pick lags 
-#***
-
-lag_res<-picklags_mat(y=pci,x1=temp,x2=calfin,maxlags=c(1,1,2))
-
-lag_res$AICwt<-exp(-.5*lag_res$DeltaAIC)
-lag_res$AICwt<-lag_res$AICwt/sum(lag_res$AICwt)
-lag_res$BICwt<-exp(-.5*lag_res$DeltaBIC)
-lag_res$BICwt<-lag_res$BICwt/sum(lag_res$BICwt)
-
-lag_res_sortAIC<-lag_res[order(lag_res$AIC),]
-lag_res_sortBIC<-lag_res[order(lag_res$BIC),]
-lag_res_sortAIC$AICwt_cumsum<-cumsum(lag_res_sortAIC$AICwt)
-lag_res_sortBIC$BICwt_cumsum<-cumsum(lag_res_sortBIC$BICwt)
-
-head(lag_res_sortAIC,20)
-head(lag_res_sortBIC,20)
-
-range(lag_res$lagAR[lag_res$DeltaAIC<=3])
[…]

### 2021-12-22 Code/Plankton_AnalysisStep_LinearModels.R

@@ -1,230 +1,0 @@
-#Does some correlation analyses to confirm influences of temp and calfin on pci, and to get lags. 
-#Also finishes the data cleaning. The linear models done for kelp are skipped here, because we
-#ended up not needing them for kelp.
-
-rm(list=ls())
-
-#***
-#External codes needed
-#***
-
-#packages needed (invoked with "::"): wsyn, parallel
-
-source("MnCor_Tiny.R")
-
-#***
-#Locations for storing results, and other prep
-#***
-
-resloc1<-"../Results/Plankton_DataAfterAllCleaning/"
-if (!dir.exists(resloc1))
-{
-  dir.create(resloc1,recursive=TRUE)
-}
-
-resloc2<-"../Results/Plankton_LinearModelResults/"
-if (!dir.exists(resloc2))
-{
-  dir.create(resloc2,recursive=TRUE)
-}
-
-theseed<-101
-numsurrog<-1000
-saveRDS(theseed,paste0(resloc2,"theseed"))
-saveRDS(numsurrog,paste0(resloc2,"numsurrog"))
-
-#***
-#Load the data
-#***
-
-datloc<-"../Results/Plankton_DataAfterBasicCleaning/"
-pci<-readRDS(paste0(datloc,"Pci_CleanedBasic.Rds")) 
-calfin<-readRDS(paste0(datloc,"Calfin_CleanedBasic.Rds")) 
-temp<-readRDS(paste0(datloc,"Temp_CleanedBasic.Rds")) 
-year<-readRDS(paste0(datloc,"Year_CleanedBasic.Rds"))
-site<-readRDS(paste0(datloc,"Site_CleanedBasic.Rds"))
-
-numts<-dim(pci)[1]
-lents<-dim(pci)[2]
-
-# #***
-# #Get surrogates and create centered, appropriately normalized versions of everything - approach 1
-# #***
-# 
-# #centered (but not detrended) versions of the "environmental" variables (considered temp and calfin), 
-# #necessary to get surrog to work below
-# calfin_c<-wsyn::cleandat(calfin,year,1)$cdat
-# temp_c<-wsyn::cleandat(temp,year,1)$cdat
-# 
-# #Get surrogates for calfin and temp - use joint surrogates of both variables, synchrony-preserving, aaft.
-# #We do NOT detrend the data before surrogating because it makes the surrogates less good.
[…]

### 2021-12-22 Code/Plankton_AnalysisStep_PlotTimeSeries.R

@@ -1,80 +1,0 @@
-#Just plots time series after the "basic" cleaning (but prior to the final cleaning done in 
-#Plankton_AnalysisStep_LinearModels.R), for visual inspection.
-
-rm(list=ls())
-
-#***
-#Locations for storing results
-#***
-
-resloc<-"../Results/Plankton_TimeSeriesPlots/"
-if (!dir.exists(resloc))
-{
-  dir.create(resloc,recursive=TRUE)
-}
-
-#***
-#Load the data
-#***
-
-#load the data
-datloc<-"../Results/Plankton_DataAfterBasicCleaning/"
-pci<-readRDS(paste0(datloc,"Pci_CleanedBasic.Rds")) 
-calfin<-readRDS(paste0(datloc,"Calfin_CleanedBasic.Rds")) 
-temp<-readRDS(paste0(datloc,"Temp_CleanedBasic.Rds")) 
-year<-readRDS(paste0(datloc,"Year_CleanedBasic.Rds"))
-site<-readRDS(paste0(datloc,"Site_CleanedBasic.Rds"))
-
-numts<-dim(pci)[1]
-lents<-dim(pci)[2]
-
-#***
-#Make the plots
-#***
-
-pan.wd<-3
-pan.ht<-pan.wd
-gap<-0.2
-xax.ht<-.6
-yax.wd<-xax.ht
-tot.wd<-3*yax.wd+3*gap+3*pan.wd
-tot.ht<-xax.ht+pan.ht+gap
-
-for (counter in 1:(dim(pci)[1]))
-{
-  #plot the pci data
-  png(paste0(resloc,"SitePlot_",counter,".png"),res=300,units="in",width = tot.wd,height = tot.ht)
-  
-  #plot pci on the first set of axes
-  par(fig=c(yax.wd/tot.wd,
-            (yax.wd+pan.wd)/tot.wd,
-            (xax.ht)/tot.ht,
-            (xax.ht+pan.ht)/tot.ht),
-      mai=c(0,0,0,0),mgp=c(3,0.75,0))
-  plot(year,pci[counter,],type="l",col="green",xlab="year",ylab="pci") 
-  mtext("year",side=1,line=1.6)
-  mtext("pci",side=2,line=1.6)
-  
-  #temp on the second axes
-  par(fig=c((2*yax.wd+pan.wd+gap)/tot.wd,
-            (2*yax.wd+2*pan.wd+gap)/tot.wd,
[…]

### 2021-12-22 Code/Plankton_AnalysisStep_SpectralDecomp.R

@@ -1,374 +1,0 @@
-#Spectral analysis steps applied to plankton.
-
-rm(list=ls())
-
-#***
-#External codes needed
-#***
-
-#packages needed (invoked with "::"): parallel, wsyn, ncf, graphics, latex2exp
-source("SpectralTools.R")
-source("AvgOffDiags.R")
-source("PlotPhaseFunc.R")
-source("WhichBreak.R")
-source("plotColLine.R")
-
-#***
-#Location for storing results and other prep
-#***
-
-resloc<-"../Results/Plankton_SpectralDecompResults/"
-if (!dir.exists(resloc))
-{
-  dir.create(resloc,recursive=TRUE)
-}
-
-theseed<-101
-BiasVariance<-0.5
-
-saveRDS(theseed,file=paste0(resloc,"theseed.Rds"))
-saveRDS(BiasVariance,file=paste0(resloc,"BiasVariance.Rds"))
-
-#***
-#Load the data
-#***
-
-#load the data
-datloc<-"../Results/Plankton_DataAfterAllCleaning/"
-pci<-readRDS(paste0(datloc,"Pci_CleanedFinal.Rds")) 
-calfin<-readRDS(paste0(datloc,"Calfin_CleanedFinal.Rds")) 
-temp<-readRDS(paste0(datloc,"Temp_CleanedFinal.Rds")) 
-locs<-readRDS(paste0(datloc,"Site_CleanedFinal.Rds"))
-year<-readRDS(paste0(datloc,"Year_CleanedFinal.Rds"))
-times<-1:length(year)
-
-numts<-dim(pci)[1]
-lents<-dim(pci)[2]
-
-#***
-#The main function for doing the analysis 
-#***
-
-#Uses the spectral mathematics to get information about fractions of synchrony in pci explained by temperature, cal fin, 
-#and interactions. Assumes all the variables created above are in the workspace when run.
-#
-#Args - a list with these named elements
-#locstouse        Indices of locations to use. These refer to rows of the pci, temp and calfin data matrices above
-#lags             Max regression lags to use, pci then temp then calfin. In years. So if you want the model to
[…]

### 2021-12-22 Code/Kelp_AnalysisStep_CleanData_MoreInvolved.R
simplified the control flow to get rid of regression prelims which will be reported in the paper of Max and Lawrence

@@ -39,7 +39,7 @@ climinds<-readRDS(paste0(datloc,"Climinds_Quarterly_CleanedSuperBasic.Rds"))
 timesQ<-1:(dim(quarters)[1])
 
 #***
-#Slightly more extensive data prep - throw out sites with too many NAs or zeroes, then fill remaining NAs 
+#Slightly more extensive data prep - throw out sites with too many NAs or zeros, then fill remaining NAs 
 #with seasonal medians
 #***
 

### 2021-12-22 Code/Kelp_AnalysisStep_CleanData_SuperBasic.R

@@ -24,7 +24,7 @@ if (!dir.exists(resloc))
 #Load the data
 #***
 
-datloc<-"../Data/KelpData/"
+datloc<-"../Data/"
 
 #Load kelp data and change NaNs to NAs. Use the biomass data, sites ordered as in the 
 #original ordering (which should match with the environmental variables), not the paddling

### 2021-12-22 Code/Kelp_AnalysisStep_LinearModelSelection.R

@@ -1,161 +1,0 @@
-#Does some linear models and information criterion analyses to select numbers of lags to use in 
-#ARMA-type models of kelp.
-
-rm(list=ls())
-
-#***
-#External codes needed
-#***
-
-#packages needed (invoked with "::"): ncf
-
-source("PickLags.R")
-
-#***
-#Locations for storing results, and other prep
-#***
-
-resloc<-"../Results/Kelp_LinearModelSelection/"
-if (!dir.exists(resloc))
-{
-  dir.create(resloc,recursive=TRUE)
-}
-
-#***
-#Load the data
-#***
-
-#load the data
-datloc<-"../Results/Kelp_DataAfterAllCleaning/"
-kelp<-readRDS(paste0(datloc,"Kelp_Quarterly_CleanedFinal.Rds")) 
-NO3<-readRDS(paste0(datloc,"NO3_Quarterly_CleanedFinal.Rds")) 
-waves<-readRDS(paste0(datloc,"Waves_Quarterly_CleanedFinal.Rds")) 
-locs<-readRDS(paste0(datloc,"Locs_CleanedFinal.Rds"))
-quarters<-readRDS(paste0(datloc,"Quarters_CleanedFinal.Rds"))
-climinds<-readRDS(paste0(datloc,"Climinds_Quarterly_CleanedFinal.Rds"))
-times<-1:(dim(quarters)[1])
-
-numts<-dim(kelp)[1]
-lents<-dim(kelp)[2]
-
-#***
-#get your regions of CA, copied from another script
-#***
-
-#from google maps
-MontBLatLon<-c(36.62393626182818, -121.8511283749223) #Monterey Bay, closer to the southern end
-CarBLatLon<-c(36.52316357182972, -121.95535513873469) #The point to the south of Carmel Bay
-MorBLatLon<-c(35.427553615598676, -120.89133343717805) #Morro Bay, closer to the northern end
-PtConcLatLon<-c(34.44830424836278,-120.47125509005721) #Pt Conception
[…]

### 2021-12-22 Code/Kelp_AnalysisStep_LinearModels.R

@@ -1,488 +1,0 @@
-#Does some linear models to confirm earlier research indicating influences of NO3 and waves 
-#on kelp. Also finishes the data cleaning.
-#
-#The linear models just involve regressing kelp against waves and NO3 from the current time
-#and last quarter, getting R^2, and then comparing to the same using surrogate data.
-
-rm(list=ls())
-
-#***
-#External codes needed
-#***
-
-#packages needed (invoked with "::"): wsyn, parallel
-
-source("MnCoefDet_Tiny2.R")
-source("MnCor_Tiny.R")
-
-#***
-#Locations for storing results, and other prep
-#***
-
-resloc1<-"../Results/Kelp_DataAfterAllCleaning/"
-if (!dir.exists(resloc1))
-{
-  dir.create(resloc1,recursive=TRUE)
-}
-
-resloc2<-"../Results/Kelp_LinearModelResults/"
-if (!dir.exists(resloc2))
-{
-  dir.create(resloc2,recursive=TRUE)
-}
-
-theseed<-101
-numsurrog<-1000
-saveRDS(theseed,paste0(resloc2,"theseed"))
-saveRDS(numsurrog,paste0(resloc2,"numsurrog"))
-
-#***
-#Load the data
-#***
-
-#load the data
-datloc<-"../Results/Kelp_DataAfterMoreInvolvedCleaning/"
-kelp<-readRDS(paste0(datloc,"Kelp_Quarterly_CleanedMoreInvolved.Rds")) 
-kelpA<-readRDS(paste0(datloc,"KelpA_Quarterly_CleanedMoreInvolved.Rds")) 
-NO3<-readRDS(paste0(datloc,"NO3_Quarterly_CleanedMoreInvolved.Rds")) 
-waves<-readRDS(paste0(datloc,"Waves_Quarterly_CleanedMoreInvolved.Rds")) 
-locs<-readRDS(paste0(datloc,"Locs_CleanedMoreInvolved.Rds"))
-quarters<-readRDS(paste0(datloc,"Quarters_CleanedMoreInvolved.Rds"))
-climinds<-readRDS(paste0(datloc,"Climinds_Quarterly_CleanedMoreInvolved.Rds"))
-times<-1:(dim(quarters)[1])
-
-numts<-dim(kelp)[1]
-lents<-dim(kelp)[2]
-
-#***
-#Get surrogates and create centered, appropriately normalized versions of everything
[…]

### 2021-12-22 Code/main.R

@@ -17,10 +17,8 @@ source("Kelp_AnalysisStep_CleanData_MoreInvolved.R")
 #Plot kelp time series for visual inspection
 source("Kelp_AnalysisStep_PlotTimeSeries.R")
 
-#Do linear models with kelp and related data to help demo the need for specific 
-#lags - this also performs and saves the final level of data cleaning
-source("Kelp_AnalysisStep_LinearModels.R")
-source("Kelp_AnalysisStep_LinearModelSelection.R")
+#Final cleaning
+source("Kelp_AnalysisStep_CleanData_Final.R")
 
 #Do the spectra decomposition analysis for kelp
 source("Kelp_AnalysisStep_SpectralDecomp.R")

### 2021-12-22 Code/MnCoefDet_Tiny2.R
removed files no longer needed because regression prelims to be done as part of other paper

@@ -1,49 +1,0 @@
-#Computes the mean coefficient of determination of a linear regression computed for each location, as well as
-#mean coefficients.
-#
-#Args
-#kelp           Locations by time matrix for kelp data
-#NO3            Same for nitrates
-#waves          Same for waves
-#
-#Output
-#For each location, does the linear regression with lags 0 and 1 for waves and NO3, and gets the coefficient of 
-#determination (R^2). Returns the average of these values across all locations. Also returns the average NO3 
-#coefficients in the regressions, also returns the average waves coefficients in the regressions.
-#
-mncoefdet_tiny2<-function(kelp,NO3,waves)
-{
-  if (any(dim(kelp)!=dim(NO3)) || any(dim(kelp)!=dim(waves)))
-  {
-    stop("Error in mncoefdet_tiny2: kelp, NO3 and waves input matrices must be the same dimensions")
-  }
-
-  numts<-dim(kelp)[1]
-  lents<-dim(kelp)[2]
-  
-  res_Rsq<-NA*numeric(numts)
-  res_NO3coefs<-matrix(NA,numts,2)
-  res_wavescoefs<-matrix(NA,numts,2)
-  for (counter in 1:numts)
-  {
-    #extract time series for this location
-    tskelp<-kelp[counter,]
-    tsNO3<-NO3[counter,]
-    tswaves<-waves[counter,]
-    
-    #implement the desired lags
-    tskelp<-tskelp[2:lents]
-    tsNO3_l0<-tsNO3[2:lents]
-    tsNO3_l1<-tsNO3[1:(lents-1)]
-    tswaves_l0<-tswaves[2:lents]
-    tswaves_l1<-tswaves[1:(lents-1)]
-    
-    #do the regressiom
-    mod<-lm(tskelp~tsNO3_l0+tsNO3_l1+tswaves_l0+tswaves_l1)
-    res_Rsq[counter]<-summary(mod)$r.squared
-    res_NO3coefs[counter,]<-unname(coef(mod)[2:3])
-    res_wavescoefs[counter,]<-unname(coef(mod)[4:5])
-  }
-  
-  return(c(mean(res_Rsq),apply(FUN=mean,X=res_NO3coefs,MARGIN=2),apply(FUN=mean,X=res_wavescoefs,MARGIN=2)))
-}

### 2021-12-22 Code/MnCor_Tiny.R

@@ -1,31 +1,0 @@
-#A function for computing the mean Pearson correlation between rows of two matrices, with a lag
-#
-#Args
-#m1, m2       The two matrices, must be the same size
-#lag          The lag. Positive values only. Means you're computing correlations between m1[i,(lag+1):(dim(m1)[2])] and m2[i,1:(dim(m2)[2]-lag)]
-#
-#Output
-#A single number, the mean correlation between rows of m1 and corresponding rows of m2
-mncor_tiny<-function(m1,m2,lag)
-{
-  numts<-dim(m1)[1]
-  lents<-dim(m1)[2]
-
-  #error catching
-  if (any(dim(m1)!=dim(m2)))
-  {
-    stop("Error in mncor_tiny: m1 and m2 must have the same dimension")
-  }
-  if (lag<0)
-  {
-    stop("Error in mncor_tiny: lag must be >=0")
-  }
-  
-  res<-NA*numeric(numts)
-  for (counter in 1:numts)
-  {
-    res[counter]<-cor(m1[counter,(lag+1):lents],m2[counter,1:(lents-lag)],method="pearson")
-  }
-  
-  return(mean(res))
-}

### 2021-12-22 Code/PickLags.R

@@ -1,255 +1,0 @@
-#This function picks lags in a linear regression setup relevant to the project. Read the specs and code for details.
-#Just picks lags for each location, one by one, using picklags_vect, below, and then aggregates across locations. 
-#
-#Args
-#y          An N by T matrix representing time series, N=number of locations, T=time series length. The population variable.
-#x1         An N by T matrix of time series. An environmental driver.
-#x2         Another such
-#maxlags    A vector of length 3. maxlags[1]>=0 is the max AR lag for the population variable, 0 meaning populations 
-#             do not depend on past populations. maxlags[2]>=-1 is the max lag for x1, -1 means x1 does not
-#             influence the populations. maxlags[3]>=-1 is similar, but for x2.
-#
-#Output - a data frame with the following columns
-#lagAR      The AR lag considered for the populations, >=0, meaning of 0 described above
-#lag1       The lag for x1 considered, >=-1, the meaning of -1 described above
-#lag2       Similar but for x2
-#Rsq        Rsq of the corresponding regression model, averaged across locations
-#AIC        AIC of the corresponding regression model, averaged across locations
-#BIC        BIC of the corresponding regression model, averaged across locations
-#
-#Notes
-#When a given lag is included in the linear regression model, all smaller lags are automatically included.
-#
-picklags_mat<-function(y,x1,x2,maxlags)
-{
-  #compute results for each location and compute average results
-  res_tot<-picklags_vect(y[1,],x1[1,],x2[1,],maxlags)
-  for (counter in 2:(dim(y)[1]))
-  {
-    res_loc<-picklags_vect(y[counter,],x1[counter,],x2[counter,],maxlags)
-    res_tot[,4:6]<-res_tot[,4:6]+res_loc[,4:6]
-  }
-  res<-res_tot
-  res[,4:6]<-res[,4:6]/(dim(y)[1])
-  
-  #put in the delta AIC and BIC columns
-  res$DeltaAIC<-res$AIC-min(res$AIC)
-  res$DeltaBIC<-res$BIC-min(res$BIC)
-  
-  return(res)
-}
-
-#This function picks lags in a linear regression setup relevant to the project. Read the specs and code for details.
-#This is for one location only, the above function iterates across locations.
-#
-#Args
[…]

### 2021-12-22 Code/tests_MnCoefDet_Tiny2.R

@@ -1,28 +1,0 @@
-context("MnCoefDet_Tiny2.R")
-
-source("MnCoefDet_Tiny2.R")
-
-test_that("test it",{
-  #test error catching
-  testthat::expect_error(mncoefdet_tiny2(kelp=array(1,c(2,2)),NO3=array(2,c(2,3)),waves=array(3,c(2,2))),
-                         "Error in mncoefdet_tiny2: kelp, NO3 and waves input matrices must be the same dimensions")
-  testthat::expect_error(mncoefdet_tiny2(kelp=array(1,c(2,2)),NO3=array(2,c(2,2)),waves=array(3,c(2,3))),
-                         "Error in mncoefdet_tiny2: kelp, NO3 and waves input matrices must be the same dimensions")
-
-  #now test some cases with legit inputs
-  set.seed(101)
-  kelp<-matrix(rnorm(1000),10,100)
-  NO3<-matrix(rnorm(1000),10,100)
-  waves<-matrix(rnorm(1000),10,100)
-  
-  res<-mncoefdet_tiny2(kelp,NO3,waves)
-  res2<-numeric(5)
-  for (counter in 1:10)
-  {
-    m<-lm(kelp[counter,2:100]~NO3[counter,2:100]+NO3[counter,1:99]+waves[counter,2:100]+waves[counter,1:99])
-    res2[1]<-res2[1]+summary(m)$r.squared
-    res2[2:3]<-res2[2:3]+unname(coef(m)[2:3])
-    res2[4:5]<-res2[4:5]+unname(coef(m)[4:5])
-  }
-  testthat::expect_equal(res,res2/10)
-})

### 2021-12-22 Code/tests_MnCor_Tiny.R

@@ -1,25 +1,0 @@
-context("MnCor_Tiny.R")
-
-source("MnCor_Tiny.R")
-
-test_that("test it",{
-  #test error catching
-  m1<-matrix(rnorm(10),2,5)
-  m2<-matrix(rnorm(10),5,2)
-  testthat::expect_error(mncor_tiny(m1,m2,0),"Error in mncor_tiny: m1 and m2 must have the same dimension")
-  m2<-t(m2)
-  testthat::expect_error(mncor_tiny(m1,m2,-1),"Error in mncor_tiny: lag must be >=0")
-  
-  #test some cases
-  set.seed(101)
-  m1<-matrix(rnorm(200),c(2,100))
-  m2<-array(rnorm(200),c(2,100))
-  
-  res<-mncor_tiny(m1,m2,0)
-  res2<-(cor(m1[1,],m2[1,])+cor(m1[2,],m2[2,]))/2
-  testthat::expect_equal(res,res2)
-  
-  res<-mncor_tiny(m1,m2,1)
-  res2<-(cor(m1[1,2:100],m2[1,1:99])+cor(m1[2,2:100],m2[2,1:99]))/2
-  testthat::expect_equal(res,res2)
-})

### 2021-12-22 Code/tests_PickLags.R

@@ -1,128 +1,0 @@
-context("PickLags.R")
-
-source("PickLags.R")
-
-test_that("test it",{
-  
-  #***test the function makedf
-  
-  #make up some data and call the function
-  set.seed(101)
-  T<-10
-  y<-rnorm(T)
-  x1<-rnorm(T)
-  x2<-rnorm(T)
-  maxlags<-c(3,2,1)
-  dat<-makedf(y,x1,x2,maxlags)
-
-  #test the column names and dimensions of the result
-  testthat::expect_equal(names(dat),c("y","y_l1","y_l2","y_l3","x1_l0","x1_l1","x1_l2","x2_l0","x2_l1"))
-  testthat::expect_equal(dim(dat)[1],7)
-
-  #test the response variable column
-  testthat::expect_equal(dat$y,y[4:10])
-
-  #test the AR columns
-  testthat::expect_equal(dat$y_l1,as.vector(y[3:9]))
-  testthat::expect_equal(dat$y_l2,as.vector(y[2:8]))
-  testthat::expect_equal(dat$y_l3,as.vector(y[1:7]))
-  
-  #test the x1 columns
-  testthat::expect_equal(dat$x1_l0,x1[4:10])
-  testthat::expect_equal(dat$x1_l1,x1[3:9])
-  testthat::expect_equal(dat$x1_l2,x1[2:8])
-  
-  #test the x2 columns
-  testthat::expect_equal(dat$x2_l0,x2[4:10])
-  testthat::expect_equal(dat$x2_l1,x2[3:9])
-  
-  #***test the function makeformula
-  
-  #test a generic case
-  res<-makeformula(3,2,1)
-  testthat::expect_equal(class(res),"formula")
-  testthat::expect_equal(res,as.formula("y~y_l1+y_l2+y_l3+x1_l0+x1_l1+x1_l2+x2_l0+x2_l1-1"))
-  
-  #test some boundary cases - one lag set at a null value
-  res<-makeformula(0,2,1)
-  testthat::expect_equal(class(res),"formula")
-  testthat::expect_equal(res,as.formula("y~x1_l0+x1_l1+x1_l2+x2_l0+x2_l1-1"))
-
-  res<-makeformula(2,-1,1)
-  testthat::expect_equal(class(res),"formula")
[…]

### 2021-12-22 Paper/Paper.Rnw
small edits while reading the paper to get back up to speed

@@
 %header material for paper
-\title{How environmental drivers of spatial synchrony interact}
+\title{How environmental drivers of synchrony interact}
 \date{}
@@
 %Environmental drivers of spatial synchrony interact
+%How environmental drivers of spatial synchrony interact
 
@@
 \section{Theory}
-
-%Tentative broad outline:
-%DONE 1) Intro spectral quantities, the spectrum, the cospectrum, and the cross spectrum of two time series/processes. And the spectral
-%matrix for putting them all in one structure.
-%DONE 2) Define your measures of timescale/frequency-specific synchrony and crosss-variable synchrony, explain why these make sense as
-%ways to measure those things.
-%DONE 3) Present the model and the theory
-%4) Present "toy" examples and fig 1 (need to think a lot about these).
-DAN: perhaps these actually belong in Methods (describe the specific-case models we will consider) and results (apply the general theory to those specific cases)
-%
-%Notes:
-%Compared to the deer paper, this is a much deeper presentation of the theory in the main text. I'm not sure that's avoidable.
-%For one thing, the theory is explicitly applied to get the results, later, and that was not the case for the deer paper.
 
@@
 four model time steps per year. This makes no real mathematical difference, but was done to 
-facilitate later comparisons with results based on kelp data, which were available on a
+facilitate later comparisons with results based on kelp data, which were available with a
 quarterly sampling frequency. 
@@
 $(\epsilon^{(1)},\epsilon^{(2)})$ is assumed to be a Gaussian white-noise process 
-for both A and B, i.e., 
+for both A and B. So 
 the random variables $(\epsilon^{(1)}(t),\epsilon^{(2)}(t))$ are independent and 
@@
 %Tentative broad outline
+%0) above stuff about the case studies 
 %1) Para from Tom on how remote sensing data are gathered and coverted into what he handed to me. Brief, mostly 
@@
 %data, especially if you make this paper also be about plankton. But do that later if you do it.
-%4) Some paras about the plankton data, similar structure to how the kelp data were presented (and using sup mat as described)
 %5) Para about the linear modelling - this is how certain parts of the main theoretical equation were estimated
@@
 also follow from the results of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not necessarily
-straightforwardly. Additionally, those are case studies, so our results also provide novel theoretical generality;
+straightforwardly. Additionally, those are case studies, so our results provide novel theoretical generality;
 and our theory can be interpreted to provide intuition about how interacting Moran effects work,
@@
 Case study B reveals how Moran effects can interact when lags are longer than 1 model time step. 
-For B, $\epsilon^{(1)}$ effects on populations were lagged by 3 quarters but 
-$\epsilon^{(2)}$ effects were unlagged (Methods). Thus interactions between Moran effects were 
+For B, $\epsilon^{(1)}$ effects on populations were lagged by $n$ quarters but 
+$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$. Thus interactions between Moran effects were 
 negative on the shortest timescales 
-(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, left side of panel). This is because the 
+(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, far left side of panel). This is because the 
 population influences of Moran effects of timescale 2
@@
 the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-is the same as that of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f); 
+is the same as that of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}g); 
 that common phase then controls the sign of 
@@
 is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
-Our theory captures these complexities because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
+Our theory captures these complexities
+because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
 quantifies the relative phase lag between the effects of $\epsilon^{(1)}$ and 
@@
 $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the phase lag between 
-oscillations in the variables themselves, relative to the timescale.
+oscillations in the variables themselves, relative to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
 The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
@@
 the variables themselves align or misalign, with the degree of alignment
-reflected in the real part of this quantity.
+reflected in the real part of this quantity. On long timescales, interactions were the
+same in the two secenaios, because on those timescales sub-annual lags in environmental 
+peaks and effects on populations make negligible difference.
 
@@
 which is population synchrony due to interacting Moran effects.
-%***DAN: More text here as we flesh out the presentation - need to describe the middle two columns of panels 
+Middle panels show the constituent parts of the theory (labelled). For these complex functions 
+of timescale, the magnitude is displayed on the vertical axis and the phase is displayed with color.
+See text for all interpretations.
 The functions plotted on d, h, l, p are those on a, e, i, m, respectively, times 
 $|f_{\mcB}|^2$, thereby showing the influence of autoregressive population effects on
-synchrony [see (\ref{rhoww_T1to3})-(\ref{rhoww_OC})].
+synchrony [see (\ref{rhoww_T1to3})].
 Model parameters were: $N=5$ and $b_1=0.4$ for all case studies; 
@@
 parameters associated with the noise for each case study,
-and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.}\label{fig:TheoryCaseA}
+and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.
+Synch. contrib. = Synchrony contribution.}\label{fig:TheoryCaseA}
 \end{figure}

### 2021-12-22 Code/Kelp_AnalysisStep_CleanData_Final.R
a file I forgot to add a couple commits ago

@@ -1,0 +1,86 @@
+#Does the final steps in cleaning. Starting from the results of the more involved cleaning,
+#just normalizes kelp time series to an estimate of density per unit usable habitat, and then
+#detrends all variables without variance standaridizing.
+
+rm(list=ls())
+
+#***
+#External codes needed
+#***
+
+#packages needed (invoked with "::"): 
+
+#***
+#Locations for storing results, and other prep
+#***
+
+resloc1<-"../Results/Kelp_DataAfterAllCleaning/"
+if (!dir.exists(resloc1))
+{
+  dir.create(resloc1,recursive=TRUE)
+}
+
+#***
+#Load the data
+#***
+
+#load the data
+datloc<-"../Results/Kelp_DataAfterMoreInvolvedCleaning/"
+kelp<-readRDS(paste0(datloc,"Kelp_Quarterly_CleanedMoreInvolved.Rds")) 
+kelpA<-readRDS(paste0(datloc,"KelpA_Quarterly_CleanedMoreInvolved.Rds")) 
+NO3<-readRDS(paste0(datloc,"NO3_Quarterly_CleanedMoreInvolved.Rds")) 
+waves<-readRDS(paste0(datloc,"Waves_Quarterly_CleanedMoreInvolved.Rds")) 
+locs<-readRDS(paste0(datloc,"Locs_CleanedMoreInvolved.Rds"))
+quarters<-readRDS(paste0(datloc,"Quarters_CleanedMoreInvolved.Rds"))
+climinds<-readRDS(paste0(datloc,"Climinds_Quarterly_CleanedMoreInvolved.Rds"))
+times<-1:(dim(quarters)[1])
+
+numts<-dim(kelp)[1]
+lents<-dim(kelp)[2]
+
+#***
+#Do the cleaning
+#***
+
+#detrend but don't variance standardize the environmental variables
+NO3<-wsyn::cleandat(NO3,times,2)$cdat
+waves<-wsyn::cleandat(waves,times,2)$cdat
+
+#detrend climate indices
+climinds$NPGO<-wsyn::cleandat(climinds$NPGO,times,2)$cdat
+climinds$MEI<-wsyn::cleandat(climinds$MEI,times,2)$cdat
+climinds$PDO<-wsyn::cleandat(climinds$PDO,times,2)$cdat
+
+#Compute a kelp density per unit useable habitat. To do this, take the max kelp area value ever
[…]

### 2021-12-22 Code/Kelp_AnalysisStep_SpectralDecomp.R
started refactoring spectral code, small steps so far, larger steps in subsequent commits

@@ -1,6 +1,4 @@
-#For computations associated with some spectral modelling math I did beginning on 2020 02 01.
-#This generalizes (through functionization and other steps) the codes in 
-#Approach_ModellingSpectral_MonterreyToMorro.R and Approach_ModellingSpectral_SantaBarbaraArea.R
+#Applies the spectral theory to the kelp data.
 
 rm(list=ls())
 
@@ -88,6 +86,35 @@ OxInd<-which(dists==min(dists))
 dists<-ncf::gcdist(x=c(LALatLon[2],locs$Lon),y=c(LALatLon[1],locs$Lat))
 dists<-dists[1,2:(dim(dists)[2])]
 LAInd<-which(dists==min(dists))
+
+#***
+#Select some regions to focus on
+#***
+
+#construct a region from Pt Conception to Oxnard
+SBlocstouse<-PtConcInd:OxInd
+plot(locs$Lon,locs$Lat,type="p",pch=20,cex=.5)
+points(locs$Lon[SBlocstouse],locs$Lat[SBlocstouse],type="p",pch=20,cex=.5,col="red")  
+length(SBlocstouse) #60 locations
+h<-c(SBlocstouse[1],SBlocstouse[length(SBlocstouse)])
+ncf::gcdist(locs$Lon[h],locs$Lat[h]) #103.8km
+
+#First look at the whole region from Carmel Bay to Morro Bay
+MorBInd-CarBInd+1 #164 locations
+h<-c(CarBInd,MorBInd)
+ncf::gcdist(locs$Lon[h],locs$Lat[h]) #152.8km
+#it's a bit too long (compared to the above region) and has too many sites, so split it
+
+CC1locstouse<-CarBInd:floor(mean(c(CarBInd,MorBInd)))
+CC2locstouse<-ceiling(mean(c(CarBInd,MorBInd))):MorBInd
+length(CC1locstouse) #82 sites
+length(CC2locstouse) #82 sites
+h<-c(CC1locstouse[1],CC1locstouse[length(CC1locstouse)])
+ncf::gcdist(locs$Lon[h],locs$Lat[h]) #73.35km
+h<-c(CC2locstouse[1],CC2locstouse[length(CC1locstouse)])
+ncf::gcdist(locs$Lon[h],locs$Lat[h]) #78.46km
+
+#So my three roughly comparable regions are SBlocstouse, CC1locstouse, CC2locstouse
 
 #***
[…]

### 2022-01-06 Code/IntroTheoryIA.R
added code for pedagogical figure for the Intro

@@ -1,0 +1,49 @@
+#This function calculates the strength of interaction effects in the simple example of SI section S1
+#of Sheppard et al (2019, Plos Comp Biol). Notation is the same as that paper.
+#
+#Args
+#a, b       The coefficients, see Sheppard et al (2019)
+#covmat     The covariance matrix of (alpha(1),alpha(2),beta(1),beta(2)). Must have 1s down the 
+#             diagonal, if only for consistency with the same simplifying assumption made in
+#             Sheppard et al (2019), SI section S1.
+#
+#Output - a length-2 named vector with these components 
+#sync       The population synchrony of the model (measured with correlation)
+#sync.ni    What the population synchrony would be if the two Moran drivers were independent.
+#
+#Note
+#Returns NA for sync if covmat is not positive definite, returns NA for sync.in if the analogous
+#covariance matrix for is is not positive definite.
+#
+IntroTheoryIA<-function(a,b,covmat)
+{
+  if (any(diag(covmat)!=1))
+  {
+    stop("Error in IntroTheoryIA: covmat should have diagonal entries equal to 1")
+  }
+
+  if (matrixcalc::is.positive.definite(covmat))
+  {
+    res1<-(a^2*covmat[1,2]+a*b*covmat[1,4]+a*b*covmat[2,3]+b^2*covmat[3,4])/
+      (sqrt(a^2+b^2+2*a*b*covmat[1,3])*sqrt(a^2+b^2+2*a*b*covmat[2,4]))
+  } else
+  {
+    res1<-NA  
+  }
+  
+  redcovmat<-covmat
+  redcovmat[1,3:4]<-0
+  redcovmat[2,3:4]<-0
+  redcovmat[3,1:2]<-0
+  redcovmat[4,1:2]<-0
+  if (matrixcalc::is.positive.definite(redcovmat))
+  {
+    res2<-(a^2*redcovmat[1,2]+a*b*redcovmat[1,4]+a*b*redcovmat[2,3]+b^2*redcovmat[3,4])/
+      (sqrt(a^2+b^2+2*a*b*redcovmat[1,3])*sqrt(a^2+b^2+2*a*b*redcovmat[2,4]))
+  } else
+  {
+    res2<-NA
+  }
+  
+  return(c(sync=res1,sync.ni=res2))
+}

### 2022-01-06 Code/Theory_IntroExample.R

@@ -1,0 +1,111 @@
+#This script implements very simple theoretical model explored in the Introduction
+
+rm(list=ls())
+
+#***
+#External codes needed
+#***
+
+#packages needed (invoked with "::"): matrixcalc, shape, latex2exp
+
+source("IntroTheoryIA.R")
+
+#***
+#Location for storing results and other prep
+#***
+
+resloc<-"../Results/Theory/"
+if (!dir.exists(resloc))
+{
+  dir.create(resloc,recursive=TRUE)
+}
+
+#***
+#Use the function to explore some examples of interactions, looking for a synergistic example
+#and a destuctive example, and save those results for possible inclusion in the Intro
+#***
+
+a<-1
+b<-1
+
+#synergistic example
+covmat<-matrix(c(1,.7,.5,.5,
+                 .7,1,.5,.5,
+                 .5,.5,1,.7,
+                 .5,.5,.7,1),4,4,byrow=TRUE)
+res_syn<-list(a=a,b=b,covmat=covmat,res=IntroTheoryIA(a,b,covmat))
+saveRDS(res_syn,file=paste0(resloc,"IntroTheory_SynergisticExample.Rds"))
+
+#destructive example
+covmat<-matrix(c(1,.7,-.5,-.5,
+                 .7,1,-.5,-.5,
+                 -.5,-.5,1,.7,
+                 -.5,-.5,.7,1),4,4,byrow=TRUE)
+res_des<-list(a=a,b=b,covmat=covmat,res=IntroTheoryIA(a,b,covmat))
+saveRDS(res_des,file=paste0(resloc,"IntroTheory_DestructiveExample.Rds"))
+
+#***
+#Now explore a one parameter family of possibilities and plot it
+#***
+
+#get the result
+lo<-101
+od<-seq(from=-1,to=1,length.out=lo)
+res<-data.frame(od=od,sync=NA,sync.in=NA)
+for (counter in 1:(dim(res)[1]))
+{
+  covmat<-matrix(c(1,.7,od[counter],od[counter],
+                   .7,1,od[counter],od[counter],
+                   od[counter],od[counter],1,.7,
+                   od[counter],od[counter],.7,1),4,4,byrow=TRUE)
+  h<-IntroTheoryIA(a=1,b=1,covmat=covmat)
[…]

### 2022-01-06 Code/Theory_PedagogFig.R

@@ -1,0 +1,168 @@
+#This script makes a pedagogical figure. No modelling here, just presentation.
+
+rm(list=ls())
+
+#***
+#External codes needed
+#***
+
+#packages needed (invoked with "::"): shape
+
+#***
+#Location for storing results and other prep
+#***
+
+resloc<-"../Results/Theory/"
+if (!dir.exists(resloc))
+{
+  dir.create(resloc,recursive=TRUE)
+}
+
+#***
+#Make the figure
+#***
+
+#dimensions for the figure to be generated, inches
+totwd<-3.5
+totht<-totwd
+gap<-0.15
+xaxht<-0.5
+panwd<-(totwd-2*gap)
+panht<-(totht-2*gap-xaxht)/2
+
+pdf(file=paste0(resloc,"PedagogFig.pdf"),height=totht,width=totwd)
+
+#***part 1, which shows synergistic effects
+
+#set up the panel
+par(fig=c((gap)/totwd,
+          (gap+panwd)/totwd,
+          (xaxht+gap+panht)/totht,
+          (xaxht+gap+2*panht)/totht),
+    mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
+
+#get and plot the sin waves
+tim<-seq(from=0,to=100,length.out=10001)
+period<-20
+phi1<-(-1.5)
+phi2<-phi1+pi/2
+sin1<-sin(2*pi*tim/period+phi1)
+sin2<-sin(2*pi*tim/period+phi2)
+offset1<-0
+offset2<-5
+y1<-sin1+offset1
+y2<-sin2+offset2
+
+xlimits<-range(tim)
+ylimits<-range(y1,y2)
+dy<-diff(ylimits)
+ylimits[2]<-ylimits[2]+.15*dy
+ylimits[1]<-ylimits[1]-.01*dy
+
+plot(tim,y1,type="l",xlim=xlimits,ylim=ylimits,xaxs="i",xaxt="n",yaxt="n")
+lines(tim,y2,type="l")
+
+#label populations
+ep<-0.4
+ctr<-mean(c(offset1,offset2))
+rect(xlimits[1],ctr-ep,xlimits[2],ctr+ep,col="lightgrey",border=NA)
+text(17,ctr,latex2exp::TeX("population, $w_i(t)"),adj=c(0,.5),cex=0.8)
+
[…]

### 2022-01-06 Code/main.R

@@ -9,6 +9,8 @@ graphics.off()
 #Develop the theoretical case studies
 source("Theory_CaseStudyAB.R")
 source("Theory_CaseStudyC.R")
+source("Theory_IntroExample.R")
+source("Theory_PedagogFig.R")
 
 #Clean the kelp data
 source("Kelp_AnalysisStep_CleanData_SuperBasic.R")

### 2022-01-06 Code/tests_IntroTheoryA.R

@@ -1,0 +1,29 @@
+context("IntroTheoryIA.R")
+
+source("IntroTheoryIA.R")
+
+test_that("test the error catching",{
+  a<-1
+  b<-1
+  covmat<-diag(4)
+  covmat[1,1]<-.9
+  testthat::expect_error(IntroTheoryIA(a,b,covmat),"Error in IntroTheoryIA: covmat should have diagonal entries equal to 1")
+})
+
+test_that("test for the values given in Sheppard et al (2019), Plos Comp Biol, SI section S1",{
+  a<-1
+  b<-1
+
+  covmat<-matrix(1,4,4)  
+  covmat[1,4]<-1.1
+  covmat[4,1]<-1.1
+  h<-IntroTheoryIA(a,b,covmat)
+  testthat::expect_true(is.na(h['sync']))
+  
+  covmat<-matrix(c(1,.9,.5,.5,
+                   .9,1,.5,.5,
+                   .5,.5,1,.5,
+                   .5,.5,.5,1),4,4,byrow = TRUE)
+  testthat::expect_equal(IntroTheoryIA(a,b,covmat),c(sync=0.8,sync.ni=0.7))
+})
+  

### 2022-01-06 Code/Theory_CaseStudyAB.R
changes to theoretical results to save numerics to they can be uploaded into the paper automatically

@@ -361,6 +361,10 @@ TheoryCaseAB<-function(b,p1,p2,N,ns,PlotName,PanLabs,XAxOn=TRUE)
   dev.off()
 }
 
+#***
+#The main code, uses the above function
+#***
+  
 #Case A
 N<-5
 ns<-c(.6,.6,.3)
@@ -368,8 +372,13 @@ p1<-c(0,1.1)
 p2<-c(0.8)
 b<-c(.4)
 PlotName<-paste0(resloc,"TheoryFigCaseA")
-PanLabs<-c("a","b","c","d")
+PanLabs<-c("(a)","(b)","(c)","(d)")
 TheoryCaseAB(b,p1,p2,N,ns,PlotName,PanLabs,XAxOn=FALSE)
+saveRDS(N,file=paste0(resloc,"Params_A_N.rds"))
+saveRDS(ns,file=paste0(resloc,"Params_A_ns.rds"))
+saveRDS(p1,file=paste0(resloc,"Params_A_p1.rds"))
+saveRDS(p2,file=paste0(resloc,"Params_A_p2.rds"))
+saveRDS(b,file=paste0(resloc,"Params_A_b.rds"))
 
 #Case B
 N<-5
@@ -378,5 +387,10 @@ p1<-c(0,0,0,1.1)
 p2<-c(0.8)
 b<-c(.4)
 PlotName<-paste0(resloc,"TheoryFigCaseB")
-PanLabs<-c("e","f","g","h")
+PanLabs<-c("(e)","(f)","(g)","(h)")
 TheoryCaseAB(b,p1,p2,N,ns,PlotName,PanLabs,XAxOn=FALSE)
+saveRDS(N,file=paste0(resloc,"Params_B_N.rds"))
+saveRDS(ns,file=paste0(resloc,"Params_B_ns.rds"))
+saveRDS(p1,file=paste0(resloc,"Params_B_p1.rds"))
+saveRDS(p2,file=paste0(resloc,"Params_B_p2.rds"))
+saveRDS(b,file=paste0(resloc,"Params_B_b.rds"))

### 2022-01-06 Code/Theory_CaseStudyC.R

@@ -26,10 +26,6 @@ if (!dir.exists(resloc))
 #The main function for doing both case studies
 #***
 
-#***
-#The main function for doing both case studies
-#***
-
 #This function calculates the analytic results and plots them for case study C.
 #The function is actually more general than the planned case study that will be 
 #displayed in the paper.
@@ -750,16 +746,25 @@ TestNoise<-function(N,ns,PlotName,numsims,BiasVariance)
   dev.off()
 }
 
+#***
+#The main code, uses the above function
+#***
+
 #Case C, parameter set 1
 N<-5
-ns<-c(0,-4/9,0,1,.6,.6,.3) #c1, c2, l1, l2, w11, w22, w12.
+ns<-c(0,-4/9,1,0,.6,.6,.3) #c1, c2, l1, l2, w11, w22, w12.
 p1<-c(0,3)
-p2<-c(1.5)
+p2<-c(-1.5)
 b<-c(.4)
 PlotName<-paste0(resloc,"TheoryFigCaseC_v01")
-PanLabs<-c("i","j","k","l")
-PanLabs2<-c("a","b","c")
+PanLabs<-c("(i)","(j)","(k)","(l)")
+PanLabs2<-c("(a)","(b)","(c)")
 TheoryCaseC(b,p1,p2,N,ns,PlotName,PanLabs,PanLabs2,SyncYAx=c(-10,20),RelSyncYAx=c(-10,25),XAxOn=FALSE)
+saveRDS(N,file=paste0(resloc,"Params_C1_N.rds"))
+saveRDS(ns,file=paste0(resloc,"Params_C1_ns.rds"))
+saveRDS(p1,file=paste0(resloc,"Params_C1_p1.rds"))
+saveRDS(p2,file=paste0(resloc,"Params_C1_p2.rds"))
+saveRDS(b,file=paste0(resloc,"Params_C1_b.rds"))
 
 #now do the test for consistency between analytic and simulation approaches for the noise, essentially
 #a test that my estimators of synchrony quantities are good estimators of my analytically calculated 
@@ -770,14 +775,19 @@ BiasVariance<-0.5
 TestNoise(N,ns,PlotName,numsims,BiasVariance)
 
 #Case study C, parameter set 2
-ns<-c(0,-4/9,1,0,.6,.6,.3) #c1, c2, l1, l2, w11, w22, w12.
[…]

### 2022-01-06 Code/Kelp_AnalysisStep_SpectralDecomp.R
fiddling with presentation of figures

@@ -1802,7 +1802,8 @@ make_plot_component<-function(freq,frg,comp,plotname,plottype)
 #Now do the analysis for these three regions and save
 #***
 
-PlotForMSArgs<-list(cov_MainPlot_PanLabs=c("A","B","C"),
+#Calculations for all three regions using lags c(4,1,0)
+PlotForMSArgs<-list(cov_MainPlot_PanLabs=c("(a)","(b)","(c)"),
                     cov_MainPlot_short_ylim=NA, #means let the data determine the ylimits
                     cov_MainPlot_mid_ylim=NA,
                     cov_MainPlot_long_ylim=NA,
@@ -1816,12 +1817,12 @@ PlotForMSArgs<-list(cov_MainPlot_PanLabs=c("A","B","C"),
                     ExplainDirectMoran_fP2_mid_ylims=c(.01,.08),
                     ExplainDirectMoran_fP2_long_ylims=c(.01,.08),
                     ExplainDirectMoran_Se2e2_short_ylims=c(0,8),
-                    ExplainDirectMoran_Se2e2_mid_ylims=c(0.06,0.53),
+                    ExplainDirectMoran_Se2e2_mid_ylims=c(0.06,0.6),
                     ExplainDirectMoran_Se2e2_long_ylims=c(0.05,0.9),
                     ExplainDirectMoran_fB_short_ylims=c(0.4,3),
                     ExplainDirectMoran_fB_mid_ylims=c(0.8,1.65),
                     ExplainDirectMoran_fB_long_ylims=c(1.2,3),
-                    ExplainDirectMoran_PanLabs=LETTERS[1:15],
+                    ExplainDirectMoran_PanLabs=paste0("(",letters[1:15],")"),
                     ExplainInteractingMoran_fP1_short_ylims=c(0.064,0.15),
                     ExplainInteractingMoran_fP1_mid_ylims=c(0.0716,0.117),
                     ExplainInteractingMoran_fP1_long_ylims=c(0.0725,0.1125),
@@ -1837,123 +1838,150 @@ PlotForMSArgs<-list(cov_MainPlot_PanLabs=c("A","B","C"),
                     ExplainInteractingMoran_Prod_short_ylims=c(0,0.7),
[…]

### 2022-01-06 Paper/Paper.Rnw
added an Intro and other modifications

@@
 \newcommand{\cov}{\text{cov}}
+\newcommand{\cor}{\text{cor}}
 \newcommand{\Rp}{\text{Re}}
@@
 %The following is where you load in the numeric results that will be embedded in the text
-%<<eval=T,echo=F,message=F,warning=F>>=
-%#place R code here for loading in necessary variables from Results
-%@
+<<eval=T,echo=F,message=F,warning=F>>=
+
+#place R code here for loading in necessary variables from Results, Theory
+resloc<-"../Results/Theory/"
+N_A<-readRDS(paste0(resloc,"Params_A_N.rds"))
+ns_A<-readRDS(paste0(resloc,"Params_A_ns.rds"))
+p1_A<-readRDS(paste0(resloc,"Params_A_p1.rds"))
+p2_A<-readRDS(paste0(resloc,"Params_A_p2.rds"))
+b_A<-readRDS(paste0(resloc,"Params_A_b.rds"))
+
+N_B<-readRDS(paste0(resloc,"Params_B_N.rds"))
+ns_B<-readRDS(paste0(resloc,"Params_B_ns.rds"))
+p1_B<-readRDS(paste0(resloc,"Params_B_p1.rds"))
+p2_B<-readRDS(paste0(resloc,"Params_B_p2.rds"))
+b_B<-readRDS(paste0(resloc,"Params_B_b.rds"))
+
+N_C1<-readRDS(paste0(resloc,"Params_C1_N.rds"))
+ns_C1<-readRDS(paste0(resloc,"Params_C1_ns.rds"))
+p1_C1<-readRDS(paste0(resloc,"Params_C1_p1.rds"))
+p2_C1<-readRDS(paste0(resloc,"Params_C1_p2.rds"))
+b_C1<-readRDS(paste0(resloc,"Params_C1_b.rds"))
+
+N_C2<-readRDS(paste0(resloc,"Params_C2_N.rds"))
+ns_C2<-readRDS(paste0(resloc,"Params_C2_ns.rds"))
+p1_C2<-readRDS(paste0(resloc,"Params_C2_p1.rds"))
+p2_C2<-readRDS(paste0(resloc,"Params_C2_p2.rds"))
+b_C2<-readRDS(paste0(resloc,"Params_C2_b.rds"))
+
+@
 
@@
 
-%***Add to these
-\keywords{kelp, synchrony, Moran effects}
+\keywords{Fourier, interactions, kelp, \emph{Macrocyctis pyrifera}, Moran effects, nutrients, synchrony, waves}
 
@@
 
-%Tentative paragraph outline written on my phone on 3/20/2021
-%1) Importance of synchrony
-%2) Classic problem of identifying causes of synchrony, but recently largely resolved due to new methods
-%  -elaborate briefly on a couple methods, mention the importance of a timescale-resolved appproach
-%3) in fact, it has recently been discovered that environmental causes of synchrony can interact, so that ...
-%  -related to, but distinct from, interaction between environmental and dispersal causes of synchrony discussed, for instance by Kendal et al
-%  -make sure you describe the two cases, cite the earlier papers, mention a few words about the two systems
-%4) the simplest form of interacting moran effects can be demonstrated when a simple model, though elaboration of the true nature of this mechanism %will take the rest of this manuscript
-%  -use Moran model
-%  -but with two drivers
-%  -moran theorem says correlation of pops equals correlation of sum of drivers, so work it out
-%5) general idea of two noises lagged or time delayed relative to each other and their effects oppositely lagged - this may be where a pedagogical %fig comes in, or maybe it is just worsd here and then the pedagog fig comes in later, at end of theory section, in form of examples.
-Yes, I
-%think I want that second approach.
-%6) brief paragraph on why we would expect interacting moran effects to be common, return to this point in the discussion
-%7) final para on goals, manuscript structure, etc.
-%  -G1) Elaborate a mathematical theory of interacting Moran effects which allows a detailed understanding 
-%      of how the mechanism works 
-%  -G2) Consider theoretical case studies, which emerge as important special cases of our general theory, and which illuminate the intuition behind how Moran effects may interact in real systems. 
-%  -G3) Apply the theory to populations giant kelp of the California coast, the intention being to further elaborate the mechanism itself
-%   Final sentence about how this paper conceptually and theoretically illuminates a new and probably common mechanism of 
-%   synchrony.
+Spatial synchrony, the tendency for geographically disjunct populations to show correlated fluctuations 
+through time, is a fundamental feature of population dynamics, linked, for instance, to important topics such as 
+population cycling \citep{Anderson2021}, extinction risk \citep{Ghosh2020}, and the 
+stability of regional populations and ecosystem functioning \citep{Wilcox2017}. Though spatial synchrony (henceforth, 
+\emph{synchrony}) has been studied for decades
+in a wide variety of species ranging from viruses to plants to mammals, and at spatial scales ranging from centimeters to over 
+1000 km \citep{Liebhold2004},
+recent advances in statistical methods and improvements in data availability have lead to several 
+major advances in our understanding of synchrony and its causes and consequences. 
+For instance, the timescale structure of synchrony is now known to be 
+important [\cite{Sheppard2016,Desharnais2018,Sheppard2019}, see also \cite{Keitt2008}],
+synchrony is now known to have a complex and pronounced geography \citep{Defriez2017,Defriez2017b,Walter2017}, 
+and patterns of synchrony are now known to be asymmetric in distribution tails in a way that
+is important for system stability \citep{Ghosh2020b,Walter2021}. Our ability to infer the causes of 
+synchrony is much improved in recent years \citep{Sheppard2016,Defriez2017b,Walter2017,Anderson2018,Sheppard2019}, 
+and there is also growing evidence
+that synchrony is changing as a newly recognized component of climate 
+change \citep{Defriez2016,Sheppard2016,Hansen2020}. Spatial and community 
+viewpoints on synchrony are becoming integrated, leading to a more wholistic understanding 
+of the importance of synchrony for ecosystem stability \citep{Wang2014}. These new developments 
+have led to an increasingly central role of the phenomenon of synchrony in many of the most 
+important research topics in ecology.
 
-%Notes:
-%1) You need to make it clear the earlier papers already identified interacting Moran effects, so what are we doing?
-%We are exploring theoreitically what the mechanism is, here, and how general it might be, and what different forms it 
-%might take, using the theory, and also using the theory to further illuminate the two examples. The emphasis here
-%is on illuminating the general mechanism of interacting Moran effects through the theory and its application, whereas
-%the earlier papers focussed on the specific cases. The Fourier theory here serves to illuminate the general mechanism,
-%whereas the earlier papers used wavelet statistics applied directly to data and did not really explore the general 
-%mechanism. Need to work some of these thoughts in without seeming defensive.
+One of the classically theorized \citep{Moran1953} and now widely empirically validated mechanisms of 
+synchrony, called the Moran effect, operates when spatially synchronized environmental signals, such
+as temperature or precipitation, help drive population dynamics and thereby induce synchrony in
+populations. Two recent papers have documented that Moran effects of distinct 
+environmental drivers can, in fact, interact, either syergistically or antagonistically,
+so that total population synchrony can be either greater or less than if the distinct Moran 
+drivers operated independently \citep{Sheppard2019,CastoraniSheppardKelp}. Effects can be strong.
+\cite{Sheppard2019} showed that long-timescale ($>4$y period) synchrony in a chlorophyll density index 
+in the seas around the United Kingdom is substantially augmented by interactions between environmental 
+drivers of synchrony and drivers linked with consumption by a copepod predator.
+\cite{CastoraniSheppardKelp} showed that nutrient dynamics and wave disturbance, two Moran drivers
+of synchrony in giant kelp populations on the California coast, interact either synergistically 
+or destructively, depending on which portion of the California coast is examined and on 
+timescale of analysis. 
+Both of these studies argued that interactions between Moran effects may be a general feature of many 
+systems, because of the large number of interrelated factors driving population fluctuations in most 
+spatially extended systems. The main purpose of this study is to establish a general theory of the mechanism
+of interacting Moran effects and to use examples and applications of the theory to provide 
+ecological intuition for how the mechanism of interacting Moran effects works. Our goals are
+distinct from those of \cite{Kendall2000}, who examined interactions between Moran and dispersal
+causes of synchrony. We relate our work to that study in the Discussion.
 
-\section{Theory}
+The basic fact that Moran drivers can interact, synergistically or destructively, can be demonstrated 
+with a very simple model which we now elaborate, though understanding the nature of the interaction
+mechanism will require the rest of this paper. Suppose $\epsilon_i^{(1)}(t)$ and $\epsilon_i^{(2)}(t)$
+are environmental random variables measured in locations $i=1,2$ at times $t$, and assume these are independent
+through time and standard normally distributed for all $i$ and $t$. If a population index $w_i(t)$ 
+follows the autoregressive process $w_i(t)=b w_i(t-1)+\epsilon_i^{(1)}(t)$ for $i=1,2$, then the classic Moran theorem 
+\citep{Moran1953} implies that temporal population correlation equals temporal environmental correlation, i.e., 
+$\cor(w_1,w_2)=\cor(\epsilon_1^{(1)},\epsilon_2^{(1)})$. If, instead, 
+$w_i(t)=b w_i(t-1)+\epsilon_i^{(1)}(t)+\epsilon_i^{(2)}(t)$, so that both environmental variables influence 
+populations, the Moran theorem then implies that population synchrony is $\cor(w_1,w_2)=
+\cor(\epsilon_1^{(1)}+\epsilon_1^{(2)},\epsilon_2^{(1)}+\epsilon_2^{(2)})=
+[\cov(\epsilon_1^{(1)},\epsilon_2^{(1)})+\cov(\epsilon_1^{(1)},\epsilon_2^{(2)})+
+\cov(\epsilon_1^{(2)},\epsilon_2^{(1)})+\cov(\epsilon_1^{(2)},\epsilon_2^{(2)})]/
+[\sqrt{2+2\cov(\epsilon_1^{(1)},\epsilon_1^{(2)})}\sqrt{2+2\cov(\epsilon_2^{(1)},\epsilon_2^{(2)})}]$,
+a quantity which depends not only on the environmental synchrony measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(a)})$, for 
+$a=1,2$, but also on the ``cross synchrony'' measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(b)})$
+for $a \neq b$. Cross synchrony measures represent synchrony between an environmental variable in one location and 
+a different variable in another location. Fig. \ref{fig:IntroTheory}a shows how population synchrony can 
+therefore differ from what synchrony would be if the environmental processes $\epsilon^{(1)}$ and 
+$\epsilon^{(2)}$ were unrelated. The difference is due to interactions
+between the two Moran effects, and can be substantially positive or negative. This example was adapted 
+from SI section S1 of \cite{Sheppard2019}.
+
+Having demonstrated that interactions between Moran effects can add to or reduce population synchrony, we now 
+begin to provide some intuition on how we can understand the underlying mechanisms. 
+A key insight is the use of a timescale-specific approach,
+i.e., environmental signals are decomposed, using Fourier analysis, into components that occur on different timescales,
+and the nature of interacting Moran effects can be understood on a timescale-by-timescale basis, with effects summed 
+across timescales afterwards if desired. Fig. \ref{fig:IntroTheory}b shows the period-20 Fourier components of two hypothetical
+environmental signals measured in sampling location $i$, and one way these can influence populations. The components are
+lagged, relative to each other, in the timing of their peaks ($l_n$ on the figure). They are also lagged in their 
+effects on populations, by the amounts $l_{e1}$ and $l_{e2}$, respectively, i.e., peaks in the environmental signals
+manifest as maximum positive influence on populations after delays of $l_{e1}$ and $l_{e2}$, respectively. In this example,
+we assume for simiplicity that larger values of both environmental variables are beneficial to populations, though the general
+theory represented below does require that assumption. If, as in Fig. \ref{fig:IntroTheory}b, $l_{e1}=l_n+l_{e2}$, then the 
+periodic maximal positive influences of environmental variables coincide. If environmental signals are then spatially synchronous
+across sampling locations, the fact of the alignment of lags produces additional synchrony in populations, 
+beyond what would manifest if environmental fluctuations were unrelated. In contrast, Fig. \ref{fig:IntroTheory}c
+shows the opposite scenario, for which $l_{e1}$ plus the half the period, $10$, equals $l_n+l_{e2}$, so that  
+the periodic maximal positive influence of environmental variable 1 coincides with the periodic maximal negative
+influence of environmental variable 2, reducing synchrony relative to the case of unrelated environmental variables. 
+Intermediate scenarios are possible, as will be revealed by the general theory, below.
+The theory also formalizes the timescale decomposition of synchrony measures, and timescale-specific lags.
+
+Thus our specific goals for thus study are the following. First, (G1) we will elaborate a general mathematical 
+theory of interacting Moran effects which allows a detailed understanding of how the mechanism works. Second, 
+(G2) we will consider theoretical case studies, which emerge as important special cases of our general theory, and 
+which illuminate the intuition behind how Moran effects may interact in real systems. Third, (G3) we will 
+apply our theory to populations of giant kelp off the California coast. Whereas \cite{CastoraniSheppardKelp}
+already carried out a detailed analysis of interacting Moran effects in California kelp populations and their importance 
+for kelp ecology, we instead use a simplified subset of the available kelp data, and our analysis is instead intended 
+to illuminate the inner workings of our theory and of the general mechanisms we describe. In the Discussion, we 
+extend the arguments of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} that interactions between Moran 
+effects are likely to be common, and hence that our general theory seems likely to pertain to numerous 
+ecosystems. Overall, this study serves to introduce a general theory of Moran interactions and to conceptually
+illuminate a newly observed but potentially quite common mechanism of spatiotemporal population dynamics.
+
+\section{General theory}
 
@@
 of a species of interest in that location), the spectrum $S_{w_i w_i}(f)$ is a function
-of frequency, $f$. For a periodic ocillation, $f$ can be defined as one over the timescale, or period, of the oscillation. 
+of frequency, $f$. For a periodic ocillation, $f$ can be defined as one over the timescale, or period, $\sigma$, of the oscillation. 
 The spectrum $S_{w_i w_i}(f)$ is larger for frequencies at which $w_i(t)$ oscillates more. So, for example, a 
@@
 The cross-spectrum takes large-magnitude values at frequencies, $f$, for which the oscillatory components
-of $w_i(t)$ and $w_j(t)$ are both strong; and the complex phase of the cross spectrum at $f$ then quantifies
-the phase lag between those components. The notation $S_{ww}(f)$ refers to the \emph{spectral matrix}, 
-the $ij^{\text{th}}$ entry of which is $S_{w_i w_j}(f)$. Spectral methods are standard tools in many fields, 
-including ecology [e.g., \cite{Vasseur2007,Reuman2008,Defriez2017}], and many background references are available 
+of $w_i(t)$ and $w_j(t)$ are strong and in a consistent phase relationship to each other; 
+and the complex phase of the cross spectrum at $f$ then quantifies
+the phase. The notation $S_{ww}(f)$ refers to the \emph{spectral matrix}, 
+which has $ij^{\text{th}}$ entry $S_{w_i w_j}(f)$. Spectral methods are standard tools 
+[e.g., \cite{Vasseur2007,Reuman2008,Defriez2017}], and many background references are available 
 [e.g., \cite{Brillinger2001}].
@@
 of distinct locations. This is a real-valued function of frequency, the integral (across frequencies) of which
-is the classic, non-frequency-specific measure $\left( \sum_{i \neq j} \cov(w_i,w_j) \right)/(N^2-N)$ (see SI section
+is the classic, non-frequency-specific synchrony 
+measure $\left( \sum_{i \neq j} \cov(w_i,w_j) \right)/(N^2-N)$ (see SI section
 \ref{SM-Sect:measuring_sync} for details, here). If two time series $\epsilon_i^{(1)}(t)$ and $\epsilon_i^{(2)}(t)$
@@
 spatially synchronous, $4$-year-period 
-oscillations, but peaks in the $\epsilon^{(1)}$ oscillations consistently lag peaks in the $\epsilon^{(2)}$ oscillations 
+oscillations, but peaks in the $\epsilon^{(2)}$ oscillations consistently lag peaks in the $\epsilon^{(1)}$ oscillations 
 by a year, then $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$, which is complex valued, will have high magnitude at
@@
 
-Our model is 
+Our population model is 
 \begin{align}
@@
 $(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ is an ergodic second-order stationary stochastic process \citep{Brillinger2001}
-with expected values of its component processes equal to $0$. We make additional mild
-regularity assumptions to ensure the model is stable 
-(SI section \ref{SM-sect:model_spec}). This model can straightforwardly be seen as 
+with expected values of its components equal to $0$. We make additional mild
+regularity assumptions for model stability 
+(SI section \ref{SM-sect:model_spec}). This model can be seen as 
 a linearization of a very general dynamical model, influenced by ``weak noise'' [see, e.g.,
 SI section S1.2 of \cite{Desharnais2018} and SI section S1 of \cite{Walter2017}].
-Linearization and ``weak noise'' assumptions are tactics which have been used very commonly to make
+Linearization and ``weak noise'' assumptions have been commonly adopted to make
 theoretical progress in ecology (REFS), and it has been demonstrated that results based on a
-weak noise assumption often also hold for noise which is quite strong (REF).
+weak noise assumption often hold for noise which is fairly strong (REF).
 %***DAN: Add refs later
@@
 due to the direct Moran effects of $\epsilon^{(1)}$, the second term is 
-the component of population synchrony due to the direct Moran effects of $\epsilon^{(2)}$, 
+the component due to the direct Moran effects of $\epsilon^{(2)}$, 
 and the third term is the component due to interactions between the Moran effects of the 
-two environmental drivers.
-The ``other contributions'' in the above equations correspond to synchronizing influences of 
+two drivers.
+The ``other contributions'' above correspond to synchronizing influences of 
 $\delta$ and of interactions between $\delta$ and the $\epsilon^{(i)}$. Such contributions would be
@@
 of (\ref{rhoww_T1to3}) gives the relative contributions of direct Moran effects and interactions 
-between Moran effects. The detailed derivation of the theory is in SI section \ref{SM-sect:derivation}.
+between Moran effects. The derivation of the theory is in SI section \ref{SM-sect:derivation}.
 
@@
 effects of $\epsilon^{(2)}$ are lagged by $n_2$ time steps, i.e.,
-$p_{n_1}^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n_1$, and 
-$p_{n_2}^{(2)}>0$ while $p_i^{(2)}=0$ for $i \neq n_2$.
+$p_{n_1}^{(1)} \neq 0$ while $p_i^{(1)}=0$ for $i \neq n_1$, and 
+$p_{n_2}^{(2)} \neq 0$ while $p_i^{(2)}=0$ for $i \neq n_2$.
 The noise process
@@
 For C, the noise processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
-each assumed to exhibit spatially synchronous periodic oscillations of period equal to one year, i.e.,
+each assumed to exhibit spatially synchronous periodic oscillations of period one year, i.e.,
 four model time steps, but with different phases. Peaks in $\epsilon^{(1)}$ are assumed to 
@@
 %other parts of the main theoretical equation were estimated. 
-%Comment) I could write these now, but then if I end up having to do something slightly different for plankton, I'll have to 
-%rewrite them. Maybe now is the time to decide if I want this to be a kelp-only or a kelp-an-plankton paper and then do the
-%plankton work.
 %Comment) Somehow need to get across what was done - different parts of main equation were examined to explore the different
@@
 comparable in strength to direct Moran effects; 2) can be either synergistic or destructive, i.e., positive or negative; and 3) 
-can depend strongly on timescale, even having different signs on different timescales 
-(Fig. \ref{fig:TheoryCaseA}a,e,i,m). First, for all case studies, the magnitude of 
-interaction effects was comparable to that of direct Moran effects (compare the dashed and solid lines on 
-Fig. \ref{fig:TheoryCaseA}a,e,i,m). So interactions can contribute substantially to overall population 
-synchrony. Second, in contrast to the direct Moran effects of a single environmental variable, which provide non-negative 
-contributions to population synchrony at all timescales for which environmental synchrony 
+can depend strongly on timescale, even having different signs on different timescales. 
+First, for all of our theoretical case studies, the magnitude of 
+interaction effects was comparable to that of direct Moran effects (Fig. \ref{fig:TheoryCaseA}a,e,i,m, 
+compare the dashed and solid lines). So interactions can contribute substantially to overall population 
+synchrony. Second, in contrast to the direct Moran effects of a 
+single environmental variable, which provide non-negative 
+contributions to population synchrony whenever, as is typical, environmental synchrony 
 is positive, interactions between Moran effects can be positive or negative. For instance, case study A
-showed negative interaction effects at short timescales (Fig. \ref{fig:TheoryCaseA}a); and, for some
-parameters, case study C showed negative interactions at annual timescales (Fig. \ref{fig:TheoryCaseA}m).
+showed negative interaction effects at short timescales (Fig. \ref{fig:TheoryCaseA}a); and 
+case study C showed negative interactions at
+long timescales (Fig. \ref{fig:TheoryCaseA}i,m), or, for some parameters, at all timescales 
+(Fig. \ref{fig:TheoryCaseA}m).
 Thus interaction effects can either augment or reduce population synchrony from what its value would be 
@@
 also follow from the results of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not necessarily
-straightforwardly. Additionally, those are case studies, so our results provide novel theoretical generality;
-and our theory can be interpreted to provide intuition about how interacting Moran effects work,
+straightforwardly, and those papers are case studies, so our results provide novel theoretical generality.
+Our theory can also be interpreted to provide intuition about how interacting Moran effects work,
 as follows.
@@
 Case study A provides an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
-Moran effects on different timescales, and how our theory successfully captures that contrast.
+Moran effects on different timescales, and how our theory captures that contrast.
 For case study A, for which $\epsilon^{(1)}$ effects on populations were lagged by 1 quarter but 
@@
 quarters and lagged by 1 quarter are exactly out of phase with unlagged Moran effects of timescale
-2 quarters, producing destructive interference, i.e., negative interactions. Thus, on short timescales, total
+2 quarters, producing destructive interference. Thus, on short timescales, total
 synchrony will be less, for this example, than it would be if only direct Moran effects operated. In contrast, 
 interaction effects on long timescales were synergistic, i.e., positive, for case study A 
-(Fig. \ref{fig:TheoryCaseA}a, right side of panel). 
-This is because, on long timescales, a lag of 1 quarter does not meaningfully alter the phase relationship between 
+(Fig. \ref{fig:TheoryCaseA}a, right side of panel) 
+because, on long timescales, a lag of 1 quarter does not meaningfully alter the phase relationship between 
 Moran drivers, and drivers are therefore almost in phase, reinforcing each other.
@@
 with each other are perfectly uncorrelated. Thus interaction effects between Moran drivers were zero 
-at timescale 1 year (Fig. \ref{fig:TheoryCaseA}a). Theory effectively captures these complexities because
+at timescale 1 year (Fig. \ref{fig:TheoryCaseA}a). Theory captures these complexities because
 the size of the 1-quarter phase lag of this scenario, relative to a given timescale,
-is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$, which has phase $-\pi$ at short timescales
+is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) = 
+p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}), 
+which has phase $-\pi$ at short timescales, $\sigma \approx 0.5 \text{y}$,
 and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b). Because 
-$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ has phase $0$ for all timescales, the phase of 
-$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}c), which controls
-the sign of the real part of this quantity, is the same as the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$.
-Thus $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, which is the
-interaction term in our theory (\ref{rhoww_T1to3}), shifts 
-from negative, for short timescales, to positive for long timescales. Comparing
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ has phase $0$ for all timescales (SI sections \ref{SM-sect:examples_cross_sync}, 
+\ref{SM-sect:details_case_studies}), the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ is the same as the phase of
+$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}c),
+which in turn controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, which is the
+interaction term in our theory (\ref{rhoww_T1to3}). Thus interactions shift 
+from negative, for short timescales, to positive, for long timescales. Comparing
 Fig. \ref{fig:TheoryCaseA}a,d shows the additional influence of autoregressive population effects, which
 simply multiply all Moran influences by the same timescale-dependent non-negative 
-quantity without altering the relative importance or sign of those effects [see (\ref{rhoww_T1to3})].
+quantity without altering the relative importance or sign of the direct and indirect Moran effects [see (\ref{rhoww_T1to3})].
 
@@
 For B, $\epsilon^{(1)}$ effects on populations were lagged by $n$ quarters but 
-$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$. Thus interactions between Moran effects were 
+$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$. Interactions between Moran effects were 
 negative on the shortest timescales 
-(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, far left side of panel). This is because the 
+(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, far left side of panel) because the 
 population influences of Moran effects of timescale 2
@@
 2 quarters, producing negative interactions. 
-Interaction effects on long timescales were instead synergistic, i.e., positive, for case study B 
+Interaction effects on long timescales were instead synergistic, i.e., positive, for B 
 (Fig. \ref{fig:TheoryCaseA}e, right side of panel), for the same reason as case study A (see above).
-At timescale 3 quarters, Moran effects are, of course, reinforcing because the lag equals the timescale,
+At timescale 3 quarters, Moran effects are reinforcing because the lag equals the timescale,
 producing Moran effects of the two drivers which are in phase 
@@
 Theory captures these complexities because 
-the size of the 3-quarter phase lag of this scenario, relative to a given timescale,
+the size of the 3-quarter phase lag of this scenario, relative to a timescale,
 is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
-Because $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number for case study B,
-the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-is the same as that of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}g); 
-that common phase then controls the sign of 
-$2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, which is the 
-interaction effect term in (\ref{rhoww_T1to3}).
+Because $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number for case study B (SI section \ref{SM-sect:details_case_studies}),
+the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ is the same as that of
+$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}g),
+which controls the sign of the interaction term,
+$2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, in our 
+theory (\ref{rhoww_T1to3}).
 
-Cases A and B assumed white noise processes and real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$,
+Cases A and B assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+(SI sections \ref{SM-sect:examples_cross_sync}, \ref{SM-sect:details_case_studies}),
 but case study C illustrates what can happen when noise processes have periodic components, e.g., 
-both noise processes oscillate with annual periodicity according to the seasons, an important scenario, ecologically. 
+both processes oscillate with annual periodicity, an important scenario because seasonality is common. 
 What matters then is whether and how the lag between the two seasonal noise processes aligns with the difference 
@@
 For the scenarios we considered under case study C (Fig. \ref{fig:TheoryCaseA}i-p), 
-$\epsilon^{(1)}$ effects on populations were lagged (i.e., delayed in their influence) by 1 quarter, but 
-$\epsilon^{(2)}$ effects were immediate (Methods and Fig. \ref{fig:TheoryCaseA} caption).
-For scenario 1 of case study C (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
+$\epsilon^{(1)}$ effects on populations were positive, and were 
+lagged (i.e., delayed in their influence) by 1 quarter; but 
+$\epsilon^{(2)}$ effects were immediate and negative (Methods and Fig. \ref{fig:TheoryCaseA} caption).
+For scenario 1 of C (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
+noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
+1 quarter. Thus lags compounded: the annual positive population effects of $\epsilon^{(1)}$ 
+were exactly 2 quarters offset from the annual negative effects of $\epsilon^{(2)}$, 
+producing reinforcing interactions 
+between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i), and more population
+synchrony on annual timescales than would have been expected if Moran effects operated independently. Contrastingly,
+for scenario 2 of case study C (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
 noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
-1 quarter. Thus lags cancelled and the annual peak population effects of $\epsilon^{(1)}$ and
-$\epsilon^{(2)}$ arrived in sync with each other each year, producing positive interactions 
-between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i), and more population
-synchrony than would have been expected if Moran effects operated independently. Contrastingly,
-for scenario 2 of case study C (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
-noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
-1 quarter. Thus lags compounded and the annual peak population effects of $\epsilon^{(1)}$
-arrived 2 quarters after those of $\epsilon^{(2)}$ every year, a half-phase shift for an 
+1 quarter. Thus lags cancelled: the annual positive population effects of $\epsilon^{(1)}$
+coincided with the negative effects of $\epsilon^{(2)}$ every year, a half-phase shift for an 
 annually oscillating effect, which corresponds to negative interactions between Moran influences
-and less population synchrony than would have been expected if Moran effects operated independently 
-(Fig. \ref{fig:TheoryCaseA}m). The only difference between scenarios 1 and 2 of case study C
+and less population synchrony, on annual timescales, than would have been expected if Moran effects operated independently 
+(Fig. \ref{fig:TheoryCaseA}m). The only difference between scenarios 1 and 2 
 is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
@@
 quantifies the relative phase lag between the effects of $\epsilon^{(1)}$ and 
-$\epsilon^{(2)}$ relative to a given timescale, while 
+$\epsilon^{(2)}$ compared to a given timescale, while 
 $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the phase lag between 
-oscillations in the variables themselves, relative to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
+oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
 The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
@@
 the variables themselves align or misalign, with the degree of alignment
-reflected in the real part of this quantity. On long timescales, interactions were the
-same in the two secenaios, because on those timescales sub-annual lags in environmental 
-peaks and effects on populations make negligible difference.
+reflected in the real part of this quantity, $\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$,
+which is half the interaction term in our theory (\ref{rhoww_T1to3}).
+On long timescales, interactions were the
+same in the two scenarios because, on those timescales, sub-annual lags 
+make negligible difference. Interactions were then negative 
+because the effects of $\epsilon^{(1)}$ on populations were 
+positive and those of $\epsilon^{(2)}$ were negative.
 
-%Current (2021 11 01) thoughts on figures:
-%1) Maybe a pedagog fig, not sure what this would look like
-%2) Combined theory fig, all 3 case studies
-%3) "main fig" for CC1 and SB systems
-%4-5) "explain direct effects" fig for CC1 and SB systems (2 figures, I think it'll have to be)
-%6-7) "explain indirect effects" figs for CC1 and SB systems (again, 2)
-%One too many figs here. What should be cut? Right now I don't know what the pedagog fig would be 
-%so that's the obvious one to cut. And then it will all fit, though the figs are really big and 
-%it seems like it is always good to have a pedagogical, warmup fig. If you do come up with a pedagog
-%fig idea, could maybe combine some figs to make it work, or put the pedagog content as the first
-%part of the theory fig? Don't worry about it for now, but return to it later.
+\section{Discussion}
+
+%Note that we get differences in intereaction effects between central and southern CA because we 
+%have lagged effects of nitrate on observed kelp density in central CA, but unlagged effects in southern
+%CA (to see this, look at the regression coefficients computed to get the models used for the
+%spectral analysis, which were saved). Careful inspection of my ms figures show this is the reason
+%for the positive versus slightly negative interaction effects seen in central v. souther CA. 
+%But I spke with Tom and Max and Kyle about this by email around 1/3/2022. They agreed that 
+%probably it has something to do with the fact that kelp is ripped out by waves in central
+%CA, so it takes a quarter to recover enough to be detectable from space (we can only see the 
+%canopy from space, on the surface of the water). In southern CA, waves rip out kelp like that
+%much less. So if kelp is always around and usually reaches the surface in southern CA, nitrate
+%effects can be observed more quickly. There ought to be a Discussion paragraph on this, basiaclly
+%saying things may be more complex that our analyses reveals with regard to some of the differences 
+%we observed between central and southern CA. We didn't explore this because our goal was to illustrate 
+%our theory, and our work does demonstrate the importance of interaction effects for the synchrony of kelp
+%as observed from space. But future work could look at this more, certainly, perhaps using state-space models
+%which can separate biological from observation processes.
+
+%We definitely need a discussion para comparing my results to those of Max and Lawrence (hopefully they are 
+%basically consistent). Note differences in methods. Note differences in results (hopefully small). 
+%Explain/explore how my theory goes beyond what they did - the intention is to get at the mechanism
+%better and how it actually works. It may be sufficient to keep this a fairly brief para.
+
+%Go back to the Kendall study of 2000 (I think that's the year), the one that looked at interactions
+%between Moran and dispersal causes of synchrony. If I remember it correctly (review it), it's a bit silly,
+%because all it basically does it turn dispersal and Moran drivers on and off in a model and show that
+%you get more or less synchrony from the both-on case that you would get by summing the synchrony 
+%you get from each of the one-on cases. But the expectation that correlations should sum is naive anyway - 
+%correlation is not a linear operator (though covariance is - check to make sure they use cor instead of
+%cov, and how would their results be different if they used cov?).
+%Our study has a better baseline to compare to than the sum - the baseline is if the two Moran drivers are
+%independent (or, rather, it's a bit less than that, and has to do with the cross synchrony measure
+%being 0). This is not to belittle the achievements of that paper, which was groundbreaking in its time,
+%instead it's just to explore the value of perspective that we now have based on 20 years of hindsight
+%and ongoing work.
+
+\bibliographystyle{ecology_letters2}
+\bibliography{refs}
+
+\clearpage
+\newpage
 
 \begin{figure}
-\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseA.pdf}\\
-\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseB.pdf}\\
-\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v01.pdf}\\
-\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v02.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Theory/IntroTheory.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Theory/PedagogFig.pdf}
+\caption{Example and basic concept of interacting Moran effects. The example (a), which 
+is based on a simple model described in the Introduction, shows that 
+interaction effects are possible, and can be substantially positive or negative. 
+We used $\cov(\epsilon_1^{(1)},\epsilon_2^{(1)})=\cov(\epsilon_1^{(2)},\epsilon_2^{(2)})=0.7$,
+and the values $\cov(\epsilon_i^{(1)},\epsilon_j^{(2)})$, for $i,j=1,2$, were all set equal to each
+other and to a parameter, $\phi$. For (b) and (c), which represent distinct scenarios, 
+the top sinusoid on each panel represents 
+the component of oscillation at a given timescale of an environmental variable, $\epsilon_i^{(1)}$,
+measured in a sampling location $i$, and the bottom sinusoid represents the 
+component at the same timescale for another environmental variable, $\epsilon_i^{(2)}$; the 
+oscillations are lagged from each other by an amount $l_n$. Environmental effects on 
+the population, $w_i(t)$, in location $i$ are also lagged, by amounts $l_{e1}$ and
+$l_{e2}$. For this setup, a positive interaction between Moran effects can occur on the
+given timescale when lags align, i.e, $l_{e1}=l_{e2}+l_n$, as in panel (a), so that environmental
+effects on the population coincide and reinforce each other. All environmental effects on 
+populations are assumed positive in these examples. A negative
+effect can occur when lags produce out-of-phase influence of environmental variables on
+populations, i.e., $l_{e1}$ differs from $l_{e2}+l_n$ by half the given timescale/period,
+as in panel (b), so that the distinct influences tend to work against each other in 
+bringing about population oscillations on the given timescale.  
+A timescale-specific approach is essential to this reasoning.}\label{fig:IntroTheory}
+\end{figure}
+
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=.85\textwidth]{../Results/Theory/TheoryFigCaseA.pdf}\\
+\includegraphics[width=.85\textwidth]{../Results/Theory/TheoryFigCaseB.pdf}\\
+\includegraphics[width=.85\textwidth]{../Results/Theory/TheoryFigCaseC_v01.pdf}\\
+\includegraphics[width=.85\textwidth]{../Results/Theory/TheoryFigCaseC_v02.pdf}
 \caption{Theoretical case studies A (panels a-d), B (panels e-h) and C (i-p, two alternative sets of 
@@
 synchrony [see (\ref{rhoww_T1to3})].
-Model parameters were: $N=5$ and $b_1=0.4$ for all case studies; 
-$p_1^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study A; 
-$p_3^{(1)}=1.1$ and $p_0^{(2)}=0.8$ for case study B;
-$p_1^{(1)}=3$, $p_0^{(2)}=1.5$ 
+Model parameters were: $N=\Sexpr{N_A}$ and $b_1=\Sexpr{b_A}$ for all case studies; 
+$p_1^{(1)}=\Sexpr{p1_A[2]}$ and $p_0^{(2)}=\Sexpr{p2_A}$ for case study A; 
+$p_3^{(1)}=\Sexpr{p1_B[4]}$ and $p_0^{(2)}=\Sexpr{p2_B[1]}$ for case study B;
+$p_1^{(1)}=\Sexpr{p1_C1[2]}$, $p_0^{(2)}=\Sexpr{p2_C1[1]}$ 
 %$c_1=0$ and $c_2=-4/9$ 
@@
 
-\section{Discussion}
+\clearpage
+\newpage
 
-\section{References}
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_MainPlot_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_MainPlot_ForMS.pdf}
+\caption{Filler.}\label{fig:SummaryKelpFig}
+\end{figure}
 
-\bibliographystyle{ecology_letters2}
-\bibliography{refs}
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainDirectMoran_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainDirectMoran_ForMS.pdf}
+\caption{Filler.}\label{fig:directMoran}
+\end{figure}
+
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
+\caption{Filler.}\label{fig:indirectMoran}
+\end{figure}
 

### 2022-01-06 Paper/SupMat.Rnw
small edits while reading the paper to get back up to speed
added an Intro and other modifications

@@
 \newcommand{\cov}{\text{cov}}
+\newcommand{\cor}{\text{cor}}
 \let\conj\overline
@@
 %header material for paper
-\title{How environmental drivers of spatial synchrony interact}
+\title{Supporting information for ``How environmental drivers of synchrony interact''}
 \date{}
@@
 quantities given previously. For theoretical case studies A and B in the main text, we used
-$N=5$, $v_{11}=v{22}=0.6$ and $v_{12}=0.3$.
+$N=\Sexpr{N_A}$, $v_{11}=v{22}=\Sexpr{ns_A[1]}$ and $v_{12}=\Sexpr{ns_A[3]}$.
 
@@
 which are automatically real-valued (SI section \ref{Sect:measuring_sync}), should be flat for a white noise
-process (as we determined above that they are), i.e., independent of timescale, because spectral quantities are flat 
-for time series with no temporal dependencies.  For the same reason, $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$,
-our cross synchrony measure, should be flat (as we determined above that it is). It makes sense
+process, i.e., independent of timescale, because spectral quantities are flat 
+for time series with no temporal dependencies.  For the same reason, $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+should be flat. It makes sense
 that cross synchrony in this example is real-valued because a complex-valued 
@@
 for $i=1,\ldots,N$ and $a=1,2$. The second step above ensures that $\epsilon^{(1)}$ and $\epsilon^{(2)}$
-are related to each other in a time-delayed manner, and the third step ensures that 
+are related to each other in a time-delayed manner. The third step ensures that 
 $\epsilon^{(1)}$ and $\epsilon^{(2)}$ can oscillate with a dominant periodic component, depending 
-on the values of $c_1$ and $c_2$ used.
+on the values of $c_1$ and $c_2$ used: the autoregressive order-two process used
+can produce dynamics with a peaked spectrum, as we will see below (see also theorem \ref{thm:useful_thm} below).
 
@@
 Because the $\zeta^{(1)}(t)$ are iid for distinct times $t$, we can again use the definitions of the spectrum and the
-cross spectrum and reasoning similar to that used above for the previous example to show that $S_{\zeta^{(1)} \zeta^{(1)}}$
+cross spectrum and reasoning similar to that used above for the previous examples to show that $S_{\zeta^{(1)} \zeta^{(1)}}$
 equals the upper-left $N \times N$ sub-matrix of $\Psi$, which is a matrix with $1$ in every diagonal entry and
@@
 manner in which the noises $\epsilon$ were defined. Autoregressive models of order $2$ are well known to be capable of
-producing oscillations for certain values of the parameters. For instance, for case study C, we use $c_1=0$ and $c_2=-4/9$, values
+producing oscillations for certain values of the parameters. For instance, for case study C, we 
+use $c_1=\Sexpr{ns_C1[1]}$ and $c_2=\Sexpr{round(ns_C1[2],4)}$, values
 which cause the expression $\left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2$ to have a single peak at
-frequency $0.25$ cycles per sampling interval (i.e., $1$ cycle per year). It makes sense that the synchrony quantities
+frequency $0.25$ cycles per sampling interval. It makes sense that the synchrony quantities
 $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and
@@
 their phase is displayed in color. Two alternative sets of parameters
-were used for panels a-c and d-f, respectively. Parameters were:
-$N=5$, $c_1=0$, $c_2=-4/9$, $w_{11}=w{22}=0.6$ and $w_{12}=0.3$ for both sets of panels;
-$l_1=0$ and $l_2=1$ for panels a-c; and 
-$l_1=1$ and $l_2=0$ for panels d-f.
+were used for panels a-c and d-f, respectively, and are also used in the two scenarios
+considered for case study C. Parameters were:
+$N=\Sexpr{N_C1}$, $c_1=\Sexpr{ns_C1[1]}$, $c_2=\Sexpr{round(ns_C1[2],4)}$, 
+$w_{11}=w{22}=\Sexpr{ns_C1[5]}$ and $w_{12}=\Sexpr{ns_C1[7]}$ for both sets of panels;
+$l_1=\Sexpr{ns_C1[3]}$ and $l_2=\Sexpr{ns_C1[4]}$ for panels a-c; and 
+$l_1=\Sexpr{ns_C2[3]}$ and $l_2=\Sexpr{ns_C2[4]}$ for panels d-f.
 }\label{fig:rhosForC}
@@
 
-%\section{Correlation analyses to help select autoregressive model lags}
-%***DAN: Not sure to what extent I may or may not even have this section any more
-%***DAN: Describe the surrogating process
-%Probably you only need to include the results in allcorfsgtd_RE, in the Results section. 
-%See line 480 and above in AnalysisStep_LinearModels.R
+\section{Supporting results for kelp}
 
-\section{References}
+\begin{figure}[!h]
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_MainPlot_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_MainPlot_ForMS.pdf} \\
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_MainPlot_ForMS.pdf}
+\caption[Extension of main text Fig. \ref{MT-fig:SummaryKelpFig}]{Same format as Fig. \ref{MT-fig:SummaryKelpFig} of the main text, but including not only results
+for the first central CA region (a-c) and the southern CA region (d-f), but also results from the 
+second central CA region (g-i).}\label{fig:SummaryKelpFig_SM_lag4}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag8_MainPlot_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag8_MainPlot_ForMS.pdf} \\
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag8_MainPlot_ForMS.pdf}
+\caption[Like Fig. \ref{fig:SummaryKelpFig_SM_lag4} but with kelp lag effects up to 8 quarters]{Same format as Fig. \ref{fig:SummaryKelpFig_SM_lag4}, but based on a model that considered 
+potential kelp lag effects from up to 8 quarters in the past (see Methods). Results were substantively
+the same as for Fig. \ref{fig:SummaryKelpFig_SM_lag4}.}\label{fig:SummaryKelpFig_SM_lag8}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag12_MainPlot_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag12_MainPlot_ForMS.pdf} \\
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag12_MainPlot_ForMS.pdf}
+\caption[Like Fig. \ref{fig:SummaryKelpFig_SM_lag4} but with kelp lag effects up to 12 quarters]{Same format as Fig. \ref{fig:SummaryKelpFig_SM_lag4}, but based on a model that considered 
+potential kelp lag effects from up to 12 quarters in the past (see Methods). Results were substantively
+the same as for Fig. \ref{fig:SummaryKelpFig_SM_lag4}.}\label{fig:SummaryKelpFig_SM_lag12}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainDirectMoran_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainDirectMoran_ForMS.pdf} \\
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainDirectMoran_ForMS.pdf}
+\caption[Extension of main text Fig. \ref{MT-fig:directMoran}]{Same format as Fig. \ref{MT-fig:directMoran} of the main text, but including not only results
+for the first central CA region (a-o) and the southern CA region (p-dd), but also results from the 
+second central CA region (ee-ss).}\label{fig:directMoran_SM_lag4}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag8_ExplainDirectMoran_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag8_ExplainDirectMoran_ForMS.pdf} \\
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag8_ExplainDirectMoran_ForMS.pdf}
+\caption[Like Fig. \ref{fig:directMoran_SM_lag4} but with kelp lag effects up to 8 quarters]{Same format as Fig. \ref{fig:directMoran_SM_lag4}, but based on a model that considered 
+potential kelp lag effects from up to 8 quarters in the past (see Methods). Results were substantively
+the same as for Fig. \ref{fig:directMoran_SM_lag4}.}\label{fig:directMoran_SM_lag8}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag12_ExplainDirectMoran_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag12_ExplainDirectMoran_ForMS.pdf} \\
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag12_ExplainDirectMoran_ForMS.pdf}
+\caption[Like Fig. \ref{fig:directMoran_SM_lag4} but with kelp lag effects up to 12 quarters]{Same format as Fig. \ref{fig:directMoran_SM_lag4}, but based on a model that considered 
+potential kelp lag effects from up to 12 quarters in the past (see Methods). Results were substantively
+the same as for Fig. \ref{fig:directMoran_SM_lag4}.}\label{fig:directMoran_SM_lag12}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf} \\
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainInteractingMoran_ForMS.pdf} 
+\caption[Extension of main text Fig. \ref{MT-fig:indirectMoran}]{Same format as Fig. \ref{MT-fig:indirectMoran} of the main text, but including not only results
+for the first central CA region (a-o) and the southern CA region (p-dd), but also results from the 
+second central CA region (ee-ss).}\label{fig:indirectMoran_SM_lag4}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag8_ExplainInteractingMoran_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag8_ExplainInteractingMoran_ForMS.pdf} \\
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag8_ExplainInteractingMoran_ForMS.pdf} 
+\caption[Like Fig. \ref{fig:indirectMoran_SM_lag4} but with kelp lag effects up to 8 quarters]{Same format as Fig. \ref{fig:indirectMoran_SM_lag4}, but based on a model that considered 
+potential kelp lag effects from up to 8 quarters in the past (see Methods). Results were substantively
+the same as for Fig. \ref{fig:indirectMoran_SM_lag4}.}\label{fig:indirectMoran_SM_lag8}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag12_ExplainInteractingMoran_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag12_ExplainInteractingMoran_ForMS.pdf} \\
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag12_ExplainInteractingMoran_ForMS.pdf} 
+\caption[Like Fig. \ref{fig:indirectMoran_SM_lag4} but with kelp lag effects up to 12 quarters]{Same format as Fig. \ref{fig:indirectMoran_SM_lag4}, but based on a model that considered 
+potential kelp lag effects from up to 12 quarters in the past (see Methods). Results were substantively
+the same as for Fig. \ref{fig:indirectMoran_SM_lag4}.}\label{fig:indirectMoran_SM_lag12}
+\end{figure}
+
+\clearpage
+\newpage
 

### 2022-01-13 Paper/Paper.Rnw
continued working on Methods

@@
 
+resloc<-"../Results/Kelp_DataAfterAllCleaning/"
+quarters<-readRDS(paste0(resloc,"Quarters_CleanedFinal.Rds"))
+
+resloc<-"../Results/Kelp_SpectralDecompResults/"
+SBlocstouse<-readRDS(paste0(resloc,"SBlocstouse.Rds"))
+CC1locstouse<-readRDS(paste0(resloc,"CC1locstouse.Rds"))
+CC2locstouse<-readRDS(paste0(resloc,"CC2locstouse.Rds"))
 @
@@
 \cite{CastoraniSheppardKelp} showed that nutrient dynamics and wave disturbance, two Moran drivers
-of synchrony in giant kelp populations on the California coast, interact either synergistically 
-or destructively, depending on which portion of the California coast is examined and on 
+of synchrony in giant kelp populations on the California (CA) coast, interact either synergistically 
+or destructively, depending on which portion of the CA coast is examined and on 
 timescale of analysis. 
@@
 which illuminate the intuition behind how Moran effects may interact in real systems. Third, (G3) we will 
-apply our theory to populations of giant kelp off the California coast. Whereas \cite{CastoraniSheppardKelp}
-already carried out a detailed analysis of interacting Moran effects in California kelp populations and their importance 
+apply our theory to populations of giant kelp off the CA coast. Whereas \cite{CastoraniSheppardKelp}
+already carried out a detailed analysis of interacting Moran effects in CA kelp populations and their importance 
 for kelp ecology, we instead use a simplified subset of the available kelp data, and our analysis is instead intended 
@@
 effects are likely to be common, and hence that our general theory seems likely to pertain to numerous 
-ecosystems. Overall, this study serves to introduce a general theory of Moran interactions and to conceptually
-illuminate a newly observed but potentially quite common mechanism of spatiotemporal population dynamics.
+ecosystems. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
+illuminate this newly observed but potentially quite common mechanism of spatiotemporal population dynamics.
 
@@
 
+%Theoretical case studies
 \noindent We now describe how we pursue goal G2 of the Introduction, to develop 
@@
 periodic environmental fluctuations sampled quarterly,
-e.g., kelp wave action in central California peaks annually in the winter, whereas surface-water
+e.g., kelp wave action in central CA peaks annually in the winter, whereas surface-water
 nitrate concentrations also fluctuate with period one year, but peak in the spring, a delay of one quarter
@@
 
+%Description of data after cleaning
+The core of this paper is the new theory, but to help illustrate that theory, we also apply
+it to an exceptional dataset on giant kelp (\emph{Macrocyctis pyrifera}) dynamics off the CA coast, and 
+associated environmental measurements. 
+The data are based on a subset of those used by \cite{CastoraniSheppardKelp} and
+\cite{Walter2021}, and data details are given in those papers and in SI section \ref{SM-sect:data}. We here
+summarize the format of the data after a preparation and cleaning process was implemented. 
+\cite{CastoraniSheppardKelp} and \cite{Walter2021} also provide an introduction to 
+kelp ecology, and information on why giant kelp is an excellent species 
+for studies of synchrony. After preparation, kelp data consisted of
+$\Sexpr{length(SBlocstouse)+length(CC1locstouse)+length(CC2locstouse)}$ quarterly kelp abundance time series
+from locations along the CA coast, each time series spanning from quarter $\Sexpr{quarters[1,3]}$ 
+of $\Sexpr{quarters[1,2]}$ to quarter $\Sexpr{quarters[dim(quarters)[1],3]}$ of 
+$\Sexpr{quarters[dim(quarters)[1],2]}$.   
+Time series were grouped into three regions which were analyzed separately: a more northerly
+central CA group of $\Sexpr{length(CC1locstouse)}$ locations (called CCA1); a more southerly central CA group of 
+$\Sexpr{length(CC2locstouse)}$ locations (CCA2); and a group of $\Sexpr{length(SBlocstouse)}$ locations 
+from southern CA, close to Santa Barbara
+(SB; see Fig. \ref{fig:map}). Each kelp measurement is an estimate of mean quarterly kelp biomass per unit 
+useable habitat in a 500m stretch of coastline. We used coastline segments where kelp was persistent
+through essentially all of the $\Sexpr{quarters[1,2]}$-$\Sexpr{quarters[dim(quarters)[1],2]}$ 
+period (SI section \ref{SM-sect:data}). We also had estimates
+of maximum wave height and mean nitrate concentration for each quarter and location; both waves
+and nutrients are known to influence kelp population dynamics and synchrony 
+\citep{Cavanaugh2013,CastoraniSheppardKelp}.
+
 %Tentative broad outline
-%0) above stuff about the case studies 
-%1) Para from Tom on how remote sensing data are gathered and coverted into what he handed to me. Brief, mostly 
-%citing other works where it is laid out in much more detail. 
-%2) Para from Tom on how he got to the wave and NO3 data
-%3) Para summarizing the data cleaning I did
-%Comment) Actually, might want to send the abive 3 paras to sup mat, and just have one para describing the final format of the
-%data, especially if you make this paper also be about plankton. But do that later if you do it.
-%5) Para about the linear modelling - this is how certain parts of the main theoretical equation were estimated
-%6) Para about spectral estimates. Brief, cites appropriate parts of Brillinger, uses sup mat if needed. This is how certain
-%other parts of the main theoretical equation were estimated. 
-%Comment) Somehow need to get across what was done - different parts of main equation were examined to explore the different
-%aspects of how the mechanism works for the focal taxa. Elaborate as appropriate.
+%2) Para about the linear modelling - this is how certain parts of the main theoretical equation were estimated,
+%about the lags we used, and justification for them etc. Might also need to cite a sup mat section. Maybe 
+%we need one single sup mat section called "supplementary methods" or "details of methods" or something. Actually
+%let's have two of these, one for data, one for stats
+%3) Para about spectral estimates. Brief, cites appropriate parts of Brillinger, uses sup mat if needed. This is how certain
+%other parts of the main theoretical equation were estimated. Basically it's just about how to do spectral
+%estimates of spectra and cross spectra, and then rest follows from the main theoretical equation.
+%4) Somehow need to get across what was done - different parts of main equation were examined to explore the different
+%aspects of how the mechanism works for the focal taxon. Elaborate as appropriate.
 
-%Para 1 from Tom
-A process was previously developed (REFS) using remote sensing to provide quarterly kelp biomass and kelp canopy area estimates,
-from quarter 1 of 1987 to quarter 4 of 2019, at $437$ locations along the California coast, from slightly north of Monterey Bay
-to San Diego. We very briefly describe the process, refering to our earlier published works for details.
-%Need to mention: The spacing and extent of each site; where the data can be downloaded (other than in our git repo, though I'll
-%a mention of that, too); lots of references as necessary; probably need slightly more formal statements of what "kelp biomass" 
-%"kelp canopy area" mean in these contextx; anything else important. Need bibliographic data for cited papers.
-
-%Para 2 from Tom
-Another process was previously developed (REFS) to provide mean nitrate concentration and maximum wave height data
-for the same quarters and locations. We again briefly describe the process, citing earlier works for details.
-
-%I asked Tom for these paragraphs by email on 2021 08 05
-
-%Need:
-%1) Some kind of map of the locations and display of some time series for the sup mat, to give a sense of the data. I think Max will
-%have made one of these for his paper with Lawrence, so maybe I could get him to make another one, same or appropriately modified?
+Data for the project are archived at X.
+%***DAN: Need to get the data location from Jon, I think he posted it, right? Maybe he did not put it all i
+%one place. Check with him. At the end of the Max/Lawrence paper, several data locations are listed.
+%I have an email out to Jon and Max.
+All computations were done in R version 3.4.4 on an ordinary laptop which was running Ubuntu Linux
+16.04. Complete codes and data are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.
 
@@
 %which can separate biological from observation processes.   
+%   Tom also told me in a meeting zoom call that in central CA growth of new fronds will be slower because 
+%there are not other fronds (left) on the same plant to provide a sugar subsidy to new fronds. In central CA,
+%basically all fronds from the same plant are ripped out.
+
+%Need a paragraph (already mentioned above) on how a frequency/timescale-specific approach
+%is necessary for all this. Need to think more about it first, i.e., how would you attempt to 
+%do this without such an approach, and how far would you get? Or maybe that is not quite the 
+%right at it (though you should try thinking that through and see what you get), maybe another
+%possibly better way would be to see that everything we list here can depend on timescale.
+%You can get a non-timescale-specific approach just by integrating our main equation across
+%frequencies. However, that will obscure the fact that you can get strong positive interaction 
+%effects on some timescales and strong negative ones on others. What would you get in our 
+%theoretical case studies if you integrated and would it reveal what's really going on? What 
+%are the implications of timescale structure for practical situations where populations and
+%environmental variables are measured over a certain typical duration (i.e, short time series)?
 
@@
 A timescale-specific approach is essential to this reasoning.}\label{fig:IntroTheory}
+\end{figure}
+
+\clearpage
+\newpage
+
+\begin{figure}
+\caption{To fill.}\label{fig:map}
 \end{figure}

### 2022-01-13 Paper/SupMat.Rnw
continued working on Methods

@@
 
+resloc<-"../Results/Kelp_DataAfterSuperBasicCleaning/"
+kelp_SB<-readRDS(paste0(resloc,"Kelp_Quarterly_CleanedSuperBasic.Rds"))
+resloc<-"../Results/Kelp_DataAfterMoreInvolvedCleaning/"
+kelp_MI<-readRDS(paste0(resloc,"Kelp_Quarterly_CleanedMoreInvolved.Rds"))
 @
@@
 
-\section{Details of data and data cleaning}
+\section{Details of data and data cleaning}\label{sect:data}
 
-%***DAN: This should include specification of the regions CC1, CC2, SB.
-%You have to get some info from Tom about the nature of the spatial sampling, and the process leading up to what he gave me.
-%You have to then describe the stages of cleaning I did.
+Floating surface canopy of kelp was measured by remote sensing, with details and
+validation reported by \cite{Cavanaugh2011} and \cite{Bell2020, Bell2021}. Multispectral Landsat
+imagery of $30$m resolution was used. Quarterly mean biomass values and spatial extent of canopy  
+within $500$m coastline segments (henceforth called ``loactions'')
+were computed by averaging. Quarters were: Q1, January-March; Q2, April-June; 
+Q3, July-September; Q4, October-December. 
+Data were thusly available from $\Sexpr{dim(kelp_SB)[1]}$ locations spanning 
+approximately from Monterey Bay to San Diego. 
+We then chose locations for analysis for which kelp was ``persistent'', in this 
+context meaning absent or missing data for no more than three years; this was to focus on 
+population dynamics of kelp rather than extinction-recolonization dynamics, which occur in some 
+locations and which can relate to sediment dynamics on the ocean floor and other factors
+not currently of primary interest. Of the original $\Sexpr{dim(kelp_SB)[1]}$ locations,
+$\Sexpr{dim(kelp_MI)[1]}$ ($\Sexpr{100*round(dim(kelp_MI)[1]/dim(kelp_SB)[1],2)}$\%) 
+were deemed persistent according to this criterion. Missing values in the remaining 
+time series were then filled with seasonal medians. The biomass and spatial extent time series
+for each location were then used together to produce one time series estimating biomass
+density per unit useable habitat at the site, as follows: every value of 
+the biomass time series was divided by the $90^{\text{th}}$ quantile of the distribution
+of the spatial extent time series at the site. The $90^{\text{th}}$ quantile was used instead
+of the maximum value of the spatial extent time series because maximum statistics have high variance. 
+This procedure served to normalize total biomass by an estimate
+of total useable habitat at a site. Kelp establish on rocky substrate, and areas with sandy or
+sedimented substrate are not useable kelp habitat. All time series were then linearly detrended,
+and only those time series were utilized which were located within the three regions pictured
+in Fig. \ref{MT-fig:map} in the main text. Regions were selected to represent distinct dynamical
+patterns known to occur in central and southern CA, and also to have another the same number
+of time series per selected region.
+
+Wave nutrient data were also available for each quarter at each location. Wave data were derived 
+by applying a swell propagation model with bouy-based wave measurements and a validation 
+techniques \citep{Wingeart2001,Hansen2009,Bell2015,OReilly2016}. Data on nitrates 
+were produced \citep{Snyder2020} by exploiting an established empirical 
+relationship between surface nitrate concentration
+and sea surface temperature (SST) in coastal CA \citep{Zimmerman1984,Palacios2013,Snyder2020}
+and SST data derived from Advanced Very High Resolution Radiometer satellite imagery. 
+Wave and nutrient time series in each location were again linearly detrended before use.
+
+\section{Details of statistical methods}
 

### 2022-01-13 Code/Kelp_AnalysisStep_SpectralDecomp.R

@@ -114,7 +114,10 @@ ncf::gcdist(locs$Lon[h],locs$Lat[h]) #73.35km
 h<-c(CC2locstouse[1],CC2locstouse[length(CC1locstouse)])
 ncf::gcdist(locs$Lon[h],locs$Lat[h]) #78.46km
 
-#So my three roughly comparable regions are SBlocstouse, CC1locstouse, CC2locstouse
+#So my three roughly comparable regions are SBlocstouse, CC1locstouse, CC2locstouse. Save them.
+saveRDS(SBlocstouse,file=paste0(resloc,"SBlocstouse.Rds"))
+saveRDS(CC1locstouse,file=paste0(resloc,"CC1locstouse.Rds"))
+saveRDS(CC2locstouse,file=paste0(resloc,"CC2locstouse.Rds"))
 
 #***
 #The main function for doing the analysis on a particular of locations

### 2022-01-14 Paper/Paper.Rnw
working on Methods and Results

@@
 %***Tentative author order, Bell last for his data contributions
-\author[a,b,*]{Daniel C. Reuman}
-\author[c]{Max Castorani}
-\author[d]{Kyle Cavanaugh}
-\author[e]{Lawrence W. Sheppard}
-\author[c]{Jonathan Walter}
-\author[f,g]{Tom Bell}
+\author[a,*]{Daniel C. Reuman}
+\author[b]{Max Castorani}
+\author[c]{Kyle Cavanaugh}
+\author[d]{Lawrence W. Sheppard}
+\author[b]{Jonathan Walter}
+\author[e,f]{Tom Bell}
 
 \affil[a]{Department of Ecology and Evolutionary Biology and Center for Ecological Research, University of Kansas}
-\affil[b]{Laboratory of Populations, Rockefeller University}
-\affil[c]{Department of Environmental Sciences, University of Virginia}
-\affil[d]{Department of Geography, University of California, Los Angeles}
-\affil[e]{Marine Biological Association of the United Kingdom}
-\affil[f]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
-\affil[g]{Earth Research Institute, University of California, Santa Barbara}
+\affil[b]{Department of Environmental Sciences, University of Virginia}
+\affil[c]{Department of Geography, University of California, Los Angeles}
+\affil[d]{Marine Biological Association of the United Kingdom}
+\affil[e]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
+\affil[f]{Earth Research Institute, University of California, Santa Barbara}
 \affil[*]{Corresponding author, reuman@ku.edu}
@@
 
-\section{General theory}
+\section{General Theory}
 
@@
 %Description of data after cleaning
-The core of this paper is the new theory, but to help illustrate that theory, we also apply
-it to an exceptional dataset on giant kelp (\emph{Macrocyctis pyrifera}) dynamics off the CA coast, and 
+The core of this paper is the new theory, but to help illustrate that theory, fulfilling goal G3 of the 
+Introduction, we also apply
+the theory to an exceptional dataset on giant kelp (\emph{Macrocyctis pyrifera}) dynamics off the CA coast, and 
 associated environmental measurements. 
@@
 
-%Tentative broad outline
-%2) Para about the linear modelling - this is how certain parts of the main theoretical equation were estimated,
-%about the lags we used, and justification for them etc. Might also need to cite a sup mat section. Maybe 
-%we need one single sup mat section called "supplementary methods" or "details of methods" or something. Actually
-%let's have two of these, one for data, one for stats
-%3) Para about spectral estimates. Brief, cites appropriate parts of Brillinger, uses sup mat if needed. This is how certain
-%other parts of the main theoretical equation were estimated. Basically it's just about how to do spectral
-%estimates of spectra and cross spectra, and then rest follows from the main theoretical equation.
-%4) Somehow need to get across what was done - different parts of main equation were examined to explore the different
-%aspects of how the mechanism works for the focal taxon. Elaborate as appropriate.
+Coefficients of the model (\ref{main_model_1})-(\ref{main_model_4}) were separately selected for each of our three regions 
+using linear regression methods and building on extensive prior work on the drivers of kelp dynamics. A no-intercept 
+regression model of kelp at time $t$ in location $i$, i.e., 
+$w_i(t)$ in (\ref{main_model_1}), against lagged values of kelp 
+($w_i(t-l)$ for $l=1,\ldots,n$ in (\ref{main_model_1})), lagged and unlagged values of nitrates
+($\epsilon_i^{(1)}(t-l)$ for $l=0,\ldots,m_1$ in (\ref{main_model_2})), and values of waves
+($\epsilon_i^{(2)}(t-l)$ for $l=0,\ldots,m_2$ in (\ref{main_model_3})) was fitted using standard regression methods.
+The same regression coefficients were used for all locations within a region.
+Waves influence kelp dynamics through direct disturbance events which can damage kelp or extirpate kelp locally 
+when waves are large \citep{Dayton1992,Cavanaugh2011,Bell2015}. Thus wave effects 
+are immediate and $m_2=0$ was used. Nitrates are known to fuel rapid kelp growth, though in some areas
+effects appear delayed by about one quarter because our kelp data quantify canopy (surface) biomass, and it can
+take time for damaged kelp to grow back to the surface 
+\citep{Zimmerman1986,Cavanaugh2011,Bell2015}. 
+Therefore $m_1=1$ was used. Kelp holdfasts on the sea floor can last for multiple years, so kelp lag effects were considered:
+we used $n=4, 8, 12$ in separate analysis, with this choice making no substantive difference to results (see Results).
+Regression coefficient determined the quantities $b_1,\ldots,b_n$, $p_0^{(1)}$, $p_1^{(1)}$, and $p_0^{(2)}$
+in (\ref{main_model_1})-(\ref{main_model_3}), and therefore the quantities $f_{\mcB}$, $f_{\mcP^{(1)}}$ and
+$f_{\mcP^{(2)}}$ in (\ref{rhoww_T1to3}), for each of our three regions (see General Theory).
 
-Data for the project are archived at X.
-%***DAN: Need to get the data location from Jon, I think he posted it, right? Maybe he did not put it all i
-%one place. Check with him. At the end of the Max/Lawrence paper, several data locations are listed.
-%I have an email out to Jon and Max.
+To estimate the components $\rho_{ww}$, $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and 
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ in (\ref{rhoww_T1to3}), where $w$ was detrended kelp time series in one of our regions, 
+$\epsilon^{(1)}$ was detrended nitrate time
+series in the region, and $\epsilon^{(2)}$ was detrended wave disturbance time series in the region, 
+we applied the definitions of these quantities (see General Theory), which required estimation from data of
+the spectra and cross spectra $S_{w_i w_j}$ and $S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}$ for $i,j=1,\ldots,N$
+and $a,b=1,2$, where $N$ is the number of locations in the region being considered. Spectral quantities were
+computed using the consistent estimator of section 7.4 of \cite{Brillinger2001}. 
+%***DAN: check this section number, and if possible give a specific formula of theorem or something, and also 
+%check the below text and maybe augment it
+The estimator is a smoothed periodogram, with the width of the smoothing kernel selected to 
+increase with the square root of time series length. Theory was interpreted in relation to kelp ecology and
+theoretical case studies by plotting the components of (\ref{rhoww_T1to3}) for each of our regions.
+
+Data for the project are publicly archived \citep{SBCLTER2021}.
 All computations were done in R version 3.4.4 on an ordinary laptop which was running Ubuntu Linux
-16.04. Complete codes and data are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.
+16.04. Complete codes for the entire project workflow are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.
 
@@
 %Para basically confirming earlier work, as a starter
-Our theoretical results demonstrate that interaction effects between Moran drivers: 1) can be 
+Helping fulfill goal G2 of the Introduction, our theoretical results 
+demonstrate that interaction effects between Moran drivers: 1) can be 
 comparable in strength to direct Moran effects; 2) can be either synergistic or destructive, i.e., positive or negative; and 3) 
@@
 
+%Case syiudy B
 Case study B reveals how Moran effects can interact when lags are longer than 1 model time step. 
@@
 
+%Case study C
 Cases A and B assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
@@
 positive and those of $\epsilon^{(2)}$ were negative.
+
+%Paralels the first para, but for kelp
+We now apply our theory to kelp, fulfilling goal G3 of the Introduction. Kelp results confirm
+the earlier theoretical results, based on our case studies 
+(see above text and Fig. \ref{fig:TheoryCaseA}a,e,i,m),
+that interaction effects between Moran drivers: 1) can be comparble in strength to direct Moran
+effects; 2) can be either synergistic or destructive; and 3) can 
+depend strongly on timescale, even having different signs on different timescales.
+First, for all of our regions and for essentially all timescales, the magintude of interaction 
+effects was comparable to that of direct Moran effects (Fig. \ref{fig:SummaryKelpFig}
+for the CCA1 and SB regions and Fig. \ref{SM-fig:SummaryKelpFig_SM_lag4} for the CCA2 region).
+Second, interaction effects could be positive (e.g., annual timescales, CCA1 region on Fig. 
+\ref{fig:SummaryKelpFig}a), or negative (e.g., annual timescales, SB region on Fig. 
+\ref{fig:SummaryKelpFig}d, or long timescales $>8$y for either region, Fig. 
+\ref{fig:SummaryKelpFig}c,f). Finally, interactions depended on timescale, e.g., for the CCA1 
+region they were positive on annual timescales (Fig. \ref{fig:SummaryKelpFig}a) and
+negative on timescales $>8$y (Fig. \ref{fig:SummaryKelpFig}c).
+
+The way direct Moran effects work in our theory is fairly straightforward. 
+For instance, the direct Moran effect
+term in (\ref{rhoww_T1to3}) for nitrates, i.e., 
+$|f_{\mcP^{(1)}}|^2 \rho_{\epsilon^{(1)} \epsilon^{(1)}}/|f_{\mcB}|^2$, 
+equals the synchrony of the nitrate time series themselves, i.e.,
+$\rho_{\epsilon^{(1)} \epsilon^{(1)}}$, times the 
+timescale-specific strength of influence of nitrates on kelp, 
+i.e., $|f_{\mcP^{(1)}}|^2$, and modulated
+by a term quantifying the autoregressive nature of kelp dynamics,
+$1/|f_{\mcB}|^2$. These components are plotted in Fig. \ref{fig:directMoran} (CCA1 and SB regions) and 
+Fig. \ref{SM-fig:directMoran_SM_lag4} (CCA2 region) so their 
+effects on synchrony can be assessed across timescales and regions. 
+Both nitrates and waves fluctuate seasonally in CA. Thus nitrate and wave 
+synchrony had a strong annual component (Fig. \ref{fig:directMoran}b,d,q,s), which produced 
+some of the annual synchrony observed in kelp 
+(Fig. \ref{fig:SummaryKelpFig}a,d, green and blue lines).
+Nitrates and waves are also synchronous on long timescales $>8$y in CA 
+(Fig. \ref{fig:directMoran}l,n,aa,cc), possibly due to the North Pacific Gyre Oscillation [NPGO; \cite{CastoraniSheppardKelp}]. The long-timescale synchrony
+in nitrates and waves produced some of the long-timescale
+synchrony observed in kelp (Fig. \ref{fig:SummaryKelpFig}c,f, green and blue lines).
+Kelp synchrony was generally stronger in the CCA1 region compared to the SB region 
+(compare Fig. \ref{fig:SummaryKelpFig}a-c to d-f) in part because the synchrony of nitrates, and of 
+waves, was more pronounced in the CCA1 region than in the SB region 
+(compare Fig. \ref{fig:directMoran}b,g,l to q,v,aa, and 
+Fig. \ref{fig:directMoran}d,i,n to s,x,cc), and also because waves
+had a stronger influence on kelp in CCA1 compared to SB (compare 
+Fig. \ref{fig:directMoran}c,h,m to r,w,bb).
+Regression coefficients which determined $f_{\mcP^{(1)}}$, $f_{\mcP^{(2)}}$
+and $f_{\mcB}$ (see Methods) are in Table \ref{SM-tab:regcoefs}.
 
@@
 bringing about population oscillations on the given timescale.  
-A timescale-specific approach is essential to this reasoning.}\label{fig:IntroTheory}
+A timescale-specific approach is essential to this reasoning. See the online version for a color
+rendering of all figures.}\label{fig:IntroTheory}
 \end{figure}
@@
 \begin{figure}
-\caption{To fill.}\label{fig:map}
+\caption{To fill. See the online version for a color
+rendering of all figures.}\label{fig:map}
 \end{figure}
@@
 and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.
-Synch. contrib. = Synchrony contribution.}\label{fig:TheoryCaseA}
+Synch. contrib. = Synchrony contribution. See the online version for a color
+rendering of all figures.}\label{fig:TheoryCaseA}
 \end{figure}
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_MainPlot_ForMS.pdf}
-\caption{Filler.}\label{fig:SummaryKelpFig}
+\caption{Main components of the new theory as applied kelp, central CA 1 (CCA1) region (a-c) and 
+Santa Barbara (SB) region (d-f). Panels show the terms on the right side of 
+(\ref{rhoww_T1to3}). Green lines show 
+$\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}$, quantifying 
+kelp population synchrony due to the direct Moran effects of $\epsilon^{(1)}$, which, in this 
+context, is nitrates. The blue line
+is $\frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}$, quantifying 
+kelp population synchrony due to the direct Moran effects of $\epsilon^{(2)}$, which,
+in this context, is waves.
+The dashed green-blue line is 
+$\frac{2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)}{|f_{\mcB}|^2}$, 
+which is population synchrony due to interacting Moran effects. The red line is the sum of the
+green, blue and green-blue lines, and is the portion of synchrony explained by nitrates,
+waves, and their interactions. Explained synchrony does not equal total kelp synchrony
+(black line) because other, unmeasured factors also help synchronize kelp dynamics.
+The CCA1 and SB regional results, respectively, approximately parallel the 
+results for theoretical case
+study C scenarios 1 and 2 (Fig. \ref{fig:TheoryCaseA}i,m; see text for details).
+See Fig. \ref{SM-fig:SummaryKelpFig_SM_lag4} for the central CA 2 (CCA2) region,
+for which results were substantially the same as for CCA1. This figure used 
+kelp lag $n=4$ (see (\ref{main_model_1}) for the meaning of $n$); 
+see Figs \ref{SM-fig:SummaryKelpFig_SM_lag8} 
+and \ref{SM-fig:SummaryKelpFig_SM_lag12} for $n=8, 12$, respectively.
+See the online version for a color rendering of this 
+figure.}\label{fig:SummaryKelpFig}
 \end{figure}
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainDirectMoran_ForMS.pdf}
-\caption{Filler.}\label{fig:directMoran}
+\caption{Components of direct Moran effects for kelp, for the CCA1 (a-o)
+and SB regions (p-dd). The index $1$ denotes nitrates, and $2$ denotes waves.
+Vertical axis extents are the same for corresponding panels for the
+two regions, to facilitate comparisons. 
+See Fig. \ref{SM-fig:directMoran_SM_lag4} for the central CCA2 region,
+for which results were similar to CCA1. This figure used 
+kelp lag $n=4$; see Figs \ref{SM-fig:directMoran_SM_lag8} 
+and \ref{SM-fig:directMoran_SM_lag12} for $n=8, 12$, respectively.
+See the online version for a color
+rendering of all figures.}\label{fig:directMoran}
 \end{figure}
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
-\caption{Filler.}\label{fig:indirectMoran}
+\caption{Filler. See the online version for a color
+rendering of all figures.}\label{fig:indirectMoran}
 \end{figure}

### 2022-01-14 Paper/SupMat.Rnw
working on Methods and Results

@@
 kelp_MI<-readRDS(paste0(resloc,"Kelp_Quarterly_CleanedMoreInvolved.Rds"))
+
+resloc<-"../Results/Kelp_SpectralDecompResults/"
+CCA1l4<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal1_KelpLag4.Rds"))
+CCA2l4<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal2_KelpLag4.Rds"))
+SBl4<-readRDS(paste0(resloc,"RegionalAnalysis_SoCal_KelpLag4.Rds"))
+CCA1l8<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal1_KelpLag8.Rds"))
+CCA2l8<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal2_KelpLag8.Rds"))
+SBl8<-readRDS(paste0(resloc,"RegionalAnalysis_SoCal_KelpLag8.Rds"))
+CCA1l12<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal1_KelpLag12.Rds"))
+CCA2l12<-readRDS(paste0(resloc,"RegionalAnalysis_CentCal2_KelpLag12.Rds"))
+SBl12<-readRDS(paste0(resloc,"RegionalAnalysis_SoCal_KelpLag12.Rds"))
 @
@@
 
-\section{Details of statistical methods}
-
 \section{Supporting results for kelp}
@@
 \caption[Extension of main text Fig. \ref{MT-fig:SummaryKelpFig}]{Same format as Fig. \ref{MT-fig:SummaryKelpFig} of the main text, but including not only results
-for the first central CA region (a-c) and the southern CA region (d-f), but also results from the 
-second central CA region (g-i).}\label{fig:SummaryKelpFig_SM_lag4}
+for the CCA1 region (a-c) and the SB region (d-f), but also results from the 
+CCA2 region (g-i).}\label{fig:SummaryKelpFig_SM_lag4}
 \end{figure}
@@
 \caption[Extension of main text Fig. \ref{MT-fig:directMoran}]{Same format as Fig. \ref{MT-fig:directMoran} of the main text, but including not only results
-for the first central CA region (a-o) and the southern CA region (p-dd), but also results from the 
-second central CA region (ee-ss).}\label{fig:directMoran_SM_lag4}
+for the CCA1 region (a-o) and the SB region (p-dd), but also results from the 
+CCA2 region (ee-ss).}\label{fig:directMoran_SM_lag4}
 \end{figure}
@@
 \caption[Extension of main text Fig. \ref{MT-fig:indirectMoran}]{Same format as Fig. \ref{MT-fig:indirectMoran} of the main text, but including not only results
-for the first central CA region (a-o) and the southern CA region (p-dd), but also results from the 
-second central CA region (ee-ss).}\label{fig:indirectMoran_SM_lag4}
+for the CCA1 region (a-o) and the SB region (p-dd), but also results from the 
+CCA2 region (ee-ss).}\label{fig:indirectMoran_SM_lag4}
 \end{figure}
@@
 \clearpage
-\newpage
+
+\begin{table}[!h]
+\caption{Regression coefficients used to parameters (\ref{MT-main_model_1})-(\ref{MT-main_model_4}) of the main text for each 
+of our three regions. All values rounded to two decimal places.}\label{tab:regcoefs}
+\begin{center}
+\begin{tabular}{r|l|l|l|l|l}
+ & $p_0^{(1)}$ & $p_1^{(1)}$ & $p_0^{(2)}$ & $b_1,b_2,\ldots,b_n$ & $R^2$\\
+ \hline
+%n=4
+CCA1, $n=4$ & \Sexpr{round(unname(CCA1l4[6]),2)} & \Sexpr{round(unname(CCA1l4[7]),2)} & \Sexpr{round(unname(CCA1l4[8]),2)} & \Sexpr{paste0(round(unname(CCA1l4[2:5]),2),collapse=",")} & \Sexpr{round(unname(CCA1l4[1]),2)} \\
+CCA2, $n=4$ & \Sexpr{round(unname(CCA2l4[6]),2)} & \Sexpr{round(unname(CCA2l4[7]),2)} & \Sexpr{round(unname(CCA2l4[8]),2)} & \Sexpr{paste0(round(unname(CCA2l4[2:5]),2),collapse=",")} & \Sexpr{round(unname(CCA2l4[1]),2)} \\
+SB, $n=4$ & \Sexpr{round(unname(SBl4[6]),2)} & \Sexpr{round(unname(SBl4[7]),2)} & \Sexpr{round(unname(SBl4[8]),2)} & \Sexpr{paste0(round(unname(SBl4[2:5]),2),collapse=",")} & \Sexpr{round(unname(SBl4[1]),2)} \\
+%n=8
+CCA1, $n=8$ & \Sexpr{round(unname(CCA1l8[10]),2)} & \Sexpr{round(unname(CCA1l8[11]),2)} & \Sexpr{round(unname(CCA1l8[12]),2)} & \Sexpr{paste0(round(unname(CCA1l8[2:9]),2),collapse=",")} & \Sexpr{round(unname(CCA1l8[1]),2)} \\
+CCA2, $n=8$ & \Sexpr{round(unname(CCA2l8[10]),2)} & \Sexpr{round(unname(CCA2l8[11]),2)} & \Sexpr{round(unname(CCA2l8[12]),2)} & \Sexpr{paste0(round(unname(CCA2l8[2:9]),2),collapse=",")} & \Sexpr{round(unname(CCA2l8[1]),2)} \\
+SB, $n=8$ & \Sexpr{round(unname(SBl8[10]),2)} & \Sexpr{round(unname(SBl8[11]),2)} & \Sexpr{round(unname(SBl8[12]),2)} & \Sexpr{paste0(round(unname(SBl8[2:9]),2),collapse=",")} & \Sexpr{round(unname(SBl8[1]),2)} \\
+%n=12
+CCA1, $n=12$ & \Sexpr{round(unname(CCA1l12[14]),2)} & \Sexpr{round(unname(CCA1l12[15]),2)} & \Sexpr{round(unname(CCA1l12[16]),2)} & \Sexpr{paste0(round(unname(CCA1l12[2:13]),2),collapse=",")} & \Sexpr{round(unname(CCA1l12[1]),2)} \\
+CCA2, $n=12$ & \Sexpr{round(unname(CCA2l12[14]),2)} & \Sexpr{round(unname(CCA2l12[15]),2)} & \Sexpr{round(unname(CCA2l12[16]),2)} & \Sexpr{paste0(round(unname(CCA2l12[2:13]),2),collapse=",")} & \Sexpr{round(unname(CCA2l12[1]),2)} \\
+SB, $n=12$ & \Sexpr{round(unname(SBl12[14]),2)} & \Sexpr{round(unname(SBl12[15]),2)} & \Sexpr{round(unname(SBl12[16]),2)} & \Sexpr{paste0(round(unname(SBl12[2:13]),2),collapse=",")} & \Sexpr{round(unname(SBl12[1]),2)} \\
+\end{tabular}
+\end{center}
+\end{table}
 

### 2022-01-14 Code/Kelp_AnalysisStep_SpectralDecomp.R

@@ -1604,7 +1604,7 @@ make_plots_ForMS<-function(freq,frg,totsync,T1_avg,T2_avg,T3_avg,T4_avg,T5_avg,T
   lines(l2timescales[inds],T4_avg[inds],type="l",lty="solid",col="green") 
   lines(l2timescales[inds],T4_avg[inds],type="l",lty="dashed",col="blue") #interactions between waves and NO3
   lines(l2timescales[inds],sumexpl[inds],type="l",col="red",lty="solid") #explained stuff - sum of all the terms relating only to NO3 and waves (direct NO3 and wave effects, and interactions between those two)
-  lines(l2timescales[inds],sumterms[inds],type="l",lty="dashed") #total of all terms, should approx equal the plot of totsync
+  #lines(l2timescales[inds],sumterms[inds],type="l",lty="dashed") #total of all terms, should approx equal the plot of totsync
   #lines(l2timescales[inds],sumunexpl[inds],type="l",lty="solid",col="red") #unexplained
   lines(l2timescales[inds],rep(0,length(timescales[inds])))
   lines(rep(log2(4),2),ylimits) #plot a vertical line at the annual frequency
@@ -1635,7 +1635,7 @@ make_plots_ForMS<-function(freq,frg,totsync,T1_avg,T2_avg,T3_avg,T4_avg,T5_avg,T
   lines(l2timescales[inds],T4_avg[inds],type="l",lty="solid",col="green") 
   lines(l2timescales[inds],T4_avg[inds],type="l",lty="dashed",col="blue") #interactions between waves and NO3
   lines(l2timescales[inds],sumexpl[inds],type="l",col="red",lty="solid") #explained stuff - sum of all the terms relating only to NO3 and waves (direct NO3 and wave effects, and interactions between those two)
-  lines(l2timescales[inds],sumterms[inds],type="l",lty="dashed") #total of all terms, should approx equal the plot of totsync
+  #lines(l2timescales[inds],sumterms[inds],type="l",lty="dashed") #total of all terms, should approx equal the plot of totsync
   #lines(l2timescales[inds],sumunexpl[inds],type="l",lty="solid",col="red") #unexplained
   lines(l2timescales[inds],rep(0,length(timescales[inds])))
   
@@ -1664,7 +1664,7 @@ make_plots_ForMS<-function(freq,frg,totsync,T1_avg,T2_avg,T3_avg,T4_avg,T5_avg,T
   lines(l2timescales[inds],T4_avg[inds],type="l",lty="solid",col="green") 
[…]

### 2022-01-16 Paper/Paper.Rnw
continuting to write and edit

@@
 [\sqrt{2+2\cov(\epsilon_1^{(1)},\epsilon_1^{(2)})}\sqrt{2+2\cov(\epsilon_2^{(1)},\epsilon_2^{(2)})}]$,
-a quantity which depends not only on the environmental synchrony measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(a)})$, for 
+a quantity which depends not only on the standard environmental synchrony measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(a)})$, for 
 $a=1,2$, but also on the ``cross synchrony'' measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(b)})$
@@
 therefore differ from what synchrony would be if the environmental processes $\epsilon^{(1)}$ and 
-$\epsilon^{(2)}$ were unrelated. The difference is due to interactions
+$\epsilon^{(2)}$ were unrelated, so that covariances between components of those processes are zero. 
+The difference is due to interactions
 between the two Moran effects, and can be substantially positive or negative. This example was adapted 
@@
 Having demonstrated that interactions between Moran effects can add to or reduce population synchrony, we now 
-begin to provide some intuition on how we can understand the underlying mechanisms. 
+begin to provide some intuition on how to understand the underlying mechanism. 
 A key insight is the use of a timescale-specific approach,
@@
 we assume for simiplicity that larger values of both environmental variables are beneficial to populations, though the general
-theory represented below does require that assumption. If, as in Fig. \ref{fig:IntroTheory}b, $l_{e1}=l_n+l_{e2}$, then the 
-periodic maximal positive influences of environmental variables coincide. If environmental signals are then spatially synchronous
+theory described below does require that assumption. If, as in Fig. \ref{fig:IntroTheory}b, $l_{e1}-l_{e2}+l_n=0$, then the 
+periodic maximal positive influences of the environmental components coincide. If environmental signals are then spatially synchronous
 across sampling locations, the fact of the alignment of lags produces additional synchrony in populations, 
-beyond what would manifest if environmental fluctuations were unrelated. In contrast, Fig. \ref{fig:IntroTheory}c
-shows the opposite scenario, for which $l_{e1}$ plus the half the period, $10$, equals $l_n+l_{e2}$, so that  
+beyond what would manifest if environmental fluctuations were unrelated, because positive influences of variable 1 in location $i$ will
+tend to coincide with positive influences of variable 2 in location $j$. In contrast, Fig. \ref{fig:IntroTheory}c
+shows the opposite scenario, for which $l_{e1}-l_{e2}+l_n$ equals half the period, $10$, so that  
 the periodic maximal positive influence of environmental variable 1 coincides with the periodic maximal negative
 influence of environmental variable 2, reducing synchrony relative to the case of unrelated environmental variables. 
-Intermediate scenarios are possible, as will be revealed by the general theory, below.
+Intermediate scenarios are possible, as will be revealed by the general theory.
 The theory also formalizes the timescale decomposition of synchrony measures, and timescale-specific lags.
 
-Thus our specific goals for thus study are the following. First, (G1) we will elaborate a general mathematical 
+Our specific goals for this study are as follows. First, (G1) we will elaborate a general mathematical 
 theory of interacting Moran effects which allows a detailed understanding of how the mechanism works. Second, 
@@
 for kelp ecology, we instead use a simplified subset of the available kelp data, and our analysis is instead intended 
-to illuminate the inner workings of our theory and of the general mechanisms we describe. In the Discussion, we 
+to illuminate the inner workings of our theory and of the general mechanism we describe. In the Discussion, we 
 extend the arguments of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} that interactions between Moran 
@@
 ecosystems. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
-illuminate this newly observed but potentially quite common mechanism of spatiotemporal population dynamics.
+illuminate this newly observed but potentially quite common and important mechanism of spatiotemporal population dynamics.
 
@@
 $w_j(t)$ are both strong and strongly correlated, i.e., they have the same or similar phase. 
-(Here, $j$ refers to another location where measurements were performed.) The cospectrum is 
+(Here, $j$ refers to another location where measurements were taken.) The cospectrum is 
 the real part of the cross spectrum, $S_{w_i w_j}(f)$, which is a complex-valued function of frequency.
@@
 and the complex phase of the cross spectrum at $f$ then quantifies
-the phase. The notation $S_{ww}(f)$ refers to the \emph{spectral matrix}, 
+that phase. The notation $S_{ww}(f)$ refers to the \emph{spectral matrix}, 
 which has $ij^{\text{th}}$ entry $S_{w_i w_j}(f)$. Spectral methods are standard tools 
@@
 
-Much prior work demonstrates the importance of a frequency-specific approach to
-synchrony [e.g., \cite{Vasseur2007,Keitt2008,Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}], 
-and it will turn out (see below, and \emph{Discussion}) that a frequency-specific 
+Much prior work demonstrates the importance of a frequency- or timescale-specific approach to
+synchrony [e.g., \cite{Vasseur2007,Keitt2008,Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}; frequency- and timescale-specific approaches
+are equivalent because frequency and timescale are reciprocal], 
+and it will turn out (see below, and \emph{Discussion}) that a timescale-specific 
 approach is essential to the development of our new theory. 
 %***DAN: Good topic for a Discussion para, the fact that this could not really be done without taking frequency specificity into account, and why, though need to think a bit more before writing
-We therefore here define, using the spectral methods outlined above, a frequency-specific measure of spatial synchrony, 
+We therefore here define, using the spectral methods outlined above, a frequency/timescale-specific measure of spatial synchrony, 
 as well as a new concept of \emph{cross-variable spatial synchrony}. If time series data $w_i(t)$ for $t=1,\ldots,T$ 
@@
 by a year, then $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$, which is complex valued, will have high magnitude at
-frequency $f=\frac{1}{4} \text{ y}^{-1}$, and will have  phase at that frequency equal to $\pi/2$, reflecting the one-year 
+timescale $4$y, equivalent to frequency $f=\frac{1}{4} \text{ y}^{-1}$, and will have  phase at that frequency equal to $\pi/2$, reflecting the one-year 
 lag between the variables.
@@
 $\epsilon^{(2)}=(\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)})$ and $\delta=(\delta_1,\ldots,\delta_N)$
-are environmental processes. The processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
-taken to be measured environmental processes, whereas $\delta$ represents the 
+are environmental processes at the same locations. The processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
+taken to be measured, whereas $\delta$ represents the 
 aggregate influence of unmeasured processes. We assume that the combined process 
@@
 $\delta$ and of interactions between $\delta$ and the $\epsilon^{(i)}$. Such contributions would be
-difficult to assess because $\delta$ was unmeasured. Comparing the three terms on the right
+difficult to assess because $\delta$ was unmeasured.
+
+Comparing the terms on the right
 of (\ref{rhoww_T1to3}) gives the relative contributions of direct Moran effects and interactions 
-between Moran effects. The derivation of the theory is in SI section \ref{SM-sect:derivation}.
+between Moran effects. 
+The magnitudes of the quantities $f_{\mcP^{(1)}}$ and $f_{\mcP^{(2)}}$ quantify the strength of 
+influence of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ on populations at the timescale $\sigma=1/f$. 
+The phases of these quantities correspond to the delays in the influences of 
+$\epsilon^{(1)}$ and $\epsilon^{(2)}$, 
+represented on Fig. \ref{fig:IntroTheory} by $l_{e1}$ and $l_{e2}$, respectively. The lag 
+$l_n$ on Fig. \ref{fig:IntroTheory} manifests in the theory through the phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$. 
+The expression $l_{e1}-l_{e2}+l_n$ from the Introduction and Fig. \ref{fig:IntroTheory} corresponds to the phase of the expression 
+$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, because the phase of
+this product equals the phase of $f_{\mcP^{(1)}}$ (equivalent to $l_{e1}$), minus the phase of $f_{\mcP^{(2)}}$
+(equivalent to $l_{e2}$), plus the phase
+of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ (equivalent to $l_n$). 
+The real part of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+is positive and larger whenever the phase of this quantity is close to zero, i.e., when    
+$l_{e1}-l_{e2}+l_n$ is close to zero, and is increasingly negative as 
+the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ gets
+close to $\pi$, e.g., when $l_{e1}-l_{e2}+l_n$ is close to half the period of oscillation.
+The factor 
+$1/|f_{\mcB}|^2$ is how the intrinsic nature of population
+dynamics modulates Moran influences.
+The derivation of the theory is in SI section \ref{SM-sect:derivation}.
 
@@
 \noindent We now describe how we pursue goal G2 of the Introduction, to develop 
-theoretical case studies that illuminate the intuition of interacting Moran effects. 
-We develop case studies A-C. For all cases, the model time step was assumed to be one quarter, i.e., 
-four model time steps per year. This makes no real mathematical difference, but was done to 
-facilitate later comparisons with results based on kelp data, which were available with a
-quarterly sampling frequency. 
-For case study A,  the environmental variable 
+theoretical case studies A-C, that illuminate the intuition of interacting Moran effects. 
+For all cases, the model time step was assumed to be one quarter (q), i.e., 
+four time steps per year. This makes no real mathematical difference, but was done to 
+facilitate later comparisons with results for kelp data, which were sampled quarterly. 
+For case study A (henceforth CaA),  the environmental variable 
 $\epsilon^{(1)}$ is assumed to have a simple positive effect on populations, but 
 lagged by one model time step, i.e.,
-$p_0^{(1)}=0$, $p_1^{(1)}>0$, and $p_i^{(1)}=0$ for $i>1$. 
-For case study B, which generalizes A, $\epsilon^{(1)}$ is again assumed to have a simple 
+$p_1^{(1)}>0$, and $p_i^{(1)}=0$ for $i \neq 1$. 
+For case study B (henceforth CaB), which generalizes A, $\epsilon^{(1)}$ is again assumed to have a simple 
 positive effect on populations,
 but now lagged by $n$ time steps, so that $p_n^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n$.
-For both A and B, the effects of $\epsilon^{(2)}$
+For both CaA and CaB, the effects of $\epsilon^{(2)}$
 are assumed to be un-lagged, i.e., $p_0^{(2)} \neq 0$ and $p_i^{(2)}=0$ for $i>0$.
-For case study C, the effects of $\epsilon^{(1)}$ are lagged by $n_1$ and the 
+For case study C (CaC), the effects of $\epsilon^{(1)}$ are lagged by $n_1$ and the 
 effects of $\epsilon^{(2)}$ are lagged by $n_2$ time steps, i.e.,
@@
 $(\epsilon^{(1)},\epsilon^{(2)})$ is assumed to be a Gaussian white-noise process 
-for both A and B. So 
+for both CaA and CaB. So 
 the random variables $(\epsilon^{(1)}(t),\epsilon^{(2)}(t))$ are independent and 
-identically distributed (iid) for distinct times, $t$. Noise may be correlated across space,
-and the components of $\epsilon^{(1)}(t)$ may be correlated with those of $\epsilon^{(2)}(t)$.
-For C, the noise processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
+identically distributed (iid) for distinct times, $t$. Noise was positively correlated across space,
+and the components of $\epsilon^{(1)}(t)$ were positively correlated with those of $\epsilon^{(2)}(t)$.
+For CaC, the noise processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
 each assumed to exhibit spatially synchronous periodic oscillations of period one year, i.e.,
 four model time steps, but with different phases. Peaks in $\epsilon^{(1)}$ are assumed to 
-either lead or lag (we consider two model scenarios) peaks in $\epsilon^{(2)}$ by one quarter.
+either lead or lag (we consider two scenarios, CaC1 and CaC2) peaks in $\epsilon^{(2)}$ by $1$q.
 Such a scenario could be realized by annually 
@@
 e.g., kelp wave action in central CA peaks annually in the winter, whereas surface-water
-nitrate concentrations also fluctuate with period one year, but peak in the spring, a delay of one quarter
+nitrate concentrations also fluctuate with period one year, but peak in the spring, a delay of $1$q
 compared to the wave peak. 
@@
 
+%***DAN: Right now sup mat section S7 is not cited. It only has figures and tables, not sure I plan to 
+%add anything to it other than that. Either get rid of it as a section, or else cite it in some of the places 
+%where you currently cite the figures and tables in it. My current hunch is that the first of those options is
+%better.
+
 %Description of data after cleaning
-The core of this paper is the new theory, but to help illustrate that theory, fulfilling goal G3 of the 
+The core of this paper is the new theory, but to help illustrate that theory, for goal G3 of the 
 Introduction, we also apply
@@
 from southern CA, close to Santa Barbara
-(SB; see Fig. \ref{fig:map}). Each kelp measurement is an estimate of mean quarterly kelp biomass per unit 
+(called SB; see Fig. \ref{fig:map}). Each kelp measurement is an estimate of mean quarterly kelp biomass per unit 
 useable habitat in a 500m stretch of coastline. We used coastline segments where kelp was persistent
@@
 period (SI section \ref{SM-sect:data}). We also had estimates
-of maximum wave height and mean nitrate concentration for each quarter and location; both waves
+of maximum wave height and mean nitrate concentration for each quarter and location. Both waves
 and nutrients are known to influence kelp population dynamics and synchrony 
@@
 are immediate and $m_2=0$ was used. Nitrates are known to fuel rapid kelp growth, though in some areas
-effects appear delayed by about one quarter because our kelp data quantify canopy (surface) biomass, and it can
+effects appear delayed by about $1$q because our kelp data quantify canopy (surface) biomass, and it can
 take time for damaged kelp to grow back to the surface 
@@
 we used $n=4, 8, 12$ in separate analysis, with this choice making no substantive difference to results (see Results).
-Regression coefficient determined the quantities $b_1,\ldots,b_n$, $p_0^{(1)}$, $p_1^{(1)}$, and $p_0^{(2)}$
-in (\ref{main_model_1})-(\ref{main_model_3}), and therefore the quantities $f_{\mcB}$, $f_{\mcP^{(1)}}$ and
+Fitted regression coefficients determined the quantities $b_1,\ldots,b_n$, $p_0^{(1)}$, $p_1^{(1)}$, and $p_0^{(2)}$
+in (\ref{main_model_1})-(\ref{main_model_3}), and therefore $f_{\mcB}$, $f_{\mcP^{(1)}}$ and
 $f_{\mcP^{(2)}}$ in (\ref{rhoww_T1to3}), for each of our three regions (see General Theory).
@@
 To estimate the components $\rho_{ww}$, $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and 
-$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ in (\ref{rhoww_T1to3}), where $w$ was detrended kelp time series in one of our regions, 
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ in (\ref{rhoww_T1to3}), where $w$ was linearly detrended kelp time series in one of our regions, 
 $\epsilon^{(1)}$ was detrended nitrate time
@@
 Data for the project are publicly archived \citep{SBCLTER2021}.
-All computations were done in R version 3.4.4 on an ordinary laptop which was running Ubuntu Linux
-16.04. Complete codes for the entire project workflow are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.
+All computations were done in R version 3.4.4 on a laptop running Ubuntu Linux
+16.04. Complete codes for the project workflow are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.
 
@@
 %Para basically confirming earlier work, as a starter
-Helping fulfill goal G2 of the Introduction, our theoretical results 
+To begin fulfilling goal G2 of the Introduction, our theoretical case studies 
 demonstrate that interaction effects between Moran drivers: 1) can be 
-comparable in strength to direct Moran effects; 2) can be either synergistic or destructive, i.e., positive or negative; and 3) 
-can depend strongly on timescale, even having different signs on different timescales. 
-First, for all of our theoretical case studies, the magnitude of 
+comparable in strength to direct Moran effects; 2) can be either synergistic or destructive; and 3) 
+can depend strongly on timescale. 
+First, for all of our case studies, the magnitude of 
 interaction effects was comparable to that of direct Moran effects (Fig. \ref{fig:TheoryCaseA}a,e,i,m, 
-compare the dashed and solid lines). So interactions can contribute substantially to overall population 
-synchrony. Second, in contrast to the direct Moran effects of a 
-single environmental variable, which provide non-negative 
-contributions to population synchrony whenever, as is typical, environmental synchrony 
-is positive, interactions between Moran effects can be positive or negative. For instance, case study A
-showed negative interaction effects at short timescales (Fig. \ref{fig:TheoryCaseA}a); and 
-case study C showed negative interactions at
-long timescales (Fig. \ref{fig:TheoryCaseA}i,m), or, for some parameters, at all timescales 
+compare the dashed and solid lines). Thus interactions can contribute substantially to overall 
+synchrony. Second, in contrast to direct Moran effects, which are positive, 
+interactions can be negative or positive. CaA and CaB
+showed negative interaction effects on short timescales (Fig. \ref{fig:TheoryCaseA}a,e); 
+CaC1 showed negative interactions on
+long timescales (Fig. \ref{fig:TheoryCaseA}i,m); and CaC2 showed negative interactions on all timescales 
 (Fig. \ref{fig:TheoryCaseA}m).
-Thus interaction effects can either augment or reduce population synchrony from what its value would be 
-if only direct Moran effects operated. Finally, interactions between Moran effects depended strongly on timescale 
-for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and differed in sign across timescales for case studies
-A and B and for one of the C scenarios. This augments earlier studies demonstrating that direct Moran effects can 
-depend on timescale \citep{Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}. The points made in this paragraph
-also follow from the results of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not necessarily
-straightforwardly, and those papers are case studies, so our results provide novel theoretical generality.
-Our theory can also be interpreted to provide intuition about how interacting Moran effects work,
-as follows.
+Thus interaction effects can either augment or reduce synchrony. Finally, interaction effects depended strongly on timescale 
+for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and even differed in sign across timescales for CaA, CaB and CaC1. 
+This result complements earlier studies that showed direct Moran effects can 
+depend on timescale \citep{Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}. The results of this paragraph
+also follow from \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not 
+straightforwardly. And those papers are case studies, so our results provide novel theoretical generality.
 
-%Case study A
-Case study A provides an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
+%***DAN: stopped here
+
+%Case study A, new try
+CaA helps provide an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
 Moran effects on different timescales, and how our theory captures that contrast.
-For case study A, for which $\epsilon^{(1)}$ effects on populations were lagged by 1 quarter but 
-$\epsilon^{(2)}$ effects were unlagged (Methods), interactions between Moran effects were negative on the shortest timescales 
-(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}a, left side of panel). This is because the 
-population influences of Moran effects of timescale 2
-quarters and lagged by 1 quarter are exactly out of phase with unlagged Moran effects of timescale
-2 quarters, producing destructive interference. Thus, on short timescales, total
-synchrony will be less, for this example, than it would be if only direct Moran effects operated. In contrast, 
-interaction effects on long timescales were synergistic, i.e., positive, for case study A 
-(Fig. \ref{fig:TheoryCaseA}a, right side of panel) 
-because, on long timescales, a lag of 1 quarter does not meaningfully alter the phase relationship between 
-Moran drivers, and drivers are therefore almost in phase, reinforcing each other.
-At an intermediate timescale of 1 year, a lag of 1 quarter causes the Moran effects
-of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be a quarter phase out of alignment (1 quarter is $1/4$ of a 1-year 
-timescale), resulting in 
-zero interaction effects, essentially because sine waves a quarter-phase out of alignment
-with each other are perfectly uncorrelated. Thus interaction effects between Moran drivers were zero 
-at timescale 1 year (Fig. \ref{fig:TheoryCaseA}a). Theory captures these complexities because
-the size of the 1-quarter phase lag of this scenario, relative to a given timescale,
-is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) = 
+For CaA, $\epsilon^{(1)}$ effects on populations were lagged by $1$q but 
+$\epsilon^{(2)}$ effects were unlagged (Methods). So, in the language of Fig. \ref{fig:IntroTheory},
+$l_{e1}=1$q and $l_{e2}=0$q. 
+The between-noise lag, $l_n$, of Fig. \ref{fig:IntroTheory} is $0$q because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
+taken to be a white noise process (see SI section \ref{SM-sect:examples_cross_sync} for details, here).
+Thus, what determines the sign of interactions between Moran effects for this example is how 
+$l_{e1}-l_{e2}+l_n=1$q compares to the timescale/period, $\sigma$, being considered. 
+On the shortest timescales ($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}a, left side of panel), $1$q
+is half the period, so interactions effects are negative. On long timescales (e.g., $\sigma>8$y), $1$q is a negligible portion of the
+period, so $l_{e1}-l_{e2}+l_n=1$q is close to zero, relative to that period, and interaction effects
+are positive: compared to such long timescales, $\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects happen close to simultaneously, so
+the two noise variables reinforce each other. Theory captures these complexities because
+the size of the $1$q lag of this scenario, relative to a timescale, $\sigma$, is captured by
+the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, which equals the phase of
+$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ because $l_n=0$ and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number
+(SI section \ref{SM-sect:examples_cross_sync}). But $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=
+p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) = 
 p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}), 
-which has phase $-\pi$ at short timescales, $\sigma \approx 0.5 \text{y}$,
-and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b). Because 
-$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ has phase $0$ for all timescales (SI sections \ref{SM-sect:examples_cross_sync}, 
-\ref{SM-sect:details_case_studies}), the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ is the same as the phase of
-$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}c),
-which in turn controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, which is the
-interaction term in our theory (\ref{rhoww_T1to3}). Thus interactions shift 
-from negative, for short timescales, to positive, for long timescales. Comparing
+which has phase $-\pi$ at short timescales, $\sigma \approx 2 \text{q}$,
+and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b).
+The phase controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, the
+interaction term in (\ref{rhoww_T1to3}), as described in General Theory. Comparing
 Fig. \ref{fig:TheoryCaseA}a,d shows the additional influence of autoregressive population effects, which
 simply multiply all Moran influences by the same timescale-dependent non-negative 
-quantity without altering the relative importance or sign of the direct and indirect Moran effects [see (\ref{rhoww_T1to3})].
+quantity, not altering their relative importance or sign [see (\ref{rhoww_T1to3})].
 
-%Case syiudy B
-Case study B reveals how Moran effects can interact when lags are longer than 1 model time step. 
-For B, $\epsilon^{(1)}$ effects on populations were lagged by $n$ quarters but 
-$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$. Interactions between Moran effects were 
-negative on the shortest timescales 
-(0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, far left side of panel) because the 
-population influences of Moran effects of timescale 2
-quarters and lagged by 3 quarters are exactly out of phase with unlagged Moran effects of timescale
-2 quarters, producing negative interactions. 
-Interaction effects on long timescales were instead synergistic, i.e., positive, for B 
-(Fig. \ref{fig:TheoryCaseA}e, right side of panel), for the same reason as case study A (see above).
-At timescale 3 quarters, Moran effects are reinforcing because the lag equals the timescale,
-producing Moran effects of the two drivers which are in phase 
-(Fig. \ref{fig:TheoryCaseA}e, peak at timescale 3 quarters = 0.75 years). 
-At a timescale of 1 year, a lag of 3 quarters causes the Moran effects of 
-$\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be $3/4$-phase out of alignment, resulting in zero interaction effects. 
-Likewise at a timescale of 3 years (12 quarters), a lag of 3 quarters 
-makes the Moran effects of 
-$\epsilon^{(1)}$ and $\epsilon^{(2)}$ be $1/4$-phase out of alignment, again producing zero interaction. 
-Finally, at timescale 1.5 years (6 quarters), the 3-quarter lag equals half the timescale, so 
-$\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects counteract each other, producing negative 
-interactions. Thus interactions between Moran drivers were zero 
-at timescales 1 and 3 years for B, were 
-negative for short timescales and for timescale 1.5 years, and were positive for 
-timescale 0.75 years and for long timescales (Fig. \ref{fig:TheoryCaseA}e).
-Theory captures these complexities because 
-the size of the 3-quarter phase lag of this scenario, relative to a timescale,
-is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
-Because $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number for case study B (SI section \ref{SM-sect:details_case_studies}),
-the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ is the same as that of
-$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}g),
-which controls the sign of the interaction term,
-$2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, in our 
-theory (\ref{rhoww_T1to3}).
+% %Case study A, old try
+% CaA helps provide an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
+% Moran effects on different timescales, and how our theory captures that contrast.
+% For CaA, for which $\epsilon^{(1)}$ effects on populations were lagged by 1 quarter but 
+% $\epsilon^{(2)}$ effects were unlagged (Methods), interactions between Moran effects were negative on the shortest timescales 
+% (0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}a, left side of panel). This is because the 
+% population influences of Moran effects of timescale 2
+% quarters and lagged by 1 quarter are exactly out of phase with unlagged Moran effects of timescale
+% 2 quarters, producing destructive interference. Thus, on short timescales, total
+% synchrony will be less, for this example, than it would be if only direct Moran effects operated. In contrast, 
+% interaction effects on long timescales were synergistic, i.e., positive, for CaA 
+% (Fig. \ref{fig:TheoryCaseA}a, right side of panel) 
+% because, on long timescales, a lag of 1 quarter does not meaningfully alter the phase relationship between 
+% Moran drivers, and drivers are therefore almost in phase, reinforcing each other.
+% At an intermediate timescale of 1 year, a lag of 1 quarter causes the Moran effects
+% of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be a quarter phase out of alignment (1 quarter is $1/4$ of a 1-year 
+% timescale), resulting in 
+% zero interaction effects, essentially because sine waves a quarter-phase out of alignment
+% with each other are perfectly uncorrelated. Thus interaction effects between Moran drivers were zero 
+% at timescale 1 year (Fig. \ref{fig:TheoryCaseA}a). Theory captures these complexities because
+% the size of the 1-quarter phase lag of this scenario, relative to a given timescale,
+% is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) = 
+% p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}), 
+% which has phase $-\pi$ at short timescales, $\sigma \approx 0.5 \text{y}$,
+% and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b). Because 
+% $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ has phase $0$ for all timescales (SI sections \ref{SM-sect:examples_cross_sync}, 
+% \ref{SM-sect:details_case_studies}), the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ is the same as the phase of
+% $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}c),
+% which in turn controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, which is the
+% interaction term in our theory (\ref{rhoww_T1to3}). Thus interactions shift 
+% from negative, for short timescales, to positive, for long timescales. Comparing
+% Fig. \ref{fig:TheoryCaseA}a,d shows the additional influence of autoregressive population effects, which
+% simply multiply all Moran influences by the same timescale-dependent non-negative 
+% quantity without altering the relative importance or sign of the direct and indirect Moran effects [see (\ref{rhoww_T1to3})].
 
-%Case study C
-Cases A and B assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+%Case study B, new version
+CaB reveals how Moran effects can interact when lags are longer than 1 model time step. 
+For CaB, $\epsilon^{(1)}$ effects on populations were lagged by $n$q but 
+$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$, so $l_{e1}=3$q and $l_{e2}=0$q. 
+As for CaA, because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
+a white noise process for CaB, $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ was a real number and $l_n=0$.
+Interaction effects were again negative on short timescales 
+($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}e, far left side of panel)
+because $l_{e1}-l_{e2}+l_n=3$q was $1.5\sigma$ on that timescale, and so effects of 
+$\epsilon^{(1)}$ and $\epsilon^{(2)}$ were in a half-phase relationship and
+counteracted each other.
+Similar to CaA, $l_{e1}-l_{e2}+l_n=3$q was again negligible compared to long timescales, so interaction effects
+were positive on long timescales (Fig. \ref{fig:TheoryCaseA}e, right side of panel).
+The quantity $l_{e1}-l_{e2}+l_n=3$q exactly equaled the timescale for $\sigma=3 \text{q} = 0.75 \text{y}$ and 
+equaled half the timescale for $\sigma=6 \text{q} = 1.5 \text{y}$, hence interactions effects
+were positive for $\sigma=0.75 \text{y}$ and negative for $\sigma=1.5$y 
+(Fig. \ref{fig:TheoryCaseA}e). Theory captures these complexities because 
+the size of the $3$q relative lag of this scenario, relative to $\sigma$,
+is again captured by the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
+
+% %Case study B, old version
+% CaB reveals how Moran effects can interact when lags are longer than 1 model time step. 
+% For B, $\epsilon^{(1)}$ effects on populations were lagged by $n$ quarters but 
+% $\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$. Interactions between Moran effects were 
+% negative on the shortest timescales 
+% (0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, far left side of panel) because the 
+% population influences of Moran effects of timescale 2
+% quarters and lagged by 3 quarters are exactly out of phase with unlagged Moran effects of timescale
+% 2 quarters, producing negative interactions. 
+% Interaction effects on long timescales were instead synergistic, i.e., positive, for B 
+% (Fig. \ref{fig:TheoryCaseA}e, right side of panel), for the same reason as CaA (see above).
+% At timescale 3 quarters, Moran effects are reinforcing because the lag equals the timescale,
+% producing Moran effects of the two drivers which are in phase 
+% (Fig. \ref{fig:TheoryCaseA}e, peak at timescale 3 quarters = 0.75 years). 
+% At a timescale of 1 year, a lag of 3 quarters causes the Moran effects of 
+% $\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be $3/4$-phase out of alignment, resulting in zero interaction effects. 
+% Likewise at a timescale of 3 years (12 quarters), a lag of 3 quarters 
+% makes the Moran effects of 
+% $\epsilon^{(1)}$ and $\epsilon^{(2)}$ be $1/4$-phase out of alignment, again producing zero interaction. 
+% Finally, at timescale 1.5 years (6 quarters), the 3-quarter lag equals half the timescale, so 
+% $\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects counteract each other, producing negative 
+% interactions. Thus interactions between Moran drivers were zero 
+% at timescales 1 and 3 years for B, were 
+% negative for short timescales and for timescale 1.5 years, and were positive for 
+% timescale 0.75 years and for long timescales (Fig. \ref{fig:TheoryCaseA}e).
+% Theory captures these complexities because 
+% the size of the 3-quarter phase lag of this scenario, relative to a timescale,
+% is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
+% Because $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number for CaB (SI section \ref{SM-sect:details_case_studies}),
+% the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ is the same as that of
+% $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}g),
+% which controls the sign of the interaction term,
+% $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, in our 
+% theory (\ref{rhoww_T1to3}).
+
+%***DAN: stopped here
+
+%***DAN: Don't forget to appropriately use q in place of quarter from here on as you edit. If you forget, do a 
+%search and replace.
+
+%Case study C, new
+CaA and CaB assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+and $l_n=0$ (SI sections \ref{SM-sect:examples_cross_sync}).
+But CaC illustrates what can happen when noise processes have periodic components, e.g., 
+both processes oscillate with annual periodicity, an important scenario because seasonality is common.
+For CaC, $\epsilon^{(1)}$ effects on populations were lagged by $n_1$q and 
+$\epsilon^{(2)}$ effects were lagged by $n_2$q (Methods). We used $n_1=1$, $n_2=0$, $p_1^{(1)}=3$ and $p_0^{(2)}=-1.5$, so 
+$\epsilon^{(1)}$ effects were lagged and positive, and $\epsilon^{(2)}$ effects were immediate and negative
+(Fig. \ref{fig:TheoryCaseA} caption). The lagged positive effects of $\epsilon^{(1)}$ correspond to 
+$l_{e1}=1$q, in the language of Fig. \ref{fig:IntroTheory}. But, for that figure, we assumed environmental 
+variables had positive effects on populations, whereas for CaC $\epsilon^{(2)}$ had negative effects.
+Immediate negative effects of an annually oscillating variable are the same as positive effects lagged 
+by $2$q, so we can take $l_{e2}=2$q for CaC.  
+%***DAN: pretty confused, come back to this
+
+%Case study C, old
+CaA and CaB assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
 (SI sections \ref{SM-sect:examples_cross_sync}, \ref{SM-sect:details_case_studies}),
-but case study C illustrates what can happen when noise processes have periodic components, e.g., 
+but CaC illustrates what can happen when noise processes have periodic components, e.g., 
 both processes oscillate with annual periodicity, an important scenario because seasonality is common. 
@@
 between the lags of their effects on populations, as we now explain. 
-For the scenarios we considered under case study C (Fig. \ref{fig:TheoryCaseA}i-p), 
+For the scenarios we considered under CaC (Fig. \ref{fig:TheoryCaseA}i-p), 
 $\epsilon^{(1)}$ effects on populations were positive, and were 
@@
 $\epsilon^{(2)}$ effects were immediate and negative (Methods and Fig. \ref{fig:TheoryCaseA} caption).
-For scenario 1 of C (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
+For CaC1 (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
 noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
@@
 synchrony on annual timescales than would have been expected if Moran effects operated independently. Contrastingly,
-for scenario 2 of case study C (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
+for CaC2 (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
 noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
@@
 and less population synchrony, on annual timescales, than would have been expected if Moran effects operated independently 
-(Fig. \ref{fig:TheoryCaseA}m). The only difference between scenarios 1 and 2 
+(Fig. \ref{fig:TheoryCaseA}m). The only difference between CaC1 and CaC2 
 is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
@@
 
+%Direct Moran effects
 The way direct Moran effects work in our theory is fairly straightforward. 
@@
 effects on synchrony can be assessed across timescales and regions. 
-Both nitrates and waves fluctuate seasonally in CA. Thus nitrate and wave 
+Both nitrates and waves fluctuate seasonally in CA (REFS). 
+%***DAN: Fugure out a ref for this later
+Thus nitrate and wave 
 synchrony had a strong annual component (Fig. \ref{fig:directMoran}b,d,q,s), which produced 
@@
 Nitrates and waves are also synchronous on long timescales $>8$y in CA 
-(Fig. \ref{fig:directMoran}l,n,aa,cc), possibly due to the North Pacific Gyre Oscillation [NPGO; \cite{CastoraniSheppardKelp}]. The long-timescale synchrony
+(Fig. \ref{fig:directMoran}l,n,aa,cc), possibly due to the North Pacific Gyre Oscillation 
+[NPGO; \cite{CastoraniSheppardKelp}]. The long-timescale synchrony
 in nitrates and waves produced some of the long-timescale
@@
 and $f_{\mcB}$ (see Methods) are in Table \ref{SM-tab:regcoefs}.
+
+%Interactions between Moran effects, CCA1
+Interactions between Moran effects in CCA1 were parallel to scenario 1 of theoretical case study C.
+In CCA1, nitrates have a one-quarter-delayed positive effect on observed 
+kelp populations and waves have immediate negative effect (REFS), 
+%***DAN: ask Tom/Kyle/Max what is a good reference for this
+just as $\epsilon^{(1)}$ effects in case study C, scenario 1, 
+were positive and delayed by a quarter and $\epsilon^{(2)}$ effects were negative and immediate.
+Delayed effects are reflected, for CCA1, in Fig. \ref{fig:indirectMoran}a, and for 
+case study C scenario 1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
+that $f_{\mcP^{(1)}}$ has phase about $-\pi/2$ at the annual timescale. The delay is also 
+reflected, for CCA1, in Table \ref{SM-tab:regcoefs}, where 
+the CCA1 entries for $p_0^{(1)}$ were close to $0$ and those for $p_1^{(1)}$ were positive.
+Immediate negative effects of waves in CCA1 are reflected in Fig. \ref{fig:indirectMoran}b,
+and for case study C scenario 1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
+that $f_{\mcP^{(2)}}$ has phase $-\pi$ at all timescales. See also the CCA1 $p_0^{(2)}$
+coefficients in Table \ref{SM-tab:regcoefs}, which are negative.
+In CCA1, annual peak nitrates tend to come in spring, whereas annual peak waves 
+tend to come in winter (REFS). 
+%***DAN: add refs later
+Thus annual nitrate peaks tend to lag annual 
+wave peaks by one quarter, just as $\epsilon^{(1)}$ peaks lagged $\epsilon^{(2)}$
+peaks by a quarter in theoretical case study C, scenario 1.
+This is reflected, for CCA1, in Fig. \ref{fig:indirectMoran}d, and for case study C scenario 1 
+in Fig. \ref{fig:TheoryCaseA}k, both of which show that $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$
+has a phase of about $-\pi/2$ at the annual timescale.
+Thus lags compound on the annual timescale, in the CCA1 case as in the theoretical case: the annual positive 
+effects of nitrates on kelp tend to come in the summer, and the annual negative effects of waves come in the winter,
+producing reinforcing interactions between Moran effects. Likewise, the $\pi/2$ phase 
+of $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:indirectMoran}c for CCA1, Fig. \ref{fig:TheoryCaseA}k for case study C scenario 1)
+approximately cancels the $-\pi/2$ phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ 
+(Fig. \ref{fig:indirectMoran}d for CCA1, Fig. \ref{fig:TheoryCaseA}k for case study C scenario 1) on annual timescales, so that  
+$f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ has phase 
+approximately $0$, and the interaction-effect term 
+$\Rp(f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}})$ is strongly positive 
+on annual timescales in both the theoretical and emprical cases. 
+%***DAN: read the below sentence again and see if it sucks
+Interactions between Moran effects on long-timescales ($>8$y) were also similar for both CCA1 and
+for case study C, scenario 1, and for the same reason in both cases: on long timescales, sub-annual lags
+are inconsequential, and the positive effects of nitrates/$\epsilon^{(1)}$ and the negative
+effects of waves/$\epsilon^{(2)}$ therefore lead to negative interactions.
+Interactions between Moran effects in CCA2 operated in a similar manner to CCA1 (Fig. \ref{SM-fig:indirectMoran_SM_lag4}).
+
+%Interactions between Moran effects, SB
 
@@
 \caption{Example and basic concept of interacting Moran effects. The example (a), which 
-is based on a simple model described in the Introduction, shows that 
+is based on the simple model described in the Introduction, shows that 
 interaction effects are possible, and can be substantially positive or negative. 
@@
 and the values $\cov(\epsilon_i^{(1)},\epsilon_j^{(2)})$, for $i,j=1,2$, were all set equal to each
-other and to a parameter, $\phi$. For (b) and (c), which represent distinct scenarios, 
+other and to a parameter, $\phi$. For (b) and (c), which present the basic concept of interaction 
+effects for two distinct scenarios, 
 the top sinusoid on each panel represents 
-the component of oscillation at a given timescale of an environmental variable, $\epsilon_i^{(1)}$,
+the component of oscillation at timescale 20 time steps of an environmental variable, $\epsilon_i^{(1)}$,
 measured in a sampling location $i$, and the bottom sinusoid represents the 
@@
 $l_{e2}$. For this setup, a positive interaction between Moran effects can occur on the
-given timescale when lags align, i.e, $l_{e1}=l_{e2}+l_n$, as in panel (a), so that environmental
+given timescale when lags align, i.e, $l_{e1}-l_{e2}+l_n=0$, as in panel (a), so that environmental
 effects on the population coincide and reinforce each other. All environmental effects on 
@@
 effect can occur when lags produce out-of-phase influence of environmental variables on
-populations, i.e., $l_{e1}$ differs from $l_{e2}+l_n$ by half the given timescale/period,
+populations, e.g., $l_{e1}-l_{e2}+l_n$ equals half the given timescale/period,
 as in panel (b), so that the distinct influences tend to work against each other in 
@@
 population synchrony due to the direct Moran effects of $\epsilon^{(2)}$.
-The dashed green-blue line is 
+The dashed line is 
 $\frac{2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)}{|f_{\mcB}|^2}$, 
-which is population synchrony due to interacting Moran effects.
-Middle panels show the constituent parts of the theory (labelled). For these complex functions 
-of timescale, the magnitude is displayed on the vertical axis and the phase is displayed with color.
-See text for all interpretations.
+quantifying population synchrony due to interacting Moran effects.
+Middle panels show the constituent parts of the theory (labelled). These constituents are complex-valued, 
+with magnitude on the vertical axis and phase displayed using color.
+See text for interpretations.
 The functions plotted on d, h, l, p are those on a, e, i, m, respectively, times 
-$|f_{\mcB}|^2$, thereby showing the influence of autoregressive population effects on
+$|f_{\mcB}|^2$, plotted to illustrate how autoregressive population effects modulate
 synchrony [see (\ref{rhoww_T1to3})].

### 2022-01-16 Paper/SupMat.Rnw
continuting to write and edit

@@
 quantities given previously. For theoretical case studies A and B in the main text, we used
-$N=\Sexpr{N_A}$, $v_{11}=v{22}=\Sexpr{ns_A[1]}$ and $v_{12}=\Sexpr{ns_A[3]}$.
+$N=\Sexpr{N_A}$, $v_{11}=v_{22}=\Sexpr{ns_A[1]}$ and $v_{12}=\Sexpr{ns_A[3]}$.
 
@@
 
-\section{Details of the theoretical case studies}\label{sect:details_case_studies}
-
-The noise processes $(\epsilon^{(1)},\epsilon^{(2)})$ used 
-in case studies A, B and C, which were only partly specified in the main text, are fully specified in 
-SI section \ref{sect:examples_cross_sync}. 
-We next derive, for our case studies, the values of the terms that appear in (\ref{MT-rhoww_T1to3}) of the main text.
-For case study A, parameter specifications presented in Methods of the main text straightforwardly imply that
-$f_{\mcB}=1-b_1 \mu$, $f_{\mcP^{(1)}}=p_1^{(1)} \mu$ and $f_{\mcP^{(2)}}=p_0^{(2)}$. In 
-SI section \ref{sect:examples_cross_sync}, we computed that 
-$\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
-$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$, and 
-$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$.
-
-For case study B, details presented in Methods of the main text imply that 
-$f_{\mcB}=1-b_1 \mu$, $f_{\mcP^{(1)}}=p_n^{(1)} \mu^n$ and $f_{\mcP^{(2)}}=p_0^{(2)}$. 
-Because the noise process for B is taken to be the same as that for A, we again have 
-$\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
-$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$ and 
-$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$.
-
-For case study C, again using details from Methods, $f_{\mcB}=1-b_1 \mu$, 
-$f_{\mcP^{(1)}}=p_{n_1}^{(1)} \mu^{n_1}$ and $f_{\mcP^{(2)}}=p_{n_2}^{(2)} \mu^{n_2}$. 
-In SI section \ref{sect:examples_cross_sync}, we computed that
-$\rho_{\epsilon^{(1)} \epsilon^{(1)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{11}$,
-$\rho_{\epsilon^{(2)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{22}$ and 
-$\rho_{\epsilon^{(1)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{12} \exp(-2 \pi \iota f (l_1-l_2))$.
-
 \section{Details of data and data cleaning}\label{sect:data}
@@
 imagery of $30$m resolution was used. Quarterly mean biomass values and spatial extent of canopy  
-within $500$m coastline segments (henceforth called ``loactions'')
+within $500$m coastline segments (henceforth called ``locations'')
 were computed by averaging. Quarters were: Q1, January-March; Q2, April-June; 
@@
 Wave and nutrient time series in each location were again linearly detrended before use.
+
+\section{Details of the theoretical case studies}\label{sect:details_case_studies}
+
+The noise processes $(\epsilon^{(1)},\epsilon^{(2)})$ used 
+in case studies A, B and C, which were only partly specified in the main text, are fully specified in 
+SI section \ref{sect:examples_cross_sync}. 
+We next derive, for our case studies, the values of the terms that appear in (\ref{MT-rhoww_T1to3}) of the main text.
+For case study A, parameter specifications presented in Methods of the main text straightforwardly imply that
+$f_{\mcB}=1-b_1 \mu$, $f_{\mcP^{(1)}}=p_1^{(1)} \mu$ and $f_{\mcP^{(2)}}=p_0^{(2)}$. In 
+SI section \ref{sect:examples_cross_sync}, we computed that 
+$\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
+$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$, and 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$.
+
+For case study B, details presented in Methods of the main text imply that 
+$f_{\mcB}=1-b_1 \mu$, $f_{\mcP^{(1)}}=p_n^{(1)} \mu^n$ and $f_{\mcP^{(2)}}=p_0^{(2)}$. 
+Because the noise process for B is taken to be the same as that for A, we again have 
+$\rho_{\epsilon^{(1)} \epsilon^{(1)}}=v_{11}$,
+$\rho_{\epsilon^{(2)} \epsilon^{(2)}}=v_{22}$ and 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}=v_{12}$.
+
+For case study C, again using details from Methods, $f_{\mcB}=1-b_1 \mu$, 
+$f_{\mcP^{(1)}}=p_{n_1}^{(1)} \mu^{n_1}$ and $f_{\mcP^{(2)}}=p_{n_2}^{(2)} \mu^{n_2}$. 
+In SI section \ref{sect:examples_cross_sync}, we computed that
+$\rho_{\epsilon^{(1)} \epsilon^{(1)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{11}$,
+$\rho_{\epsilon^{(2)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{22}$ and 
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{12} \exp(-2 \pi \iota f (l_1-l_2))$.
 

### 2022-01-16 Code/IntroTheoryIA.R

@@ -13,7 +13,7 @@
 #
 #Note
 #Returns NA for sync if covmat is not positive definite, returns NA for sync.in if the analogous
-#covariance matrix for is is not positive definite.
+#covariance matrix for it is not positive definite.
 #
 IntroTheoryIA<-function(a,b,covmat)
 {

### 2022-01-16 Code/Theory_PedagogFig.R

@@ -45,7 +45,7 @@ par(fig=c((gap)/totwd,
 tim<-seq(from=0,to=100,length.out=10001)
 period<-20
 phi1<-(-1.5)
-phi2<-phi1+pi/2
+phi2<-phi1-pi/2
 sin1<-sin(2*pi*tim/period+phi1)
 sin2<-sin(2*pi*tim/period+phi2)
 offset1<-0
@@ -74,21 +74,21 @@ lines(rep((pi/2-phi1)*period/2/pi,2),c(1,3.8),type="l",lty="dashed")
 text(mean(c((pi/2-phi2)*period/2/pi,(pi/2-phi1)*period/2/pi)),3.5,latex2exp::TeX("$l_n$"),adj=c(0.5,0.5),cex=0.8)
 
 #illustrate the delay of env effects on population
-mutdel<-7
+mutdel<-2
 lines(rep((pi/2-phi2)*period/2/pi+3*period,2),c(6,3.5),type="l",lty="solid")
 lines(c((pi/2-phi2)*period/2/pi+3*period,(pi/2-phi2)*period/2/pi+3*period+period/4+mutdel),rep(3.5,2))
 shape::Arrows((pi/2-phi2)*period/2/pi+3*period+period/4+mutdel,3.5,
               (pi/2-phi2)*period/2/pi+3*period+period/4+mutdel,ctr+.2,
               arr.length=.2)
-text(.75*((pi/2-phi2)*period/2/pi+3*period+.75)+.25*((pi/2-phi2)*period/2/pi+3*period+period/4+mutdel),
+text(.55*((pi/2-phi2)*period/2/pi+3*period+.75)+.45*((pi/2-phi2)*period/2/pi+3*period+period/4+mutdel),
       3.5,latex2exp::TeX("$l_{e1}$"),cex=0.8,adj=c(0.5,-.2))
 
 lines(rep((pi/2-phi1)*period/2/pi+3*period,2),c(1,1.25),type="l",lty="solid")
-lines(c((pi/2-phi1)*period/2/pi+3*period,(pi/2-phi1)*period/2/pi+3*period+mutdel),rep(1.25,2))
-shape::Arrows((pi/2-phi1)*period/2/pi+3*period+mutdel,1.25,
-              (pi/2-phi1)*period/2/pi+3*period+mutdel,ctr-.2,
[…]

### 2022-01-16 Code/Kelp_AnalysisStep_MakeMapFig.R
new script stub for Max to add code to to make a map fig for the kelp data

@@ -1,0 +1,122 @@
+#Makes a map figure for the paper that orients the reader to the kelp data we used
+
+#***
+#External codes needed
+#***
+
+#packages needed (invoked with "::"): 
+
+#***
+#Location for storing results and other prep
+#***
+
+resloc<-"../Results/Kelp_MapFig/"
+if (!dir.exists(resloc))
+{
+  dir.create(resloc,recursive=TRUE)
+}
+
+#***
+#Load the data
+#***
+
+#load the data
+datloc<-"../Results/Kelp_DataAfterAllCleaning/"
+kelp<-readRDS(paste0(datloc,"Kelp_Quarterly_CleanedFinal.Rds")) 
+NO3<-readRDS(paste0(datloc,"NO3_Quarterly_CleanedFinal.Rds")) 
+waves<-readRDS(paste0(datloc,"Waves_Quarterly_CleanedFinal.Rds")) 
+locs<-readRDS(paste0(datloc,"Locs_CleanedFinal.Rds"))
+quarters<-readRDS(paste0(datloc,"Quarters_CleanedFinal.Rds"))
+climinds<-readRDS(paste0(datloc,"Climinds_Quarterly_CleanedFinal.Rds"))
+times<-1:(dim(quarters)[1])
+
+numts<-dim(kelp)[1]
+lents<-dim(kelp)[2]
+
+#***
+#Which site is closest to pt conception, same for SB and some other locations, for later use
+#***
+
+#from google maps
+MontBLatLon<-c(36.62393626182818, -121.8511283749223) #Monterey Bay, closer to the southern end
+CarBLatLon<-c(36.52316357182972, -121.95535513873469) #The point to the south of Carmel Bay
+MorBLatLon<-c(35.427553615598676, -120.89133343717805) #Morro Bay, closer to the northern end
+PtConcLatLon<-c(34.44830424836278,-120.47125509005721) #Pt Conception
+SBLatLon<-c(34.407876790966135, -119.68428533917637) #Santa Barbara pier
+OxLatLon<-c(34.2310543912163, -119.2726684255394) #Oxnard, where the Santa Clara River empties
+LALatLon<-c(34.00907538429378, -118.50422514984244) #LA, near the Santa Monica pier
+
[…]

### 2022-01-16 Code/Kelp_AnalysisStep_SpectralDecomp.R

@@ -47,77 +47,13 @@ numts<-dim(kelp)[1]
 lents<-dim(kelp)[2]
 
 #***
-#Which site is closest to pt conception, same for SB and some other locations, for later use
+#Load the indices of the regions we are using
 #***
 
-#from google maps
-MontBLatLon<-c(36.62393626182818, -121.8511283749223) #Monterey Bay, closer to the southern end
-CarBLatLon<-c(36.52316357182972, -121.95535513873469) #The point to the south of Carmel Bay
-MorBLatLon<-c(35.427553615598676, -120.89133343717805) #Morro Bay, closer to the northern end
-PtConcLatLon<-c(34.44830424836278,-120.47125509005721) #Pt Conception
-SBLatLon<-c(34.407876790966135, -119.68428533917637) #Santa Barbara pier
-OxLatLon<-c(34.2310543912163, -119.2726684255394) #Oxnard, where the Santa Clara River empties
-LALatLon<-c(34.00907538429378, -118.50422514984244) #LA, near the Santa Monica pier
-
-dists<-ncf::gcdist(x=c(MontBLatLon[2],locs$Lon),y=c(MontBLatLon[1],locs$Lat))
-dists<-dists[1,2:(dim(dists)[2])]
-MontBInd<-which(dists==min(dists))
-
-dists<-ncf::gcdist(x=c(CarBLatLon[2],locs$Lon),y=c(CarBLatLon[1],locs$Lat))
-dists<-dists[1,2:(dim(dists)[2])]
-CarBInd<-which(dists==min(dists))
-
-dists<-ncf::gcdist(x=c(MorBLatLon[2],locs$Lon),y=c(MorBLatLon[1],locs$Lat))
-dists<-dists[1,2:(dim(dists)[2])]
-MorBInd<-which(dists==min(dists))
-
-dists<-ncf::gcdist(x=c(PtConcLatLon[2],locs$Lon),y=c(PtConcLatLon[1],locs$Lat))
-dists<-dists[1,2:(dim(dists)[2])]
-PtConcInd<-which(dists==min(dists))
-
-dists<-ncf::gcdist(x=c(SBLatLon[2],locs$Lon),y=c(SBLatLon[1],locs$Lat))
-dists<-dists[1,2:(dim(dists)[2])]
-SBInd<-which(dists==min(dists))
-
[…]

### 2022-01-16 Code/main.R

@@ -22,6 +22,10 @@ source("Kelp_AnalysisStep_PlotTimeSeries.R")
 #Final cleaning
 source("Kelp_AnalysisStep_CleanData_Final.R")
 
+#Make the map fig, to help orient the reader of the paper to the dataset
+source("Kelp_AnalysisStep_MakeMapFig.R")
+
 #Do the spectra decomposition analysis for kelp
 source("Kelp_AnalysisStep_SpectralDecomp.R")
+
 

### 2022-01-16 Code/Kelp_AnalysisStep_CleanData_SuperBasic.R
commented some codes which are no longer necessary to speed up overall run times

@@ -155,15 +155,15 @@ rm(paddle)
 #check visually to make sure it is paddle order
 #***
 
-pl<-.1
-plot(locs$Lon[1],locs$Lat[1],type="p",pch=20,cex=.1,ylim=range(locs$Lat),xlim=range(locs$Lon))
-profvis::pause(pl)
-for (counter in 2:(dim(locs)[1]))
-{
-  points(locs$Lon[counter],locs$Lat[counter],type="p",pch=20,cex=.1)
-  profvis::pause(pl)
-}
-#paddling order visually confirmed
+# pl<-.1
+# plot(locs$Lon[1],locs$Lat[1],type="p",pch=20,cex=.1,ylim=range(locs$Lat),xlim=range(locs$Lon))
+# profvis::pause(pl)
+# for (counter in 2:(dim(locs)[1]))
+# {
+#   points(locs$Lon[counter],locs$Lat[counter],type="p",pch=20,cex=.1)
+#   profvis::pause(pl)
+# }
+# #paddling order visually confirmed
 
 #***
 #Do some early-stage cleaning

### 2022-01-16 Code/main.R

@@ -17,7 +17,7 @@ source("Kelp_AnalysisStep_CleanData_SuperBasic.R")
 source("Kelp_AnalysisStep_CleanData_MoreInvolved.R")
 
 #Plot kelp time series for visual inspection
-source("Kelp_AnalysisStep_PlotTimeSeries.R")
+#source("Kelp_AnalysisStep_PlotTimeSeries.R")
 
 #Final cleaning
 source("Kelp_AnalysisStep_CleanData_Final.R")

### 2022-01-17 Paper/Paper.Rnw
thorough edit of theoretical case study results now complete

@@
 %where you currently cite the figures and tables in it. My current hunch is that the first of those options is
-%better.
+%better. Have to decide what to do with Fig. S1 in that case.
 
@@
 
-%***DAN: stopped here
-
 %Case study A, new try
@@
 which has phase $-\pi$ at short timescales, $\sigma \approx 2 \text{q}$,
-and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b).
+and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b,c).
 The phase controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, the
@@
 (Fig. \ref{fig:TheoryCaseA}e). Theory captures these complexities because 
-the size of the $3$q relative lag of this scenario, relative to $\sigma$,
-is again captured by the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
+the size of the $3$q lag of this scenario, relative to $\sigma$,
+is again captured by the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f,g).
 
@@
 
-%***DAN: stopped here
+%Case study C, new
+CaA and CaB assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+and $l_n=0$.
+But CaC illustrates what can happen when noise processes have periodic components and lagged interdependences, e.g., 
+both processes oscillate with annual periodicity and distinct phenology, an important scenario because seasonality is common.
+For CaC, $\epsilon^{(1)}$ effects on populations were lagged by $n_1$q and 
+$\epsilon^{(2)}$ effects were lagged by $n_2$q (Methods). We used $n_1=1$, $n_2=0$, $p_1^{(1)}=3$ and $p_0^{(2)}=-1.5$, so 
+$\epsilon^{(1)}$ effects were lagged and positive, and $\epsilon^{(2)}$ effects were immediate and negative. The language and logic of 
+Fig. \ref{fig:IntroTheory} does not directly apply, because we there assumed, in order to introduce concepts simply,
+that environmental effects were positive, and we now take $p_0^{(2)}=-1.5$. But similar logic does apply, with
+both the current and Fig. \ref{fig:IntroTheory} paradigms covered by our general theory.
+For CaC1 (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
+noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
+$1$q (Methods, SI section \ref{SM-sect:examples_cross_sync}). Thus lags compounded: the annual positive population effects of $\epsilon^{(1)}$ 
+were exactly $2$q offset from the annual negative effects of $\epsilon^{(2)}$, 
+producing reinforcing interactions 
+between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i).
+Contrastingly, for CaC2 (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
+noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
+$1$q. Thus lags cancelled: the annual positive population effects of $\epsilon^{(1)}$
+coincided with the negative effects of $\epsilon^{(2)}$ every year, producing 
+negative interactions between Moran influences (Fig. \ref{fig:TheoryCaseA}m).
+Our theory captures these complexities
+because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
+quantifies the degree of alignment between the effects of $\epsilon^{(1)}$ and 
+$\epsilon^{(2)}$ compared to a given timescale, while 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the degree of aslignment between 
+oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
+The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+therefore captures whether delays in effects and lags between 
+variables align or misalign, with alignment
+reflected in the real part of this quantity, $\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$,
+which is half the interaction term in our theory (\ref{rhoww_T1to3}).
+On long timescales, interactions were the
+same in the two scenarios, and were negative, because, on those timescales, sub-annual lags 
+make negligible difference, and the effects of $\epsilon^{(1)}$ on populations were 
+positive and those of $\epsilon^{(2)}$ were negative.
+
+% %Case study C, old
+% CaA and CaB assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+% (SI sections \ref{SM-sect:examples_cross_sync}, \ref{SM-sect:details_case_studies}),
+% but CaC illustrates what can happen when noise processes have periodic components, e.g., 
+% both processes oscillate with annual periodicity, an important scenario because seasonality is common. 
+% What matters then is whether and how the lag between the two seasonal noise processes aligns with the difference 
+% between the lags of their effects on populations, as we now explain. 
+% For the scenarios we considered under CaC (Fig. \ref{fig:TheoryCaseA}i-p), 
+% $\epsilon^{(1)}$ effects on populations were positive, and were 
+% lagged (i.e., delayed in their influence) by 1 quarter; but 
+% $\epsilon^{(2)}$ effects were immediate and negative (Methods and Fig. \ref{fig:TheoryCaseA} caption).
+% For CaC1 (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
+% noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
+% 1 quarter. Thus lags compounded: the annual positive population effects of $\epsilon^{(1)}$ 
+% were exactly 2 quarters offset from the annual negative effects of $\epsilon^{(2)}$, 
+% producing reinforcing interactions 
+% between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i), and more population
+% synchrony on annual timescales than would have been expected if Moran effects operated independently. Contrastingly,
+% for CaC2 (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
+% noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
+% 1 quarter. Thus lags cancelled: the annual positive population effects of $\epsilon^{(1)}$
+% coincided with the negative effects of $\epsilon^{(2)}$ every year, a half-phase shift for an 
+% annually oscillating effect, which corresponds to negative interactions between Moran influences
+% and less population synchrony, on annual timescales, than would have been expected if Moran effects operated independently 
+% (Fig. \ref{fig:TheoryCaseA}m). The only difference between CaC1 and CaC2 
+% is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
+% Our theory captures these complexities
+% because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
+% quantifies the relative phase lag between the effects of $\epsilon^{(1)}$ and 
+% $\epsilon^{(2)}$ compared to a given timescale, while 
+% $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the phase lag between 
+% oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
+% The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+% therefore captures whether relative lags of effects and lags between 
+% the variables themselves align or misalign, with the degree of alignment
+% reflected in the real part of this quantity, $\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$,
+% which is half the interaction term in our theory (\ref{rhoww_T1to3}).
+% On long timescales, interactions were the
+% same in the two scenarios because, on those timescales, sub-annual lags 
+% make negligible difference. Interactions were then negative 
+% because the effects of $\epsilon^{(1)}$ on populations were 
+% positive and those of $\epsilon^{(2)}$ were negative.
 
@@
 %search and replace.
-
-%Case study C, new
-CaA and CaB assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-and $l_n=0$ (SI sections \ref{SM-sect:examples_cross_sync}).
-But CaC illustrates what can happen when noise processes have periodic components, e.g., 
-both processes oscillate with annual periodicity, an important scenario because seasonality is common.
-For CaC, $\epsilon^{(1)}$ effects on populations were lagged by $n_1$q and 
-$\epsilon^{(2)}$ effects were lagged by $n_2$q (Methods). We used $n_1=1$, $n_2=0$, $p_1^{(1)}=3$ and $p_0^{(2)}=-1.5$, so 
-$\epsilon^{(1)}$ effects were lagged and positive, and $\epsilon^{(2)}$ effects were immediate and negative
-(Fig. \ref{fig:TheoryCaseA} caption). The lagged positive effects of $\epsilon^{(1)}$ correspond to 
-$l_{e1}=1$q, in the language of Fig. \ref{fig:IntroTheory}. But, for that figure, we assumed environmental 
-variables had positive effects on populations, whereas for CaC $\epsilon^{(2)}$ had negative effects.
-Immediate negative effects of an annually oscillating variable are the same as positive effects lagged 
-by $2$q, so we can take $l_{e2}=2$q for CaC.  
-%***DAN: pretty confused, come back to this
-
-%Case study C, old
-CaA and CaB assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-(SI sections \ref{SM-sect:examples_cross_sync}, \ref{SM-sect:details_case_studies}),
-but CaC illustrates what can happen when noise processes have periodic components, e.g., 
-both processes oscillate with annual periodicity, an important scenario because seasonality is common. 
-What matters then is whether and how the lag between the two seasonal noise processes aligns with the difference 
-between the lags of their effects on populations, as we now explain. 
-For the scenarios we considered under CaC (Fig. \ref{fig:TheoryCaseA}i-p), 
-$\epsilon^{(1)}$ effects on populations were positive, and were 
-lagged (i.e., delayed in their influence) by 1 quarter; but 
-$\epsilon^{(2)}$ effects were immediate and negative (Methods and Fig. \ref{fig:TheoryCaseA} caption).
-For CaC1 (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
-noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
-1 quarter. Thus lags compounded: the annual positive population effects of $\epsilon^{(1)}$ 
-were exactly 2 quarters offset from the annual negative effects of $\epsilon^{(2)}$, 
-producing reinforcing interactions 
-between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i), and more population
-synchrony on annual timescales than would have been expected if Moran effects operated independently. Contrastingly,
-for CaC2 (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
-noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
-1 quarter. Thus lags cancelled: the annual positive population effects of $\epsilon^{(1)}$
-coincided with the negative effects of $\epsilon^{(2)}$ every year, a half-phase shift for an 
-annually oscillating effect, which corresponds to negative interactions between Moran influences
-and less population synchrony, on annual timescales, than would have been expected if Moran effects operated independently 
-(Fig. \ref{fig:TheoryCaseA}m). The only difference between CaC1 and CaC2 
-is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
-Our theory captures these complexities
-because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
-quantifies the relative phase lag between the effects of $\epsilon^{(1)}$ and 
-$\epsilon^{(2)}$ compared to a given timescale, while 
-$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the phase lag between 
-oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
-The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-therefore captures whether relative lags of effects and lags between 
-the variables themselves align or misalign, with the degree of alignment
-reflected in the real part of this quantity, $\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$,
-which is half the interaction term in our theory (\ref{rhoww_T1to3}).
-On long timescales, interactions were the
-same in the two scenarios because, on those timescales, sub-annual lags 
-make negligible difference. Interactions were then negative 
-because the effects of $\epsilon^{(1)}$ on populations were 
-positive and those of $\epsilon^{(2)}$ were negative.
 

### 2022-01-18 Paper/Paper.Rnw
finished writing draft except for Discussion and Abstract

@@
 
-%***DAN: Don't forget to appropriately use q in place of quarter from here on as you edit. If you forget, do a 
-%search and replace.
-
 %Paralels the first para, but for kelp
@@
 the earlier theoretical results, based on our case studies 
-(see above text and Fig. \ref{fig:TheoryCaseA}a,e,i,m),
+(above text and Fig. \ref{fig:TheoryCaseA}a,e,i,m),
 that interaction effects between Moran drivers: 1) can be comparble in strength to direct Moran
 effects; 2) can be either synergistic or destructive; and 3) can 
-depend strongly on timescale, even having different signs on different timescales.
-First, for all of our regions and for essentially all timescales, the magintude of interaction 
+depend strongly on timescale.
+First, for all of our regions and for essentially all timescales, the magnitude of interaction 
 effects was comparable to that of direct Moran effects (Fig. \ref{fig:SummaryKelpFig}
@@
 \ref{fig:SummaryKelpFig}a), or negative (e.g., annual timescales, SB region on Fig. 
-\ref{fig:SummaryKelpFig}d, or long timescales $>8$y for either region, Fig. 
-\ref{fig:SummaryKelpFig}c,f). Finally, interactions depended on timescale, e.g., for the CCA1 
-region they were positive on annual timescales (Fig. \ref{fig:SummaryKelpFig}a) and
+\ref{fig:SummaryKelpFig}d; or long timescales $>8$y for either region, Fig. 
+\ref{fig:SummaryKelpFig}c,f). Finally, interactions depended on timescale, e.g., for CCA1 
+they were positive on annual timescales (Fig. \ref{fig:SummaryKelpFig}a) and
 negative on timescales $>8$y (Fig. \ref{fig:SummaryKelpFig}c).
@@
 i.e., $|f_{\mcP^{(1)}}|^2$, and modulated
-by a term quantifying the autoregressive nature of kelp dynamics,
-$1/|f_{\mcB}|^2$. These components are plotted in Fig. \ref{fig:directMoran} (CCA1 and SB regions) and 
-Fig. \ref{SM-fig:directMoran_SM_lag4} (CCA2 region) so their 
+by the autoregressive nature of kelp dynamics,
+$1/|f_{\mcB}|^2$. These components are plotted in Fig. \ref{fig:directMoran} (CCA1 and SB) and 
+Fig. \ref{SM-fig:directMoran_SM_lag4} (CCA2) so their 
 effects on synchrony can be assessed across timescales and regions.
+Nitrates are henceforth identified with $\epsilon^{(1)}$ and waves with $\epsilon^{(2)}$. 
 Both nitrates and waves fluctuate seasonally in CA (REFS). 
@@
 (Fig. \ref{fig:SummaryKelpFig}a,d, green and blue lines).
-Nitrates and waves are also synchronous on long timescales $>8$y in CA 
+Nitrates and waves are also synchronous on long timescales $>8$y  
 (Fig. \ref{fig:directMoran}l,n,aa,cc), possibly due to the North Pacific Gyre Oscillation 
@@
 in nitrates and waves produced some of the long-timescale
-synchrony observed in kelp (Fig. \ref{fig:SummaryKelpFig}c,f, green and blue lines).
-Kelp synchrony was generally stronger in the CCA1 region compared to the SB region 
-(compare Fig. \ref{fig:SummaryKelpFig}a-c to d-f) in part because the synchrony of nitrates, and of 
-waves, was more pronounced in the CCA1 region than in the SB region 
-(compare Fig. \ref{fig:directMoran}b,g,l to q,v,aa, and 
-Fig. \ref{fig:directMoran}d,i,n to s,x,cc), and also because waves
-had a stronger influence on kelp in CCA1 compared to SB (compare 
-Fig. \ref{fig:directMoran}c,h,m to r,w,bb).
-Regression coefficients which determined $f_{\mcP^{(1)}}$, $f_{\mcP^{(2)}}$
-and $f_{\mcB}$ (see Methods) are in Table \ref{SM-tab:regcoefs}.
+synchrony in kelp (Fig. \ref{fig:SummaryKelpFig}c,f, green and blue lines).
+Kelp synchrony was stronger in CCA1 than in SB 
+(Fig. \ref{fig:SummaryKelpFig}a-c v. d-f, black lines) in part because the synchrony of nitrates, and of 
+waves, was more pronounced in CCA1 than in SB 
+(Fig. \ref{fig:directMoran}b,g,l v. q,v,aa, and 
+Fig. \ref{fig:directMoran}d,i,n v. s,x,cc), and also because waves
+had a stronger influence on kelp in CCA1 than in SB (Fig. \ref{fig:directMoran}c,h,m v. r,w,bb).
+Regression coefficients determining $f_{\mcP^{(1)}}$, $f_{\mcP^{(2)}}$
+and $f_{\mcB}$ (Methods) are in Table \ref{SM-tab:regcoefs}.
 
 %Interactions between Moran effects, CCA1
-Interactions between Moran effects in CCA1 were parallel to scenario 1 of theoretical case study C.
-In CCA1, nitrates have a one-quarter-delayed positive effect on observed 
-kelp populations and waves have immediate negative effect (REFS), 
+Interactions between Moran effects in CCA1 were parallel to one of the theoretical case studies, CaC1.
+In CCA1, nitrates had a $1$q-delayed positive effect on observed 
+kelp populations and waves had immediate negative effect (REFS), 
 %***DAN: ask Tom/Kyle/Max what is a good reference for this
-just as $\epsilon^{(1)}$ effects in case study C, scenario 1, 
-were positive and delayed by a quarter and $\epsilon^{(2)}$ effects were negative and immediate.
+just as $\epsilon^{(1)}$ effects in CaC 
+were positive and delayed by $1$q and $\epsilon^{(2)}$ effects were negative and immediate.
 Delayed effects are reflected, for CCA1, in Fig. \ref{fig:indirectMoran}a, and for 
-case study C scenario 1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
+CaC1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
 that $f_{\mcP^{(1)}}$ has phase about $-\pi/2$ at the annual timescale. The delay is also 
 reflected, for CCA1, in Table \ref{SM-tab:regcoefs}, where 
-the CCA1 entries for $p_0^{(1)}$ were close to $0$ and those for $p_1^{(1)}$ were positive.
-Immediate negative effects of waves in CCA1 are reflected in Fig. \ref{fig:indirectMoran}b,
-and for case study C scenario 1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
-that $f_{\mcP^{(2)}}$ has phase $-\pi$ at all timescales. See also the CCA1 $p_0^{(2)}$
-coefficients in Table \ref{SM-tab:regcoefs}, which are negative.
+the entries for $p_0^{(1)}$ for CCA1 are close to $0$ and those for $p_1^{(1)}$ are positive.
+Immediate negative effects of waves in CCA1 are reflected in Fig. \ref{fig:indirectMoran}b,g,l
+and for CaC1 in Fig. \ref{fig:TheoryCaseA}j, both of which show 
+that $f_{\mcP^{(2)}}$ has phase $-\pi$ at all timescales. See also the $p_0^{(2)}$
+coefficients for CCA1 in Table \ref{SM-tab:regcoefs}, which are negative.
 In CCA1, annual peak nitrates tend to come in spring, whereas annual peak waves 
@@
 Thus annual nitrate peaks tend to lag annual 
-wave peaks by one quarter, just as $\epsilon^{(1)}$ peaks lagged $\epsilon^{(2)}$
-peaks by a quarter in theoretical case study C, scenario 1.
-This is reflected, for CCA1, in Fig. \ref{fig:indirectMoran}d, and for case study C scenario 1 
+wave peaks by $1$q, just as $\epsilon^{(1)}$ peaks lagged $\epsilon^{(2)}$
+peaks by $1$q in CaC1.
+This is reflected, for CCA1, in Fig. \ref{fig:indirectMoran}d, and for CaC1 
 in Fig. \ref{fig:TheoryCaseA}k, both of which show that $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$
@@
 Thus lags compound on the annual timescale, in the CCA1 case as in the theoretical case: the annual positive 
-effects of nitrates on kelp tend to come in the summer, and the annual negative effects of waves come in the winter,
+effects of nitrates on kelp tend to come in summer, and the annual negative effects of waves come in winter,
 producing reinforcing interactions between Moran effects. Likewise, the $\pi/2$ phase 
-of $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:indirectMoran}c for CCA1, Fig. \ref{fig:TheoryCaseA}k for case study C scenario 1)
-approximately cancels the $-\pi/2$ phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ 
-(Fig. \ref{fig:indirectMoran}d for CCA1, Fig. \ref{fig:TheoryCaseA}k for case study C scenario 1) on annual timescales, so that  
+of $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}$ on annual timescales (Fig. \ref{fig:indirectMoran}c for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaC1)
+approximately cancels the $-\pi/2$ phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ on annual timescales
+(Fig. \ref{fig:indirectMoran}d for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaC1), so that  
 $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ has phase 
-approximately $0$, and the interaction-effect term 
-$\Rp(f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}})$ is strongly positive 
-on annual timescales in both the theoretical and emprical cases. 
+approximately $0$, making the interaction-effect term 
+$\Rp(f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}})$ strongly positive 
+on annual timescales for both the theoretical and emprical cases 
+(Fig. \ref{fig:indirectMoran}e for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaC1). 
 %***DAN: read the below sentence again and see if it sucks
 Interactions between Moran effects on long-timescales ($>8$y) were also similar for both CCA1 and
-for case study C, scenario 1, and for the same reason in both cases: on long timescales, sub-annual lags
+CaC1, and for the same reason in both cases: on long timescales, sub-annual lags
 are inconsequential, and the positive effects of nitrates/$\epsilon^{(1)}$ and the negative
 effects of waves/$\epsilon^{(2)}$ therefore lead to negative interactions.
-Interactions between Moran effects in CCA2 operated in a similar manner to CCA1 (Fig. \ref{SM-fig:indirectMoran_SM_lag4}).
+Interactions between Moran effects in CCA2 operated similarly to CCA1 (Fig. \ref{SM-fig:indirectMoran_SM_lag4}).
+
+%***DAN: Don't forget to appropriately use q in place of quarter from here on as you edit. If you forget, do a 
+%search and replace.
 
 %Interactions between Moran effects, SB
+Due to different nitrate effects on observable kelp growth in SB compared to 
+CCA1, interactions between Moran effects followed a different 
+mechanism in SB compared to CCA1, leading to the slightly negative versus positive interactions already 
+documented on annual timescales for the two regions in Fig. \ref{fig:SummaryKelpFig}d v. a. 
+Whereas in CCA1, nitrate effects on observed kelp density were delayed by $1$q, in SB nitrate effects 
+were observable within the same quarter (Fig. \ref{fig:indirectMoran}a v. p; see also Table 
+\ref{SM-tab:regcoefs}, where $p_0^{(1)}$ terms were close to zero for CCA1 and positive for SB,
+whereas $p_1^{(1)}$ terms were positive for CCA1 and close to zero for SB). Wave effects were negative and
+immediate in both regions, though they were less pronounced in SB (Fig. \ref{fig:indirectMoran}b,g,l v. q,v,aa; see also 
+Table \ref{SM-tab:regcoefs}). Annual peaks in nitrates still tended to lag annual peaks in waves by $1$q in SB,
+as in CCA1, though the seasonal periodicity of both variables was reduced in SB compared to 
+CCA1 (Fig. \ref{fig:indirectMoran}s v. d). 
+Whereas in CCA1, the tendency of nitrate annual peaks to follow wave annual peaks by $1$q, added to the tendency of 
+nitrate effects on kelp to be delayed by $1$q, resulted in summertime positive nitrate effects and wintertime 
+negative wave effects which reinforced sycnhrony; in SB nitrate instead both peaked and had its positive effects on kelp in spring,
+while waves still had negative effects on kelp in wintertime. Thus the effects of nitrates and waves in SB
+were approximately a quarter-cycle separated from each other with respect to the annual timescale (Fig \ref{fig:indirectMoran}t), 
+and so produced neither
+much reinforcement nor much destructive interference of synchrony on annual timescales (Fig. \ref{fig:SummaryKelpFig}d). Slightly negative 
+interactions were observed (Fig. \ref{fig:SummaryKelpFig}d) because of slight deviations from the approximate phase
+alignments described above. On long timescales, interactions between Moran effects were analogous in CCA1 and SB
+(though weaker in SB; Fig. \ref{fig:indirectMoran}o v. dd and Fig. \ref{fig:SummaryKelpFig}c v. f) because, again, 
+sub-annual lags are inconsequential on long timescales and what mattered instead was the oppositely signed 
+influences of the two environmental variables.
 
 \section{Discussion}
+
+%Maybe a para somewwhere about how it can work that synchrony explained by one factor (or the sum of 
+%synchrony explained by two) can add up to more than 100% and then interaction effects can bring the total
+%down. Basically this means if the two factors were independent, synchrony would be much greater. E.g.,
+%CCA1 on long timescales. Might want to use the analogy with two parents pushing a kid on a swing,
+%and then multiple sets of such setups. Or maybe make it more official sounding by talking about 
+%oscillators instead of swings. Not sure a swing is a good example because it has a very specific 
+%intrinsic dynamics. Likewise for a oscillator. Think about it, not sure what to do here.
 
@@
 %Explain/explore how my theory goes beyond what they did - the intention is to get at the mechanism
-%better and how it actually works. It may be sufficient to keep this a fairly brief para.
+%better and how it actually works. It may be sufficient to keep this a fairly brief para. One diff is that
+%my Fourier approach facilitates analytic examination of interaction effects for the model I consider.
+%Their approach is better for dealing with non-stationarity and other features of some datasets. My
+%approach facilitates understanding the inner workings of the mechanism of interacting Moran effects, 
+%their approach facilitates identifying such effects in datasets.
 
@@
 %instead it's just to explore the value of perspective that we now have based on 20 years of hindsight
-%and ongoing work.
+%and ongoing work. Note that you cannot really make dispersal and Moran effects independent/uncorrelated
+%in the same way you can make two Moran effects independent/uncorrelated.
 
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_MainPlot_ForMS.pdf}
-\caption{Main components of the new theory as applied kelp, central CA 1 (CCA1) region (a-c) and 
+\caption{Main components of the new theory as applied to kelp, central CA 1 (CCA1) region (a-c) and 
 Santa Barbara (SB) region (d-f). Panels show the terms on the right side of 
@@
 (black line) because other, unmeasured factors also help synchronize kelp dynamics.
-The CCA1 and SB regional results, respectively, approximately parallel the 
+The CCA1 results approximately parallel the 
 results for theoretical case
-study C scenarios 1 and 2 (Fig. \ref{fig:TheoryCaseA}i,m; see text for details).
+study C scenario 1 (Fig. \ref{fig:TheoryCaseA}i; see text for details).
 See Fig. \ref{SM-fig:SummaryKelpFig_SM_lag4} for the central CA 2 (CCA2) region,
 for which results were substantially the same as for CCA1. This figure used 
-kelp lag $n=4$ (see (\ref{main_model_1}) for the meaning of $n$); 
+kelp lag $n=4$ [see (\ref{main_model_1})]; 
 see Figs \ref{SM-fig:SummaryKelpFig_SM_lag8} 
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainDirectMoran_ForMS.pdf}
-\caption{Components of direct Moran effects for kelp, for the CCA1 (a-o)
-and SB regions (p-dd). The index $1$ denotes nitrates, and $2$ denotes waves.
+\caption{Components of direct Moran effects for kelp, for CCA1 (a-o)
+and SB (p-dd). Nitrates are identified with $\epsilon^{(1)}$ and waves with $\epsilon^{(2)}$. 
 Vertical axis extents are the same for corresponding panels for the
 two regions, to facilitate comparisons. 
-See Fig. \ref{SM-fig:directMoran_SM_lag4} for the central CCA2 region,
+See Fig. \ref{SM-fig:directMoran_SM_lag4} for CCA2,
 for which results were similar to CCA1. This figure used 
 kelp lag $n=4$; see Figs \ref{SM-fig:directMoran_SM_lag8} 
-and \ref{SM-fig:directMoran_SM_lag12} for $n=8, 12$, respectively.
-See the online version for a color
-rendering of all figures.}\label{fig:directMoran}
+and \ref{SM-fig:directMoran_SM_lag12} for $n=8, 12$, respectively.}\label{fig:directMoran}
 \end{figure}
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
-\caption{Filler. See the online version for a color
+\caption{Components of interactions between Moran effects for kelp, for CCA1 (a-o) and SB (b-dd). 
+Nitrates are identified with $\epsilon^{(1)}$ and waves with $\epsilon^{(2)}$. 
+Vertical axis extents are the same for corresponding panels for the
+two regions, to facilitate comparisons. 
+See Fig. \ref{SM-fig:indirectMoran_SM_lag4} for CCA2,
+for which results were similar to CCA1. This figure used 
+kelp lag $n=4$; see Figs \ref{SM-fig:indirectMoran_SM_lag8} 
+and \ref{SM-fig:indirectMoran_SM_lag12} for $n=8, 12$, respectively.
+See the online version for a color
 rendering of all figures.}\label{fig:indirectMoran}

### 2022-01-18 Code/Kelp_AnalysisStep_MakeMapFig.R

@@ -119,4 +119,12 @@ CC2locs<-locs[CC2locstouse,] #lat lon of locations corresponding to rows of CC2k
 SBkelpdat<-kelp[SBlocstouse,] #each row is a location with time running left to right, columns corresponding to quarters as in the variable "quarters"
 SBlocs<-locs[SBlocstouse,] #lat lon of locations corresponding to rows of SBkelpdat
 
+#temp save files just so I can send these all to Max without him even accessing the repo
+saveRDS(CC1kelpdat,file=paste0(resloc,"CC1kelpdat.Rds"))
+saveRDS(CC2kelpdat,file=paste0(resloc,"CC2kelpdat.Rds"))
+saveRDS(SBkelpdat,file=paste0(resloc,"SBkelpdat.Rds"))
+saveRDS(CC1locs,file=paste0(resloc,"CC1locs.Rds"))
+saveRDS(CC2locs,file=paste0(resloc,"CC2locs.Rds"))
+saveRDS(SBlocs,file=paste0(resloc,"SBlocs.Rds"))
+
 #OK, MAX, TAKE IT FROM HERE...

### 2022-01-27 Paper/Paper.Rnw
draft of Discussion done

@@
 \usepackage{authblk}
+\usepackage{lineno}
 
@@
 \externaldocument[SM-]{SupMat}
+
+%Get line numbers to work, also in figure captions
+\linenumbers
+\DeclareCaptionFont{linenumbers}{\internallinenumbers}
+\captionsetup[figure]{font+=linenumbers} % use \internallinenumbers inside figure caption
 
@@
 distinct from those of \cite{Kendall2000}, who examined interactions between Moran and dispersal
-causes of synchrony. We relate our work to that study in the Discussion.
+causes of synchrony.
 
@@
 are equivalent because frequency and timescale are reciprocal], 
-and it will turn out (see below, and \emph{Discussion}) that a timescale-specific 
+and it will turn out (see below, and Discussion) that a timescale-specific 
 approach is essential to the development of our new theory. 
-%***DAN: Good topic for a Discussion para, the fact that this could not really be done without taking frequency specificity into account, and why, though need to think a bit more before writing
 We therefore here define, using the spectral methods outlined above, a frequency/timescale-specific measure of spatial synchrony, 
@@
 Our population model is 
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 where $w_i(t)$ is an index of the population in location $i=1,\ldots,N$ at time
@@
 Linearization and ``weak noise'' assumptions have been commonly adopted to make
-theoretical progress in ecology (REFS), and it has been demonstrated that results based on a
-weak noise assumption often hold for noise which is fairly strong (REF).
-%***DAN: Add refs later
+theoretical progress in ecology, and it has been demonstrated that results based on a
+weak noise assumption often hold for noise which is fairly strong \citep{Nisbet1977,Reuman2006,Desharnais2018}.
 See \cite{Brillinger2001} for background on stochastic processes.
@@
 synchrony between those processes: 
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 Here, $f_{\mcB} = 1-b_1 \mu - b_2 \mu^2 - \cdots - b_n \mu^n$, 
@@
 Waves influence kelp dynamics through direct disturbance events which can damage kelp or extirpate kelp locally 
-when waves are large \citep{Dayton1992,Cavanaugh2011,Bell2015}. Thus wave effects 
+when waves are large \citep{Cavanaugh2011,Bell2015,Schiel2015}. Thus wave effects 
 are immediate and $m_2=0$ was used. Nitrates are known to fuel rapid kelp growth, though in some areas
@@
 take time for damaged kelp to grow back to the surface 
-\citep{Zimmerman1986,Cavanaugh2011,Bell2015}. 
+\citep{Cavanaugh2011,Bell2015,Schiel2015}. 
 Therefore $m_1=1$ was used. Kelp holdfasts on the sea floor can last for multiple years, so kelp lag effects were considered:
@@
 computed using the consistent estimator of section 7.4 of \cite{Brillinger2001}. 
-%***DAN: check this section number, and if possible give a specific formula of theorem or something, and also 
-%check the below text and maybe augment it
 The estimator is a smoothed periodogram, with the width of the smoothing kernel selected to 
@@
 In CCA1, nitrates had a $1$q-delayed positive effect on observed 
-kelp populations and waves had immediate negative effect (REFS), 
-%***DAN: ask Tom/Kyle/Max what is a good reference for this
+kelp populations and waves had immediate negative effect \citep{Cavanaugh2011,Bell2015,Schiel2015}, 
 just as $\epsilon^{(1)}$ effects in CaC 
@@
 (Fig. \ref{fig:indirectMoran}e for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaC1). 
-%***DAN: read the below sentence again and see if it sucks
 Interactions between Moran effects on long-timescales ($>8$y) were also similar for both CCA1 and
@@
 Interactions between Moran effects in CCA2 operated similarly to CCA1 (Fig. \ref{SM-fig:indirectMoran_SM_lag4}).
-
-%***DAN: Don't forget to appropriately use q in place of quarter from here on as you edit. If you forget, do a 
-%search and replace.
 
@@
 
-%Maybe a para somewwhere about how it can work that synchrony explained by one factor (or the sum of 
-%synchrony explained by two) can add up to more than 100% and then interaction effects can bring the total
-%down. Basically this means if the two factors were independent, synchrony would be much greater. E.g.,
-%CCA1 on long timescales. Might want to use the analogy with two parents pushing a kid on a swing,
-%and then multiple sets of such setups. Or maybe make it more official sounding by talking about 
-%oscillators instead of swings. Not sure a swing is a good example because it has a very specific 
-%intrinsic dynamics. Likewise for a oscillator. Think about it, not sure what to do here.
-
+%*DONE Summary
+We provided a general mathematical theory of interactions between Moran effects. When two related spatially 
+synchronous environmental drivers both influence a set of populations across a landscape, interactions can make 
+synchrony in the populations either substantially stronger or markedly weaker 
+than would be expected if the 
+two drivers were unrelated, a new mechanism of population spatial synchrony.
+Our general theory illuminates precisely how timings of 
+influences of the Moran drivers on populations can interact with relationships between the
+drivers to produce interaction effects, which may vary by timescale in both their strength and
+sign. Moran effects are very common. Interactions between Moran effects have been tested for, and observed, twice previously, and may
+be common [\cite{Sheppard2019,CastoraniSheppardKelp}; see also below]. 
+Our study is the first to provide a comprehensive theory of the new mechanism. 
+We used our theory and several case studies based on models and 
+California kelp populations to provide improved intuition about the mechanism.
+
+%*DONE arguments about likelihood that interactions will be common, extending earlier arguments
+We argue that it is likely that interactions between Moran effect are common. Most species
+are influenced by more than one environmental driver. Drivers are frequently spatially autocorrelated,
+and are also often related to each other because of their common origin in underlying meteorological 
+phenomena including major climatic oscillations such as 
+the NPGO or El Ni\~{n}o Southern Oscillation (ENSO).
+Driver pairs which may commonly produce interacting Moran effects include instances where the same quantity 
+is measured in distinct parts of the year (e.g., spring and summer temperatures, or March and April rainfall,
+or December and January snowfall); or when distinct variables are measured in the same part of the 
+year (e.g., spring temperature and precipitation). Future work should systematically 
+investigate cross synchrony of environmental variables, and could then use our theory
+to directly assess the potential for interacting Moran effects to occur under a variety of biological scenarios. Since 
+interactions between Moran effects have been discovered in both of the two systems so far 
+tested \citep{Sheppard2019,CastoraniSheppardKelp}, future work should also, when
+data are available, test directly for interactions.
+
+%*Revist intuition
+We revisit and extend the intuition behind interacting Moran effects, using the central CA kelp
+example as a vehicle \citep{CastoraniSheppardKelp}. Large winter waves 
+have immediate negative effects on kelp in central CA, whereas the positive
+effects of spring nitrates manifest in summer. So nutrient and wave effects can 
+reinforce each other in producing annual oscillations: large kelp increases in summer due 
+to abundant nutrients can be followed by big crashes in winter due to waves, both factors combining
+to accentuate the annual cycle. Thus positive
+interactions between Moran effects on annual timescales occur whenever 
+years with above-average waves coincide with years with plentiful nutrients in other locations:
+if a large-wave year in location A coincides with a high-nutrient year in location B, 
+both locations will tend to have bigger annual fluctuations that year, accentuating 
+annual-timescale synchrony. Sub-annual lags and delays make essentially no difference, 
+however, on long timescales. On long timescales, large-wave and abundant-nutrient years
+counteract each other. If a multi-year period of large waves in location A coincides with
+a multi-year period of abundant nutrients in location B, the multi-year-average kelp abundance
+in A will tend to be reduced, whereas the multi-year-average kelp abundance in B will
+tend to be augmented, reducing long-timescale synchrony. Lags and interactions between
+drivers must always be compared to the timescale of interest to determine interaction effects,
+as in Fig. \ref{fig:IntroTheory}, and as captured mathematically in (\ref{rhoww_T1to3}).
+
+%*DONE Comparison to Castorani et al paper, and maybe also a bit on comparison to the Sheppard et al 2019 paper
+Our results about kelp were consistent with those of \cite{CastoraniSheppardKelp}, though
+the two papers have complementary emphases. In spite of numerous methodological choices which differed between the 
+two studies, both our results and \cite{CastoraniSheppardKelp}
+show positive effects of both nutrients and waves on synchrony on both annual and long ($>4$y)
+timescales in both central and southern CA. And both studies
+show positive interactions between nutrient and wave Moran effects on annual timescales
+in central CA, but negative interactions in southern CA at annual timescales and in both central
+and southern CA at long timescales. The Fourier approach of this study was designed to facilitate 
+mathematical examination of interactions between Moran effect as a general mechanism, whereas the 
+wavelet approach of \cite{CastoraniSheppardKelp} was instead optimized for detecting interactions 
+and identifying Moran mechanisms in data, in spite of non-stationarity and
+other complicating features which are present in many ecological datasets. The study of
+\citep{Sheppard2019} developed the wavelet methods applied by \cite{CastoraniSheppardKelp}.
+The approach of this study relates indirectly to the approach of \cite{Anderson2021}, though that
+study concerned different research questions.
+
+%*DONE relationship to earlier papers comparing synchrony of drivers and pops
+It has been a frequent topic of research why populations are often less synchronous, or sychronous over a smaller
+spatial extent, than might be expected given the strength or extent of 
+synchrony of environmental drivers \citep{Herfindal2022}.
+Antagonistic (negative) interaction effects between Moran drivers may be one common and previously
+unrecognized mechanism contributing to this discrepancy. On the other hand,
+a recent paper descibed an ``enhanced Moran effect'' by which specific 
+patterns of temporal autocorrelation
+in Moran drivers can theoretically cause greater synchrony in populations 
+than in the drivers themselves \citep{Massie2015}.
+Synergistic (positive) interaction effects is another mechanism by which populations can
+be more synchronous than expected.
+
+%*DONE How/why a frequency-specific approach is necessary for all this
+An ever growing body of work shows that synchrony is, and should be analyzed as, a timescale-specific 
+phenomenon \citep{Vasseur2007,Keitt2008,Vasseur2014,Sheppard2016,Defriez2016,Defriez2017,Defriez2017b,Desharnais2018,Anderson2021}.
+The results of our present study extend this earlier work by showing that interactions between Moran effects 
+are also timescale-specific, differing stongly in strength, and in sign, across timescales. 
+A non-timescale-specific version of our theory could be obtained by appropriately integrating our main equation 
+(\ref{rhoww_T1to3}) across frequencies. But doing so would reduce insight by conflating distinct and
+often disparate processes. For instance, if the terms of (\ref{rhoww_T1to3}) were integrated 
+across frequencies for our central CA kelp example, positive interactions between Moran drivers on short timescales 
+(Fig. \ref{fig:SummaryKelpFig}a) would be summed with and partially cancel 
+negative interactions on long timescales (Fig. \ref{fig:SummaryKelpFig}c),
+falsely suggesting reduced importance of interaction effects. Our results also have implications given the 
+short duration of most time series in ecology. 
+Analyses based on short time series can only be expected to reveal Moran effects or interactions between Moran
+effects occurring on short timescales, which our results show can differ substantially from long-timescale results. 
+It has not escaped our notice that kelp time series even longer than ours, once 
+available, may reveal Moran effects and interactions on timescales longer than ours which differ from
+our results. In our view this provides another strong argument, among many others, in support of 
+long-term monitoring of ecosystems.
+
+%DECIDED NOT TO WRITE*Revisitation of diffs between SoCal and CentCal
 %Note that we get differences in intereaction effects between central and southern CA because we 
@@
 %basically all fronds from the same plant are ripped out.
-
-%Need a paragraph (already mentioned above) on how a frequency/timescale-specific approach
-%is necessary for all this. Need to think more about it first, i.e., how would you attempt to 
-%do this without such an approach, and how far would you get? Or maybe that is not quite the 
-%right at it (though you should try thinking that through and see what you get), maybe another
-%possibly better way would be to see that everything we list here can depend on timescale.
-%You can get a non-timescale-specific approach just by integrating our main equation across
-%frequencies. However, that will obscure the fact that you can get strong positive interaction 
-%effects on some timescales and strong negative ones on others. What would you get in our 
-%theoretical case studies if you integrated and would it reveal what's really going on? What 
-%are the implications of timescale structure for practical situations where populations and
-%environmental variables are measured over a certain typical duration (i.e, short time series)?
-
-%We definitely need a discussion para comparing my results to those of Max and Lawrence (hopefully they are 
-%basically consistent). Note differences in methods. Note differences in results (hopefully small). 
-%Explain/explore how my theory goes beyond what they did - the intention is to get at the mechanism
-%better and how it actually works. It may be sufficient to keep this a fairly brief para. One diff is that
-%my Fourier approach facilitates analytic examination of interaction effects for the model I consider.
-%Their approach is better for dealing with non-stationarity and other features of some datasets. My
-%approach facilitates understanding the inner workings of the mechanism of interacting Moran effects, 
-%their approach facilitates identifying such effects in datasets.
-
-%Go back to the Kendall study of 2000 (I think that's the year), the one that looked at interactions
-%between Moran and dispersal causes of synchrony. If I remember it correctly (review it), it's a bit silly,
-%because all it basically does it turn dispersal and Moran drivers on and off in a model and show that
-%you get more or less synchrony from the both-on case that you would get by summing the synchrony 
-%you get from each of the one-on cases. But the expectation that correlations should sum is naive anyway - 
-%correlation is not a linear operator (though covariance is - check to make sure they use cor instead of
-%cov, and how would their results be different if they used cov?).
-%Our study has a better baseline to compare to than the sum - the baseline is if the two Moran drivers are
-%independent (or, rather, it's a bit less than that, and has to do with the cross synchrony measure
-%being 0). This is not to belittle the achievements of that paper, which was groundbreaking in its time,
-%instead it's just to explore the value of perspective that we now have based on 20 years of hindsight
-%and ongoing work. Note that you cannot really make dispersal and Moran effects independent/uncorrelated
-%in the same way you can make two Moran effects independent/uncorrelated.
 

### 2022-01-27 Paper/SupMat.Rnw
draft of Discussion done

@@
 \externaldocument[MT-]{Paper}
+
+%Get line numbers to work, also in figure captions
+\linenumbers
+\DeclareCaptionFont{linenumbers}{\internallinenumbers}
+\captionsetup[figure]{font+=linenumbers} % use \internallinenumbers inside figure caption
 
@@
 to sampling locations, then we measure the synchrony of $w$ as 
+\begin{linenomath}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath}
 where $S_{ww}$ is the spectral matrix of the process $w$. This is a function of frequency or timescale because
@@
 Therefore 
+\begin{linenomath}
 \begin{align}
@@
 \end{align} 
+\end{linenomath}
 The quantity $\rho_{ww}$ makes sense as a measure of synchrony in part because 
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 The quantity $\Rp(S_{w_i w_j})$ is the cospectrum of the component processes $w_i$ and $w_j$, 
@@
 Thus an integral of $\rho_{ww}$ across frequencies equals 
+\begin{linenomath}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath}
 which is a standard quantity that can be used to measure non-frequency-specific synchrony.
@@
 (or, simply, the \emph{cross synchrony}) between $\epsilon^{(1)}$ and $\epsilon^{(2)}$ as
+\begin{linenomath}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath}
 This is a complex-valued function of frequency or timescale. Why is this interpretable as cross-variable 
@@
 $\epsilon_i^{(a)}(t+u)$ and $\epsilon_j^{(b)}(t)$ were assumed independent for $u \neq 0$, we have 
+\begin{linenomath}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath}
 Thus the spectra and cross spectra in this example are constant. It
@@
 Here, 
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 So $S_{\zeta_i^{(1)}\zeta_j^{(2)}}=w_{12} \exp(-2 \pi \iota f (l_1 - l_2))$. It then follows 
@@
 The population model is specified starting with the equation
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 where $i=1,\ldots,N$ indexes habitat patches or sampling locations, $w_i(t)$ is a measure of the 
@@
 we define the linear filters 
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 Using the language of filters and stochastic processes, the original model 
 (\ref{main_model_1}-\ref{main_model_4}) can then be written as
+\begin{linenomath}
 \begin{equation}
@@
 \end{equation}
+\end{linenomath}
 Writing the model in the language of stochastic processes and linear filters makes it straightforward
@@
 Computing the spectral matrix of both sides of the model (\ref{main_model_stochproc}) gives
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 where $T(\mcB)$, $T(\mcP^{(1)})$ and $T(\mcP^{(2)})$ are the tranfer functions of the filters
@@
 $f$ is frequency in units of cycles per time step, we have
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 \noindent These are all scalar matrices, so (\ref{spectmat_eq1_t1}-\ref{spectmat_eq1_t6}) simplifies to
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 where 
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 We here made use of the assumption that the complex roots of the
@@
 Now, averaging the off-diagonal entries of both sides of (\ref{Sww_T1}-\ref{Sww_T6}) gives
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 \noindent where $\Rp(\cdot)$ denotes the real part, and where (\ref{SumSww_T4}-\ref{SumSww_T6})
@@
 Now, making use of the definitions of section \ref{Sect:measuring_sync}, this becomes
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 which expresses population synchrony in terms of noise synchrony and cross synchrony.
@@
 the aggregate effects of unknown processes, then we can write
+\begin{linenomath}
 \begin{align}
@@
 \end{align}
+\end{linenomath}
 The term (\ref{rhoww_T1}) is contributions to the synchrony of $w$ due to direct Moran effects of $\epsilon^{(1)}$.

### 2022-01-31 Paper/Paper.Rnw
various edits to Discussion and getting repo ready

@@
 
-%***DAN: Right now sup mat section S7 is not cited. It only has figures and tables, not sure I plan to 
-%add anything to it other than that. Either get rid of it as a section, or else cite it in some of the places 
-%where you currently cite the figures and tables in it. My current hunch is that the first of those options is
-%better. Have to decide what to do with Fig. S1 in that case.
-
 %Description of data after cleaning
@@
 from southern CA, close to Santa Barbara
-(called SB; see Fig. \ref{fig:map}). Each kelp measurement is an estimate of mean quarterly kelp biomass per unit 
-useable habitat in a 500m stretch of coastline. We used coastline segments where kelp was persistent
+(called SB; see Fig. \ref{fig:map}). Each kelp measurement is an estimate of 
+mean quarterly kelp biomass (kg wet) per unit 
+useable habitat ($\text{m}^2$) in a 500m stretch of coastline. We used coastline segments where kelp was persistent
 through essentially all of the $\Sexpr{quarters[1,2]}$-$\Sexpr{quarters[dim(quarters)[1],2]}$ 
@@
 of maximum wave height and mean nitrate concentration for each quarter and location. Both waves
-and nutrients are known to influence kelp population dynamics and synchrony 
+and nutrients are known to influence kelp dynamics and synchrony 
 \citep{Cavanaugh2013,CastoraniSheppardKelp}.
@@
 Data for the project are publicly archived \citep{SBCLTER2021}.
-All computations were done in R version 3.4.4 on a laptop running Ubuntu Linux
+All computations were done in R version 3.6.3 on a laptop running Ubuntu Linux
 16.04. Complete codes for the project workflow are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.
@@
 Nitrates are henceforth identified with $\epsilon^{(1)}$ and waves with $\epsilon^{(2)}$. 
-Both nitrates and waves fluctuate seasonally in CA (REFS). 
-%***DAN: Fugure out a ref for this later
+Both nitrates and waves fluctuate seasonally in CA \citep{Schiel2015}. 
 Thus nitrate and wave 
@@
 In CCA1, annual peak nitrates tend to come in spring, whereas annual peak waves 
-tend to come in winter (REFS). 
-%***DAN: add refs later
+tend to come in winter \citep{Schiel2015,Bell2015}. 
 Thus annual nitrate peaks tend to lag annual 
@@
 %*DONE Summary
-We provided a general mathematical theory of interactions between Moran effects. When two related spatially 
+We provided a general mathematical theory of the new mechanism of 
+interactions between Moran effects. When two related spatially 
 synchronous environmental drivers both influence a set of populations across a landscape, interactions can make 
 synchrony in the populations either substantially stronger or markedly weaker 
-than would be expected if the 
-two drivers were unrelated, a new mechanism of population spatial synchrony.
+than would otherwise be expected.
 Our general theory illuminates precisely how timings of 
-influences of the Moran drivers on populations can interact with relationships between the
-drivers to produce interaction effects, which may vary by timescale in both their strength and
-sign. Moran effects are very common. Interactions between Moran effects have been tested for, and observed, twice previously, and may
-be common [\cite{Sheppard2019,CastoraniSheppardKelp}; see also below]. 
-Our study is the first to provide a comprehensive theory of the new mechanism. 
-We used our theory and several case studies based on models and 
-California kelp populations to provide improved intuition about the mechanism.
+influences of the drivers on populations can interact with relationships between the
+drivers to alter Moran effects. Interactions may vary by timescale in both their strength and
+sign. We used our theory and several case studies based on models and 
+kelp populations to provide intuition about the new mechanism (see also below).
+Moran effects are ubiquitous. Interactions 
+between Moran effects have been tested for only twice previously, and were observed
+on both occassions; for this and other reasons (see below) interactions may
+also be common \citep{Sheppard2019,CastoraniSheppardKelp}. 
+Moran interactions are therefore
+a newly recognized and potentially widespread aspect of a fundamental means (Moran effects) by which 
+environmental factors influence populations
+and diverse phenomena such as ecosystem stability \citep{Wilcox2017}, population cycling \citep{Anderson2021} 
+and extinction \citep{Ghosh2020}.
+Climate change is altering many aspects of environmental variables, including their means, variances and
+spatial correlations \citep{Lyon2019,Keelings2020}, as well as 
+relationships between environmental variables and the nature of 
+their influences on populations. There is therefore potential for climate change to alter interactions 
+between Moran effects, in ways our new theory may help researchers to understand (see below).  
+The augmented fundamental understanding of Moran effects which we have provided may 
+substantially benefit both basic \citep{Liebhold2004} 
+and applied \citep{Hansen2020,Larsen2021,Herfindal2022} ecological research.
 
@@
 and are also often related to each other because of their common origin in underlying meteorological 
-phenomena including major climatic oscillations such as 
+phenomena, including major climatic oscillations such as 
 the NPGO or El Ni\~{n}o Southern Oscillation (ENSO).
@@
 year (e.g., spring temperature and precipitation). Future work should systematically 
-investigate cross synchrony of environmental variables, and could then use our theory
+investigate cross synchrony ($\rho_{\epsilon^{(1)} \epsilon^{(2)}}$) of environmental variables, and could then use our theory
 to directly assess the potential for interacting Moran effects to occur under a variety of biological scenarios. Since 
@@
 tested \citep{Sheppard2019,CastoraniSheppardKelp}, future work should also, when
-data are available, test directly for interactions.
-
-%*Revist intuition
+sufficient data are available, test directly for interactions. Temperature and precipitation variables
+measured in the same season may be particularly promising candidates for producing interactions because of
+the well known joint influence of these variables on plants. 
+\cite{Defriez2017b} tested for interactions between Moran influences
+of temperature and precipitation on aspects of synchrony of a terrestrial plant biomass index, finding no 
+evidence for interactions. But their data were global in extent and spatially coarse, their time series were 
+short, and they used statistical methods more aligned with their goals of 
+understanding the geography of synchrony than with our present goals of understanding 
+synchrony itself. So temperature-precipitation interactions may still be found in future work
+by applying our theory and the methods of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} to
+suitable data. Likewise, rainfall variables measured in distinct portions
+of the growing season may also be promising candidates for finding interaction effects, for similar reasons.
+
+%*DONE Revist intuition
 We revisit and extend the intuition behind interacting Moran effects, using the central CA kelp
-example as a vehicle \citep{CastoraniSheppardKelp}. Large winter waves 
+example as a vehicle. Large winter waves 
 have immediate negative effects on kelp in central CA, whereas the positive
@@
 both locations will tend to have bigger annual fluctuations that year, accentuating 
-annual-timescale synchrony. Sub-annual lags and delays make essentially no difference, 
+annual-timescale synchrony between the locations \citep{CastoraniSheppardKelp}. 
+Sub-annual lags and delays make essentially no difference, 
 however, on long timescales. On long timescales, large-wave and abundant-nutrient years
-counteract each other. If a multi-year period of large waves in location A coincides with
+counteract each other: if a multi-year period of large waves in location A coincides with
 a multi-year period of abundant nutrients in location B, the multi-year-average kelp abundance
@@
 drivers must always be compared to the timescale of interest to determine interaction effects,
-as in Fig. \ref{fig:IntroTheory}, and as captured mathematically in (\ref{rhoww_T1to3}).
+as in Fig. \ref{fig:IntroTheory}, and as captured rigorously in our theory.
 
@@
 Our results about kelp were consistent with those of \cite{CastoraniSheppardKelp}, though
-the two papers have complementary emphases. In spite of numerous methodological choices which differed between the 
+these two studies use distinct methods; the studies complement each other in important ways. In spite of 
+numerous methodological choices which differed between the 
 two studies, both our results and \cite{CastoraniSheppardKelp}
@@
 show positive interactions between nutrient and wave Moran effects on annual timescales
-in central CA, but negative interactions in southern CA at annual timescales and in both central
-and southern CA at long timescales. The Fourier approach of this study was designed to facilitate 
-mathematical examination of interactions between Moran effect as a general mechanism, whereas the 
+in central CA, but negative interactions in southern CA on annual timescales and in both central
+and southern CA on long timescales. The Fourier approach of our study was designed to facilitate 
+mathematical examination of interactions between Moran effect as a general mechanism and development
+of general theory, whereas the 
 wavelet approach of \cite{CastoraniSheppardKelp} was instead optimized for detecting interactions 
@@
 \citep{Sheppard2019} developed the wavelet methods applied by \cite{CastoraniSheppardKelp}.
-The approach of this study relates indirectly to the approach of \cite{Anderson2021}, though that
+The modelling approach of this study relates indirectly to the approach of \cite{Anderson2021}, though that
 study concerned different research questions.
@@
 It has been a frequent topic of research why populations are often less synchronous, or sychronous over a smaller
-spatial extent, than might be expected given the strength or extent of 
+spatial extent, than might be expected given the strength and extent of 
 synchrony of environmental drivers \citep{Herfindal2022}.
-Antagonistic (negative) interaction effects between Moran drivers may be one common and previously
+Antagonistic interaction effects between Moran drivers may be one common and previously
 unrecognized mechanism contributing to this discrepancy. On the other hand,
-a recent paper descibed an ``enhanced Moran effect'' by which specific 
+two recent papers descibed ``enhanced Moran effects'' by which specific 
 patterns of temporal autocorrelation
 in Moran drivers can theoretically cause greater synchrony in populations 
-than in the drivers themselves \citep{Massie2015}.
-Synergistic (positive) interaction effects is another mechanism by which populations can
+than in the drivers themselves \citep{Massie2015,Desharnais2018}.
+Synergistic interaction effects is another mechanism by which populations can
 be more synchronous than expected.
+
+%*DONE Revisit the kendall paper, briefly. Talk about possible three-way interactions.
+\cite{Kendall2000} considered interactions between mechanisms of synchrony. Their research 
+questions were distinct from ours, but combining their viewpoints and ours may lead to valuable
+future work. \cite{Kendall2000} considered interactions between dispersal and Moran causes of 
+synchrony, whereas we analyzed interactions between two Moran drivers, a distinct context.
+But it seems likely that numuerous ecological systems involve important dispersal effects on 
+synchrony as well as influences of at least two Moran effects. It may be necessary in future
+work to consider three-way interactions between dispersal and two Moran drivers and their 
+influence on synchrony. Dispersal can readily be added to our modelling framework.
+\cite{Desharnais2018} performed a spectral analysis on a model similar to ours
+which included dispersal but which only considered two habitat patches and one environmental driver.
+
+%*DONE Climate change
+Climate change has the potential to influence interactions between Moran effects in two specific ways
+which can be illuminated by our theory, and this potential should be investigated in
+future work. Examining the third term of (\ref{rhoww_T1to3}), climate change could
+alter interaction effects if it: 1) alters the term $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+quantifying cross synchrony between Moran drivers; or 2) alters one of the terms
+$f_{\mcP^{(1)}}$ or $f_{\mcP^{(2)}}$ specifying the nature of the influence of the 
+environmental variables $\epsilon^{(1)}$ and $\epsilon^{(2)}$, respectively, on populations. As advocated above,
+the term $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ should be systematically computed in future work, for a variety 
+of environmental variables, to assess whether interactions between Moran effects are likely to
+be general. As part of that process, the potential for changes in $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+could also be assessed, by using either time-windowed versions or wavelet adaptations 
+of this statistic, applied to either long-term climate records or future climate scanarios. 
+Differences in $f_{\mcP^{(1)}}$ were responsible for differences
+in the nature of Moran interactions between central CA and southern CA (Results), specifically
+differences between the two regions in the lag of nitrate effects on kelp populations. 
+If climate change modifies environmental effects on populations in a related way it would 
+be expected to produce similarly large changes in Moran interactions. Nitrate effects on kelp 
+are probably delayed by a quarter in central CA because winter waves commonly remove kelp,
+and it takes time for kelp to grow to the surface and become detectable by satellite, only
+then revealing the effects of elevated nitrates. In southern CA, kelp are less likely to be
+removed by winter waves. Effects of elevated nitrates on the growth of kelp stands already reaching
+the surface can be immediately observed. Climate change may similarly alter 
+lags and delays associated with environmental effects on populations in at least two ways:
+1) by altering species phenology; and 2) by increasing or decreasing growth rates and thereby
+decreasing or increasing delays (respectively). Though it is too early to conclude that
+effects on Moran interactions are among the important impacts of climate change, we feel the mechanisms
+outlined above are sufficient to warrant further investigation.
 
@@
 It has not escaped our notice that kelp time series even longer than ours, once 
-available, may reveal Moran effects and interactions on timescales longer than ours which differ from
+available, may reveal Moran effects and interactions, on timescales longer than ours, which differ from
 our results. In our view this provides another strong argument, among many others, in support of 
-long-term monitoring of ecosystems.
-
-%DECIDED NOT TO WRITE*Revisitation of diffs between SoCal and CentCal
-%Note that we get differences in intereaction effects between central and southern CA because we 
-%have lagged effects of nitrate on observed kelp density in central CA, but unlagged effects in southern
-%CA (to see this, look at the regression coefficients computed to get the models used for the
-%spectral analysis, which were saved). Careful inspection of my ms figures show this is the reason
-%for the positive versus slightly negative interaction effects seen in central v. souther CA. 
-%But I spke with Tom and Max and Kyle about this by email around 1/3/2022. They agreed that 
-%probably it has something to do with the fact that kelp is ripped out by waves in central
-%CA, so it takes a quarter to recover enough to be detectable from space (we can only see the 
-%canopy from space, on the surface of the water). In southern CA, waves rip out kelp like that
-%much less. So if kelp is always around and usually reaches the surface in southern CA, nitrate
-%effects can be observed more quickly. There ought to be a Discussion paragraph on this, basiaclly
-%saying things may be more complex that our analyses reveals with regard to some of the differences 
-%we observed between central and southern CA. We didn't explore this because our goal was to illustrate 
-%our theory, and our work does demonstrate the importance of interaction effects for the synchrony of kelp
-%as observed from space. But future work could look at this more, certainly, perhaps using state-space models
-%which can separate biological from observation processes.   
-%   Tom also told me in a meeting zoom call that in central CA growth of new fronds will be slower because 
-%there are not other fronds (left) on the same plant to provide a sugar subsidy to new fronds. In central CA,
-%basically all fronds from the same plant are ripped out.
+long-term on-the-ground or remote-sensing efforts to monitor ecosystems, such as the efforts 
+from which our kelp data are derived.
 
@@
 \begin{figure}
-\caption{To fill. See the online version for a color
-rendering of all figures.}\label{fig:map}
+\caption{Kelp sampling sites and example time series. Sampling sites (a) were from three regions,
+central California 1 (CCA1, blue points), central california 2 (CCA2, green points) and the region
+around Santa Barbara (SB, red points). Kelp density is shown with color intensity in (b)-(d),
+and those panels correspond to the regions. One example time series from each region is shown in
+(e)-(g), with locations at which these time series were measured labeled on panels (a)-(d). 
+See the online version for a color rendering of all figures.}\label{fig:map}
 \end{figure}

### 2022-01-31 Code/Kelp_AnalysisStep_CleanData_Final.R

@@ -58,11 +58,15 @@ for (counter in 1:(dim(kelp)[1]))
 {
   kelp[counter,]<-kelp[counter,]/(unname(quantile(kelpA[counter,],prob=.9))) #use the 90th percentile for robustness, instead of the max
 }
+kelpNDT<-kelp
 kelp<-wsyn::cleandat(kelp,times,2)$cdat
 
 #***
-#Save the final version of cleaned data
+#Save the final version of cleaned data, save the non-detrended data separately
 #***
+
+saveRDS(kelpNDT,file=paste0(resloc1,"Kelp_Quarterly_CleanedFinal_NDT.Rds"))
+write.table(kelpNDT,file=paste0(resloc1,"Kelp_Quarterly_CleanedFinal_NDT.csv"),row.names = FALSE,col.names = FALSE,sep=",")
 
 saveRDS(kelp,file=paste0(resloc1,"Kelp_Quarterly_CleanedFinal.Rds"))
 write.table(kelp,file=paste0(resloc1,"Kelp_Quarterly_CleanedFinal.csv"),row.names = FALSE,col.names = FALSE,sep=",")

### 2022-01-31 Code/Kelp_AnalysisStep_MakeMapFigDat.R

@@ -1,5 +1,7 @@
 #Makes a map figure for the paper that orients the reader to the kelp data we used
 
+rm(list=ls())
+
 #***
 #External codes needed
 #***
@@ -22,12 +24,9 @@ if (!dir.exists(resloc))
 
 #load the data
 datloc<-"../Results/Kelp_DataAfterAllCleaning/"
-kelp<-readRDS(paste0(datloc,"Kelp_Quarterly_CleanedFinal.Rds")) 
-NO3<-readRDS(paste0(datloc,"NO3_Quarterly_CleanedFinal.Rds")) 
-waves<-readRDS(paste0(datloc,"Waves_Quarterly_CleanedFinal.Rds")) 
+kelp<-readRDS(paste0(datloc,"Kelp_Quarterly_CleanedFinal_NDT.Rds")) 
 locs<-readRDS(paste0(datloc,"Locs_CleanedFinal.Rds"))
 quarters<-readRDS(paste0(datloc,"Quarters_CleanedFinal.Rds"))
-climinds<-readRDS(paste0(datloc,"Climinds_Quarterly_CleanedFinal.Rds"))
 times<-1:(dim(quarters)[1])
 
 numts<-dim(kelp)[1]
@@ -127,4 +126,3 @@ saveRDS(CC1locs,file=paste0(resloc,"CC1locs.Rds"))
 saveRDS(CC2locs,file=paste0(resloc,"CC2locs.Rds"))
 saveRDS(SBlocs,file=paste0(resloc,"SBlocs.Rds"))
 
-#OK, MAX, TAKE IT FROM HERE...

### 2022-01-31 Code/main.R

@@ -6,26 +6,64 @@
 rm(list=ls())
 graphics.off()
 
+#***
+#checkpoint, to help ensure it continues to work into the future
+#***
+
+print("Running checkpoint, this may take a while the first time...")
+if (!dir.exists("./.checkpoint/")){
+  dir.create("./.checkpoint/")
+}
+checkpoint::checkpoint(snapshot_date="2022-01-01",r_version=getRversion(),checkpoint_location=getwd(),
+                       scan_now = TRUE) #mine is R-3.6.3
+#The version of the checkpoint package I used was installed from CRAN on 2022 01 31
+
+#***
 #Develop the theoretical case studies
+#***
+
+print("Theory case studies A and B...")
 source("Theory_CaseStudyAB.R")
+
+print("Theory case study C...")
 source("Theory_CaseStudyC.R")
+
+print("Theory example in intro...")
 source("Theory_IntroExample.R")
+
+print("Pedagogical figure...")
 source("Theory_PedagogFig.R")
 
-#Clean the kelp data
+#***
+#Clean the data
+#***
+
+print("Super basic data cleaning...")
 source("Kelp_AnalysisStep_CleanData_SuperBasic.R")
+
+print("More involved data cleaning...")
 source("Kelp_AnalysisStep_CleanData_MoreInvolved.R")
 
-#Plot kelp time series for visual inspection
+#print("Plot kelp time series for visual inspection...")
 #source("Kelp_AnalysisStep_PlotTimeSeries.R")
 
-#Final cleaning
+print("Final cleaning...")
 source("Kelp_AnalysisStep_CleanData_Final.R")
 
-#Make the map fig, to help orient the reader of the paper to the dataset
-source("Kelp_AnalysisStep_MakeMapFig.R")
+#***
+#Make the map figure, to help orient the reader of the paper to the dataset 
+#***
 
+print("Mapping the system...")
+source("Kelp_AnalysisStep_MakeMapFigDat.R") #preps the data for Max Castorani code
+#source("Kelp_AnalysisStep_MakeMapFig.R") #this is Max's code
+
+#***
 #Do the spectra decomposition analysis for kelp
+#***
+
+print("Spectral analysis...")
 source("Kelp_AnalysisStep_SpectralDecomp.R")
+print("Done!")
 
-
+checkpoint::uncheckpoint()

### 2022-01-31 Paper/Paper.Rnw
added an abstract, worked on the readme

@@
 \begin{abstract}
-Place abstract here
+\noindent Spatial synchrony, the tendency for populations across space to show correlated fluctuations, is a fundamental 
+feature of population dynamics, linked to central topics of ecology such as population cycling, extinction 
+risk, and ecosystem stability. 
+A common mechanism of spatial synchrony is the Moran effect, whereby spatially synchronized environmental signals 
+drive population dynamics and hence induce population synchrony. We here elaborate a general theory of how 
+Moran effects of different environmental drivers acting on the same populations can interact, either 
+synergistically or destructively, to produce either substantially more or markedly less population synchrony 
+than would otherwise occur. 
+We provide intuition for how this newly recognized mechanism works through theoretical case studies and 
+application of our theory to California populations of giant kelp.
+We argue that Moran interactions should be common.     
+Our general theory explains an important new aspect of a fundamental feature of spatiotemporal population dynamics. 
 \end{abstract}
@@
 
+\section{Acknowledgments}
+This study was partly funded by the U.S. National Science Foundation (NSF) through linked NSF-OCE awards 
+2023555, 2023523, 2140335, and 2023474 to M.C.N.C, K.C.C., T.W.B, and D.C.R., respectively; and by NSF-Math Bio award 
+1714195 to D.C.R.; and by support to D.C.R. from the James S. McDonnell Foundation and the California Department of Fish
+and Wildlife Delta Science Program. This project used data developed through the Santa Barbara Coastal 
+Long Term Ecological Research project, funded through NSF-OCE award 1831937. The authors thank Vadim Karatayev, 
+Maowei Liang, Kyle Emery, Nat Coombs, Adeola Adeboje and Ethan Kadiyala for helpful discussions.
+%***DAN: If you end up adding any of these people to the author list, take them off here.
+
 \bibliographystyle{ecology_letters2}
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{./Fig2.pdf}
 \caption{Kelp sampling sites and example time series. Sampling sites (a) were from three regions,

### 2022-01-31 Code/main.R

@@ -55,8 +55,8 @@ source("Kelp_AnalysisStep_CleanData_Final.R")
 #***
 
 print("Mapping the system...")
-source("Kelp_AnalysisStep_MakeMapFigDat.R") #preps the data for Max Castorani code
-#source("Kelp_AnalysisStep_MakeMapFig.R") #this is Max's code
+source("Kelp_AnalysisStep_MakeMapFigDat.R") #preps the data for Max Castorani code,
+#run separately on another machine to produce Fig. 2, which is purely descriptive
 
 #***
 #Do the spectra decomposition analysis for kelp

### 2022-02-02 Paper/Paper.Rnw
fiddling with formatting in the README
added Fig 2 pdf from Max and minor tweaks

@@
 \documentclass[letterpaper,11pt]{article}
+%\documentclass[letterpaper,12pt]{article}
 
@@
 \usepackage{graphicx}
-\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
+%\usepackage[left=2cm,top=2cm,right=2cm,bottom=2cm,head=.5cm,foot=.5cm]{geometry}
+\usepackage[left=1in,top=1in,right=1in,bottom=1in,head=.5cm,foot=.5cm]{geometry}
 \usepackage{url}
@@
 \usepackage{lineno}
+\usepackage{times}
 
@@
 \captionsetup[figure]{font+=linenumbers} % use \internallinenumbers inside figure caption
+
+%\doublespacing
 
@@
 Spatial synchrony, the tendency for geographically disjunct populations to show correlated fluctuations 
-through time, is a fundamental feature of population dynamics, linked, for instance, to important topics such as 
+through time, is a fundamental feature of population dynamics, linked to important topics such as 
 population cycling \citep{Anderson2021}, extinction risk \citep{Ghosh2020}, and the 
@@
 populations. Two recent papers have documented that Moran effects of distinct 
-environmental drivers can, in fact, interact, either syergistically or antagonistically,
+environmental drivers can, in fact, interact, either synergistically or antagonistically,
 so that total population synchrony can be either greater or less than if the distinct Moran 
@@
 in the seas around the United Kingdom is substantially augmented by interactions between environmental 
-drivers of synchrony and drivers linked with consumption by a copepod predator.
+drivers of synchrony and drivers linked with consumption by a copepod consumer.
 \cite{CastoraniSheppardKelp} showed that nutrient dynamics and wave disturbance, two Moran drivers
@@
 Maowei Liang, Kyle Emery, Nat Coombs, Adeola Adeboje and Ethan Kadiyala for helpful discussions.
-%***DAN: If you end up adding any of these people to the author list, take them off here.
+%***DAN: If you end up adding any of these people to the author list, take them off here. Also in the ack for the readme.
 
@@
 $l_{e2}$. For this setup, a positive interaction between Moran effects can occur on the
-given timescale when lags align, i.e, $l_{e1}-l_{e2}+l_n=0$, as in panel (a), so that environmental
+given timescale when lags align, i.e, $l_{e1}-l_{e2}+l_n=0$, as in panel (b), so that environmental
 effects on the population coincide and reinforce each other. All environmental effects on 
@@
 populations, e.g., $l_{e1}-l_{e2}+l_n$ equals half the given timescale/period,
-as in panel (b), so that the distinct influences tend to work against each other in 
+as in panel (c), so that the distinct influences tend to work against each other in 
 bringing about population oscillations on the given timescale.  

### 2022-11-07 Paper/Paper.Rnw
changes to the writeup from a long time ago that I forgot to commit. comming back to this for the first time in a while now.

@@
-\documentclass[letterpaper,11pt]{article}
-%\documentclass[letterpaper,12pt]{article}
+%\documentclass[letterpaper,11pt]{article}
+\documentclass[letterpaper,12pt]{article}
 
@@
 %header material for paper
-\title{How environmental drivers of synchrony interact}
+\title{Environmental drivers of synchrony and their interactions}
 \date{}
@@
 %How environmental drivers of spatial synchrony interact
+%how environmental drivers of synchrony interact
+%Interactions between environmental drivers of spatial synchrony
+%On interactions between environmental drivers of spatial synchrony
+%Interactions between environmental drivers of synchrony
+%On interactions between environmental drivers of synchrony
+%A general theory for interactions between environmental drivers of synchrony
+%A general theory on how environmental drivers of synchrony interact
+%A general theory on environmental drivers of synchrony and their interactions
+%Environmental drivers of synchrony and their interactions
 
@@
 1000 km \citep{Liebhold2004},
-recent advances in statistical methods and improvements in data availability have lead to several 
+recent advances in statistical methods and improvements in data availability have led to several 
 major advances in our understanding of synchrony and its causes and consequences. 
@@
 for $a \neq b$. Cross synchrony measures represent synchrony between an environmental variable in one location and 
-a different variable in another location. Fig. \ref{fig:IntroTheory}a shows how population synchrony can 
+a different variable in another location. Fig. \ref{fig:IntroTheory} illustrates how population synchrony can 
 therefore differ from what synchrony would be if the environmental processes $\epsilon^{(1)}$ and 
@@
 from SI section S1 of \cite{Sheppard2019}.
-
-Having demonstrated that interactions between Moran effects can add to or reduce population synchrony, we now 
+%***DAN: If we decide to go longer, we could add some brief commentary here about how common it should be for 
+%spatially extended population systems to be influenced by more than one env var. Maybe also mention these
+%env vars may be different vars measured at the same time of year, or the same or different vars measured
+%at different times of year, i.e., separated phenologically. Keep it brief, it's just intended to make them 
+%this is is going to be important and common, and to make them sorta wonder why this has not been examined already.
+%Actually that second thing is done in the Discussion, either don't do it here or else if you do, then take it
+%out of there.
+
+%***DAN: Another potentially good Discussion para: make connections to the work (or Wang and Loreau and
+%others) that is fusing spatial and community ecology. What are the implications of this work for that work?
+%Making such additions may help it appeal to a broader audience, which is something we need to work on as 
+%much as we can for Ecol Lett.
+
+Having demonstrated that interactions between Moran effects can augment or reduce population synchrony, we now 
 begin to provide some intuition on how to understand the underlying mechanism. 
@@
 and the nature of interacting Moran effects can be understood on a timescale-by-timescale basis, with effects summed 
-across timescales afterwards if desired. Fig. \ref{fig:IntroTheory}b shows the period-20 Fourier components of two hypothetical
-environmental signals measured in sampling location $i$, and one way these can influence populations. The components are
+across timescales afterwards if desired. Fig. \ref{fig:PedagogAlt}a,b shows the period-20 Fourier components of two hypothetical,
+spatially synchronous environmental variables measured in sampling locations 
+$i=1,2,3$, and one way these can influence populations. The components are
 lagged, relative to each other, in the timing of their peaks ($l_n$ on the figure). They are also lagged in their 
@@
 manifest as maximum positive influence on populations after delays of $l_{e1}$ and $l_{e2}$, respectively. In this example,
-we assume for simiplicity that larger values of both environmental variables are beneficial to populations, though the general
-theory described below does require that assumption. If, as in Fig. \ref{fig:IntroTheory}b, $l_{e1}-l_{e2}+l_n=0$, then the 
-periodic maximal positive influences of the environmental components coincide. If environmental signals are then spatially synchronous
-across sampling locations, the fact of the alignment of lags produces additional synchrony in populations, 
+we assume for simplicity that larger values of both environmental variables are beneficial to populations, though the general
+theory described below does not require that assumption. If, as in Fig. \ref{fig:PedagogAlt}a, b, $l_n+l_{e1}=l_{e2}$, then the 
+periodic maximal positive influences of the two environmental variables coincide with each other as well as being 
+spatial synchronous. This alignment of produces additional synchrony in populations, 
 beyond what would manifest if environmental fluctuations were unrelated, because positive influences of variable 1 in location $i$ will
-tend to coincide with positive influences of variable 2 in location $j$. In contrast, Fig. \ref{fig:IntroTheory}c
-shows the opposite scenario, for which $l_{e1}-l_{e2}+l_n$ equals half the period, $10$, so that  
+tend to coincide with positive influences of variable 2 in location $j$. In contrast, Fig. \ref{fig:PedagogAlt}c, d
+shows the opposite scenario, for which $l_n+l_{e1}=l_{e2}+\sigma/2$, where $\sigma=20$ is the period. Thus
 the periodic maximal positive influence of environmental variable 1 coincides with the periodic maximal negative
@@
 for kelp ecology, we instead use a simplified subset of the available kelp data, and our analysis is instead intended 
-to illuminate the inner workings of our theory and of the general mechanism we describe. In the Discussion, we 
+to illuminate the inner workings of the general mechanism we describe. In the Discussion, we 
 extend the arguments of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} that interactions between Moran 
@@
 ecosystems. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
-illuminate this newly observed but potentially quite common and important mechanism of spatiotemporal population dynamics.
+illuminate this newly observed but potentially quite important mechanism of spatiotemporal population dynamics.
 
@@
 of a species of interest in that location), the spectrum $S_{w_i w_i}(f)$ is a function
-of frequency, $f$. For a periodic ocillation, $f$ can be defined as one over the timescale, or period, $\sigma$, of the oscillation. 
+of frequency, $f$. For a periodic ocillation, $f$ can be defined as one over 
+the timescale, or period, $\sigma$, of the oscillation. 
 The spectrum $S_{w_i w_i}(f)$ is larger for frequencies at which $w_i(t)$ oscillates more. So, for example, a 
@@
 $S_{w_i w_i}(f)$ for $f=\frac{1}{5} \text{ y}^{-1}$. The spectrum separates the overall variance of a time series into
-contributions which occur on different frequencies, in the sense that an integral of $S_{w_i w_i}(f)$ across all 
+contributions which occur at different frequencies, in the sense that an integral of $S_{w_i w_i}(f)$ across all 
 frequencies equals $\var(w_i)$. In a similar way, the cospectrum of two time series, $w_i(t)$ and $w_j(t)$,
@@
 and the complex phase of the cross spectrum at $f$ then quantifies
-that phase. The notation $S_{ww}(f)$ refers to the \emph{spectral matrix}, 
-which has $ij^{\text{th}}$ entry $S_{w_i w_j}(f)$. Spectral methods are standard tools 
+that relationship. The notation $S_{ww}(f)$ refers to the \emph{spectral matrix}, 
+which has $ij^{\text{th}}$ entry $S_{w_i w_j}(f)$. Spectral methods are standard  
 [e.g., \cite{Vasseur2007,Reuman2008,Defriez2017}], and many background references are available 
@@
 approach is essential to the development of our new theory. 
-We therefore here define, using the spectral methods outlined above, a frequency/timescale-specific measure of spatial synchrony, 
-as well as a new concept of \emph{cross-variable spatial synchrony}. If time series data $w_i(t)$ for $t=1,\ldots,T$ 
-were gathered at locations $i=1,\ldots,N$, our spatial synchrony measure is simply
+We therefore here define, using the spectral methods outlined above, a frequency/timescale-specific measure of synchrony, 
+as well as a new concept of \emph{cross-variable synchrony}. If time series data $w_i(t)$ for $t=1,\ldots,T$ 
+were gathered at locations $i=1,\ldots,N$, our synchrony measure is simply
 $\rho_{ww}=\left( \sum_{i \neq j} S_{w_i w_j} \right)/(N^2-N)$, the average of the cross spectra for all pairs
@@
 ($t=1,\ldots,T$ and $i=1,\ldots,N$) were measured at each sampling location (e.g., two environmental variables),
-cross-variable spatial synchrony (or, simply, \emph{cross synchrony}) between the variables
+cross-variable synchrony (or, simply, \emph{cross synchrony}) between the variables
 is $\rho_{\epsilon^{(1)}\epsilon^{(2)}}=\left( \sum_{i \neq j} S_{\epsilon_i^{(1)}\epsilon_j^{(2)}} \right)/(N^2-N)$. 
@@
 by a year, then $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$, which is complex valued, will have high magnitude at
-timescale $4$y, equivalent to frequency $f=\frac{1}{4} \text{ y}^{-1}$, and will have  phase at that frequency equal to $\pi/2$, reflecting the one-year 
-lag between the variables.
+timescale $4$y, equivalent to frequency $f=\frac{1}{4} \text{ y}^{-1}$, and will have  
+phase at that frequency equal to $\pi/2$, reflecting the one-year lag.
 See SI section \ref{SM-sect:examples_cross_sync} for further examples of the new cross synchrony index.
+%***DAN: If we go longer, you could bring these examples here and it would probably be a positive addition of
+%about one para and one new Fig. Would be a positive addition because it helps people understand our new metric.
+
+%***DAN: If you go longer, you might think about adding a para to Discussion about how general theories can bind together
+%multiple effects that are often understood idiosyncratically or as special cases by ecologists. Not sure this could be done 
+%effectively, but I'm always reading papers testing specific things and understanding phenomena in ways like "A can happen
+%and B can also happen". For instance, X (which is a good paper) says "We show that environmental fluctuations can reduce 
+%synchrony between species and increase stability, yet they can also decouple asynchrony between species and increase 
+%population and metapopulation variability." (https://esajournals-onlinelibrary-wiley-com.www2.lib.ku.edu/doi/pdfdirect/10.1002/ecm.1484) 
+%Sounds like they are saying both things can occur, we want a general theory
+%to understand when each outcome happens. Likewise could cite the old debate on red noise and effects on exitinction risk. 
+%There were papers show conflicting results until someone looked for the general theory and then it was resolved. Ecologists
+%have a natural proclivity for specific examples, and like results like "A can happen" but as scientists we ought to be 
+%seeking general theories. Not sure how to say all this in a constructive and useful as opposed to an annoying way. 
+%The assumptions that your dynamical models have an equilibrium and then an approach that assumes weak noise and linearizes 
+%about that equilibrium can provide a general theory (as we did here, but also cite several papers of myself and others using 
+%this appraoch). Whereas the theory is not completely general, because it does not cover strong noise and strong nonlinearities,
+%which do happen in ecology, we consider it better to produce a a general theory based on linearization and then examine
+%departures from it under strong noise than it is to use a series of scattershot simulation approaches, which has more than 
+%led to conflicting results in the literature which then had to be resolved. In our opinion, generality should be pursued in
+%theoretical ecological contexts rather sooner than it is our impression that it is typically pursued, wrking on it 
+%early on instead of only after several case-study-type studies based on simulations. The weak-noise linearization approach
+%is widely applicable and often more fruitful than simulations (REFS).
+
+%***DAN: Potential reviewers or editors:
+%Maybe Holt as editor if we go to Ecography
+%Gonzalez
+%Loreau?
+%Bjornstad
+%Holt's postdoc, name slipping my mind
+%Barfield? 
+%Holt as reviewer? Will he even do it?
+%Cohen? Probably a conflict.
+%Arnoldi? Bart Haegeman?
+%Montoya?
+
+%***DAN: Question for later: Suppose we considered a model with three measured env drivers, would you get
+%three-way interactions? You might add the question to the Discussion para about adding dispersal to this model.
+%Not sure actually doing that is a good use of my time, though.
 
@@
 Fulfilling goal G1 of the Introduction (i.e., to elaborate theory of interacting Moran effects), 
-the outcome of our theory is an equation which expresses population spatial synchrony 
+the outcome of our theory is an equation which expresses population synchrony 
 in terms of synchrony of the environmenal processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$, and cross
@@
 $\mu=\exp(-2 \pi \iota f)$, and where $\iota$, the Greek letter iota, represents the imaginary unit.
-The first term on the right of (\ref{rhoww_T1to3}) is the component of population spatial synchrony 
+The first term on the right of (\ref{rhoww_T1to3}) is the component of population synchrony 
 due to the direct Moran effects of $\epsilon^{(1)}$, the second term is 
@@
 influence of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ on populations at the timescale $\sigma=1/f$. 
-The phases of these quantities correspond to the delays in the influences of 
+The phases of these quantities encode the lags in the influences of 
 $\epsilon^{(1)}$ and $\epsilon^{(2)}$, 
-represented on Fig. \ref{fig:IntroTheory} by $l_{e1}$ and $l_{e2}$, respectively. The lag 
-$l_n$ on Fig. \ref{fig:IntroTheory} manifests in the theory through the phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$. 
-The expression $l_{e1}-l_{e2}+l_n$ from the Introduction and Fig. \ref{fig:IntroTheory} corresponds to the phase of the expression 
+represented on Fig. \ref{fig:PedagogAlt} by $l_{e1}$ and $l_{e2}$,
+relative to the timescale $\sigma$. The lag 
+$l_n$ on Fig. \ref{fig:PedagogAlt} manifests in the theory 
+through the phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$. 
+The expression $l_{e1}-l_{e2}+l_{n}$, which quantifies the degree of mismatch among the lags
+(see Introduction and Fig. \ref{fig:PedagogAlt}) corresponds to the phase of the expression 
 $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, because the phase of
-this product equals the phase of $f_{\mcP^{(1)}}$ (equivalent to $l_{e1}$), minus the phase of $f_{\mcP^{(2)}}$
-(equivalent to $l_{e2}$), plus the phase
-of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ (equivalent to $l_n$). 
+this product equals the phase of $f_{\mcP^{(1)}}$ (which corresponds to $l_{e1}$), minus 
+the phase of $f_{\mcP^{(2)}}$
+(which corresponds to $l_{e2}$), plus the phase
+of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ (which corresponds to $l_n$). 
 The real part of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
 is positive and larger whenever the phase of this quantity is close to zero, i.e., when    
-$l_{e1}-l_{e2}+l_n$ is close to zero, and is increasingly negative as 
+$l_{e1}-l_{e2}+l_n$ is close to zero (Fig. \ref{fig:PedagogAlt}a, b), and is increasingly negative as 
 the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ gets
-close to $\pi$, e.g., when $l_{e1}-l_{e2}+l_n$ is close to half the period of oscillation.
-The factor 
-$1/|f_{\mcB}|^2$ is how the intrinsic nature of population
+close to $\pi$, e.g., when $l_{e1}-l_{e2}+l_n$ is close to half the period of oscillation
+(Fig. \ref{fig:PedagogAlt}c, d). The factor $1/|f_{\mcB}|^2$ is how the intrinsic nature of population
 dynamics modulates Moran influences.
@@
 \noindent We now describe how we pursue goal G2 of the Introduction, to develop 
-theoretical case studies A-C, that illuminate the intuition of interacting Moran effects. 
+three theoretical case studies, that illuminate the intuition of interacting Moran effects. 
 For all cases, the model time step was assumed to be one quarter (q), i.e., 
@@
 facilitate later comparisons with results for kelp data, which were sampled quarterly. 
-For case study A (henceforth CaA),  the environmental variable 
+For case study A (henceforth CaseA),  the environmental variable 
 $\epsilon^{(1)}$ is assumed to have a simple positive effect on populations, but 
@@
 $p_1^{(1)}>0$, and $p_i^{(1)}=0$ for $i \neq 1$. 
-For case study B (henceforth CaB), which generalizes A, $\epsilon^{(1)}$ is again assumed to have a simple 
+For case study B (henceforth CaseB), which generalizes A, $\epsilon^{(1)}$ is again assumed to have a simple 
 positive effect on populations,
 but now lagged by $n$ time steps, so that $p_n^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n$.
-For both CaA and CaB, the effects of $\epsilon^{(2)}$
+For both CaseA and CaseB, the effects of $\epsilon^{(2)}$
 are assumed to be un-lagged, i.e., $p_0^{(2)} \neq 0$ and $p_i^{(2)}=0$ for $i>0$.
-For case study C (CaC), the effects of $\epsilon^{(1)}$ are lagged by $n_1$ and the 
+For case study C (CaseC), the effects of $\epsilon^{(1)}$ are lagged by $n_1$ and the 
 effects of $\epsilon^{(2)}$ are lagged by $n_2$ time steps, i.e.,
@@
 $(\epsilon^{(1)},\epsilon^{(2)})$ is assumed to be a Gaussian white-noise process 
-for both CaA and CaB. So 
+for both CaseA and CaseB. So 
 the random variables $(\epsilon^{(1)}(t),\epsilon^{(2)}(t))$ are independent and 
@@
 and the components of $\epsilon^{(1)}(t)$ were positively correlated with those of $\epsilon^{(2)}(t)$.
-For CaC, the noise processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
+For CaseC, the noise processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ are
 each assumed to exhibit spatially synchronous periodic oscillations of period one year, i.e.,
 four model time steps, but with different phases. Peaks in $\epsilon^{(1)}$ are assumed to 
-either lead or lag (we consider two scenarios, CaC1 and CaC2) peaks in $\epsilon^{(2)}$ by $1$q.
-Such a scenario could be realized by annually 
+either lead or lag peaks in $\epsilon^{(2)}$ by $1$q (we consider two scenarios, CaseC1 and CaseC2).
+Such scenarios could be realized by annually 
 periodic environmental fluctuations sampled quarterly,
@@
 compared to the wave peak. 
-Details of the noise processes are specified in SI section \ref{SM-sect:examples_cross_sync}.
+Details of the noise processes are in SI section \ref{SM-sect:examples_cross_sync}.
 For all case studies,
@@
 %Description of data after cleaning
-The core of this paper is the new theory, but to help illustrate that theory, for goal G3 of the 
+The core of this paper is the theory, but to help illustrate that theory, for goal G3 of the 
 Introduction, we also apply
@@
 The data are based on a subset of those used by \cite{CastoraniSheppardKelp} and
-\cite{Walter2021}, and data details are given in those papers and in SI section \ref{SM-sect:data}. We here
-summarize the format of the data after a preparation and cleaning process was implemented. 
+\cite{Walter2021}, and data details are given in those papers and in SI section \ref{SM-sect:data};
 \cite{CastoraniSheppardKelp} and \cite{Walter2021} also provide an introduction to 
 kelp ecology, and information on why giant kelp is an excellent species 
-for studies of synchrony. After preparation, kelp data consisted of
+for studies of synchrony. We here summarize the format of the data after a 
+preparation and cleaning process was implemented. After preparation, kelp data consisted of
 $\Sexpr{length(SBlocstouse)+length(CC1locstouse)+length(CC2locstouse)}$ quarterly kelp abundance time series
@@
 mean quarterly kelp biomass (kg wet) per unit 
-useable habitat ($\text{m}^2$) in a 500m stretch of coastline. We used coastline segments where kelp was persistent
+useable habitat ($\text{m}^2$) along a 500m stretch of coastline. We used coastline segments where kelp was persistent
 through essentially all of the $\Sexpr{quarters[1,2]}$-$\Sexpr{quarters[dim(quarters)[1],2]}$ 
@@
 of maximum wave height and mean nitrate concentration for each quarter and location. Both waves
-and nutrients are known to influence kelp dynamics and synchrony 
+and nutrients influence kelp dynamics and synchrony 
 \citep{Cavanaugh2013,CastoraniSheppardKelp}.
@@
 synchrony. Second, in contrast to direct Moran effects, which are positive, 
-interactions can be negative or positive. CaA and CaB
+interactions can be negative or positive. CaseA and CaseB
 showed negative interaction effects on short timescales (Fig. \ref{fig:TheoryCaseA}a,e); 
-CaC1 showed negative interactions on
-long timescales (Fig. \ref{fig:TheoryCaseA}i,m); and CaC2 showed negative interactions on all timescales 
+CaseC1 showed negative interactions on
+long timescales (Fig. \ref{fig:TheoryCaseA}i); and CaseC2 showed negative interactions on all timescales 
 (Fig. \ref{fig:TheoryCaseA}m).
 Thus interaction effects can either augment or reduce synchrony. Finally, interaction effects depended strongly on timescale 
-for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and even differed in sign across timescales for CaA, CaB and CaC1. 
+for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and even differed in sign across timescales for CaseA, CaseB and CaseC1. 
 This result complements earlier studies that showed direct Moran effects can 
@@
 %Case study A, new try
-CaA helps provide an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
+CaseA helps provide an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
 Moran effects on different timescales, and how our theory captures that contrast.
-For CaA, $\epsilon^{(1)}$ effects on populations were lagged by $1$q but 
-$\epsilon^{(2)}$ effects were unlagged (Methods). So, in the language of Fig. \ref{fig:IntroTheory},
+For CaseA, $\epsilon^{(1)}$ effects on populations were lagged by $1$q but 
+$\epsilon^{(2)}$ effects were unlagged (Methods). So, in the language of Fig. \ref{fig:PedagogAlt},
 $l_{e1}=1$q and $l_{e2}=0$q. 
-The between-noise lag, $l_n$, of Fig. \ref{fig:IntroTheory} is $0$q because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
+The between-noise lag, $l_n$, of Fig. \ref{fig:PedagogAlt} is $0$q because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
 taken to be a white noise process (see SI section \ref{SM-sect:examples_cross_sync} for details, here).
@@
 p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}), 
-which has phase $-\pi$ at short timescales, $\sigma \approx 2 \text{q}$,
+which has phase $-\pi$ at short timescales
 and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b,c).
@@
 interaction term in (\ref{rhoww_T1to3}), as described in General Theory. Comparing
-Fig. \ref{fig:TheoryCaseA}a,d shows the additional influence of autoregressive population effects, which
+Fig. \ref{fig:TheoryCaseA}a and d shows the additional influence of autoregressive population effects, which
 simply multiply all Moran influences by the same timescale-dependent non-negative 
@@
 
-% %Case study A, old try
-% CaA helps provide an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
-% Moran effects on different timescales, and how our theory captures that contrast.
-% For CaA, for which $\epsilon^{(1)}$ effects on populations were lagged by 1 quarter but 
-% $\epsilon^{(2)}$ effects were unlagged (Methods), interactions between Moran effects were negative on the shortest timescales 
-% (0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}a, left side of panel). This is because the 
-% population influences of Moran effects of timescale 2
-% quarters and lagged by 1 quarter are exactly out of phase with unlagged Moran effects of timescale
-% 2 quarters, producing destructive interference. Thus, on short timescales, total
-% synchrony will be less, for this example, than it would be if only direct Moran effects operated. In contrast, 
-% interaction effects on long timescales were synergistic, i.e., positive, for CaA 
-% (Fig. \ref{fig:TheoryCaseA}a, right side of panel) 
-% because, on long timescales, a lag of 1 quarter does not meaningfully alter the phase relationship between 
-% Moran drivers, and drivers are therefore almost in phase, reinforcing each other.
-% At an intermediate timescale of 1 year, a lag of 1 quarter causes the Moran effects
-% of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be a quarter phase out of alignment (1 quarter is $1/4$ of a 1-year 
-% timescale), resulting in 
-% zero interaction effects, essentially because sine waves a quarter-phase out of alignment
-% with each other are perfectly uncorrelated. Thus interaction effects between Moran drivers were zero 
-% at timescale 1 year (Fig. \ref{fig:TheoryCaseA}a). Theory captures these complexities because
-% the size of the 1-quarter phase lag of this scenario, relative to a given timescale,
-% is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) = 
-% p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}), 
-% which has phase $-\pi$ at short timescales, $\sigma \approx 0.5 \text{y}$,
-% and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b). Because 
-% $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ has phase $0$ for all timescales (SI sections \ref{SM-sect:examples_cross_sync}, 
-% \ref{SM-sect:details_case_studies}), the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ is the same as the phase of
-% $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}c),
-% which in turn controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, which is the
-% interaction term in our theory (\ref{rhoww_T1to3}). Thus interactions shift 
-% from negative, for short timescales, to positive, for long timescales. Comparing
-% Fig. \ref{fig:TheoryCaseA}a,d shows the additional influence of autoregressive population effects, which
-% simply multiply all Moran influences by the same timescale-dependent non-negative 
-% quantity without altering the relative importance or sign of the direct and indirect Moran effects [see (\ref{rhoww_T1to3})].
-
 %Case study B, new version
-CaB reveals how Moran effects can interact when lags are longer than 1 model time step. 
-For CaB, $\epsilon^{(1)}$ effects on populations were lagged by $n$q but 
+CaseB reveals how Moran effects can interact when lags are longer than 1 model time step. 
+For CaseB, $\epsilon^{(1)}$ effects on populations were lagged by $n$q but 
 $\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$, so $l_{e1}=3$q and $l_{e2}=0$q. 
-As for CaA, because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
-a white noise process for CaB, $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ was a real number and $l_n=0$.
+As for CaseA, because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
+a white noise process for CaseB, $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ was a real number and $l_n=0$.
 Interaction effects were again negative on short timescales 
@@
 counteracted each other.
-Similar to CaA, $l_{e1}-l_{e2}+l_n=3$q was again negligible compared to long timescales, so interaction effects
+Similar to CaseA, $l_{e1}-l_{e2}+l_n=3$q was again negligible compared to long timescales, so interaction effects
 were positive on long timescales (Fig. \ref{fig:TheoryCaseA}e, right side of panel).
@@
 (Fig. \ref{fig:TheoryCaseA}e). Theory captures these complexities because 
-the size of the $3$q lag of this scenario, relative to $\sigma$,
-is again captured by the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f,g).
-
-% %Case study B, old version
-% CaB reveals how Moran effects can interact when lags are longer than 1 model time step. 
-% For B, $\epsilon^{(1)}$ effects on populations were lagged by $n$ quarters but 
-% $\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$. Interactions between Moran effects were 
-% negative on the shortest timescales 
-% (0.5 years = 2 quarters; Fig. \ref{fig:TheoryCaseA}e, far left side of panel) because the 
-% population influences of Moran effects of timescale 2
-% quarters and lagged by 3 quarters are exactly out of phase with unlagged Moran effects of timescale
-% 2 quarters, producing negative interactions. 
-% Interaction effects on long timescales were instead synergistic, i.e., positive, for B 
-% (Fig. \ref{fig:TheoryCaseA}e, right side of panel), for the same reason as CaA (see above).
-% At timescale 3 quarters, Moran effects are reinforcing because the lag equals the timescale,
-% producing Moran effects of the two drivers which are in phase 
-% (Fig. \ref{fig:TheoryCaseA}e, peak at timescale 3 quarters = 0.75 years). 
-% At a timescale of 1 year, a lag of 3 quarters causes the Moran effects of 
-% $\epsilon^{(1)}$ and $\epsilon^{(2)}$ to be $3/4$-phase out of alignment, resulting in zero interaction effects. 
-% Likewise at a timescale of 3 years (12 quarters), a lag of 3 quarters 
-% makes the Moran effects of 
-% $\epsilon^{(1)}$ and $\epsilon^{(2)}$ be $1/4$-phase out of alignment, again producing zero interaction. 
-% Finally, at timescale 1.5 years (6 quarters), the 3-quarter lag equals half the timescale, so 
-% $\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects counteract each other, producing negative 
-% interactions. Thus interactions between Moran drivers were zero 
-% at timescales 1 and 3 years for B, were 
-% negative for short timescales and for timescale 1.5 years, and were positive for 
-% timescale 0.75 years and for long timescales (Fig. \ref{fig:TheoryCaseA}e).
-% Theory captures these complexities because 
-% the size of the 3-quarter phase lag of this scenario, relative to a timescale,
-% is captured by $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ (Fig. \ref{fig:TheoryCaseA}f).
-% Because $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number for CaB (SI section \ref{SM-sect:details_case_studies}),
-% the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ is the same as that of
-% $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ (Fig. \ref{fig:TheoryCaseA}g),
-% which controls the sign of the interaction term,
-% $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, in our 
-% theory (\ref{rhoww_T1to3}).
+the size of the $3$q sum of lags for this scenario, relative to $\sigma$,
+is again captured by the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$; the phase of that
+quantity, represented by colors on Fig. \ref{fig:TheoryCaseA}f,g, fluctuates analogously to the
+interaction effects on Fig. \ref{fig:TheoryCaseA}a.
 
 %Case study C, new
-CaA and CaB assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-and $l_n=0$.
-But CaC illustrates what can happen when noise processes have periodic components and lagged interdependences, e.g., 
+CaseA and CaseB assumed white noise processes, and therefore 
+real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ and $l_n=0$.
+But CaseC illustrates what can happen when noise processes have periodic components and lagged interdependences, e.g., 
 both processes oscillate with annual periodicity and distinct phenology, an important scenario because seasonality is common.
-For CaC, $\epsilon^{(1)}$ effects on populations were lagged by $n_1$q and 
+For CaseC, $\epsilon^{(1)}$ effects on populations were lagged by $n_1$q and 
 $\epsilon^{(2)}$ effects were lagged by $n_2$q (Methods). We used $n_1=1$, $n_2=0$, $p_1^{(1)}=3$ and $p_0^{(2)}=-1.5$, so 
-$\epsilon^{(1)}$ effects were lagged and positive, and $\epsilon^{(2)}$ effects were immediate and negative. The language and logic of 
-Fig. \ref{fig:IntroTheory} does not directly apply, because we there assumed, in order to introduce concepts simply,
-that environmental effects were positive, and we now take $p_0^{(2)}=-1.5$. But similar logic does apply, with
-both the current and Fig. \ref{fig:IntroTheory} paradigms covered by our general theory.
-For CaC1 (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
+$\epsilon^{(1)}$ effects were lagged and positive, and $\epsilon^{(2)}$ effects 
+were immediate and negative. The language and logic of 
+Fig. \ref{fig:PedagogAlt} does not directly apply, because we there assumed, in order to introduce concepts simply,
+that environmental effects were positive, and we now take $p_0^{(2)}=-1.5$. But similar logic does apply 
+(Fig. \ref{fig:PedagogAlt_SM}), with
+both the new and previous paradigms covered by our general theory.
+For CaseC1 (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
 noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
-$1$q (Methods, SI section \ref{SM-sect:examples_cross_sync}). Thus lags compounded: the annual positive population effects of $\epsilon^{(1)}$ 
+$1$q (Methods, SI section \ref{SM-sect:examples_cross_sync}). Thus lags compounded: the 
+annual positive population effects of $\epsilon^{(1)}$ 
 were exactly $2$q offset from the annual negative effects of $\epsilon^{(2)}$, 
@@
 between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i).
-Contrastingly, for CaC2 (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
+Contrastingly, for CaseC2 (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
 noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
@@
 $\epsilon^{(2)}$ compared to a given timescale, while 
-$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the degree of aslignment between 
+$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the degree of alignment between 
 oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
+%***DAN: You might even consider a new panel or two somehow showing how the theory captures these effects more directly than 
+%Fig. 3 currently shows, though not sure exactly how that would be done or if it's actually a good idea.
 The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
@@
 variables align or misalign, with alignment
-reflected in the real part of this quantity, $\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$,
-which is half the interaction term in our theory (\ref{rhoww_T1to3}).
+reflected in the real part of this quantity, which is half the interaction term in our theory (\ref{rhoww_T1to3}).
 On long timescales, interactions were the
@@
 
-% %Case study C, old
-% CaA and CaB assumed white noise processes, and therefore real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-% (SI sections \ref{SM-sect:examples_cross_sync}, \ref{SM-sect:details_case_studies}),
-% but CaC illustrates what can happen when noise processes have periodic components, e.g., 
-% both processes oscillate with annual periodicity, an important scenario because seasonality is common. 
-% What matters then is whether and how the lag between the two seasonal noise processes aligns with the difference 
-% between the lags of their effects on populations, as we now explain. 
-% For the scenarios we considered under CaC (Fig. \ref{fig:TheoryCaseA}i-p), 
-% $\epsilon^{(1)}$ effects on populations were positive, and were 
-% lagged (i.e., delayed in their influence) by 1 quarter; but 
-% $\epsilon^{(2)}$ effects were immediate and negative (Methods and Fig. \ref{fig:TheoryCaseA} caption).
-% For CaC1 (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
-% noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
-% 1 quarter. Thus lags compounded: the annual positive population effects of $\epsilon^{(1)}$ 
-% were exactly 2 quarters offset from the annual negative effects of $\epsilon^{(2)}$, 
-% producing reinforcing interactions 
-% between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i), and more population
-% synchrony on annual timescales than would have been expected if Moran effects operated independently. Contrastingly,
-% for CaC2 (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
-% noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
-% 1 quarter. Thus lags cancelled: the annual positive population effects of $\epsilon^{(1)}$
-% coincided with the negative effects of $\epsilon^{(2)}$ every year, a half-phase shift for an 
-% annually oscillating effect, which corresponds to negative interactions between Moran influences
-% and less population synchrony, on annual timescales, than would have been expected if Moran effects operated independently 
-% (Fig. \ref{fig:TheoryCaseA}m). The only difference between CaC1 and CaC2 
-% is the relative timing of annual peaks in $\epsilon^{(1)}$ and $\epsilon^{(2)}$.
-% Our theory captures these complexities
-% because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
-% quantifies the relative phase lag between the effects of $\epsilon^{(1)}$ and 
-% $\epsilon^{(2)}$ compared to a given timescale, while 
-% $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the phase lag between 
-% oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
-% The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-% therefore captures whether relative lags of effects and lags between 
-% the variables themselves align or misalign, with the degree of alignment
-% reflected in the real part of this quantity, $\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$,
-% which is half the interaction term in our theory (\ref{rhoww_T1to3}).
-% On long timescales, interactions were the
-% same in the two scenarios because, on those timescales, sub-annual lags 
-% make negligible difference. Interactions were then negative 
-% because the effects of $\epsilon^{(1)}$ on populations were 
-% positive and those of $\epsilon^{(2)}$ were negative.
-
 %Paralels the first para, but for kelp
@@
 (above text and Fig. \ref{fig:TheoryCaseA}a,e,i,m),
-that interaction effects between Moran drivers: 1) can be comparble in strength to direct Moran
+that interaction effects between Moran drivers: 1) can be comparable in strength to direct Moran
 effects; 2) can be either synergistic or destructive; and 3) can 
@@
 For instance, the direct Moran effect
-term in (\ref{rhoww_T1to3}) for nitrates, i.e., 
+term in (\ref{rhoww_T1to3}) for nitrates, 
 $|f_{\mcP^{(1)}}|^2 \rho_{\epsilon^{(1)} \epsilon^{(1)}}/|f_{\mcB}|^2$, 
-equals the synchrony of the nitrate time series themselves, i.e.,
+equals the synchrony of the nitrate time series themselves, 
 $\rho_{\epsilon^{(1)} \epsilon^{(1)}}$, times the 
 timescale-specific strength of influence of nitrates on kelp, 
-i.e., $|f_{\mcP^{(1)}}|^2$, and modulated
+$|f_{\mcP^{(1)}}|^2$, and modulated
 by the autoregressive nature of kelp dynamics,
@@
 %Interactions between Moran effects, CCA1
-Interactions between Moran effects in CCA1 were parallel to one of the theoretical case studies, CaC1.
+Interactions between Moran effects in CCA1 were parallel to one of the theoretical case studies, CaseC1.
 In CCA1, nitrates had a $1$q-delayed positive effect on observed 
 kelp populations and waves had immediate negative effect \citep{Cavanaugh2011,Bell2015,Schiel2015}, 
-just as $\epsilon^{(1)}$ effects in CaC 
+just as $\epsilon^{(1)}$ effects in CaseC 
 were positive and delayed by $1$q and $\epsilon^{(2)}$ effects were negative and immediate.
 Delayed effects are reflected, for CCA1, in Fig. \ref{fig:indirectMoran}a, and for 
-CaC1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
+CaseC1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
 that $f_{\mcP^{(1)}}$ has phase about $-\pi/2$ at the annual timescale. The delay is also 
@@
 Immediate negative effects of waves in CCA1 are reflected in Fig. \ref{fig:indirectMoran}b,g,l
-and for CaC1 in Fig. \ref{fig:TheoryCaseA}j, both of which show 
+and for CaseC1 in Fig. \ref{fig:TheoryCaseA}j, both of which show 
 that $f_{\mcP^{(2)}}$ has phase $-\pi$ at all timescales. See also the $p_0^{(2)}$
@@
 wave peaks by $1$q, just as $\epsilon^{(1)}$ peaks lagged $\epsilon^{(2)}$
-peaks by $1$q in CaC1.
-This is reflected, for CCA1, in Fig. \ref{fig:indirectMoran}d, and for CaC1 
+peaks by $1$q in CaseC1.
+This is reflected, for CCA1, in Fig. \ref{fig:indirectMoran}d, and for CaseC1 
 in Fig. \ref{fig:TheoryCaseA}k, both of which show that $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$
@@
 producing reinforcing interactions between Moran effects. Likewise, the $\pi/2$ phase 
-of $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}$ on annual timescales (Fig. \ref{fig:indirectMoran}c for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaC1)
+of $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}$ on annual timescales (Fig. \ref{fig:indirectMoran}c for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1)
 approximately cancels the $-\pi/2$ phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ on annual timescales
-(Fig. \ref{fig:indirectMoran}d for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaC1), so that  
+(Fig. \ref{fig:indirectMoran}d for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1), so that  
 $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ has phase 
@@
 on annual timescales for both the theoretical and emprical cases 
-(Fig. \ref{fig:indirectMoran}e for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaC1). 
+(Fig. \ref{fig:indirectMoran}e for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1). 
 Interactions between Moran effects on long-timescales ($>8$y) were also similar for both CCA1 and
-CaC1, and for the same reason in both cases: on long timescales, sub-annual lags
+CaseC1, and for the same reason in both cases: on long timescales, sub-annual lags
 are inconsequential, and the positive effects of nitrates/$\epsilon^{(1)}$ and the negative
@@
 Interactions between Moran effects in CCA2 operated similarly to CCA1 (Fig. \ref{SM-fig:indirectMoran_SM_lag4}).
+%***DAN: This para is a lot to track. It would probably help to have the new figures described above.
 
@@
 Whereas in CCA1, the tendency of nitrate annual peaks to follow wave annual peaks by $1$q, added to the tendency of 
-nitrate effects on kelp to be delayed by $1$q, resulted in summertime positive nitrate effects and wintertime 
+nitrate effects on kelp to be delayed by $1$q, resulted in summer positive nitrate effects and winter 
 negative wave effects which reinforced sycnhrony; in SB nitrate instead both peaked and had its positive effects on kelp in spring,
-while waves still had negative effects on kelp in wintertime. Thus the effects of nitrates and waves in SB
+while waves still had negative effects on kelp in winter. Thus the effects of nitrates and waves in SB
 were approximately a quarter-cycle separated from each other with respect to the annual timescale (Fig \ref{fig:indirectMoran}t), 
 and so produced neither
-much reinforcement nor much destructive interference of synchrony on annual timescales (Fig. \ref{fig:SummaryKelpFig}d). Slightly negative 
+much reinforcement nor much destructive interference of synchrony on 
+annual timescales (Fig. \ref{fig:SummaryKelpFig}d). Slightly negative 
 interactions were observed (Fig. \ref{fig:SummaryKelpFig}d) because of slight deviations from the approximate phase
@@
 drivers must always be compared to the timescale of interest to determine interaction effects,
-as in Fig. \ref{fig:IntroTheory}, and as captured rigorously in our theory.
+as in Fig. \ref{fig:PedagogAlt}, and as captured formally in our theory.
 
@@
 other complicating features which are present in many ecological datasets. The study of
-\citep{Sheppard2019} developed the wavelet methods applied by \cite{CastoraniSheppardKelp}.
+\cite{Sheppard2019} developed the wavelet methods applied by \cite{CastoraniSheppardKelp}.
 The modelling approach of this study relates indirectly to the approach of \cite{Anderson2021}, though that
@@
 Maowei Liang, Kyle Emery, Nat Coombs, Adeola Adeboje and Ethan Kadiyala for helpful discussions.
-%***DAN: If you end up adding any of these people to the author list, take them off here. Also in the ack for the readme.
+%***DAN: If you end up adding any of these people to the author list, take them off here. Also in the acknowledgments for the README.
 
@@
 \includegraphics[width=.5\textwidth]{../Results/Theory/IntroTheory.pdf}
-\includegraphics[width=.5\textwidth]{../Results/Theory/PedagogFig.pdf}
-\caption{Example and basic concept of interacting Moran effects. The example (a), which 
-is based on the simple model described in the Introduction, shows that 
+\caption{%\doublespacing 
+Example of interacting Moran effects. The example, which 
+is based on the very simple model described in the Introduction, shows that 
 interaction effects are possible, and can be substantially positive or negative. 
@@
 and the values $\cov(\epsilon_i^{(1)},\epsilon_j^{(2)})$, for $i,j=1,2$, were all set equal to each
-other and to a parameter, $\phi$. For (b) and (c), which present the basic concept of interaction 
-effects for two distinct scenarios, 
-the top sinusoid on each panel represents 
-the component of oscillation at timescale 20 time steps of an environmental variable, $\epsilon_i^{(1)}$,
-measured in a sampling location $i$, and the bottom sinusoid represents the 
-component at the same timescale for another environmental variable, $\epsilon_i^{(2)}$; the 
-oscillations are lagged from each other by an amount $l_n$. Environmental effects on 
-the population, $w_i(t)$, in location $i$ are also lagged, by amounts $l_{e1}$ and
-$l_{e2}$. For this setup, a positive interaction between Moran effects can occur on the
-given timescale when lags align, i.e, $l_{e1}-l_{e2}+l_n=0$, as in panel (b), so that environmental
-effects on the population coincide and reinforce each other. All environmental effects on 
-populations are assumed positive in these examples. A negative
-effect can occur when lags produce out-of-phase influence of environmental variables on
-populations, e.g., $l_{e1}-l_{e2}+l_n$ equals half the given timescale/period,
-as in panel (c), so that the distinct influences tend to work against each other in 
-bringing about population oscillations on the given timescale.  
-A timescale-specific approach is essential to this reasoning. See the online version for a color
-rendering of all figures.}\label{fig:IntroTheory}
+other and to a parameter, $\phi$. 
+See the online version for color
+renderings of all figures.}\label{fig:IntroTheory}
 \end{figure}
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt.pdf}
+\caption{%\doublespacing 
+Illustration of the main concept of synergistically (a, b) or antagonistically (c, d) interacting Moran effects.
+Interactions require that each environmental variable itself be spatially synchronous; and then 
+the alignment or misalignment of three types of lag (see below) determine the sign and strength of the interactions. 
+Solid-line sinusoids represent the period-20 components of an environmental 
+variable in three locations ($\epsilon_i^{(1)}$ for $i=1,2,3$) and dashed-line sinusoids represent the period-20 components of 
+a different environmental variable in the same locations
+($\epsilon_i^{(2)}$ for $i=1,2,3$). Black arrows represent peak positive influences of environmental 
+variables on populations, which are lagged by an amount $l_{e1}$ for $\epsilon_i^{(1)}$ and by 
+an amount $l_{e2}$ for $\epsilon_i^{(2)}$, where these lags differ across the scenarios but are the same 
+in all locations within a scenario.
+Analogously, red arrows represent maximally negative effects.
+Oscillations are separately spatially synchronous for each environmental variable in 
+each scenario, illustrated with blue and pink rectangles: due to this environmental synchrony, peak positive population effects 
+of the same variable occur at similar times across locations, and likewise for peak negative effects.
+In the synergistic scenario, the lag between the environmental variables ($l_n$) and the lags of the population effects
+of the variables ($l_{e1}$ and $l_{e2}$) are aligned, i.e., $l_n+l_{e1}=l_{e2}$, so that peak positive 
+effects of $\epsilon_i^{(1)}$ coincide with peak positive effects of $\epsilon_i^{(2)}$ (the pink rectangles are
+aligned on a, b), augmenting synchrony. Likewise negative effects are aligned (blue rectangles).
+In the antagonistic scenario, lags are misaligned, i.e., $l_n+l_{e1}=l_{e2}+\sigma/2$, where $\sigma=20$ is the period, 
+so that peak positive 
+effects of $\epsilon_i^{(1)}$ coincide with maximally negative effects of $\epsilon_i^{(2)}$, and  maximally 
+negative effects of $\epsilon_i^{(1)}$ coincide with peak positive effects of $\epsilon_i^{(2)}$ 
+(pink rectangles on c are aligned with blue ones on d, and vice versa), reducing synchrony. 
+See the online version for color
+renderings of all figures.}\label{fig:PedagogAlt}
+\end{figure}
+
+\clearpage
+\newpage
+
+\begin{figure}
 \includegraphics[width=\textwidth]{./Fig2.pdf}
-\caption{Kelp sampling sites and example time series. Sampling sites (a) were from three regions,
+\caption{%\doublespacing 
+Kelp sampling sites and example time series. Sampling sites (a) were from three regions,
 central California 1 (CCA1, blue points), central california 2 (CCA2, green points) and the region
@@
 (e)-(g), with locations at which these time series were measured labeled on panels (a)-(d). 
-See the online version for a color rendering of all figures.}\label{fig:map}
+See the online version for color renderings of all figures.}\label{fig:map}
 \end{figure}
@@
 \includegraphics[width=.85\textwidth]{../Results/Theory/TheoryFigCaseC_v02.pdf}
-\caption{Theoretical case studies A (panels a-d), B (panels e-h) and C (i-p, two alternative sets of 
+\caption{%\doublespacing 
+Theoretical case studies A (panels a-d), B (panels e-h) and C (i-p, two alternative sets of 
 parameters used for i-l and m-p, respectively). Left panels (a, e, i, m) show the terms on the right side
@@
 quantifying population synchrony due to interacting Moran effects.
-Middle panels show the constituent parts of the theory (labelled). These constituents are complex-valued, 
+Middle panels show the constituent parts of the theory (labeled). These constituents are complex-valued, 
 with magnitude on the vertical axis and phase displayed using color.
@@
 and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.
-Synch. contrib. = Synchrony contribution. See the online version for a color
-rendering of all figures.}\label{fig:TheoryCaseA}
+Synch. contrib. = Synchrony contribution. See the online version for color
+renderings of all figures.}\label{fig:TheoryCaseA}
+\end{figure}
+
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt_SM.pdf}
+\caption{Similar to Fig. \ref{fig:PedagogAlt}, but for slightly altered 
+scenarios. For Fig. \ref{fig:PedagogAlt}, for simplicity, we assumed that larger values of environmental variables had 
+more positive influences on populations. We here assume that larger values of $\epsilon_i^{(1)}$ have more 
+positive influences, but that smaller values of $\epsilon_i^{(2)}$ have more positive influences.
+In scenario 1 here, $l_n+l_{e1}=l_{e2}$, but Moran interactions are nonetheless antagonistic because the 
+altered direction of effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
+variable correspond to the peak negative effects of $\epsilon_i^{(1)}$. In scenario 2 here, 
+$l_n+l_{e1}=l_{e2}+\sigma/2$ for $\sigma=20$ the period, but Moran interactions are synergistic because the 
+altered direction of effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
+variable correspond to the peak positive effects of $\epsilon_i^{(1)}$. See the online version for color
+renderings of all figures.}\label{fig:PedagogAlt_SM}
 \end{figure}
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_MainPlot_ForMS.pdf}
-\caption{Main components of the new theory as applied to kelp, central CA 1 (CCA1) region (a-c) and 
+\caption{%\doublespacing 
+Main components of the new theory as applied to kelp, central CA 1 (CCA1) region (a-c) and 
 Santa Barbara (SB) region (d-f). Panels show the terms on the right side of 
@@
 and \ref{SM-fig:SummaryKelpFig_SM_lag12} for $n=8, 12$, respectively.
-See the online version for a color rendering of this 
-figure.}\label{fig:SummaryKelpFig}
+See the online version for color renderings of all 
+figures.}\label{fig:SummaryKelpFig}
 \end{figure}
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainDirectMoran_ForMS.pdf}
-\caption{Components of direct Moran effects for kelp, for CCA1 (a-o)
+\caption{%\doublespacing 
+Components of direct Moran effects for kelp, for CCA1 (a-o)
 and SB (p-dd). Nitrates are identified with $\epsilon^{(1)}$ and waves with $\epsilon^{(2)}$. 
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
-\caption{Components of interactions between Moran effects for kelp, for CCA1 (a-o) and SB (b-dd). 
+\caption{%\doublespacing 
+Components of interactions between Moran effects for kelp, for CCA1 (a-o) and SB (b-dd). 
 Nitrates are identified with $\epsilon^{(1)}$ and waves with $\epsilon^{(2)}$. 
@@
 and \ref{SM-fig:indirectMoran_SM_lag12} for $n=8, 12$, respectively.
-See the online version for a color
-rendering of all figures.}\label{fig:indirectMoran}
+See the online version for color
+renderings of all figures.}\label{fig:indirectMoran}
 \end{figure}

### 2022-11-08 Paper/Paper.Rnw
mainly added a bunch of comments which suggest how to restructure the write-up for simplicity, which is what I am going to do next

@@
 or destructively, depending on which portion of the CA coast is examined and on 
-timescale of analysis. 
+the timescale of analysis. 
 Both of these studies argued that interactions between Moran effects may be a general feature of many 
@@
 
+%***DAN: The following para is a lot to ask the reader to absorb in the Intro, way too much for 
+%an ecological audience. Move it Theory. Can I replace it with something more intuitive, like the
+%swing example? Or is that a bad combination of too involved and too informal/non-ecological? 
+%If I were to use the swing example I'd have to do it in limited space, to prevent it from 
+%spiraling out of control as an overly extended metaphor.
+%***DAN: You also might want to add a para to the Intro on why interactions between Moran effects
+%should be important, to take up some space upon moving the below para to Theory (and it would also 
+%help motivate).
 Having demonstrated that interactions between Moran effects can augment or reduce population synchrony, we now 
@@
 periodic maximal positive influences of the two environmental variables coincide with each other as well as being 
-spatial synchronous. This alignment of produces additional synchrony in populations, 
+spatial synchronous. This alignment of influences produces additional synchrony in populations, 
 beyond what would manifest if environmental fluctuations were unrelated, because positive influences of variable 1 in location $i$ will
@@
 
-Our specific goals for this study are as follows. First, (G1) we will elaborate a general mathematical 
+Our specific goals for this study are as follows. First, (G1) we will elaborate a general 
 theory of interacting Moran effects which allows a detailed understanding of how the mechanism works. Second, 
@@
 already carried out a detailed analysis of interacting Moran effects in CA kelp populations and their importance 
-for kelp ecology, we instead use a simplified subset of the available kelp data, and our analysis is instead intended 
-to illuminate the inner workings of the general mechanism we describe. In the Discussion, we 
+for kelp ecology, we instead use a simplified subset of the available kelp data. Our analysis is intended 
+to illuminate the inner workings of the general mechanism we describe rather than being a complete exploration
+of kelp dynamics. In the Discussion, we 
 extend the arguments of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} that interactions between Moran 
-effects are likely to be common, and hence that our general theory seems likely to pertain to numerous 
+effects are likely to be common, and hence that our theory seems likely to pertain to numerous 
 ecosystems. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
@@
 The derivation of the theory is in SI section \ref{SM-sect:derivation}.
+%***DAN: Some of the content of the above para is quite technical, and is more about interpreting the
+%inner working of the theory as opposed to just comparing its predictions to the intution from
+%Figs 2 and 5. So maybe some of this should be move to a sup mat where I would really
+%delve into how the inner workings of the theory correspond to all these lags and alignments
+%of lags and so on.
 
 \section{Methods}
+
+%***DAN: Consider renaming our three scenarios as follows: A->A1, B->A2, C1->B1, C2->B2. After all,
+%A and B are quite related, so it makes sense to consider the as scenarios within a case study.
+%Then it would be more symmetric, and the four rows of panels of Fig. 4 would each correspond to
+%a scenario within a case study.
 
@@
 $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ in (\ref{rhoww_T1to3}), where $w$ was linearly detrended kelp time series in one of our regions, 
+%***DAN: Probably should shift this definition of what w stands for to the previous paragraph, since I was 
+%already using w in this context, there.
 $\epsilon^{(1)}$ was detrended nitrate time
 series in the region, and $\epsilon^{(2)}$ was detrended wave disturbance time series in the region, 
+%***DAN: likewise these ought to have been defined in the previous paragraph, I think
 we applied the definitions of these quantities (see General Theory), which required estimation from data of
@@
 also follow from \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not 
-straightforwardly. And those papers are case studies, so our results provide novel theoretical generality.
+straightforwardly. And those papers are case studies, whereas our results provide general theory.
 
@@
 On the shortest timescales ($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}a, left side of panel), $1$q
-is half the period, so interactions effects are negative. On long timescales (e.g., $\sigma>8$y), $1$q is a negligible portion of the
+is half the period, so interaction effects are negative. On long timescales (e.g., $\sigma>8$y), $1$q is a negligible portion of the
 period, so $l_{e1}-l_{e2}+l_n=1$q is close to zero, relative to that period, and interaction effects
 are positive: compared to such long timescales, $\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects happen close to simultaneously, so
-the two noise variables reinforce each other. Theory captures these complexities because
+the two noise variables reinforce each other. 
+%***DAN: Upon re-reading aftre taking several months break from this project, I find myself skimming
+%the following sentences which are supposed to be showing how the theory captures what is going on.
+%It makes me wonder if explaining how theory corresponds to intuition is maybe beyond the audience,
+%and that part should be consistent relegated to sup mat, and instead in the main text we just 
+%present the results of theory, and explain how they correspond to the intuition developed in Fig. 2.
+%I think that would also allow us to put Fig. 4 panels b,c,f,g,j,k,n,o into sup mat, massively
+%simplifying the presentation. Keep the last sentence of this para here, though, along with panels 
+%d,h,l,p.
+Theory captures these complexities because
 the size of the $1$q lag of this scenario, relative to a timescale, $\sigma$, is captured by
@@
 %Case study B, new version
+%***DAN: Note that case B was introduced in Methods as having a lag of length n in the effects
+%of epsilon 1, but we immediate specify n=3 in Results. Do we want to instead intro it as n=3
+%from the beginning?
 CaseB reveals how Moran effects can interact when lags are longer than 1 model time step. 
@@
 were positive for $\sigma=0.75 \text{y}$ and negative for $\sigma=1.5$y 
-(Fig. \ref{fig:TheoryCaseA}e). Theory captures these complexities because 
+(Fig. \ref{fig:TheoryCaseA}e). 
+%***DAN: We could again put this "theory captures these complexities" business in sup mat, where
+%we could anyway explain in greater detail. Also put the middle panels there.
+Theory captures these complexities because 
 the size of the $3$q sum of lags for this scenario, relative to $\sigma$,
@@
 For CaseC, $\epsilon^{(1)}$ effects on populations were lagged by $n_1$q and 
-$\epsilon^{(2)}$ effects were lagged by $n_2$q (Methods). We used $n_1=1$, $n_2=0$, $p_1^{(1)}=3$ and $p_0^{(2)}=-1.5$, so 
+$\epsilon^{(2)}$ effects were lagged by $n_2$q (Methods). 
+%***DAN: give the exact lag values used in methods. Also the coefficient values, probably. 
+We used $n_1=1$, $n_2=0$, $p_1^{(1)}=3$ and $p_0^{(2)}=-1.5$, so 
 $\epsilon^{(1)}$ effects were lagged and positive, and $\epsilon^{(2)}$ effects 
@@
 both the new and previous paradigms covered by our general theory.
+%***DAN: It is asking way too much of the reader to spend the labor to go through Fig. 5 and update their
+%understanding to become more general than the original Fig. 2 just because of this one sentence. I am
+%asking them to spend maybe 15-20 minutes puzzling over Fig. 5 based on this one sentence of main text.
+%People won't do that, instead they'll just accept a more surface-level understanding and blunder onward.
+%Suppose I moved the text referring to Fig. 2 from the Intro to the Theory, and then also cited what is
+%currently Fig. 5 there (it would become Fig. 3). So the reader would be prepared for both positive and
+%negative effects, and when they got here they would be ready to absorb this example. I think this is a 
+%smart idea.
 For CaseC1 (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
@@
 negative interactions between Moran influences (Fig. \ref{fig:TheoryCaseA}m).
+%***DAN: More "theory captures these complexities" stuff that should be sent to sup mat.
+%No one will absorb this kind of stuff who isn't incredibly dedicated, and therefore
+%will be willing to read sup mat. Make the main text more about comparing the intution 
+%from fig 2 and 5 with the predictions of theory, while black-boxing how the theory
+%actually works.
 Our theory captures these complexities
@@
 oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
-%***DAN: You might even consider a new panel or two somehow showing how the theory captures these effects more directly than 
-%Fig. 3 currently shows, though not sure exactly how that would be done or if it's actually a good idea.
 The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
@@
 %Direct Moran effects
+%***DAN: Do I really want to spend all this space (and reader mental energy), including 
+%a very complex figure, on direct Moran effects, when the paper is not even about that?
+%***DAN: And this paragraph also requires plotting all these inner portions of the
+%theory, which I am starting to think is not really going to work for an ecological 
+%audience.
 The way direct Moran effects work in our theory is fairly straightforward. 
+%***DAN: Why is the following sentence not in the Theory section?
 For instance, the direct Moran effect
@@
 (Fig. \ref{fig:directMoran}l,n,aa,cc), possibly due to the North Pacific Gyre Oscillation 
-[NPGO; \cite{CastoraniSheppardKelp}]. The long-timescale synchrony
+\citep{CastoraniSheppardKelp}. The long-timescale synchrony
 in nitrates and waves produced some of the long-timescale
@@
 %Interactions between Moran effects, CCA1
+%***DAN: In previous comments I have begun outlining how I might simplify
+%and make the paper more about comparing the intuition of Fig. 2 and 5 to 
+%the *predictions* of theory, rather than taking the reader through all the
+%inner workings of theory. I like those thoughts so far, but they present some chanllenges
+%when we come to this paragraph. Much of what is stated below can be demonstrated
+%by looking at the estimates of the constituents of the theory, but I think that'll
+%be too confusing. So we have to come up with other ways to demo it. I think the 
+%solution is: we feel free to plot our synchrony and cross synchrony measures,
+%rho_epsiloni_epsilonj, even though in the case of i differnt from j, we have to
+%plot a complex quantity. But instead of plotting the fs, we just talk about
+%lags garnered from regression.
 Interactions between Moran effects in CCA1 were parallel to one of the theoretical case studies, CaseC1.
@@
 were positive and delayed by $1$q and $\epsilon^{(2)}$ effects were negative and immediate.
+%***DAN: So we would drop the following sentence, or send it to a new section in 
+%sup mat where we systematically explored the constituents of the theory. Instead
+%we would just rely on the regression results which follow.
 Delayed effects are reflected, for CCA1, in Fig. \ref{fig:indirectMoran}a, and for 
 CaseC1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
-that $f_{\mcP^{(1)}}$ has phase about $-\pi/2$ at the annual timescale. The delay is also 
+that $f_{\mcP^{(1)}}$ has phase about $-\pi/2$ at the annual timescale. 
+%***DAN: The above is pretty horrible. The reader has to locate two panels on two differently
+%formatted and very complex figures and compare them to see this! They will not do that!
+The delay is also 
 reflected, for CCA1, in Table \ref{SM-tab:regcoefs}, where 
 the entries for $p_0^{(1)}$ for CCA1 are close to $0$ and those for $p_1^{(1)}$ are positive.
+%***DAN: On the other hand, the contents of the table are pretty clear and intutitive to
+%an ecological audience, and can be used to match the intution of lags (Figs 2 and 5) with the
+%*outputs* of the theory, without plotting all the inner working in the main text. Even just
+%plotting both phase and magnitude on the same axes is going to be a stretch, and we have been
+%doing that a LOT! I'm really thinking that dragging the reader through all the inner workings
+%of this theory is just not going to work, unless I want to publish in Theoretical Population
+%Biology or something.
+%***DAN: So we would also drop the following sentence, relying instead
+%on the regression results which follow.
 Immediate negative effects of waves in CCA1 are reflected in Fig. \ref{fig:indirectMoran}b,g,l
@@
 peaks by $1$q in CaseC1.
+%***DAN: We would keep the portion of the following sentence that is about CCA1,
+%and we would keep a version of Fig 8d, which shows the cross synchrony between 
+%these drivers corresponds to a lag of 1 quarter. The reference to 4k will just 
+%be dropped, since anyway we built that theoretical case study to have that lag 
+%so don't need a confusing spectral reflection of the lag. In the sup mat section 
+%where we go through the constituents of the theory in detail can have this kinda 
+%stuff.
 This is reflected, for CCA1, in Fig. \ref{fig:indirectMoran}d, and for CaseC1 
@@
 effects of nitrates on kelp tend to come in summer, and the annual negative effects of waves come in winter,
-producing reinforcing interactions between Moran effects. Likewise, the $\pi/2$ phase 
+producing reinforcing interactions between Moran effects. 
+%***DAN: Following sentence moved to sup mat.
+Likewise, the $\pi/2$ phase 
 of $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}$ on annual timescales (Fig. \ref{fig:indirectMoran}c for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1)
@@
 Interactions between Moran effects in CCA2 operated similarly to CCA1 (Fig. \ref{SM-fig:indirectMoran_SM_lag4}).
-%***DAN: This para is a lot to track. It would probably help to have the new figures described above.
 
@@
 Whereas in CCA1, nitrate effects on observed kelp density were delayed by $1$q, in SB nitrate effects 
-were observable within the same quarter (Fig. \ref{fig:indirectMoran}a v. p; see also Table 
+were observable within the same quarter 
+%***DAN: Just rely on the regression results here, how the complex spectral stuff
+(Fig. \ref{fig:indirectMoran}a v. p; see also Table 
 \ref{SM-tab:regcoefs}, where $p_0^{(1)}$ terms were close to zero for CCA1 and positive for SB,
 whereas $p_1^{(1)}$ terms were positive for CCA1 and close to zero for SB). Wave effects were negative and
-immediate in both regions, though they were less pronounced in SB (Fig. \ref{fig:indirectMoran}b,g,l v. q,v,aa; see also 
-Table \ref{SM-tab:regcoefs}). Annual peaks in nitrates still tended to lag annual peaks in waves by $1$q in SB,
+immediate in both regions, though they were less pronounced in SB 
+%***DAN: Same, not only for the negativity, but also for the fact that wave effects were less pronounced in SB.
+(Fig. \ref{fig:indirectMoran}b,g,l v. q,v,aa; see also 
+Table \ref{SM-tab:regcoefs}). 
+%***DAN: The following setence will stay in main text, so we will have to keep a version of Fig. 8d and s,
+%although most of the rest of that figure will be simplified away.
+Annual peaks in nitrates still tended to lag annual peaks in waves by $1$q in SB,
 as in CCA1, though the seasonal periodicity of both variables was reduced in SB compared to 
@@
 between Moran effects have been tested for only twice previously, and were observed
-on both occassions; for this and other reasons (see below) interactions may
-also be common \citep{Sheppard2019,CastoraniSheppardKelp}. 
+on both occassions \citep{Sheppard2019,CastoraniSheppardKelp}
+; for this and other reasons (see below) interactions may also be common. 
 Moran interactions are therefore
@@
 phenomena, including major climatic oscillations such as 
-the NPGO or El Ni\~{n}o Southern Oscillation (ENSO).
+the North Pacific Gyre Oscillation or El Ni\~{n}o Southern Oscillation (ENSO).
 Driver pairs which may commonly produce interacting Moran effects include instances where the same quantity 
@@
 year (e.g., spring temperature and precipitation). Future work should systematically 
-investigate cross synchrony ($\rho_{\epsilon^{(1)} \epsilon^{(2)}}$) of environmental variables, and could then use our theory
-to directly assess the potential for interacting Moran effects to occur under a variety of biological scenarios. Since 
+investigate cross synchrony ($\rho_{\epsilon^{(1)} \epsilon^{(2)}}$) of environmental 
+variables, and could then use our theory
+to assess the potential for interacting Moran effects under a variety of biological scenarios. Since 
 interactions between Moran effects have been discovered in both of the two systems so far 
@@
 evidence for interactions. But their data were global in extent and spatially coarse, their time series were 
-short, and they used statistical methods more aligned with their goals of 
+short, and they used methods more aligned with their goals of 
 understanding the geography of synchrony than with our present goals of understanding 
@@
 by applying our theory and the methods of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} to
-suitable data. Likewise, rainfall variables measured in distinct portions
-of the growing season may also be promising candidates for finding interaction effects, for similar reasons.
+suitable data.
 
@@
 drivers must always be compared to the timescale of interest to determine interaction effects,
-as in Fig. \ref{fig:PedagogAlt}, and as captured formally in our theory.
+as in Figs \ref{fig:PedagogAlt}, \ref{fig:PedagogAlt_SM}, and as captured formally in our theory.
 
@@
 synchrony of environmental drivers \citep{Herfindal2022}.
-Antagonistic interaction effects between Moran drivers may be one common and previously
+Antagonistic interactions between Moran drivers may be one common and previously
 unrecognized mechanism contributing to this discrepancy. On the other hand,
@@
 in Moran drivers can theoretically cause greater synchrony in populations 
-than in the drivers themselves \citep{Massie2015,Desharnais2018}.
+than in drivers \citep{Massie2015,Desharnais2018}.
 Synergistic interaction effects is another mechanism by which populations can
@@
 1) by altering species phenology; and 2) by increasing or decreasing growth rates and thereby
-decreasing or increasing delays (respectively). Though it is too early to conclude that
+decreasing or increasing delays. Though it is too early to conclude that
 effects on Moran interactions are among the important impacts of climate change, we feel the mechanisms
@@
 It has not escaped our notice that kelp time series even longer than ours, once 
-available, may reveal Moran effects and interactions, on timescales longer than ours, which differ from
-our results. In our view this provides another strong argument, among many others, in support of 
-long-term on-the-ground or remote-sensing efforts to monitor ecosystems, such as the efforts 
-from which our kelp data are derived.
+available, may reveal novel Moran effects and interactions on timescales longer than ours. 
+This provides another argument, among many others, in support of 
+long-term ecosystem monitoring.
 
@@
 2023555, 2023523, 2140335, and 2023474 to M.C.N.C, K.C.C., T.W.B, and D.C.R., respectively; and by NSF-Math Bio award 
-1714195 to D.C.R.; and by support to D.C.R. from the James S. McDonnell Foundation and the California Department of Fish
+1714195 to D.C.R.; and by support to D.C.R. from the James S. McDonnell Foundation, the Humboldt Foundation, 
+and the California Department of Fish
 and Wildlife Delta Science Program. This project used data developed through the Santa Barbara Coastal 

### 2022-11-08 Paper/SupMat.Rnw
various edits to Discussion and getting repo ready
changes to the writeup from a long time ago that I forgot to commit. comming back to this for the first time in a while now.
mainly added a bunch of comments which suggest how to restructure the write-up for simplicity, which is what I am going to do next

@@
 %header material for paper
-\title{Supporting information for ``How environmental drivers of synchrony interact''}
+\title{Supporting information for ``Environmental drivers of synchrony and their interactions''}
 \date{}
@@
 For pedagogical clarity, we start with a white-noise example that does not illustrate the full value of the
-idea of cross synchrony, but which is simple; we subsequently move on to an example which is more
+idea of cross synchrony, but which is simple; we then move on to an example which is more
 complex but also more illustrative. These examples are also used in the theoretical case studies (main text).
@@
 The term $\exp(-2 \pi \iota f (l_1-l_2))$ in the expression for 
-$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ reflects that fact that cross variable synchrony for this example is lagged
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ reflects the fact that cross variable synchrony for this example is lagged
 by an amount $l_1-l_2$. The lack of a similar term in the expressions for $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$
@@
 synchrony in this example.
-
-\begin{figure}
-\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v01_rhos.pdf}\\
-\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v02_rhos.pdf}
-\caption[Spectral quantities for noise used in case study C]{The quantities $\rho_{\epsilon^{(1)} \epsilon^{(1)}}$,
-$\rho_{\epsilon^{(2)} \epsilon^{(2)}}$, and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ for the noise
-described in SI section \ref{sect:examples_cross_sync} and used for theoretical case study C. 
-These are complex-valued functions of timescale; their magnitude is displayed on vertical axes and
-their phase is displayed in color. Two alternative sets of parameters
-were used for panels a-c and d-f, respectively, and are also used in the two scenarios
-considered for case study C. Parameters were:
-$N=\Sexpr{N_C1}$, $c_1=\Sexpr{ns_C1[1]}$, $c_2=\Sexpr{round(ns_C1[2],4)}$, 
-$w_{11}=w{22}=\Sexpr{ns_C1[5]}$ and $w_{12}=\Sexpr{ns_C1[7]}$ for both sets of panels;
-$l_1=\Sexpr{ns_C1[3]}$ and $l_2=\Sexpr{ns_C1[4]}$ for panels a-c; and 
-$l_1=\Sexpr{ns_C2[3]}$ and $l_2=\Sexpr{ns_C2[4]}$ for panels d-f.
-}\label{fig:rhosForC}
-\end{figure}
 
@@
 in Fig. \ref{MT-fig:map} in the main text. Regions were selected to represent distinct dynamical
-patterns known to occur in central and southern CA, and also to have another the same number
+patterns known to occur in central and southern CA, and also to have about the same number
 of time series per selected region.
 
-Wave nutrient data were also available for each quarter at each location. Wave data were derived 
+Wave and nutrient data were also available for each quarter at each location. Wave data were derived 
 by applying a swell propagation model with bouy-based wave measurements and a validation 
@@
 
-\section{Supporting results for kelp}
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v01_rhos.pdf}\\
+\includegraphics[width=.75\textwidth]{../Results/Theory/TheoryFigCaseC_v02_rhos.pdf}
+\caption[Spectral quantities for noise used in case study C]{The quantities $\rho_{\epsilon^{(1)} \epsilon^{(1)}}$,
+$\rho_{\epsilon^{(2)} \epsilon^{(2)}}$, and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ for the noise
+described in SI section \ref{sect:examples_cross_sync} and used for theoretical case study C. 
+These are complex-valued functions of timescale; their magnitude is displayed on vertical axes and
+their phase is displayed in color. Two alternative sets of parameters
+were used for panels a-c and d-f, respectively, and are also used in the two scenarios
+considered for case study C. Parameters were:
+$N=\Sexpr{N_C1}$, $c_1=\Sexpr{ns_C1[1]}$, $c_2=\Sexpr{round(ns_C1[2],4)}$, 
+$w_{11}=w{22}=\Sexpr{ns_C1[5]}$ and $w_{12}=\Sexpr{ns_C1[7]}$ for both sets of panels;
+$l_1=\Sexpr{ns_C1[3]}$ and $l_2=\Sexpr{ns_C1[4]}$ for panels a-c; and 
+$l_1=\Sexpr{ns_C2[3]}$ and $l_2=\Sexpr{ns_C2[4]}$ for panels d-f.
+}\label{fig:rhosForC}
+\end{figure}
 

### 2022-11-15 Paper/Paper.Rnw
lots of edits working toward simplification of the ms

@@
 %header material for paper
-\title{Environmental drivers of synchrony and their interactions}
+\title{How environmental drivers of spatial synchrony interact}
 \date{}
@@
 %Environmental drivers of spatial synchrony interact
-%How environmental drivers of spatial synchrony interact
+%Environmental drivers of synchrony and their interactions
 %how environmental drivers of synchrony interact
@@
 
+%Intro to synchrony
 Spatial synchrony, the tendency for geographically disjunct populations to show correlated fluctuations 
@@
 \emph{synchrony}) has been studied for decades
-in a wide variety of species ranging from viruses to plants to mammals, and at spatial scales ranging from centimeters to over 
-1000 km \citep{Liebhold2004},
+in a wide variety of species ranging from viruses to plants to mammals, and at 
+spatial scales ranging from centimeters to over 1000 km \citep{Liebhold2004},
 recent advances in statistical methods and improvements in data availability have led to several 
@@
 and patterns of synchrony are now known to be asymmetric in distribution tails in a way that
-is important for system stability \citep{Ghosh2020b,Walter2021}. Our ability to infer the causes of 
+is important for system stability \citep{Ghosh2020b,Ghosh2020c,Walter2021}. Our ability to infer the causes of 
 synchrony is much improved in recent years \citep{Sheppard2016,Defriez2017b,Walter2017,Anderson2018,Sheppard2019}, 
@@
 
+%Intro to Moran effects and interactions
 One of the classically theorized \citep{Moran1953} and now widely empirically validated mechanisms of 
@@
 so that total population synchrony can be either greater or less than if the distinct Moran 
-drivers operated independently \citep{Sheppard2019,CastoraniSheppardKelp}. Effects can be strong.
+drivers operated independently \citep{Sheppard2019,CastoraniSheppardKelp}. 
 \cite{Sheppard2019} showed that long-timescale ($>4$y period) synchrony in a chlorophyll density index 
@@
 the timescale of analysis. 
-Both of these studies argued that interactions between Moran effects may be a general feature of many 
-systems, because of the large number of interrelated factors driving population fluctuations in most 
-spatially extended systems. The main purpose of this study is to establish a general theory of the mechanism
+The main purpose of this study is to establish a general theory of the mechanism
 of interacting Moran effects and to use examples and applications of the theory to provide 
@@
 
+%Demo that interactions can happen
 The basic fact that Moran drivers can interact, synergistically or destructively, can be demonstrated 
-with a very simple model which we now elaborate, though understanding the nature of the interaction
+with a very simple model which we now elaborate, though understanding the full nature of the interaction
 mechanism will require the rest of this paper. Suppose $\epsilon_i^{(1)}(t)$ and $\epsilon_i^{(2)}(t)$
@@
 [\sqrt{2+2\cov(\epsilon_1^{(1)},\epsilon_1^{(2)})}\sqrt{2+2\cov(\epsilon_2^{(1)},\epsilon_2^{(2)})}]$,
-a quantity which depends not only on the standard environmental synchrony measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(a)})$, for 
+a quantity which depends not only on the standard environmental synchrony 
+measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(a)})$, for 
 $a=1,2$, but also on the ``cross synchrony'' measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(b)})$
@@
 from SI section S1 of \cite{Sheppard2019}.
-%***DAN: If we decide to go longer, we could add some brief commentary here about how common it should be for 
-%spatially extended population systems to be influenced by more than one env var. Maybe also mention these
-%env vars may be different vars measured at the same time of year, or the same or different vars measured
-%at different times of year, i.e., separated phenologically. Keep it brief, it's just intended to make them 
-%this is is going to be important and common, and to make them sorta wonder why this has not been examined already.
-%Actually that second thing is done in the Discussion, either don't do it here or else if you do, then take it
-%out of there.
-
-%***DAN: Another potentially good Discussion para: make connections to the work (or Wang and Loreau and
-%others) that is fusing spatial and community ecology. What are the implications of this work for that work?
-%Making such additions may help it appeal to a broader audience, which is something we need to work on as 
-%much as we can for Ecol Lett.
-
-%***DAN: The following para is a lot to ask the reader to absorb in the Intro, way too much for 
-%an ecological audience. Move it Theory. Can I replace it with something more intuitive, like the
-%swing example? Or is that a bad combination of too involved and too informal/non-ecological? 
-%If I were to use the swing example I'd have to do it in limited space, to prevent it from 
-%spiraling out of control as an overly extended metaphor.
-%***DAN: You also might want to add a para to the Intro on why interactions between Moran effects
-%should be important, to take up some space upon moving the below para to Theory (and it would also 
-%help motivate).
-Having demonstrated that interactions between Moran effects can augment or reduce population synchrony, we now 
-begin to provide some intuition on how to understand the underlying mechanism. 
-A key insight is the use of a timescale-specific approach,
-i.e., environmental signals are decomposed, using Fourier analysis, into components that occur on different timescales,
-and the nature of interacting Moran effects can be understood on a timescale-by-timescale basis, with effects summed 
-across timescales afterwards if desired. Fig. \ref{fig:PedagogAlt}a,b shows the period-20 Fourier components of two hypothetical,
-spatially synchronous environmental variables measured in sampling locations 
-$i=1,2,3$, and one way these can influence populations. The components are
-lagged, relative to each other, in the timing of their peaks ($l_n$ on the figure). They are also lagged in their 
-effects on populations, by the amounts $l_{e1}$ and $l_{e2}$, respectively, i.e., peaks in the environmental signals
-manifest as maximum positive influence on populations after delays of $l_{e1}$ and $l_{e2}$, respectively. In this example,
-we assume for simplicity that larger values of both environmental variables are beneficial to populations, though the general
-theory described below does not require that assumption. If, as in Fig. \ref{fig:PedagogAlt}a, b, $l_n+l_{e1}=l_{e2}$, then the 
-periodic maximal positive influences of the two environmental variables coincide with each other as well as being 
-spatial synchronous. This alignment of influences produces additional synchrony in populations, 
-beyond what would manifest if environmental fluctuations were unrelated, because positive influences of variable 1 in location $i$ will
-tend to coincide with positive influences of variable 2 in location $j$. In contrast, Fig. \ref{fig:PedagogAlt}c, d
-shows the opposite scenario, for which $l_n+l_{e1}=l_{e2}+\sigma/2$, where $\sigma=20$ is the period. Thus
-the periodic maximal positive influence of environmental variable 1 coincides with the periodic maximal negative
-influence of environmental variable 2, reducing synchrony relative to the case of unrelated environmental variables. 
-Intermediate scenarios are possible, as will be revealed by the general theory.
-The theory also formalizes the timescale decomposition of synchrony measures, and timescale-specific lags.
-
+
+%Intutitive idea with the swing analogy
+Having demonstrated that interactions between Moran effects can occur, we now 
+use a simple analogy from everyday experience to begin to provide intuition. 
+Consider $N$ children, each riding on their own playground swing and each
+being pushed both by their own mother and father. If the fathers from seperate families
+were to synchronize their pushes, they would, of course, act as a Moran-like influence, tending to synchronize
+the swinging motions of the children. Likewise, if the mothers from seperate families were to
+synchronize their pushes, it would tend to have a separate synchronizing effect. If the pushes
+of the fathers were appropriately coordinated with the pushes of the mothers, the children would swing higher, and
+would also swing more synchronously, demonstrating a synergistic interaction between synchronizing
+effects, i.e., a tendency for the two synchronizing effects to reinforce each other. On the 
+other hand, if the maternal and paternal pushes
+were not appropriately coordinated, the children 
+would swing more or less randomly and would become asynchronous with each other, demonstrating
+an antagonistic interaction between synchronoizing effects, i.e., a tendency for the two synchronizing
+effects to cancel. As oscillators, children on swings are distinct in many ways from ecological
+populations, not least because swings oscillate on one frequency/on one timescale, whereas populations
+typically oscillate simultaneously on many timescales, having a broadband spectrum. 
+We will see that a key insight of our theory is the use of a timescale-specific approach: 
+environmental signals are decomposed, using Fourier analysis, into components that occur 
+on different timescales,
+and the nature of interacting Moran effects can be understood on a timescale-by-timescale basis.
+The simple swing analogy turns out to supply, in much idealized form, the basic intuition our formal
+theory will extend.
+
+%Importance of interactions between Moran effects
+Both \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} argued that interactions between Moran 
+effects may be a general feature of many 
+systems, because of the large number of interrelated factors driving population fluctuations in most 
+spatially extended systems. Interaction effects can also be strong. For these reasons and others, development of a 
+general theoretical and intutitive understanding of how interactions between Moran effects work is an 
+important research goal. We explore in the Discussion why climate change may also influence Moran
+interactions, and we 
+extend the arguments of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} that interactions between Moran 
+effects are likely to be common.
+
+%Final intro para with specific goals
 Our specific goals for this study are as follows. First, (G1) we will elaborate a general 
@@
 already carried out a detailed analysis of interacting Moran effects in CA kelp populations and their importance 
-for kelp ecology, we instead use a simplified subset of the available kelp data. Our analysis is intended 
+for kelp ecology, we instead use a simplified subset of the available kelp data. Our kelp analysis is intended 
 to illuminate the inner workings of the general mechanism we describe rather than being a complete exploration
-of kelp dynamics. In the Discussion, we 
-extend the arguments of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} that interactions between Moran 
-effects are likely to be common, and hence that our theory seems likely to pertain to numerous 
-ecosystems. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
+of kelp dynamics. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
 illuminate this newly observed but potentially quite important mechanism of spatiotemporal population dynamics.
@@
 
-\noindent Our main theoretical work requires a conceptual understanding of the \emph{spectrum} of an environmental 
+\noindent Prior to presenting our formal theory, we extend and formalize the intution behind it
+that we began to develop with the swing example in the Introduction. We again focus on a single
+timescale of oscillation, later combining timescales mathematically. 
+Fig. \ref{fig:PedagogAlt}a,b shows the period-20 Fourier components of two hypothetical,
+spatially synchronous environmental variables measured in sampling locations 
+$i=1,2,3$, and one way these can influence populations. The components are
+lagged, relative to each other, in the timing of their peaks ($l_n$ on the figure). They are also lagged in their 
+effects on populations, by the amounts $l_{e1}$ and $l_{e2}$, respectively, i.e., peaks in the environmental signals
+manifest as maximum positive influence on populations after delays of $l_{e1}$ and $l_{e2}$, respectively. In this example,
+we assume for simplicity that larger values of both environmental variables are beneficial to 
+populations, though the general
+theory described below does not require that assumption, and see also the next example which makes a
+different assumption. If, as in Fig. \ref{fig:PedagogAlt}a, b, $l_n+l_{e1}=l_{e2}$, then the 
+periodic maximal positive influences of the two environmental variables coincide with each other as well as being 
+spatial synchronous. This alignment of influences produces additional synchrony in populations, 
+beyond what would manifest if environmental fluctuations were unrelated, because positive 
+influences of variable 1 in location $i$ will
+tend to coincide with positive influences of variable 2 in location $j$. In contrast, Fig. \ref{fig:PedagogAlt}c, d
+shows the opposite scenario, for which $l_n+l_{e1}=l_{e2}+\sigma/2$, where $\sigma=20$ is the period. Thus
+the periodic maximal positive influence of environmental variable 1 coincides with the periodic maximal negative
+influence of environmental variable 2, reducing synchrony relative to the case of unrelated environmental variables. 
+Intermediate scenarios are possible, as will be revealed by the theory. If we replace the assumption that
+larger values of both environmental variables are beneficial to populations with an assuption that larger 
+values of the first variable and smaller values of the second are beneficial, scenarios of synergistic versus
+antagonistic Moran interactions are reversed, but with the same general concepts nevertheless still operating 
+(Fig. \ref{fig:PedagogAlt_SM}).
+
+Our formal theory requires a conceptual understanding of the \emph{spectrum} of an environmental 
 or population time series, as well as of the \emph{cospectrum} and \emph{cross spectrum} of two time series,
@@
 
-%***DAN: If you go longer, you might think about adding a para to Discussion about how general theories can bind together
-%multiple effects that are often understood idiosyncratically or as special cases by ecologists. Not sure this could be done 
-%effectively, but I'm always reading papers testing specific things and understanding phenomena in ways like "A can happen
-%and B can also happen". For instance, X (which is a good paper) says "We show that environmental fluctuations can reduce 
-%synchrony between species and increase stability, yet they can also decouple asynchrony between species and increase 
-%population and metapopulation variability." (https://esajournals-onlinelibrary-wiley-com.www2.lib.ku.edu/doi/pdfdirect/10.1002/ecm.1484) 
-%Sounds like they are saying both things can occur, we want a general theory
-%to understand when each outcome happens. Likewise could cite the old debate on red noise and effects on exitinction risk. 
-%There were papers show conflicting results until someone looked for the general theory and then it was resolved. Ecologists
-%have a natural proclivity for specific examples, and like results like "A can happen" but as scientists we ought to be 
-%seeking general theories. Not sure how to say all this in a constructive and useful as opposed to an annoying way. 
-%The assumptions that your dynamical models have an equilibrium and then an approach that assumes weak noise and linearizes 
-%about that equilibrium can provide a general theory (as we did here, but also cite several papers of myself and others using 
-%this appraoch). Whereas the theory is not completely general, because it does not cover strong noise and strong nonlinearities,
-%which do happen in ecology, we consider it better to produce a a general theory based on linearization and then examine
-%departures from it under strong noise than it is to use a series of scattershot simulation approaches, which has more than 
-%led to conflicting results in the literature which then had to be resolved. In our opinion, generality should be pursued in
-%theoretical ecological contexts rather sooner than it is our impression that it is typically pursued, wrking on it 
-%early on instead of only after several case-study-type studies based on simulations. The weak-noise linearization approach
-%is widely applicable and often more fruitful than simulations (REFS).
-
 %***DAN: Potential reviewers or editors:
@@
 $\epsilon^{(1)}$ and $\epsilon^{(2)}$, 
-represented on Fig. \ref{fig:PedagogAlt} by $l_{e1}$ and $l_{e2}$,
+represented on Figs \ref{fig:PedagogAlt} and \ref{fig:PedagogAlt_SM} by $l_{e1}$ and $l_{e2}$,
 relative to the timescale $\sigma$. The lag 
-$l_n$ on Fig. \ref{fig:PedagogAlt} manifests in the theory 
+$l_n$ on Figs \ref{fig:PedagogAlt} and \ref{fig:PedagogAlt_SM}  manifests in the theory 
 through the phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$. 
 The expression $l_{e1}-l_{e2}+l_{n}$, which quantifies the degree of mismatch among the lags
-(see Introduction and Fig. \ref{fig:PedagogAlt}) corresponds to the phase of the expression 
+(see above text and Figs \ref{fig:PedagogAlt}, \ref{fig:PedagogAlt_SM}) corresponds to the phase of the expression 
 $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, because the phase of
@@
 dynamics modulates Moran influences.
-The derivation of the theory is in SI section \ref{SM-sect:derivation}.
-%***DAN: Some of the content of the above para is quite technical, and is more about interpreting the
-%inner working of the theory as opposed to just comparing its predictions to the intution from
-%Figs 2 and 5. So maybe some of this should be move to a sup mat where I would really
-%delve into how the inner workings of the theory correspond to all these lags and alignments
-%of lags and so on.
+The derivation of the theory and details supporting the above interpretations of 
+constituents of the theory are in SI section \ref{SM-sect:derivation}. 
+%***DAN: I need to add some stuff to the SI section mentioned to support the claim that 
+%"details supporting the above assertions are in SI section X"
 
 \section{Methods}
-
-%***DAN: Consider renaming our three scenarios as follows: A->A1, B->A2, C1->B1, C2->B2. After all,
-%A and B are quite related, so it makes sense to consider the as scenarios within a case study.
-%Then it would be more symmetric, and the four rows of panels of Fig. 4 would each correspond to
-%a scenario within a case study.
 
@@
 \noindent We now describe how we pursue goal G2 of the Introduction, to develop 
-three theoretical case studies, that illuminate the intuition of interacting Moran effects. 
+three theoretical case studies that illuminate the intuition of interacting Moran effects. 
 For all cases, the model time step was assumed to be one quarter (q), i.e., 
@@
 $\epsilon^{(1)}$ is assumed to have a simple positive effect on populations, but 
-lagged by one model time step, i.e.,
+lagged by one model time step (1q), i.e.,
 $p_1^{(1)}>0$, and $p_i^{(1)}=0$ for $i \neq 1$. 
-For case study B (henceforth CaseB), which generalizes A, $\epsilon^{(1)}$ is again assumed to have a simple 
+For case study B (henceforth CaseB), $\epsilon^{(1)}$ is again assumed to have a simple 
 positive effect on populations,
-but now lagged by $n$ time steps, so that $p_n^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq n$.
+but now lagged by $3$q, so that $p_3^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq 3$.
 For both CaseA and CaseB, the effects of $\epsilon^{(2)}$
-are assumed to be un-lagged, i.e., $p_0^{(2)} \neq 0$ and $p_i^{(2)}=0$ for $i>0$.
-For case study C (CaseC), the effects of $\epsilon^{(1)}$ are lagged by $n_1$ and the 
-effects of $\epsilon^{(2)}$ are lagged by $n_2$ time steps, i.e.,
-$p_{n_1}^{(1)} \neq 0$ while $p_i^{(1)}=0$ for $i \neq n_1$, and 
-$p_{n_2}^{(2)} \neq 0$ while $p_i^{(2)}=0$ for $i \neq n_2$.
+are assumed to be un-lagged and positive, i.e., $p_0^{(2)} > 0$ and $p_i^{(2)}=0$ for $i>0$.
+For case study C (CaseC), the effects of $\epsilon^{(1)}$ are positive and lagged by $1$q and the 
+effects of $\epsilon^{(2)}$ are unlagged and negative, i.e.,
+$p_{1}^{(1)} > 0$ while $p_i^{(1)}=0$ for $i \neq 1$, and 
+$p_{0}^{(2)} < 0$ while $p_i^{(2)}=0$ for $i \neq n_2$.
 The noise process
@@
 The same regression coefficients were used for all locations within a region.
+Here, the $w_i$ represent linearly detrended kelp time series in one of our regions, 
+$\epsilon_i^{(1)}$ were detrended nitrate time
+series in the region, and $\epsilon_i^{(2)}$ were detrended wave disturbance time series.
 Waves influence kelp dynamics through direct disturbance events which can damage kelp or extirpate kelp locally 
@@
 \citep{Cavanaugh2011,Bell2015,Schiel2015}. 
-Therefore $m_1=1$ was used. Kelp holdfasts on the sea floor can last for multiple years, so kelp lag effects were considered:
+Therefore $m_1=1$ was used. Kelp holdfasts on the sea floor can 
+last for multiple years, so kelp lag effects were considered:
 we used $n=4, 8, 12$ in separate analysis, with this choice making no substantive difference to results (see Results).
@@
 To estimate the components $\rho_{ww}$, $\rho_{\epsilon^{(1)}\epsilon^{(1)}}$, $\rho_{\epsilon^{(2)}\epsilon^{(2)}}$ and 
-$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ in (\ref{rhoww_T1to3}), where $w$ was linearly detrended kelp time series in one of our regions, 
-%***DAN: Probably should shift this definition of what w stands for to the previous paragraph, since I was 
-%already using w in this context, there.
-$\epsilon^{(1)}$ was detrended nitrate time
-series in the region, and $\epsilon^{(2)}$ was detrended wave disturbance time series in the region, 
-%***DAN: likewise these ought to have been defined in the previous paragraph, I think
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ in (\ref{rhoww_T1to3}), 
 we applied the definitions of these quantities (see General Theory), which required estimation from data of
@@
 First, for all of our case studies, the magnitude of 
-interaction effects was comparable to that of direct Moran effects (Fig. \ref{fig:TheoryCaseA}a,e,i,m, 
+interaction effects was comparable to that of direct Moran effects (Fig. \ref{fig:TheoryCaseA}, 
 compare the dashed and solid lines). Thus interactions can contribute substantially to overall 
@@
 interactions can be negative or positive. CaseA and CaseB
-showed negative interaction effects on short timescales (Fig. \ref{fig:TheoryCaseA}a,e); 
+showed negative interaction effects on short timescales (Fig. \ref{fig:TheoryCaseA}a,c); 
 CaseC1 showed negative interactions on
-long timescales (Fig. \ref{fig:TheoryCaseA}i); and CaseC2 showed negative interactions on all timescales 
-(Fig. \ref{fig:TheoryCaseA}m).
+long timescales (Fig. \ref{fig:TheoryCaseA}e); and CaseC2 showed negative interactions on all timescales 
+(Fig. \ref{fig:TheoryCaseA}g).
 Thus interaction effects can either augment or reduce synchrony. Finally, interaction effects depended strongly on timescale 
-for all case studies (Fig. \ref{fig:TheoryCaseA}a,e,i,m), and even differed in sign across timescales for CaseA, CaseB and CaseC1. 
-This result complements earlier studies that showed direct Moran effects can 
+for all case studies. This result complements earlier studies that showed direct Moran effects can 
 depend on timescale \citep{Defriez2016,Sheppard2016,Desharnais2018,Anderson2021}. The results of this paragraph
 also follow from \cite{Sheppard2019} and \cite{CastoraniSheppardKelp}, though not 
-straightforwardly. And those papers are case studies, whereas our results provide general theory.
-
-%Case study A, new try
+straightforwardly, and those papers are case studies, whereas our results provide general theory.
+
+%Case study A
 CaseA helps provide an intuitive understanding about how lags in Moran drivers can produce contrasting interactions between
@@
 The between-noise lag, $l_n$, of Fig. \ref{fig:PedagogAlt} is $0$q because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
-taken to be a white noise process (see SI section \ref{SM-sect:examples_cross_sync} for details, here).
+taken to be a white noise process (see SI section \ref{SM-sect:examples_cross_sync} for details).
 Thus, what determines the sign of interactions between Moran effects for this example is how 
@@
 the two noise variables reinforce each other. 
-%***DAN: Upon re-reading aftre taking several months break from this project, I find myself skimming
-%the following sentences which are supposed to be showing how the theory captures what is going on.
-%It makes me wonder if explaining how theory corresponds to intuition is maybe beyond the audience,
-%and that part should be consistent relegated to sup mat, and instead in the main text we just 
-%present the results of theory, and explain how they correspond to the intuition developed in Fig. 2.
-%I think that would also allow us to put Fig. 4 panels b,c,f,g,j,k,n,o into sup mat, massively
-%simplifying the presentation. Keep the last sentence of this para here, though, along with panels 
-%d,h,l,p.
-Theory captures these complexities because
-the size of the $1$q lag of this scenario, relative to a timescale, $\sigma$, is captured by
-the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, which equals the phase of
-$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ because $l_n=0$ and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number
-(SI section \ref{SM-sect:examples_cross_sync}). But $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=
-p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) = 
-p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}), 
-which has phase $-\pi$ at short timescales
-and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b,c).
-The phase controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, the
-interaction term in (\ref{rhoww_T1to3}), as described in General Theory. Comparing
+%***DAN: I comented the below because I want to put it in sup mat. Once you write that section, consider
+%referring to it here with something like: "See SI section X for a detailed breakdown of how the terms in 
+%our theory capture these complexities."
+%Theory captures these complexities because
+%the size of the $1$q lag of this scenario, relative to a timescale, $\sigma$, is captured by
+%the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, which equals the phase of
+%$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ because $l_n=0$ and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number
+%(SI section \ref{SM-sect:examples_cross_sync}). But $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=
+%p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) = 
+%p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}), 
+%which has phase $-\pi$ at short timescales
+%and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b,c).
+%The phase controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, the
+%interaction term in (\ref{rhoww_T1to3}), as described in General Theory. 
+Comparing
 Fig. \ref{fig:TheoryCaseA}a and d shows the additional influence of autoregressive population effects, which
@@
 
-%Case study B, new version
-%***DAN: Note that case B was introduced in Methods as having a lag of length n in the effects
-%of epsilon 1, but we immediate specify n=3 in Results. Do we want to instead intro it as n=3
-%from the beginning?
+%Case study B
 CaseB reveals how Moran effects can interact when lags are longer than 1 model time step. 
-For CaseB, $\epsilon^{(1)}$ effects on populations were lagged by $n$q but 
-$\epsilon^{(2)}$ effects were unlagged (Methods). We used $n=3$, so $l_{e1}=3$q and $l_{e2}=0$q. 
+For CaseB, recall that $\epsilon^{(1)}$ effects on populations were lagged by $3$q but 
+$\epsilon^{(2)}$ effects were unlagged (Methods), so $l_{e1}=3$q and $l_{e2}=0$q in the language of Fig. \ref{fig:PedagogAlt}. 
 As for CaseA, because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
-a white noise process for CaseB, $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ was a real number and $l_n=0$.
+a white noise process for CaseB, $l_n=0$.
 Interaction effects were again negative on short timescales 
-($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}e, far left side of panel)
+($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}c, far left side of panel)
 because $l_{e1}-l_{e2}+l_n=3$q was $1.5\sigma$ on that timescale, and so effects of 
@@
 Similar to CaseA, $l_{e1}-l_{e2}+l_n=3$q was again negligible compared to long timescales, so interaction effects
-were positive on long timescales (Fig. \ref{fig:TheoryCaseA}e, right side of panel).
+were positive on long timescales (Fig. \ref{fig:TheoryCaseA}c, right side of panel), though timescales
+had to be a bit longer than in CaseA for this approximation to be as good (compare the rates at which the dashed
+lines level off in Fig. \ref{fig:TheoryCaseA}a, c).
 The quantity $l_{e1}-l_{e2}+l_n=3$q exactly equaled the timescale for $\sigma=3 \text{q} = 0.75 \text{y}$ and 
-equaled half the timescale for $\sigma=6 \text{q} = 1.5 \text{y}$, hence interactions effects
-were positive for $\sigma=0.75 \text{y}$ and negative for $\sigma=1.5$y 
-(Fig. \ref{fig:TheoryCaseA}e). 
-%***DAN: We could again put this "theory captures these complexities" business in sup mat, where
-%we could anyway explain in greater detail. Also put the middle panels there.
-Theory captures these complexities because 
-the size of the $3$q sum of lags for this scenario, relative to $\sigma$,
-is again captured by the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$; the phase of that
-quantity, represented by colors on Fig. \ref{fig:TheoryCaseA}f,g, fluctuates analogously to the
-interaction effects on Fig. \ref{fig:TheoryCaseA}a.
-
-%Case study C, new
-CaseA and CaseB assumed white noise processes, and therefore 
-real-valued $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ and $l_n=0$.
-But CaseC illustrates what can happen when noise processes have periodic components and lagged interdependences, e.g., 
+equaled half the timescale for $\sigma=6 \text{q} = 1.5 \text{y}$, hence interactions effect
+were, respectively, positive and negative for these timescales (Fig. \ref{fig:TheoryCaseA}c). 
+%***DAN: May again want to cite a new section of sup mat, in a similar way suggested in a comment in the previous paragraph
+%Theory captures these complexities because 
+%the size of the $3$q sum of lags for this scenario, relative to $\sigma$,
+%is again captured by the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$; the phase of that
+%quantity, represented by colors on Fig. \ref{fig:TheoryCaseA}f,g, fluctuates analogously to the
+%interaction effects on Fig. \ref{fig:TheoryCaseA}a.
+
+%Case study C
+CaseA and CaseB assumed white noise processes, and therefore $l_n$ was $0$.
+But CaseC illustrates what can happen when noise processes have periodic components and lagged associations, e.g., 
 both processes oscillate with annual periodicity and distinct phenology, an important scenario because seasonality is common.
-For CaseC, $\epsilon^{(1)}$ effects on populations were lagged by $n_1$q and 
-$\epsilon^{(2)}$ effects were lagged by $n_2$q (Methods). 
-%***DAN: give the exact lag values used in methods. Also the coefficient values, probably. 
-We used $n_1=1$, $n_2=0$, $p_1^{(1)}=3$ and $p_0^{(2)}=-1.5$, so 
-$\epsilon^{(1)}$ effects were lagged and positive, and $\epsilon^{(2)}$ effects 
-were immediate and negative. The language and logic of 
-Fig. \ref{fig:PedagogAlt} does not directly apply, because we there assumed, in order to introduce concepts simply,
-that environmental effects were positive, and we now take $p_0^{(2)}=-1.5$. But similar logic does apply 
-(Fig. \ref{fig:PedagogAlt_SM}), with
-both the new and previous paradigms covered by our general theory.
-%***DAN: It is asking way too much of the reader to spend the labor to go through Fig. 5 and update their
-%understanding to become more general than the original Fig. 2 just because of this one sentence. I am
-%asking them to spend maybe 15-20 minutes puzzling over Fig. 5 based on this one sentence of main text.
-%People won't do that, instead they'll just accept a more surface-level understanding and blunder onward.
-%Suppose I moved the text referring to Fig. 2 from the Intro to the Theory, and then also cited what is
-%currently Fig. 5 there (it would become Fig. 3). So the reader would be prepared for both positive and
-%negative effects, and when they got here they would be ready to absorb this example. I think this is a 
-%smart idea.
-For CaseC1 (Fig. \ref{fig:TheoryCaseA}i-l), peaks in the periodic
-noise process $\epsilon^{(1)}$ lagged peaks in the periodic process $\epsilon^{(2)}$ by
-$1$q (Methods, SI section \ref{SM-sect:examples_cross_sync}). Thus lags compounded: the 
+Recall that, for CaseC, $\epsilon^{(1)}$ effects on populations were positive and lagged by $1$q, i.e., $l_{e1}=1$ in the 
+language of Fig. \ref{fig:PedagogAlt_SM}; whereas 
+$\epsilon^{(2)}$ effects were unlagged and negative, i.e., $l_{e2}=0$ (Methods). 
+We compare CaseC to Fig. \ref{fig:PedagogAlt_SM} instead of Fig. \ref{fig:PedagogAlt} because
+$\epsilon^{(2)}$ effects were negative for CaseC, the situation considered by Fig. \ref{fig:PedagogAlt_SM}.
+For CaseC1, peaks in the periodic
+noise process $\epsilon^{(1)}$ were set up to lag peaks in the periodic process $\epsilon^{(2)}$ by
+$1$q (Methods, SI section \ref{SM-sect:examples_cross_sync}), so that $l_n=1$. 
+Thus $l_n+l_{e1}-l_{e2}=2 \text{q}=0.5 \text{y}$, and, on annual timescales, lags compounded, similar to 
+Fig. \ref{fig:PedagogAlt_SM}b: the 
 annual positive population effects of $\epsilon^{(1)}$ 
 were exactly $2$q offset from the annual negative effects of $\epsilon^{(2)}$,
-producing reinforcing interactions 
-between Moran effects on annual timescales (Fig. \ref{fig:TheoryCaseA}i).
-Contrastingly, for CaseC2 (Fig. \ref{fig:TheoryCaseA}m-p), peaks in the periodic
-noise process $\epsilon^{(2)}$ lagged peaks in the periodic process $\epsilon^{(1)}$ by
-$1$q. Thus lags cancelled: the annual positive population effects of $\epsilon^{(1)}$
-coincided with the negative effects of $\epsilon^{(2)}$ every year, producing 
-negative interactions between Moran influences (Fig. \ref{fig:TheoryCaseA}m).
-%***DAN: More "theory captures these complexities" stuff that should be sent to sup mat.
-%No one will absorb this kind of stuff who isn't incredibly dedicated, and therefore
-%will be willing to read sup mat. Make the main text more about comparing the intution 
-%from fig 2 and 5 with the predictions of theory, while black-boxing how the theory
-%actually works.
-Our theory captures these complexities
-because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
-quantifies the degree of alignment between the effects of $\epsilon^{(1)}$ and 
-$\epsilon^{(2)}$ compared to a given timescale, while 
-$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the degree of alignment between 
-oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
-The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-therefore captures whether delays in effects and lags between 
-variables align or misalign, with alignment
-reflected in the real part of this quantity, which is half the interaction term in our theory (\ref{rhoww_T1to3}).
+one quarter because of the lag of $\epsilon^{(1)}$ peaks behind $\epsilon^{(2)}$ peaks,
+and one additional quarter because of the delayed influence of $\epsilon^{(1)}$ peaks on
+populations. This produced reinforcing interactions 
+between the Moran effects of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ on annual timescales,
+as reflected by our theory (Fig. \ref{fig:TheoryCaseA}e,f).
+Contrastingly, for CaseC2, peaks in the periodic
+noise process $\epsilon^{(2)}$ were set up to lag peaks in the periodic process $\epsilon^{(1)}$ by
+$1$q, so that $l_n=-1$. Thus $l_n+l_{e1}-l_{e2}=0$ and, on annual timescales, lags cancelled,
+similar to Fig. \ref{fig:PedagogAlt_SM}a: the annual positive population effects of $\epsilon^{(1)}$
+coincided with the negative effects of $\epsilon^{(2)}$ every year, because $\epsilon^{(1)}$
+peaks came $1$q ahead of $\epsilon^{(2)}$ peaks each year, but population influences of $\epsilon^{(1)}$ were 
+delayed by $1$q. This produced negative interactions between the Moran influences of 
+$\epsilon^{(1)}$ and $\epsilon^{(2)}$ on annual timescales, as also reflected by our theory
+(Fig. \ref{fig:TheoryCaseA}g,h). 
+%***DAN: Maybe cite a new sup mat section here, once written
+%Our theory captures these complexities
+%because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
+%quantifies the degree of alignment between the effects of $\epsilon^{(1)}$ and 
+%$\epsilon^{(2)}$ compared to a given timescale, while 
+%$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the degree of alignment between 
+%oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
+%The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
+%therefore captures whether delays in effects and lags between 
+%variables align or misalign, with alignment
+%reflected in the real part of this quantity, which is half the interaction term in our theory (\ref{rhoww_T1to3}).
 On long timescales, interactions were the
@@
 the earlier theoretical results, based on our case studies 
-(above text and Fig. \ref{fig:TheoryCaseA}a,e,i,m),
+(text above and Fig. \ref{fig:TheoryCaseA}),
 that interaction effects between Moran drivers: 1) can be comparable in strength to direct Moran
@@
 negative on timescales $>8$y (Fig. \ref{fig:SummaryKelpFig}c).
+
+%***DAN: start from here after making the new version of the figure sketched in a separate planning doc
 
@@
 and $f_{\mcB}$ (Methods) are in Table \ref{SM-tab:regcoefs}.
+
+%***DAN: You need to combine the current figs 2 and 3, saving that for later.
+
+%***DAN: Stopped here 2022 11 15, 230pm
 
@@
 
+%***DAN: Another potentially good Discussion para: make connections to the work (or Wang and Loreau and
+%others) that is fusing spatial and community ecology. What are the implications of this work for that work?
+%Making such additions may help it appeal to a broader audience, which is something we need to work on as 
+%much as we can for Ecol Lett.
+
 %*DONE Summary
@@
 between Moran effects have been tested for only twice previously, and were observed
-on both occassions \citep{Sheppard2019,CastoraniSheppardKelp}
-; for this and other reasons (see below) interactions may also be common. 
+on both occassions \citep{Sheppard2019,CastoraniSheppardKelp}; 
+for this and other reasons (see below) interactions may also be common. 
 Moran interactions are therefore
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{./Fig2.pdf}
-\caption{%\doublespacing 
-Kelp sampling sites and example time series. Sampling sites (a) were from three regions,
-central California 1 (CCA1, blue points), central california 2 (CCA2, green points) and the region
-around Santa Barbara (SB, red points). Kelp density is shown with color intensity in (b)-(d),
-and those panels correspond to the regions. One example time series from each region is shown in
-(e)-(g), with locations at which these time series were measured labeled on panels (a)-(d). 
-See the online version for color renderings of all figures.}\label{fig:map}
-\end{figure}
-
-\clearpage
-\newpage
-
-\begin{figure}
-\includegraphics[width=.85\textwidth]{../Results/Theory/TheoryFigCaseA.pdf}\\
-\includegraphics[width=.85\textwidth]{../Results/Theory/TheoryFigCaseB.pdf}\\
-\includegraphics[width=.85\textwidth]{../Results/Theory/TheoryFigCaseC_v01.pdf}\\
-\includegraphics[width=.85\textwidth]{../Results/Theory/TheoryFigCaseC_v02.pdf}
-\caption{%\doublespacing 
-Theoretical case studies A (panels a-d), B (panels e-h) and C (i-p, two alternative sets of 
-parameters used for i-l and m-p, respectively). Left panels (a, e, i, m) show the terms on the right side
-of (\ref{rhoww_T1to3}). On those panels, the green line is 
-$\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}$, quantifying 
-population synchrony due to the direct Moran effects of $\epsilon^{(1)}$. The blue line
-is $\frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}$, quantifying 
-population synchrony due to the direct Moran effects of $\epsilon^{(2)}$.
-The dashed line is 
-$\frac{2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)}{|f_{\mcB}|^2}$, 
-quantifying population synchrony due to interacting Moran effects.
-Middle panels show the constituent parts of the theory (labeled). These constituents are complex-valued, 
-with magnitude on the vertical axis and phase displayed using color.
-See text for interpretations.
-The functions plotted on d, h, l, p are those on a, e, i, m, respectively, times 
-$|f_{\mcB}|^2$, plotted to illustrate how autoregressive population effects modulate
-synchrony [see (\ref{rhoww_T1to3})].
-Model parameters were: $N=\Sexpr{N_A}$ and $b_1=\Sexpr{b_A}$ for all case studies; 
-$p_1^{(1)}=\Sexpr{p1_A[2]}$ and $p_0^{(2)}=\Sexpr{p2_A}$ for case study A; 
-$p_3^{(1)}=\Sexpr{p1_B[4]}$ and $p_0^{(2)}=\Sexpr{p2_B[1]}$ for case study B;
-$p_1^{(1)}=\Sexpr{p1_C1[2]}$, $p_0^{(2)}=\Sexpr{p2_C1[1]}$ 
-%$c_1=0$ and $c_2=-4/9$ 
-for case study C. For C, peaks in the periodic noise process $\epsilon^{(2)}$
-lagged peaks in the periodic process $\epsilon^{(1)}$ by 1 quarter
-in panels i-l, and $\epsilon^{(1)}$ lagged $\epsilon^{(2)}$ by the same
-amount in m-p.
-See SI section \ref{SM-sect:examples_cross_sync} and Fig. \ref{SM-fig:rhosForC} for 
-parameters associated with the noise for each case study,
-and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.
-Synch. contrib. = Synchrony contribution. See the online version for color
-renderings of all figures.}\label{fig:TheoryCaseA}
-\end{figure}
-
-\clearpage
-\newpage
-
-\begin{figure}
 \includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt_SM.pdf}
@@
 \begin{figure}
+\includegraphics[width=\textwidth]{./Fig2.pdf}
+\caption{%\doublespacing 
+Kelp sampling sites and example time series. Sampling sites (a) were from three regions,
+central California 1 (CCA1, blue points), central california 2 (CCA2, green points) and the region
+around Santa Barbara (SB, red points). Kelp density is shown with color intensity in (b)-(d),
+and those panels correspond to the regions. One example time series from each region is shown in
+(e)-(g), with locations at which these time series were measured labeled on panels (a)-(d). 
+See the online version for color renderings of all figures.}\label{fig:map}
+\end{figure}
+
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=.4\textwidth]{../Results/Theory/TheoryFigCaseA_MT.pdf}\\
+\includegraphics[width=.4\textwidth]{../Results/Theory/TheoryFigCaseB_MT.pdf}\\
+\includegraphics[width=.4\textwidth]{../Results/Theory/TheoryFigCaseC_v01_MT.pdf}\\
+\includegraphics[width=.4\textwidth]{../Results/Theory/TheoryFigCaseC_v02_MT.pdf}
+\caption{%\doublespacing 
+Theoretical case studies A (panels a-b), B (panels c-d) and C (e-h, two alternative sets of 
+parameters used for e-f and g-h, respectively). Left panels (a, c, e, g) show the terms on the right side
+of (\ref{rhoww_T1to3}). On those panels, the green line is 
+$\frac{|f_{\mcP^{(1)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(1)} \epsilon^{(1)}}$, quantifying 
+population synchrony due to the direct Moran effects of $\epsilon^{(1)}$. The blue line
+is $\frac{|f_{\mcP^{(2)}}|^2}{|f_{\mcB}|^2} \rho_{\epsilon^{(2)} \epsilon^{(2)}}$, quantifying 
+population synchrony due to the direct Moran effects of $\epsilon^{(2)}$.
+The dashed line is 
+$\frac{2 \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right)}{|f_{\mcB}|^2}$, 
+quantifying population synchrony due to interacting Moran effects.
+The functions plotted on b, d, f, h are those on a, c, e, g, respectively, times 
+$|f_{\mcB}|^2$, plotted to illustrate how autoregressive population effects modulate
+synchrony [see (\ref{rhoww_T1to3})].
+Model parameters were: $N=\Sexpr{N_A}$ and $b_1=\Sexpr{b_A}$ for all case studies; 
+$p_1^{(1)}=\Sexpr{p1_A[2]}$ and $p_0^{(2)}=\Sexpr{p2_A}$ for case study A; 
+$p_3^{(1)}=\Sexpr{p1_B[4]}$ and $p_0^{(2)}=\Sexpr{p2_B[1]}$ for case study B;
+$p_1^{(1)}=\Sexpr{p1_C1[2]}$, $p_0^{(2)}=\Sexpr{p2_C1[1]}$ 
+%$c_1=0$ and $c_2=-4/9$ 
+for case study C. For C, peaks in the periodic noise process $\epsilon^{(2)}$
+lagged peaks in the periodic process $\epsilon^{(1)}$ by 1 quarter
+in panels e-f, and $\epsilon^{(1)}$ lagged $\epsilon^{(2)}$ by the same
+amount in g-h.
+See SI section \ref{SM-sect:examples_cross_sync} and Fig. \ref{SM-fig:rhosForC} for 
+parameters associated with the noise for each case study,
+and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.
+Synch. contrib. = Synchrony contribution; Rel. synch. contrib. = Relative synchrony contribution. 
+See the online version for color
+renderings of all figures.}\label{fig:TheoryCaseA}
+\end{figure}
+
+\clearpage
+\newpage
+
+\begin{figure}
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_MainPlot_ForMS.pdf}
@@
 \caption{%\doublespacing 
-Main components of the new theory as applied to kelp, central CA 1 (CCA1) region (a-c) and 
+The new theory as applied to kelp, central CA 1 (CCA1) region (a-c) and 
 Santa Barbara (SB) region (d-f). Panels show the terms on the right side of 
@@
 (black line) because other, unmeasured factors also help synchronize kelp dynamics.
+The timescale bands $0.5-2$, $2-4$ and $>4$ are separated on different panels because of the 
+very different $y$-axis ranges.
 The CCA1 results approximately parallel the 
 results for theoretical case
-study C scenario 1 (Fig. \ref{fig:TheoryCaseA}i; see text for details).
+study C scenario 1 (Fig. \ref{fig:TheoryCaseA}e; see text for details).
 See Fig. \ref{SM-fig:SummaryKelpFig_SM_lag4} for the central CA 2 (CCA2) region,
@@
 \begin{figure}
-\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainDirectMoran_ForMS.pdf}
-\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainDirectMoran_ForMS.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainDirectMoran_ForMS_short.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainDirectMoran_ForMS_short.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainInteractingMoran_ForMS_short.pdf}
+\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS_short.pdf}
 \caption{%\doublespacing 
-Components of direct Moran effects for kelp, for CCA1 (a-o)
-and SB (p-dd). Nitrates are identified with $\epsilon^{(1)}$ and waves with $\epsilon^{(2)}$. 
-Vertical axis extents are the same for corresponding panels for the
-two regions, to facilitate comparisons. 
-See Fig. \ref{SM-fig:directMoran_SM_lag4} for CCA2,
-for which results were similar to CCA1. This figure used 
-kelp lag $n=4$; see Figs \ref{SM-fig:directMoran_SM_lag8} 
-and \ref{SM-fig:directMoran_SM_lag12} for $n=8, 12$, respectively.}\label{fig:directMoran}
-\end{figure}
-
-\clearpage
-\newpage
-
-\begin{figure}
-\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
-\includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_SoCal_KelpLag4_ExplainInteractingMoran_ForMS.pdf}
-\caption{%\doublespacing 
-Components of interactions between Moran effects for kelp, for CCA1 (a-o) and SB (b-dd). 
+Synchrony (top two rows of panels) and cross synchrony (bottom panels) between environmental 
+variables influencing kelp for CCA1 (a-i) and SB (j-r). 
 Nitrates are identified with $\epsilon^{(1)}$ and waves with $\epsilon^{(2)}$. 
@@
 two regions, to facilitate comparisons.
-See Fig. \ref{SM-fig:indirectMoran_SM_lag4} for CCA2,
+Cross synchrony is complex valued, with magnitude plotted on the vertical axis and phase displayed using color. 
+See Figs \ref{SM-fig:directMoran_SM_lag4} and \ref{SM-fig:indirectMoran_SM_lag4} for CCA2,
 for which results were similar to CCA1. This figure used 
-kelp lag $n=4$; see Figs \ref{SM-fig:indirectMoran_SM_lag8} 
-and \ref{SM-fig:indirectMoran_SM_lag12} for $n=8, 12$, respectively.
+kelp lag $n=4$; see Figs \ref{SM-fig:directMoran_SM_lag8}, \ref{SM-fig:indirectMoran_SM_lag8}, 
+\ref{SM-fig:directMoran_SM_lag12} and \ref{SM-fig:indirectMoran_SM_lag12} for $n=8$ and$ 12$.
 See the online version for color
-renderings of all figures.}\label{fig:indirectMoran}
+renderings of all figures.}\label{fig:noise}
 \end{figure}

### 2022-11-15 Code/Kelp_AnalysisStep_SpectralDecomp.R

@@ -323,6 +323,10 @@ do_analysis<-function(Args)
 explain_interacting_Moran<-function(freq,frg,fP1,fP2,Se1e2_avg,
                           fnpre,plottype,PlotForMSArgs)
 {
+  #***
+  #preparation
+  #***
+  
   #unpack plotting params
   fP1_short_ylims<-PlotForMSArgs$ExplainInteractingMoran_fP1_short_ylims
   fP1_mid_ylims<-PlotForMSArgs$ExplainInteractingMoran_fP1_mid_ylims
@@ -340,6 +344,7 @@ explain_interacting_Moran<-function(freq,frg,fP1,fP2,Se1e2_avg,
   Prod_mid_ylims<-PlotForMSArgs$ExplainInteractingMoran_Prod_mid_ylims
   Prod_long_ylims<-PlotForMSArgs$ExplainInteractingMoran_Prod_long_ylims
   ExplainInteractingMoran_PanLabs<-PlotForMSArgs$ExplainInteractingMoran_PanLabs
+  ExplainInteractingMoran_short_PanLabs<-PlotForMSArgs$ExplainInteractingMoran_short_PanLabs
   
   #constraint to certain frequencies, for plotting
   fP1<-fP1[frg[1]<freq & freq<=frg[2]]
@@ -355,6 +360,10 @@ explain_interacting_Moran<-function(freq,frg,fP1,fP2,Se1e2_avg,
   xlimits_short<-c(xlimits[1],log2(8))
   xlimits_mid<-c(log2(8),log2(16))
   xlimits_long<-c(log2(16),xlimits[2])
+  
+  #***
+  #the long version, with panels for all the parts of the theory
+  #***
   
   #panel layout quantities
   gap<-.15
@@ -800,12 +809,483 @@ explain_interacting_Moran<-function(freq,frg,fP1,fP2,Se1e2_avg,
   text(xlimits_long[1],ylimits[2],ExplainInteractingMoran_PanLabs[15],adj=c(-0.1,1))
   
   dev.off()
+  
+  #***
+  #the short version, with only the panels for the noise
+  #***
+  
+  #panel layout quantities
+  gap<-.15
[…]

### 2022-11-15 Code/Theory_CaseStudyAB.R

@@ -362,6 +362,350 @@ TheoryCaseAB<-function(b,p1,p2,N,ns,PlotName,PanLabs,XAxOn=TRUE)
 }
 
 #***
+#A simplified main text version of the same thing. I just copied and cut a couple panels.
+#***
+
+#***
+#The main function for doing both case studies
+#***
+
+#This function calculates the analytic results and plots them for case studies A and B.
+#The function is actually more general than the planned case studies that will be 
+#displayed in the paper. Saem as above but omits a couple panels.
+#
+#Args
+#b            The parameters (b_1,...,b_n) of the model
+#p1           The parameters (p_0^{(1)},...,p_{m_1}^{(1)}) of the model
+#p2           The parameters (p_0^{(2)},...,p_{m_1}^{(2)}) of the model
+#N            The number of habitat patches
+#ns           Parameters for the noise. A length-3 vector with some constraints.
+#               See the code or the paper for information on how the noise is
+#               constructed based on these parameters. Using notation from the paper,
+#               the entries are v11, v22, v12.
+#PlotName     A file name of the saved plots, including path, excluding file extension.
+#PanLabs      A vector of panel labels (typically c("A","B"), or similar)
+#XAxOn        Default TRUE means plot the x axis label and numbers, FALSE means don't,
+#               with the size of the plot correspondingly adjusted.
+#
+#Output - none, but plots saved.
+#
+#Notes
+#Very little error checking, so watch out!
+#
+TheoryCaseAB_MT<-function(b,p1,p2,N,ns,PlotName,PanLabs,XAxOn=TRUE)
+{
+  #construct Sig, the covariance matrix for the noise, basically just to find out if it is positive definite
[…]

### 2022-11-15 Code/Theory_CaseStudyC.R

@@ -747,7 +747,524 @@ TestNoise<-function(N,ns,PlotName,numsims,BiasVariance)
 }
 
 #***
-#The main code, uses the above function
+#The main function for doing both case studies, but this is a truncated version of the above that
+#only plots some of the panels, for the main text of the paper.
+#***
+
+#This function calculates the analytic results and plots them for case study C.
+#The function is actually more general than the planned case study that will be 
+#displayed in the paper.
+#
+#Args
+#b            The parameters (b_1,...,b_n) of the model
+#p1           The parameters (p_0^{(1)},...,p_{m_1}^{(1)}) of the model
+#p2           The parameters (p_0^{(2)},...,p_{m_1}^{(2)}) of the model
+#N            The number of habitat patches
+#ns           Parameters for the noise. A length-7 vector with some constraints.
+#               See the code or the paper for information on how the noise is
+#               constructed based on these parameters. Using notation from the paper,
+#               the entries are c1, c2, l1, l2, w11, w22, w12.
+#PlotName     A file name of the saved plots, including path, excluding file extension.
+#PanLabs      A vector of panel labels (typically c("A","B","C","D"), or similar)
+#PanLabs2     Another similar, this one just 3 long, for the sup mat figure showing the 
+#               rhos
+#SyncYAx      A length 2 vector with the y axis extent for the synchrony plot. This is 
+#               just for the publication-quality plotting, set to NA if you want to 
+#               have this determined automatically.
+#RelSyncYAx   Similar but for the relative synchrony plot.
[…]

### 2022-11-16 Paper/Paper.Rnw
continued edits for simplifying

@@
 \usepackage{times}
+
+%Found this online. Makes figures on pages by themselves appear at the top instead of 
+%vertically centered
+\makeatletter
+\setlength{\@fptop}{0pt}
+\setlength{\@fpbot}{0pt plus 1fil}
+\makeatother
 
@@
 values of the first variable and smaller values of the second are beneficial, scenarios of synergistic versus
-antagonistic Moran interactions are reversed, but with the same general concepts nevertheless still operating 
-(Fig. \ref{fig:PedagogAlt_SM}).
+antagonistic Moran interactions are reversed, but with the same general concepts still operating 
+(Fig. \ref{fig:PedagogAlt}).
 
@@
 $f_{\mcP^{(2)}} = p_0^{(2)} + p_1^{(2)} \mu + \cdots + p_{m_2}^{(2)} \mu^{m_2}$, where 
-$\mu=\exp(-2 \pi \iota f)$, and where $\iota$, the Greek letter iota, represents the imaginary unit.
+$\mu=\exp(-2 \pi \iota f)$, and where $\iota$, the Greek letter iota, is the imaginary unit.
+
+Comparing the terms on the right
+of (\ref{rhoww_T1to3}) gives the relative contributions of direct Moran effects and interactions 
+between Moran effects. 
 The first term on the right of (\ref{rhoww_T1to3}) is the component of population synchrony 
@@
 
-Comparing the terms on the right
-of (\ref{rhoww_T1to3}) gives the relative contributions of direct Moran effects and interactions 
-between Moran effects. 
+The way direct Moran effects work in our theory is fairly straightforward.
 The magnitudes of the quantities $f_{\mcP^{(1)}}$ and $f_{\mcP^{(2)}}$ quantify the strength of 
 influence of $\epsilon^{(1)}$ and $\epsilon^{(2)}$ on populations at the timescale $\sigma=1/f$. 
-The phases of these quantities encode the lags in the influences of 
+The direct Moran effect
+term in (\ref{rhoww_T1to3}) for $\epsilon^{(1)}$, i.e., 
+$|f_{\mcP^{(1)}}|^2 \rho_{\epsilon^{(1)} \epsilon^{(1)}}/|f_{\mcB}|^2$, 
+equals the synchrony of the $\epsilon^{(1)}$ time series themselves, 
+$\rho_{\epsilon^{(1)} \epsilon^{(1)}}$, times the 
+timescale-specific strength of influence of $\epsilon^{(1)}$ on the populations, 
+$|f_{\mcP^{(1)}}|^2$, and modulated
+by the autoregressive nature of kelp dynamics,
+$1/|f_{\mcB}|^2$. The term for the direct Moran effects of $\epsilon^{(2)}$
+is interpreted similarly.
+
+The components of the interacting Moran effects term are also interpretable.
+The phases of the quantities $f_{\mcP^{(1)}}$ and $f_{\mcP^{(2)}}$ quantify the lags in the population influences of 
 $\epsilon^{(1)}$ and $\epsilon^{(2)}$, 
-represented on Figs \ref{fig:PedagogAlt} and \ref{fig:PedagogAlt_SM} by $l_{e1}$ and $l_{e2}$,
+represented on Fig. \ref{fig:PedagogAlt} by $l_{e1}$ and $l_{e2}$,
 relative to the timescale $\sigma$. The lag 
-$l_n$ on Figs \ref{fig:PedagogAlt} and \ref{fig:PedagogAlt_SM}  manifests in the theory 
+$l_n$ on Fig. \ref{fig:PedagogAlt} manifests in the theory 
 through the phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$. 
-The expression $l_{e1}-l_{e2}+l_{n}$, which quantifies the degree of mismatch among the lags
-(see above text and Figs \ref{fig:PedagogAlt}, \ref{fig:PedagogAlt_SM}) corresponds to the phase of the expression 
+The expression $l_{e1}-l_{e2}+l_{n}$, which quantifies the degree of mismatch among the lags,
+corresponds to the phase of the expression 
 $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, because the phase of
@@
 is positive and larger whenever the phase of this quantity is close to zero, i.e., when    
-$l_{e1}-l_{e2}+l_n$ is close to zero (Fig. \ref{fig:PedagogAlt}a, b), and is increasingly negative as 
+$l_{e1}-l_{e2}+l_n$ is close to zero (e.g., Fig. \ref{fig:PedagogAlt}a, b), and is increasingly negative as 
 the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ gets
 close to $\pi$, e.g., when $l_{e1}-l_{e2}+l_n$ is close to half the period of oscillation
-(Fig. \ref{fig:PedagogAlt}c, d). The factor $1/|f_{\mcB}|^2$ is how the intrinsic nature of population
+(Fig. \ref{fig:PedagogAlt}c, d). The factor $1/|f_{\mcB}|^2$ is again how the intrinsic nature of population
 dynamics modulates Moran influences.
-The derivation of the theory and details supporting the above interpretations of 
-constituents of the theory are in SI section \ref{SM-sect:derivation}. 
+The derivation of the theory and further information supporting the above interpretations 
+are in SI section \ref{SM-sect:derivation}. 
 %***DAN: I need to add some stuff to the SI section mentioned to support the claim that 
@@
 four model time steps, but with different phases. Peaks in $\epsilon^{(1)}$ are assumed to 
-either lead or lag peaks in $\epsilon^{(2)}$ by $1$q (we consider two scenarios, CaseC1 and CaseC2).
-Such scenarios could be realized by annually 
+either lead or lag peaks in $\epsilon^{(2)}$ by $1$q (we consider two sub-cases, CaseC1 and CaseC2).
+Such a situation could be realized by annually 
 periodic environmental fluctuations sampled quarterly,
@@
 Recall that, for CaseC, $\epsilon^{(1)}$ effects on populations were positive and lagged by $1$q, i.e., $l_{e1}=1$ in the 
-language of Fig. \ref{fig:PedagogAlt_SM}; whereas 
+language of Fig. \ref{fig:PedagogAlt}; whereas 
 $\epsilon^{(2)}$ effects were unlagged and negative, i.e., $l_{e2}=0$ (Methods). 
-We compare CaseC to Fig. \ref{fig:PedagogAlt_SM} instead of Fig. \ref{fig:PedagogAlt} because
-$\epsilon^{(2)}$ effects were negative for CaseC, the situation considered by Fig. \ref{fig:PedagogAlt_SM}.
+We compare CaseC to the Fig. \ref{fig:PedagogAlt} $\beta$ scenarios instead of the Fig. \ref{fig:PedagogAlt} 
+$\alpha$ scenarios because
+$\epsilon^{(2)}$ effects were negative for CaseC, the situation considered by the $\beta$ scenarios.
 For CaseC1, peaks in the periodic
@@
 Thus $l_n+l_{e1}-l_{e2}=2 \text{q}=0.5 \text{y}$, and, on annual timescales, lags compounded, similar to 
-Fig. \ref{fig:PedagogAlt_SM}b: the 
+Fig. \ref{fig:PedagogAlt}g, h: the 
 annual positive population effects of $\epsilon^{(1)}$ 
@@
 $1$q, so that $l_n=-1$. Thus $l_n+l_{e1}-l_{e2}=0$ and, on annual timescales, lags cancelled,
-similar to Fig. \ref{fig:PedagogAlt_SM}a: the annual positive population effects of $\epsilon^{(1)}$
+similar to Fig. \ref{fig:PedagogAlt}e, f: the annual positive population effects of $\epsilon^{(1)}$
 coincided with the negative effects of $\epsilon^{(2)}$ every year, because $\epsilon^{(1)}$
@@
 On long timescales, interactions were the
-same in the two scenarios, and were negative, because, on those timescales, sub-annual lags 
+same in the two sub-cases, and were negative, because, on those timescales, sub-annual lags 
 make negligible difference, and the effects of $\epsilon^{(1)}$ on populations were 
@@
 
-%***DAN: start from here after making the new version of the figure sketched in a separate planning doc
-
 %Direct Moran effects
@@
 %audience.
-The way direct Moran effects work in our theory is fairly straightforward. 
-%***DAN: Why is the following sentence not in the Theory section?
-For instance, the direct Moran effect
-term in (\ref{rhoww_T1to3}) for nitrates, 
-$|f_{\mcP^{(1)}}|^2 \rho_{\epsilon^{(1)} \epsilon^{(1)}}/|f_{\mcB}|^2$, 
-equals the synchrony of the nitrate time series themselves, 
-$\rho_{\epsilon^{(1)} \epsilon^{(1)}}$, times the 
-timescale-specific strength of influence of nitrates on kelp, 
-$|f_{\mcP^{(1)}}|^2$, and modulated
-by the autoregressive nature of kelp dynamics,
-$1/|f_{\mcB}|^2$. These components are plotted in Fig. \ref{fig:directMoran} (CCA1 and SB) and 
-Fig. \ref{SM-fig:directMoran_SM_lag4} (CCA2) so their 
-effects on synchrony can be assessed across timescales and regions.
+Direct Moran effects for kelp and how they manifest in our theopry are fairly straightforward. 
 Nitrates are henceforth identified with $\epsilon^{(1)}$ and waves with $\epsilon^{(2)}$. 
@@
 Thus nitrate and wave 
-synchrony had a strong annual component (Fig. \ref{fig:directMoran}b,d,q,s), which produced 
+synchrony had a strong annual component (Fig. \ref{fig:noise}a,b,j,k), which produced 
 some of the annual synchrony observed in kelp 
@@
 Nitrates and waves are also synchronous on long timescales $>8$y  
-(Fig. \ref{fig:directMoran}l,n,aa,cc), possibly due to the North Pacific Gyre Oscillation 
+(Fig. \ref{fig:noise}g,h,p,q), possibly due to the North Pacific Gyre Oscillation 
 \citep{CastoraniSheppardKelp}. The long-timescale synchrony
@@
 waves, was more pronounced in CCA1 than in SB 
-(Fig. \ref{fig:directMoran}b,g,l v. q,v,aa, and 
-Fig. \ref{fig:directMoran}d,i,n v. s,x,cc), and also because waves
-had a stronger influence on kelp in CCA1 than in SB (Fig. \ref{fig:directMoran}c,h,m v. r,w,bb).
-Regression coefficients determining $f_{\mcP^{(1)}}$, $f_{\mcP^{(2)}}$
+(Fig. \ref{fig:noise}), and also because waves
+had a stronger influence on kelp in CCA1 than in SB:
+regression coefficients determining $f_{\mcP^{(1)}}$, $f_{\mcP^{(2)}}$
 and $f_{\mcB}$ (Methods) are in Table \ref{SM-tab:regcoefs}.
 
-%***DAN: You need to combine the current figs 2 and 3, saving that for later.
-
-%***DAN: Stopped here 2022 11 15, 230pm
+%***DAN: start from here
 
@@
 drivers must always be compared to the timescale of interest to determine interaction effects,
-as in Figs \ref{fig:PedagogAlt}, \ref{fig:PedagogAlt_SM}, and as captured formally in our theory.
+as in Fig. \ref{fig:PedagogAlt}, and as captured formally in our theory.
 
@@
 \includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt.pdf}
+\includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt_SM.pdf}
 \caption{%\doublespacing 
-Illustration of the main concept of synergistically (a, b) or antagonistically (c, d) interacting Moran effects.
+Illustration of the main concept of synergistically or antagonistically interacting Moran effects.
 Interactions require that each environmental variable itself be spatially synchronous; and then 
-the alignment or misalignment of three types of lag (see below) determine the sign and strength of the interactions. 
+the alignment or misalignment of three types of lag determine the sign and strength of the interactions. 
 Solid-line sinusoids represent the period-20 components of an environmental 
@@
 a different environmental variable in the same locations
-($\epsilon_i^{(2)}$ for $i=1,2,3$). Black arrows represent peak positive influences of environmental 
+($\epsilon_i^{(2)}$ for $i=1,2,3$). For the scenarios $\alpha 1$ (a-b) and $\alpha 2$ (c-d), 
+black arrows represent}\label{fig:PedagogAlt}
+\end{figure}
+
+\clearpage
+
+\noindent peak positive influences of environmental 
 variables on populations, which are lagged by an amount $l_{e1}$ for $\epsilon_i^{(1)}$ and by 
-an amount $l_{e2}$ for $\epsilon_i^{(2)}$, where these lags differ across the scenarios but are the same 
-in all locations within a scenario.
+an amount $l_{e2}$ for $\epsilon_i^{(2)}$, where these lags differ across the scenarios $\alpha 1$ and $\alpha 2$, 
+but are the same in all locations within one of these scenarios.
 Analogously, red arrows represent maximally negative effects.
-Oscillations are separately spatially synchronous for each environmental variable in 
-each scenario, illustrated with blue and pink rectangles: due to this environmental synchrony, peak positive population effects 
-of the same variable occur at similar times across locations, and likewise for peak negative effects.
-In the synergistic scenario, the lag between the environmental variables ($l_n$) and the lags of the population effects
+Due to environmental synchrony, peak positive population effects 
+of the same variable occur at similar times across locations, and likewise for peak negative effects,
+illustrated with blue and pink rectangles.
+In the synergistic scenario ($\alpha 1$), the lag between 
+the environmental variables ($l_n$) and the lags of the population effects
 of the variables ($l_{e1}$ and $l_{e2}$) are aligned, i.e., $l_n+l_{e1}=l_{e2}$, so that peak positive 
@@
 aligned on a, b), augmenting synchrony. Likewise negative effects are aligned (blue rectangles).
-In the antagonistic scenario, lags are misaligned, i.e., $l_n+l_{e1}=l_{e2}+\sigma/2$, where $\sigma=20$ is the period, 
-so that peak positive 
+In the antagonistic scenario ($\alpha 2$), lags are misaligned, i.e., 
+$l_n+l_{e1}=l_{e2}+\sigma/2$, where $\sigma=20$ is the period, so that peak positive 
 effects of $\epsilon_i^{(1)}$ coincide with maximally negative effects of $\epsilon_i^{(2)}$, and  maximally 
@@
 (pink rectangles on c are aligned with blue ones on d, and vice versa), reducing synchrony.
-See the online version for color
-renderings of all figures.}\label{fig:PedagogAlt}
-\end{figure}
-
-\clearpage
-\newpage
-
-\begin{figure}
-\includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt_SM.pdf}
-\caption{Similar to Fig. \ref{fig:PedagogAlt}, but for slightly altered 
-scenarios. For Fig. \ref{fig:PedagogAlt}, for simplicity, we assumed that larger values of environmental variables had 
-more positive influences on populations. We here assume that larger values of $\epsilon_i^{(1)}$ have more 
-positive influences, but that smaller values of $\epsilon_i^{(2)}$ have more positive influences.
-In scenario 1 here, $l_n+l_{e1}=l_{e2}$, but Moran interactions are nonetheless antagonistic because the 
-altered direction of effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
-variable correspond to the peak negative effects of $\epsilon_i^{(1)}$. In scenario 2 here, 
+For the scenarios $\beta 1$ (e-f) and $\beta 2$ (g-h), we assume that larger values of $\epsilon_i^{(1)}$ have more 
+positive population influences (as in the $\alpha$ scenarios), but that larger values of 
+$\epsilon_i^{(2)}$ have more negative population influences (contrasting
+with the $\alpha$ scenarios).
+In scenario $\beta 1$, $l_n+l_{e1}=l_{e2}$, as in $\alpha 1$; but Moran interactions are nonetheless antagonistic because the 
+negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
+variable correspond to the peak negative effects of $\epsilon_i^{(1)}$. In scenario $\beta 2$, as in $\alpha 2$,
 $l_n+l_{e1}=l_{e2}+\sigma/2$ for $\sigma=20$ the period, but Moran interactions are synergistic because the 
-altered direction of effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
-variable correspond to the peak positive effects of $\epsilon_i^{(1)}$. See the online version for color
-renderings of all figures.}\label{fig:PedagogAlt_SM}
-\end{figure}
+negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
+variable correspond to the peak positive effects of $\epsilon_i^{(1)}$. 
+See the online version for color renderings of all figures.
 

### 2022-11-16 Code/Theory_PedagogFig_Alt.R

@@ -170,12 +170,13 @@ text(sinpks_1_x[1,4]+le1/2,1+.5,latex2exp::TeX("$l_{e1}$"),cex=0.7,adj=c(0.5,-.1
 #panel label and title
 if (maintext)
 {
-  mtext("Scenario 1: Synergistic interactions",side=3,line=0.1,cex=1,at=xlimits[1],adj=0)
+  mtext(expression(paste("Scenario ",alpha,"1: Synergistic interactions")),side=3,line=0,cex=1,at=xlimits[1],adj=0)
+  text(xlimits[1],ylimits[2],"(a)",adj=c(0,1))
 }else
 {
-  mtext("Scenario 1: Antagonistic interactions",side=3,line=0.1,cex=1,at=xlimits[1],adj=0)
+  mtext(expression(paste("Scenario ",beta,"1: Antagonistic interactions")),side=3,line=0,cex=1,at=xlimits[1],adj=0)
+  text(xlimits[1],ylimits[2],"(e)",adj=c(0,1))
 }
-text(xlimits[1],ylimits[2],"(a)",adj=c(0,1))
 
 #label ln
 par(xpd=NA)
@@ -312,7 +313,13 @@ text(sintro_2_x[2,3]+le2/2,vertoffset-1+2.2,latex2exp::TeX("$l_{e2}$"),cex=0.7,a
 text(sinpks_2_x[1,4]+le2/2,1+.5,latex2exp::TeX("$l_{e2}$"),cex=0.7,adj=c(0.5,-.1))
 
 #panel label
-text(xlimits[1],ylimits[2],"(b)",adj=c(0,1))
+if (maintext)
+{
+  text(xlimits[1],ylimits[2],"(b)",adj=c(0,1))
+} else
+{
+  text(xlimits[1],ylimits[2],"(f)",adj=c(0,1))
+}
 
 #label ln
 par(xpd=NA)
@@ -440,12 +447,13 @@ text(sinpks_1_x[1,4]+le1/4,1+.5,latex2exp::TeX("$l_{e1}$"),cex=0.7,adj=c(0.5,-.1
 #panel label and title
 if (maintext)
 {
-  mtext("Scenario 2: Antagonistic interactions",side=3,line=0.1,cex=1,at=xlimits[1],adj=0)
[…]

### 2022-11-22 Paper/Paper.Rnw
a bunch more edits

@@
 We argue that Moran interactions should be common.     
-Our general theory explains an important new aspect of a fundamental feature of spatiotemporal population dynamics. 
+Our theory and analysis explain an important new aspect of a fundamental feature of spatiotemporal population dynamics. 
 \end{abstract}
@@
 being pushed both by their own mother and father. If the fathers from seperate families
-were to synchronize their pushes, they would, of course, act as a Moran-like influence, tending to synchronize
+were to synchronize their pushes, they would act as a Moran-like influence, tending to synchronize
 the swinging motions of the children. Likewise, if the mothers from seperate families were to
@@
 were not appropriately coordinated, the children 
-would swing more or less randomly and would become asynchronous with each other, demonstrating
-an antagonistic interaction between synchronoizing effects, i.e., a tendency for the two synchronizing
+would swing more or less randomly, with small amplitude, and would become asynchronous with each other. 
+This second case demonstrates
+an antagonistic interaction between synchronizing effects, i.e., a tendency for the two synchronizing
 effects to cancel. As oscillators, children on swings are distinct in many ways from ecological
-populations, not least because swings oscillate on one frequency/on one timescale, whereas populations
+populations, not least because swings oscillate on one frequency/timescale, whereas populations
 typically oscillate simultaneously on many timescales, having a broadband spectrum. 
@@
 systems, because of the large number of interrelated factors driving population fluctuations in most 
-spatially extended systems. Interaction effects can also be strong. For these reasons and others, development of a 
+spatially extended systems. The same two studies demonstrated that 
+interaction effects can be strong. For these reasons and others, development of a 
 general theoretical and intutitive understanding of how interactions between Moran effects work is an 
@@
 for kelp ecology, we instead use a simplified subset of the available kelp data. Our kelp analysis is intended 
-to illuminate the inner workings of the general mechanism we describe rather than being a complete exploration
-of kelp dynamics. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
+to illuminate the inner workings of the general mechanism we describe rather than being an exploration
+of kelp dynamics, specifically. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
 illuminate this newly observed but potentially quite important mechanism of spatiotemporal population dynamics.
@@
 \noindent Prior to presenting our formal theory, we extend and formalize the intution behind it
-that we began to develop with the swing example in the Introduction. We again focus on a single
+that began with the swing analogy in the Introduction. We again focus on a single
 timescale of oscillation, later combining timescales mathematically. 
@@
 antagonistic Moran interactions are reversed, but with the same general concepts still operating 
-(Fig. \ref{fig:PedagogAlt}).
+(Fig. \ref{fig:PedagogAlt2}).
 
@@
 $w_j(t)$ are both strong and strongly correlated, i.e., they have the same or similar phase. 
-(Here, $j$ refers to another location where measurements were taken.) The cospectrum is 
+Here, $j$ refers to another location where measurements were taken. The cospectrum is 
 the real part of the cross spectrum, $S_{w_i w_j}(f)$, which is a complex-valued function of frequency.
@@
 $\rho_{ww}=\left( \sum_{i \neq j} S_{w_i w_j} \right)/(N^2-N)$, the average of the cross spectra for all pairs
-of distinct locations. This is a real-valued function of frequency, the integral (across frequencies) of which
+of distinct locations. This is a real-valued function of frequency, an integral (across frequencies) of which
 is the classic, non-frequency-specific synchrony 
@@
 cross-variable synchrony (or, simply, \emph{cross synchrony}) between the variables
-is $\rho_{\epsilon^{(1)}\epsilon^{(2)}}=\left( \sum_{i \neq j} S_{\epsilon_i^{(1)}\epsilon_j^{(2)}} \right)/(N^2-N)$. 
+is defined as $\rho_{\epsilon^{(1)}\epsilon^{(2)}}=\left( \sum_{i \neq j} S_{\epsilon_i^{(1)}\epsilon_j^{(2)}} \right)/(N^2-N)$. 
 This is interpretable as pertaining to spatial synchrony because it makes comparisons across 
@@
 phase at that frequency equal to $\pi/2$, reflecting the one-year lag.
-See SI section \ref{SM-sect:examples_cross_sync} for further examples of the new cross synchrony index.
+See SI section \ref{SM-sect:examples_cross_sync} for detailed examples.
 %***DAN: If we go longer, you could bring these examples here and it would probably be a positive addition of
 %about one para and one new Fig. Would be a positive addition because it helps people understand our new metric.
+%***DAN: Upon further reflection, I'm not totally sure of that. Those examples in the sup mat are pretty technical.
+%If I brought a version of them here, it would have to be a stripped down, intuitive, version, with the math removed.
+%And the new fig would have to just be panel c or f from Fig. S1. That's still possible to do, but does it really
+%help more than just the verbal description above?
 
@@
 %Holt as reviewer? Will he even do it?
-%Cohen? Probably a conflict.
+%Cohen? 
 %Arnoldi? Bart Haegeman?
 %Montoya?
-
-%***DAN: Question for later: Suppose we considered a model with three measured env drivers, would you get
-%three-way interactions? You might add the question to the Discussion para about adding dispersal to this model.
-%Not sure actually doing that is a good use of my time, though.
 
@@
 $\mu=\exp(-2 \pi \iota f)$, and where $\iota$, the Greek letter iota, is the imaginary unit.
+The derivation of the theory is in SI section \ref{SM-sect:derivation}.
 
@@
 
-The way direct Moran effects work in our theory is fairly straightforward.
+The way direct Moran effects in our theory are interpreted is fairly straightforward.
 The magnitudes of the quantities $f_{\mcP^{(1)}}$ and $f_{\mcP^{(2)}}$ quantify the strength of 
@@
 
-The components of the interacting Moran effects term are also interpretable.
+The components of the interacting Moran effects term in our theory are also interpretable.
 The phases of the quantities $f_{\mcP^{(1)}}$ and $f_{\mcP^{(2)}}$ quantify the lags in the population influences of 
@@
 represented on Fig. \ref{fig:PedagogAlt} by $l_{e1}$ and $l_{e2}$,
-relative to the timescale $\sigma$. The lag 
+relative to the timescale $\sigma=1/f$. The lag 
 $l_n$ on Fig. \ref{fig:PedagogAlt} manifests in the theory 
@@
 close to $\pi$, e.g., when $l_{e1}-l_{e2}+l_n$ is close to half the period of oscillation
-(Fig. \ref{fig:PedagogAlt}c, d). The factor $1/|f_{\mcB}|^2$ is again how the intrinsic nature of population
-dynamics modulates Moran influences.
-The derivation of the theory and further information supporting the above interpretations 
-are in SI section \ref{SM-sect:derivation}. 
+(Fig. \ref{fig:PedagogAlt}c, d). The factor $1/|f_{\mcB}|^2$ again captures how 
+the intrinsic nature of population dynamics modulates Moran influences.
 %***DAN: I need to add some stuff to the SI section mentioned to support the claim that 
 %"details supporting the above assertions are in SI section X"
+%***DAN: I am currently (2022 11 17) unsure of how much of this belongs in the main text
+%versus the sup mat, and how to structure new material I will add to the sup mat, which will 
+%influence what section I site above (currently not the right section). 
+%***DAN: One way to go may be to put the 2 paras immediately above at the end of the theory
+%derivation section in sup mat, and just state here that interpretations of theory 
+%consituents, as well as the derivation itself, are in that section. The you could add 
+%a detailed elaboration of the theoretical constituents for the theory case studies to S6,
+%including a reinstatement of the components of Fig. 5 that were cut. And then I guess you could
+%add a new SI section S7 which would have a detailed exploration of the theory constituents
+%for kelp.
+%***DAN: Decision, and also started implementing a bit: Keep the above paras, no extension of
+%them in sup mat, but do add 
+%a detailed elaboration of the theoretical constituents for the theory case studies to S6,
+%including a reinstatement of the components of Fig. 5 that were cut. And then 
+%add a new SI section S7 which would have a detailed exploration of the theory constituents
+%for kelp.
 
@@
 $p_{1}^{(1)} > 0$ while $p_i^{(1)}=0$ for $i \neq 1$, and 
-$p_{0}^{(2)} < 0$ while $p_i^{(2)}=0$ for $i \neq n_2$.
+$p_{0}^{(2)} < 0$ while $p_i^{(2)}=0$ for $i \neq 0$.
 The noise process
@@
 in (\ref{main_model_1})-(\ref{main_model_3}), and therefore $f_{\mcB}$, $f_{\mcP^{(1)}}$ and
-$f_{\mcP^{(2)}}$ in (\ref{rhoww_T1to3}), for each of our three regions (see General Theory).
+$f_{\mcP^{(2)}}$ in (\ref{rhoww_T1to3}), for each of our three regions.
 
@@
 $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ in (\ref{rhoww_T1to3}), 
-we applied the definitions of these quantities (see General Theory), which required estimation from data of
+we applied the definitions of these quantities (General Theory), which required estimation from data of
 the spectra and cross spectra $S_{w_i w_j}$ and $S_{\epsilon_i^{(a)}\epsilon_j^{(b)}}$ for $i,j=1,\ldots,N$
@@
 the two noise variables reinforce each other.
-%***DAN: I comented the below because I want to put it in sup mat. Once you write that section, consider
-%referring to it here with something like: "See SI section X for a detailed breakdown of how the terms in 
-%our theory capture these complexities."
-%Theory captures these complexities because
-%the size of the $1$q lag of this scenario, relative to a timescale, $\sigma$, is captured by
-%the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, which equals the phase of
-%$f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ because $l_n=0$ and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number
-%(SI section \ref{SM-sect:examples_cross_sync}). But $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=
-%p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) = 
-%p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}), 
-%which has phase $-\pi$ at short timescales
-%and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b,c).
-%The phase controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, the
-%interaction term in (\ref{rhoww_T1to3}), as described in General Theory. 
 Comparing
-Fig. \ref{fig:TheoryCaseA}a and d shows the additional influence of autoregressive population effects, which
+Fig. \ref{fig:TheoryCaseA}a and b shows the additional influence of autoregressive population effects, which
 simply multiply all Moran influences by the same timescale-dependent non-negative 
 quantity, not altering their relative importance or sign [see (\ref{rhoww_T1to3})].
+%See SI section \ref{SM-sect:breakdown_case_studies} for a detailed technical breakdown of how the terms in 
+%our theory capture the complexities described above.
 
@@
 were, respectively, positive and negative for these timescales (Fig. \ref{fig:TheoryCaseA}c). 
+%See SI section \ref{SM-sect:breakdown_case_studies} for a detailed technical breakdown of how the terms in 
+%our theory capture the complexities described above.
 %***DAN: May again want to cite a new section of sup mat, in a similar way suggested in a comment in the previous paragraph
@@
 CaseA and CaseB assumed white noise processes, and therefore $l_n$ was $0$.
-But CaseC illustrates what can happen when noise processes have periodic components and lagged associations, e.g., 
+But CaseC illustrates what can happen when noise processes have lagged associations, e.g., 
 both processes oscillate with annual periodicity and distinct phenology, an important scenario because seasonality is common.
 Recall that, for CaseC, $\epsilon^{(1)}$ effects on populations were positive and lagged by $1$q, i.e., $l_{e1}=1$ in the 
-language of Fig. \ref{fig:PedagogAlt}; whereas 
+language of Fig. \ref{fig:PedagogAlt2}; whereas 
 $\epsilon^{(2)}$ effects were unlagged and negative, i.e., $l_{e2}=0$ (Methods). 
-We compare CaseC to the Fig. \ref{fig:PedagogAlt} $\beta$ scenarios instead of the Fig. \ref{fig:PedagogAlt} 
+We compare CaseC to the Fig. \ref{fig:PedagogAlt2} $\beta$ scenarios instead of the Fig. \ref{fig:PedagogAlt} 
 $\alpha$ scenarios because
@@
 Thus $l_n+l_{e1}-l_{e2}=2 \text{q}=0.5 \text{y}$, and, on annual timescales, lags compounded, similar to 
-Fig. \ref{fig:PedagogAlt}g, h: the 
+Fig. \ref{fig:PedagogAlt2}c, d: the 
 annual positive population effects of $\epsilon^{(1)}$ 
@@
 $1$q, so that $l_n=-1$. Thus $l_n+l_{e1}-l_{e2}=0$ and, on annual timescales, lags cancelled,
-similar to Fig. \ref{fig:PedagogAlt}e, f: the annual positive population effects of $\epsilon^{(1)}$
+similar to Fig. \ref{fig:PedagogAlt2}a, b: the annual positive population effects of $\epsilon^{(1)}$
 coincided with the negative effects of $\epsilon^{(2)}$ every year, because $\epsilon^{(1)}$
@@
 On long timescales, interactions were the
-same in the two sub-cases, and were negative, because, on those timescales, sub-annual lags 
+same for C1 and C2, and were negative, because, on those timescales, sub-annual lags 
 make negligible difference, and the effects of $\epsilon^{(1)}$ on populations were 
 positive and those of $\epsilon^{(2)}$ were negative.
+%See SI section \ref{SM-sect:breakdown_case_studies} for a detailed technical breakdown of how the terms in 
+%our theory capture the complexities described above.
 
@@
 %Direct Moran effects
-%***DAN: Do I really want to spend all this space (and reader mental energy), including 
-%a very complex figure, on direct Moran effects, when the paper is not even about that?
-%***DAN: And this paragraph also requires plotting all these inner portions of the
-%theory, which I am starting to think is not really going to work for an ecological 
-%audience.
-Direct Moran effects for kelp and how they manifest in our theopry are fairly straightforward. 
+Direct Moran effects for kelp and how they manifest in our theory are fairly straightforward. 
 Nitrates are henceforth identified with $\epsilon^{(1)}$ and waves with $\epsilon^{(2)}$. 
@@
 Thus nitrate and wave 
-synchrony had a strong annual component (Fig. \ref{fig:noise}a,b,j,k), which produced 
+synchrony had a strong annual component (Fig. \ref{fig:noise}a,b,j,k for the CCA1 and SB regions,
+Fig. \ref{SM-fig:directMoran_SM_lag4} for CCA2), which produced 
 some of the annual synchrony observed in kelp 
@@
 Nitrates and waves are also synchronous on long timescales $>8$y  
-(Fig. \ref{fig:noise}g,h,p,q), possibly due to the North Pacific Gyre Oscillation 
+(Fig. \ref{fig:noise}g,h,p,q for the CCA1 and SB regions, Fig. \ref{SM-fig:directMoran_SM_lag4} 
+for CCA2), possibly due to the North Pacific Gyre Oscillation 
 \citep{CastoraniSheppardKelp}. The long-timescale synchrony
@@
 (Fig. \ref{fig:noise}), and also because waves
-had a stronger influence on kelp in CCA1 than in SB:
+had a stronger influence on kelp in central California than in SB:
 regression coefficients determining $f_{\mcP^{(1)}}$, $f_{\mcP^{(2)}}$
-and $f_{\mcB}$ (Methods) are in Table \ref{SM-tab:regcoefs}.
-
-%***DAN: start from here
+and $f_{\mcB}$ are in Table \ref{SM-tab:regcoefs}.
+%***DAN: Later you will have to make sure all rules are followed for citing all sup mat figures and 
+%tables and sections from the main text, in order. I think currently they are not all cited, in part
+%because I have been shifting around what information is where in the whole paper.
 
 %Interactions between Moran effects, CCA1
-%***DAN: In previous comments I have begun outlining how I might simplify
-%and make the paper more about comparing the intuition of Fig. 2 and 5 to 
-%the *predictions* of theory, rather than taking the reader through all the
-%inner workings of theory. I like those thoughts so far, but they present some chanllenges
-%when we come to this paragraph. Much of what is stated below can be demonstrated
-%by looking at the estimates of the constituents of the theory, but I think that'll
-%be too confusing. So we have to come up with other ways to demo it. I think the 
-%solution is: we feel free to plot our synchrony and cross synchrony measures,
-%rho_epsiloni_epsilonj, even though in the case of i differnt from j, we have to
-%plot a complex quantity. But instead of plotting the fs, we just talk about
-%lags garnered from regression.
 Interactions between Moran effects in CCA1 were parallel to one of the theoretical case studies, CaseC1.
-In CCA1, nitrates had a $1$q-delayed positive effect on observed 
-kelp populations and waves had immediate negative effect \citep{Cavanaugh2011,Bell2015,Schiel2015}, 
-just as $\epsilon^{(1)}$ effects in CaseC 
+In CCA1, nitrates had $1$q-delayed positive effects on observed 
+kelp populations and waves had immediate negative effects \citep{Cavanaugh2011,Bell2015,Schiel2015}, 
+just as $\epsilon^{(1)}$ effects in CaseC1 
 were positive and delayed by $1$q and $\epsilon^{(2)}$ effects were negative and immediate.
-%***DAN: So we would drop the following sentence, or send it to a new section in 
-%sup mat where we systematically explored the constituents of the theory. Instead
-%we would just rely on the regression results which follow.
-Delayed effects are reflected, for CCA1, in Fig. \ref{fig:indirectMoran}a, and for 
-CaseC1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
-that $f_{\mcP^{(1)}}$ has phase about $-\pi/2$ at the annual timescale. 
-%***DAN: The above is pretty horrible. The reader has to locate two panels on two differently
-%formatted and very complex figures and compare them to see this! They will not do that!
-The delay is also 
-reflected, for CCA1, in Table \ref{SM-tab:regcoefs}, where 
-the entries for $p_0^{(1)}$ for CCA1 are close to $0$ and those for $p_1^{(1)}$ are positive.
-%***DAN: On the other hand, the contents of the table are pretty clear and intutitive to
-%an ecological audience, and can be used to match the intution of lags (Figs 2 and 5) with the
-%*outputs* of the theory, without plotting all the inner working in the main text. Even just
-%plotting both phase and magnitude on the same axes is going to be a stretch, and we have been
-%doing that a LOT! I'm really thinking that dragging the reader through all the inner workings
-%of this theory is just not going to work, unless I want to publish in Theoretical Population
-%Biology or something.
-%***DAN: So we would also drop the following sentence, relying instead
-%on the regression results which follow.
-Immediate negative effects of waves in CCA1 are reflected in Fig. \ref{fig:indirectMoran}b,g,l
-and for CaseC1 in Fig. \ref{fig:TheoryCaseA}j, both of which show 
-that $f_{\mcP^{(2)}}$ has phase $-\pi$ at all timescales. See also the $p_0^{(2)}$
-coefficients for CCA1 in Table \ref{SM-tab:regcoefs}, which are negative.
+%Delayed effects are reflected, for CCA1, in Fig. \ref{fig:indirectMoran}a, and for 
+%CaseC1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
+%that $f_{\mcP^{(1)}}$ has phase about $-\pi/2$ at the annual timescale. 
+The delayed effects of nitrates in CCA1 are reflected in Table \ref{SM-tab:regcoefs}, where 
+regression results for $p_0^{(1)}$ are close to $0$ and those for $p_1^{(1)}$ are positive.
+%Immediate negative effects of waves in CCA1 are reflected in Fig. \ref{fig:indirectMoran}b,g,l
+%and for CaseC1 in Fig. \ref{fig:TheoryCaseA}j, both of which show 
+%that $f_{\mcP^{(2)}}$ has phase $-\pi$ at all timescales. 
+Immediate negative effects of waves in CCA1 are also reflected in Table \ref{SM-tab:regcoefs},
+where entries for $p_0^{(2)}$ are negative.
 In CCA1, annual peak nitrates tend to come in spring, whereas annual peak waves 
@@
 %stuff.
-This is reflected, for CCA1, in Fig. \ref{fig:indirectMoran}d, and for CaseC1 
-in Fig. \ref{fig:TheoryCaseA}k, both of which show that $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$
+For CCA1, this is reflected in Fig. \ref{fig:noise}c, which shows that 
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$
 has a phase of about $-\pi/2$ at the annual timescale.
-Thus lags compound on the annual timescale, in the CCA1 case as in the theoretical case: the annual positive 
+Thus lags compound on the annual timescale, in the CCA1 case as in CaseC1: the annual positive 
 effects of nitrates on kelp tend to come in summer, and the annual negative effects of waves come in winter,
 producing reinforcing interactions between Moran effects. 
-%***DAN: Following sentence moved to sup mat.
-Likewise, the $\pi/2$ phase 
-of $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}$ on annual timescales (Fig. \ref{fig:indirectMoran}c for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1)
-approximately cancels the $-\pi/2$ phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ on annual timescales
-(Fig. \ref{fig:indirectMoran}d for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1), so that  
-$f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ has phase 
-approximately $0$, making the interaction-effect term 
-$\Rp(f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}})$ strongly positive 
-on annual timescales for both the theoretical and emprical cases 
-(Fig. \ref{fig:indirectMoran}e for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1). 
+%Likewise, the $\pi/2$ phase 
+%of $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}$ on annual timescales 
+%(Fig. \ref{fig:indirectMoran}c for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1)
+%approximately cancels the $-\pi/2$ phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ on annual timescales
+%(Fig. \ref{fig:indirectMoran}d for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1), so that  
+%$f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ has phase 
+%approximately $0$, making the interaction-effect term 
+%$\Rp(f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}})$ strongly positive 
+%on annual timescales for both the theoretical and emprical cases 
+%(Fig. \ref{fig:indirectMoran}e for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1). 
 Interactions between Moran effects on long-timescales ($>8$y) were also similar for both CCA1 and
-CaseC1, and for the same reason in both cases: on long timescales, sub-annual lags
+CaseC1 (compare Figs \ref{fig:TheoryCaseA}e and \ref{fig:SummaryKelpFig}c), 
+and for the same reason in both cases: on long timescales, sub-annual lags
 are inconsequential, and the positive effects of nitrates/$\epsilon^{(1)}$ and the negative
@@
 mechanism in SB compared to CCA1, leading to the slightly negative versus positive interactions already 
-documented on annual timescales for the two regions in Fig. \ref{fig:SummaryKelpFig}d v. a. 
+documented on annual timescales for the two regions (Fig. \ref{fig:SummaryKelpFig}d v. a). 
 Whereas in CCA1, nitrate effects on observed kelp density were delayed by $1$q, in SB nitrate effects 
-were observable within the same quarter 
-%***DAN: Just rely on the regression results here, how the complex spectral stuff
-(Fig. \ref{fig:indirectMoran}a v. p; see also Table 
-\ref{SM-tab:regcoefs}, where $p_0^{(1)}$ terms were close to zero for CCA1 and positive for SB,
-whereas $p_1^{(1)}$ terms were positive for CCA1 and close to zero for SB). Wave effects were negative and
-immediate in both regions, though they were less pronounced in SB 
-%***DAN: Same, not only for the negativity, but also for the fact that wave effects were less pronounced in SB.
-(Fig. \ref{fig:indirectMoran}b,g,l v. q,v,aa; see also 
-Table \ref{SM-tab:regcoefs}). 
-%***DAN: The following setence will stay in main text, so we will have to keep a version of Fig. 8d and s,
-%although most of the rest of that figure will be simplified away.
+were observable within the same quarter: 
+%(Fig. \ref{fig:indirectMoran}a v. p; see also 
+see Table \ref{SM-tab:regcoefs}, where $p_0^{(1)}$ terms were close to zero for CCA1 and positive for SB,
+whereas $p_1^{(1)}$ terms were positive for CCA1 and close to zero for SB. Wave effects were negative and
+immediate in both regions, i.e., $p_0^{(2)}<0$ in Table \ref{SM-tab:regcoefs}. 
+%(Fig. \ref{fig:indirectMoran}b,g,l v. q,v,aa; see also 
+%Table \ref{SM-tab:regcoefs}). 
 Annual peaks in nitrates still tended to lag annual peaks in waves by $1$q in SB,
 as in CCA1, though the seasonal periodicity of both variables was reduced in SB compared to 
-CCA1 (Fig. \ref{fig:indirectMoran}s v. d). 
-Whereas in CCA1, the tendency of nitrate annual peaks to follow wave annual peaks by $1$q, added to the tendency of 
+CCA1 (Fig. \ref{fig:noise}l v. c). 
+Whereas in CCA1, the tendency of nitrate annual peaks to follow wave annual peaks by $1$q, together with to the tendency of 
 nitrate effects on kelp to be delayed by $1$q, resulted in summer positive nitrate effects and winter 
@@
 while waves still had negative effects on kelp in winter. Thus the effects of nitrates and waves in SB
-were approximately a quarter-cycle separated from each other with respect to the annual timescale (Fig \ref{fig:indirectMoran}t), 
+were approximately a quarter-cycle separated from each other with respect to the annual timescale, 
 and so produced neither
@@
 alignments described above. On long timescales, interactions between Moran effects were analogous in CCA1 and SB
-(though weaker in SB; Fig. \ref{fig:indirectMoran}o v. dd and Fig. \ref{fig:SummaryKelpFig}c v. f) because, again, 
+(though weaker in SB; Fig. \ref{fig:SummaryKelpFig}c v. f) because, again, 
 sub-annual lags are inconsequential on long timescales and what mattered instead was the oppositely signed 
@@
 sign. We used our theory and several case studies based on models and 
-kelp populations to provide intuition about the new mechanism (see also below).
-Moran effects are ubiquitous. Interactions 
-between Moran effects have been tested for only twice previously, and were observed
-on both occassions \citep{Sheppard2019,CastoraniSheppardKelp}; 
-for this and other reasons (see below) interactions may also be common. 
+kelp populations to provide intuition about the new mechanism.
+Because Moran effects are ubiquitous and interactions 
+between Moran effects were detected on both of the two occassions they have been tested 
+for \citep{Sheppard2019,CastoraniSheppardKelp}, interactions may be common. 
 Moran interactions are therefore
 a newly recognized and potentially widespread aspect of a fundamental means (Moran effects) by which 
-environmental factors influence populations
-and diverse phenomena such as ecosystem stability \citep{Wilcox2017}, population cycling \citep{Anderson2021} 
-and extinction \citep{Ghosh2020}.
+environmental factors influence populations.
 Climate change is altering many aspects of environmental variables, including their means, variances and
@@
 relationships between environmental variables and the nature of 
-their influences on populations. There is therefore potential for climate change to alter interactions 
-between Moran effects, in ways our new theory may help researchers to understand (see below).  
+their influences on populations. There is therefore also potential for climate change to alter interactions 
+between Moran effects, in ways our new theory may help researchers to understand.  
 The augmented fundamental understanding of Moran effects which we have provided may 
@@
 are influenced by more than one environmental driver. Drivers are frequently spatially autocorrelated,
-and are also often related to each other because of their common origin in underlying meteorological 
-phenomena, including major climatic oscillations such as 
-the North Pacific Gyre Oscillation or El Ni\~{n}o Southern Oscillation (ENSO).
+and are also often related to each other because of their common origin in underlying climatic 
+phenomena such as the North Pacific Gyre Oscillation or El Ni\~{n}o Southern Oscillation (ENSO).
 Driver pairs which may commonly produce interacting Moran effects include instances where the same quantity 
-is measured in distinct parts of the year (e.g., spring and summer temperatures, or March and April rainfall,
-or December and January snowfall); or when distinct variables are measured in the same part of the 
+is measured in distinct parts of the year (e.g., spring and summer temperatures, or March and April rainfall); 
+or when distinct variables are measured in the same part of the 
 year (e.g., spring temperature and precipitation). Future work should systematically 
 investigate cross synchrony ($\rho_{\epsilon^{(1)} \epsilon^{(2)}}$) of environmental 
-variables, and could then use our theory
-to assess the potential for interacting Moran effects under a variety of biological scenarios. Since 
-interactions between Moran effects have been discovered in both of the two systems so far 
-tested \citep{Sheppard2019,CastoraniSheppardKelp}, future work should also, when
-sufficient data are available, test directly for interactions. Temperature and precipitation variables
-measured in the same season may be particularly promising candidates for producing interactions because of
+variables. Temperature and precipitation variables
+measured in the same season may be particularly important candidates for interactions because of
 the well known joint influence of these variables on plants.
-\cite{Defriez2017b} tested for interactions between Moran influences
-of temperature and precipitation on aspects of synchrony of a terrestrial plant biomass index, finding no 
-evidence for interactions. But their data were global in extent and spatially coarse, their time series were 
-short, and they used methods more aligned with their goals of 
-understanding the geography of synchrony than with our present goals of understanding 
-synchrony itself. So temperature-precipitation interactions may still be found in future work
-by applying our theory and the methods of \cite{Sheppard2019} and \cite{CastoraniSheppardKelp} to
-suitable data.
 
 %*DONE Revist intuition
-We revisit and extend the intuition behind interacting Moran effects, using the central CA kelp
+%***DAN: Do we still need this if we're cutting? Are we cutting?
+We again revisit the intuition behind interacting Moran effects, using the central CA kelp
 example as a vehicle. Large winter waves 
@@
 %*DONE Comparison to Castorani et al paper, and maybe also a bit on comparison to the Sheppard et al 2019 paper
+%***DAN: Could we put this in sup mat if we're cutting? Are we cutting?
 Our results about kelp were consistent with those of \cite{CastoraniSheppardKelp}, though
-these two studies use distinct methods; the studies complement each other in important ways. In spite of 
+that study uses distinct methods; our results complement the results of \cite{CastoraniSheppardKelp} 
+in important ways. In spite of 
 numerous methodological choices which differed between the 
@@
 %*DONE relationship to earlier papers comparing synchrony of drivers and pops
+%***DAN: I guess you could move this to sup mat if we're cutting
 It has been a frequent topic of research why populations are often less synchronous, or sychronous over a smaller
@@
 
-%*DONE Revisit the kendall paper, briefly. Talk about possible three-way interactions.
-\cite{Kendall2000} considered interactions between mechanisms of synchrony. Their research 
-questions were distinct from ours, but combining their viewpoints and ours may lead to valuable
-future work. \cite{Kendall2000} considered interactions between dispersal and Moran causes of 
-synchrony, whereas we analyzed interactions between two Moran drivers, a distinct context.
-But it seems likely that numuerous ecological systems involve important dispersal effects on 
-synchrony as well as influences of at least two Moran effects. It may be necessary in future
-work to consider three-way interactions between dispersal and two Moran drivers and their 
-influence on synchrony. Dispersal can readily be added to our modelling framework.
-\cite{Desharnais2018} performed a spectral analysis on a model similar to ours
-which included dispersal but which only considered two habitat patches and one environmental driver.
+% %*DONE Revisit the kendall paper, briefly. Talk about possible three-way interactions.
+% %***DAN: Could def put this in sup mat if we're cutting
+% \cite{Kendall2000} considered interactions between mechanisms of synchrony. Their research 
+% questions were distinct from ours, but combining their viewpoints and ours may lead to valuable
+% future work. \cite{Kendall2000} considered interactions between dispersal and Moran causes of 
+% synchrony, whereas we analyzed interactions between two Moran drivers, a distinct context.
+% But it seems likely that numuerous ecological systems involve important dispersal effects on 
+% synchrony as well as influences of at least two Moran effects. It may be necessary in future
+% work to consider three-way interactions between dispersal and two Moran drivers and their 
+% influence on synchrony. Dispersal can readily be added to our modelling framework.
+% \cite{Desharnais2018} performed a spectral analysis on a model similar to ours
+% which included dispersal but which only considered two habitat patches and one environmental driver.
 
@@
 then revealing the effects of elevated nitrates. In southern CA, kelp are less likely to be
-removed by winter waves. Effects of elevated nitrates on the growth of kelp stands already reaching
+completely removed by winter waves. Effects of elevated nitrates on the growth of kelp stands already reaching
 the surface can be immediately observed. Climate change may similarly alter 
@@
 
-%*DONE How/why a frequency-specific approach is necessary for all this
-An ever growing body of work shows that synchrony is, and should be analyzed as, a timescale-specific 
-phenomenon \citep{Vasseur2007,Keitt2008,Vasseur2014,Sheppard2016,Defriez2016,Defriez2017,Defriez2017b,Desharnais2018,Anderson2021}.
-The results of our present study extend this earlier work by showing that interactions between Moran effects 
-are also timescale-specific, differing stongly in strength, and in sign, across timescales. 
-A non-timescale-specific version of our theory could be obtained by appropriately integrating our main equation 
-(\ref{rhoww_T1to3}) across frequencies. But doing so would reduce insight by conflating distinct and
-often disparate processes. For instance, if the terms of (\ref{rhoww_T1to3}) were integrated 
-across frequencies for our central CA kelp example, positive interactions between Moran drivers on short timescales 
-(Fig. \ref{fig:SummaryKelpFig}a) would be summed with and partially cancel 
-negative interactions on long timescales (Fig. \ref{fig:SummaryKelpFig}c),
-falsely suggesting reduced importance of interaction effects. Our results also have implications given the 
-short duration of most time series in ecology. 
-Analyses based on short time series can only be expected to reveal Moran effects or interactions between Moran
-effects occurring on short timescales, which our results show can differ substantially from long-timescale results. 
-It has not escaped our notice that kelp time series even longer than ours, once 
-available, may reveal novel Moran effects and interactions on timescales longer than ours. 
-This provides another argument, among many others, in support of 
-long-term ecosystem monitoring.
+% %*DONE How/why a frequency-specific approach is necessary for all this
+% %***DAN: You could also move this to sup mat, as the point that "timescale structure is important"
+% %has been made many other places. Though I DO like the ending of this para, the ending of the prior
+% %para is also a good place to stop, maybe better. 
+% An ever growing body of work shows that synchrony is, and should be analyzed as, a timescale-specific 
+% phenomenon \citep{Vasseur2007,Keitt2008,Vasseur2014,Sheppard2016,Defriez2016,Defriez2017,Defriez2017b,Desharnais2018,Anderson2021}.
+% The results of our present study extend this earlier work by showing that interactions between Moran effects 
+% are also timescale-specific, differing stongly in strength, and in sign, across timescales. 
+% A non-timescale-specific version of our theory could be obtained by appropriately integrating our main equation 
+% (\ref{rhoww_T1to3}) across frequencies. But doing so would reduce insight by conflating distinct and
+% often disparate processes. For instance, if the terms of (\ref{rhoww_T1to3}) were integrated 
+% across frequencies for our central CA kelp example, positive interactions between Moran drivers on short timescales 
+% (Fig. \ref{fig:SummaryKelpFig}a) would be summed with and partially cancel 
+% negative interactions on long timescales (Fig. \ref{fig:SummaryKelpFig}c),
+% falsely suggesting reduced importance of interaction effects. Our results also have implications given the 
+% short duration of most time series in ecology. 
+% Analyses based on short time series can only be expected to reveal Moran effects or interactions between Moran
+% effects occurring on short timescales, which our results show can differ substantially from long-timescale results. 
+% It has not escaped our notice that kelp time series even longer than ours, once 
+% available, may reveal novel Moran effects and interactions on timescales longer than ours, 
+% providing another argument, among many, in support of 
+% long-term ecosystem monitoring.
 
@@
 \includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt.pdf}
-\includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt_SM.pdf}
 \caption{%\doublespacing 
@@
 a different environmental variable in the same locations
-($\epsilon_i^{(2)}$ for $i=1,2,3$). For the scenarios $\alpha 1$ (a-b) and $\alpha 2$ (c-d), 
-black arrows represent}\label{fig:PedagogAlt}
-\end{figure}
-
-\clearpage
-
-\noindent peak positive influences of environmental 
+($\epsilon_i^{(2)}$ for $i=1,2,3$). Black arrows represent peak positive influences of environmental 
 variables on populations, which are lagged by an amount $l_{e1}$ for $\epsilon_i^{(1)}$ and by 
@@
 (pink rectangles on c are aligned with blue ones on d, and vice versa), reducing synchrony.
-For the scenarios $\beta 1$ (e-f) and $\beta 2$ (g-h), we assume that larger values of $\epsilon_i^{(1)}$ have more 
-positive population influences (as in the $\alpha$ scenarios), but that larger values of 
-$\epsilon_i^{(2)}$ have more negative population influences (contrasting
-with the $\alpha$ scenarios).
-In scenario $\beta 1$, $l_n+l_{e1}=l_{e2}$, as in $\alpha 1$; but Moran interactions are nonetheless antagonistic because the 
+See the online version for color renderings of all figures.}\label{fig:PedagogAlt}
+\end{figure}
+
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt_SM.pdf}
+\caption{%\doublespacing 
+Similar to Fig. \ref{fig:PedagogAlt}, but for slightly altered scenarios. For Fig. \ref{fig:PedagogAlt},
+for simplicity, we assumed that larger values of environmental variables had more positive influences
+on populations. We here assume that larger values of $\epsilon_i^{(1)}$ have more 
+positive population influences (as in the $\alpha$ scenarios of Fig. \ref{fig:PedagogAlt}), 
+but that larger values of $\epsilon_i^{(2)}$ have more negative population influences (contrasting
+with the Fig. \ref{fig:PedagogAlt} scenarios). In scenario $\beta 1$, 
+$l_n+l_{e1}=l_{e2}$, as in $\alpha 1$; but Moran interactions are nonetheless antagonistic because the 
 negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
@@
 variable correspond to the peak positive effects of $\epsilon_i^{(1)}$. 
-See the online version for color renderings of all figures.
+See the online version for color renderings of all figures.}\label{fig:PedagogAlt2}
+\end{figure}
 
@@
 \caption{%\doublespacing 
-Theoretical case studies A (panels a-b), B (panels c-d) and C (e-h, two alternative sets of 
-parameters used for e-f and g-h, respectively). Left panels (a, c, e, g) show the terms on the right side
+Theoretical case studies A (panels a-b), B (c-d), C1 (e-h) and C2 (g-h). 
+Left panels (a, c, e, g) show the terms on the right side
 of (\ref{rhoww_T1to3}). On those panels, the green line is 
@@
 %$c_1=0$ and $c_2=-4/9$ 
-for case study C. For C, peaks in the periodic noise process $\epsilon^{(2)}$
+for case study C. For C1, peaks in the periodic noise process $\epsilon^{(2)}$
 lagged peaks in the periodic process $\epsilon^{(1)}$ by 1 quarter
-in panels e-f, and $\epsilon^{(1)}$ lagged $\epsilon^{(2)}$ by the same
-amount in g-h.
+(e-f), and for C2, $\epsilon^{(1)}$ lagged $\epsilon^{(2)}$ by the same
+amount (g-h).
 See SI section \ref{SM-sect:examples_cross_sync} and Fig. \ref{SM-fig:rhosForC} for 
@@
 kelp lag $n=4$; see Figs \ref{SM-fig:directMoran_SM_lag8}, \ref{SM-fig:indirectMoran_SM_lag8}, 
-\ref{SM-fig:directMoran_SM_lag12} and \ref{SM-fig:indirectMoran_SM_lag12} for $n=8$ and$ 12$.
+\ref{SM-fig:directMoran_SM_lag12} and \ref{SM-fig:indirectMoran_SM_lag12} for $n=8$ and $12$.
 See the online version for color

### 2022-11-22 Paper/SupMat.Rnw
lots of edits working toward simplification of the ms
a bunch more edits

@@
 %header material for paper
-\title{Supporting information for ``Environmental drivers of synchrony and their interactions''}
+\title{Supporting information for ``How environmental drivers of spatial synchrony interact''}
 \date{}
@@
 For case study B, details presented in Methods of the main text imply that 
-$f_{\mcB}=1-b_1 \mu$, $f_{\mcP^{(1)}}=p_n^{(1)} \mu^n$ and $f_{\mcP^{(2)}}=p_0^{(2)}$. 
+$f_{\mcB}=1-b_1 \mu$, $f_{\mcP^{(1)}}=p_n^{(1)} \mu^n$ for $n=3$, and $f_{\mcP^{(2)}}=p_0^{(2)}$. 
 Because the noise process for B is taken to be the same as that for A, we again have 
@@
 For case study C, again using details from Methods, $f_{\mcB}=1-b_1 \mu$, 
-$f_{\mcP^{(1)}}=p_{n_1}^{(1)} \mu^{n_1}$ and $f_{\mcP^{(2)}}=p_{n_2}^{(2)} \mu^{n_2}$. 
+$f_{\mcP^{(1)}}=p_{n_1}^{(1)} \mu^{n_1}$ for $n_1=1$, and $f_{\mcP^{(2)}}=p_{n_2}^{(2)} \mu^{n_2}$ for $n_2=0$. 
 In SI section \ref{sect:examples_cross_sync}, we computed that
@@
 $\rho_{\epsilon^{(1)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{12} \exp(-2 \pi \iota f (l_1-l_2))$.
+
+% \section{Detailed breakdown of terms in the general theory for the case studies}\label{sect:breakdown_case_studies}
+% 
+% We here provide a detailed technical breakdown how the terms in our main theoretical equation, 
+% (\ref{MT-rhoww_T1to3}), capture the complexities of CaseA described in the main text.
+% The size of the $1$q lag of the effects of 
+% 
+% 
+% Theory captures these complexities because
+% the size of the $1$q lag of this scenario, relative to a timescale, $\sigma$, is captured by
+% the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, which equals the phase of
+% $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ because $l_n=0$ and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number
+% (SI section \ref{SM-sect:examples_cross_sync}). But $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=
+% p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) =
+% p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}),
+% which has phase $-\pi$ at short timescales
+% and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b,c).
+% The phase controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, the
+% interaction term in (\ref{rhoww_T1to3}), as described in General Theory.
+% 
+% 
+% \section{Detailed breakdown of terms in the general theory for kelp}\label{sect:breakdown_kelp}
 
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_MainPlot_ForMS.pdf}
-\caption[Extension of main text Fig. \ref{MT-fig:SummaryKelpFig}]{Same format as Fig. \ref{MT-fig:SummaryKelpFig} of the main text, but including not only results
+\caption[Extension of main text Fig. \ref{MT-fig:SummaryKelpFig} to include the CCA2 region]{Same format as Fig. \ref{MT-fig:SummaryKelpFig} of the main text, but including not only results
 for the CCA1 region (a-c) and the SB region (d-f), but also results from the 
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainDirectMoran_ForMS.pdf}
-\caption[Extension of main text Fig. \ref{MT-fig:directMoran}]{Same format as Fig. \ref{MT-fig:directMoran} of the main text, but including not only results
-for the CCA1 region (a-o) and the SB region (p-dd), but also results from the 
-CCA2 region (ee-ss).}\label{fig:directMoran_SM_lag4}
+\caption[All theory constituents for kelp for direct Moran effects terms, kelp lags up to 4 quarters]{Plots of all constituents of 
+the terms in theory that pertain to direct Moran effects, for kelp, for the CCA1 region (a-o), the SB region
+(p-dd) and the CCA2 region (ee-ss), based on a model that considered 
+potential kelp lag effects from up to 4 quarters in the past (see Methods).}\label{fig:directMoran_SM_lag4}
 \end{figure}
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag8_ExplainDirectMoran_ForMS.pdf}
-\caption[Like Fig. \ref{fig:directMoran_SM_lag4} but with kelp lag effects up to 8 quarters]{Same format as Fig. \ref{fig:directMoran_SM_lag4}, but based on a model that considered 
+\caption[Like Fig. \ref{fig:directMoran_SM_lag4} but with kelp lag effects up to 8 quarters]{Same format 
+as Fig. \ref{fig:directMoran_SM_lag4}, but based on a model that considered 
 potential kelp lag effects from up to 8 quarters in the past (see Methods). Results were substantively
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag12_ExplainDirectMoran_ForMS.pdf}
-\caption[Like Fig. \ref{fig:directMoran_SM_lag4} but with kelp lag effects up to 12 quarters]{Same format as Fig. \ref{fig:directMoran_SM_lag4}, but based on a model that considered 
+\caption[Like Fig. \ref{fig:directMoran_SM_lag4} but with kelp lag effects up to 12 quarters]{Same format 
+as Fig. \ref{fig:directMoran_SM_lag4}, but based on a model that considered 
 potential kelp lag effects from up to 12 quarters in the past (see Methods). Results were substantively
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainInteractingMoran_ForMS.pdf} 
-\caption[Extension of main text Fig. \ref{MT-fig:indirectMoran}]{Same format as Fig. \ref{MT-fig:indirectMoran} of the main text, but including not only results
-for the CCA1 region (a-o) and the SB region (p-dd), but also results from the 
-CCA2 region (ee-ss).}\label{fig:indirectMoran_SM_lag4}
+\caption[All theory constituents for kelp for Moran interaction terms, kelp lags up to 4 quarters]{Plots of 
+all constituents of the terms in theory that pertain to interactions between Moran effects, for kelp, for CCA1 region
+(a-o), the SB region (p-dd), and the CCA2 region (ee-ss), based on a model that considered 
+potential kelp lag effects from up to 4 quarters in the past (see Methods).}\label{fig:indirectMoran_SM_lag4}
 \end{figure}

### 2022-11-22 Code/Kelp_AnalysisStep_SpectralDecomp.R

@@ -393,15 +393,15 @@ explain_interacting_Moran<-function(freq,frg,fP1,fP2,Se1e2_avg,
   #technical stuff for the circular colorbar for phases
   breaks<-seq(from=-pi,to=pi,length.out=201)
   colbarQuad1<-rgb(red=1,green=seq(from=1,to=0,length.out=50),blue=seq(from=1,to=0,length.out=50))
-  #plot(1:50,1:50,col=colbarQuad1,pch=20,cex=2)
+  #plot(1:50,1:50,col=colbarQuad1,pch=20,cex=1.5)
   colbarQuad2<-rgb(red=seq(from=1,to=0,length.out=50),green=0,blue=0)
-  #plot(1:50,1:50,col=colbarQuad2,pch=20,cex=2)
+  #plot(1:50,1:50,col=colbarQuad2,pch=20,cex=1.5)
   colbarQuad3<-rgb(red=0,green=0,blue=seq(from=0,to=1,length.out=50))
-  #plot(1:50,1:50,col=colbarQuad3,pch=20,cex=2)
+  #plot(1:50,1:50,col=colbarQuad3,pch=20,cex=1.5)
   colbarQuad4<-rgb(red=seq(from=0,to=1,length.out=50),green=seq(from=0,to=1,length.out=50),blue=1)
-  #plot(1:50,1:50,col=colbarQuad4,pch=20,cex=2)
+  #plot(1:50,1:50,col=colbarQuad4,pch=20,cex=1.5)
   colbar<-c(colbarQuad3,colbarQuad4,colbarQuad1,colbarQuad2)
-  #plot(1:200,1:200,col=colbar,pch=20,cex=2)
+  #plot(1:200,1:200,col=colbar,pch=20,cex=1.5)
 
   #***panels for fP1 
   htind<-5
@@ -428,7 +428,7 @@ explain_interacting_Moran<-function(freq,frg,fP1,fP2,Se1e2_avg,
   h[h==-pi]<-pi
   cols<-colbar[WhichBreak(breaks=breaks,h)] 
   plotColLine(x=l2timescales[inds],y=y,cols=cols,lnwid=lnwid,ylim=ylimits,xlim=xlimits_short,xaxs="i",xaxt="n")
-  mtext(latex2exp::TeX("$f_{P^{(1)}}$"),2,1.7,cex=2) 
[…]

### 2022-11-22 Code/Theory_PedagogFig_Alt.R

@@ -175,7 +175,7 @@ if (maintext)
 }else
 {
   mtext(expression(paste("Scenario ",beta,"1: Antagonistic interactions")),side=3,line=0,cex=1,at=xlimits[1],adj=0)
-  text(xlimits[1],ylimits[2],"(e)",adj=c(0,1))
+  text(xlimits[1],ylimits[2],"(a)",adj=c(0,1))
 }
 
 #label ln
@@ -318,7 +318,7 @@ if (maintext)
   text(xlimits[1],ylimits[2],"(b)",adj=c(0,1))
 } else
 {
-  text(xlimits[1],ylimits[2],"(f)",adj=c(0,1))
+  text(xlimits[1],ylimits[2],"(b)",adj=c(0,1))
 }
 
 #label ln
@@ -452,7 +452,7 @@ if (maintext)
 }else
 {
   mtext(expression(paste("Scenario ",beta,"2: Synergistic interactions")),side=3,line=0,cex=1,at=xlimits[1],adj=0)
-  text(xlimits[1],ylimits[2],"(g)",adj=c(0,1))
+  text(xlimits[1],ylimits[2],"(c)",adj=c(0,1))
 }
 
 #label ln
@@ -594,7 +594,7 @@ if (maintext)
   text(xlimits[1],ylimits[2],"(d)",adj=c(0,1))
 } else
 {
-  text(xlimits[1],ylimits[2],"(h)",adj=c(0,1))
+  text(xlimits[1],ylimits[2],"(d)",adj=c(0,1))
 }
 
 #label ln

### 2022-12-09 Paper/Paper.Rnw
edits to paper

@@
 important [\cite{Sheppard2016,Desharnais2018,Sheppard2019}, see also \cite{Keitt2008}],
-synchrony is now known to have a complex and pronounced geography \citep{Defriez2017,Defriez2017b,Walter2017}, 
+synchrony is now known to have a complex and pronounced geography \citep{Defriez2017,Defriez2017b,Walter2017,Walter2020}, 
 and patterns of synchrony are now known to be asymmetric in distribution tails in a way that
@@
 between the two Moran effects, and can be substantially positive or negative. This example was adapted 
-from SI section S1 of \cite{Sheppard2019}.
+from SI section S1 of \cite{Sheppard2019}. All notation and abbreviations used throughout the paper 
+are summarized in Table \ref{tab:notation}.
 
@@
 theory described below does not require that assumption, and see also the next example which makes a
-different assumption. If, as in Fig. \ref{fig:PedagogAlt}a, b, $l_n+l_{e1}=l_{e2}$, then the 
+different assumption. If, as in Fig. \ref{fig:PedagogAlt}a, b, $l_n+l_{e1}=l_{e2}$ (or $l_n+l_{e1}-l_{e2}$ is any integer
+multiple of $\sigma=20$, the period), then the 
 periodic maximal positive influences of the two environmental variables coincide with each other as well as being 
@@
 tend to coincide with positive influences of variable 2 in location $j$. In contrast, Fig. \ref{fig:PedagogAlt}c, d
-shows the opposite scenario, for which $l_n+l_{e1}=l_{e2}+\sigma/2$, where $\sigma=20$ is the period. Thus
+shows the opposite scenario, for which $l_n+l_{e1}=l_{e2}-\sigma/2$. Thus
 the periodic maximal positive influence of environmental variable 1 coincides with the periodic maximal negative
 influence of environmental variable 2, reducing synchrony relative to the case of unrelated environmental variables. 
-Intermediate scenarios are possible, as will be revealed by the theory. If we replace the assumption that
+(The same outcome would occur if $l_n+l_{e1}-l_{e2}$ were any integer multiple of $\sigma$ plus or minus $\sigma/2$.)
+Intermediate scenarios between the two scenarios of Fig. \ref{fig:PedagogAlt} are also possible, as will be revealed by 
+the theory. If we replace the assumption that
 larger values of both environmental variables are beneficial to populations with an assuption that larger 
@@
 See SI section \ref{SM-sect:examples_cross_sync} for detailed examples.
-%***DAN: If we go longer, you could bring these examples here and it would probably be a positive addition of
-%about one para and one new Fig. Would be a positive addition because it helps people understand our new metric.
-%***DAN: Upon further reflection, I'm not totally sure of that. Those examples in the sup mat are pretty technical.
-%If I brought a version of them here, it would have to be a stripped down, intuitive, version, with the math removed.
-%And the new fig would have to just be panel c or f from Fig. S1. That's still possible to do, but does it really
-%help more than just the verbal description above?
 
@@
 The real part of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-is positive and larger whenever the phase of this quantity is close to zero, i.e., when    
-$l_{e1}-l_{e2}+l_n$ is close to zero (e.g., Fig. \ref{fig:PedagogAlt}a, b), and is increasingly negative as 
+is positive whenever the phase of this quantity is close to zero, e.g., 
+when environmental influences are positive and    
+$l_{e1}-l_{e2}+l_n$ is close to 0 (Fig. \ref{fig:PedagogAlt}a, b); and is increasingly negative as 
 the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$ gets
-close to $\pi$, e.g., when $l_{e1}-l_{e2}+l_n$ is close to half the period of oscillation
+close to $\pi$, e.g., when $l_{e1}-l_{e2}+l_n$ is close to $-\sigma/2$
 (Fig. \ref{fig:PedagogAlt}c, d). The factor $1/|f_{\mcB}|^2$ again captures how 
@@
 $1$q (Methods, SI section \ref{SM-sect:examples_cross_sync}), so that $l_n=1$. 
-Thus $l_n+l_{e1}-l_{e2}=2 \text{q}=0.5 \text{y}$, and, on annual timescales, lags compounded, similar to 
+Thus $l_{e1}-l_{e2}+l_n=2 \text{q}=0.5 \text{y}$, and, on annual timescales, lags compounded, similar to 
 Fig. \ref{fig:PedagogAlt2}c, d: the 
@@
 noise process $\epsilon^{(2)}$ were set up to lag peaks in the periodic process $\epsilon^{(1)}$ by
-$1$q, so that $l_n=-1$. Thus $l_n+l_{e1}-l_{e2}=0$ and, on annual timescales, lags cancelled,
+$1$q, so that $l_n=-1$. Thus $l_{e1}-l_{e2}+l_n=0$ and, on annual timescales, lags cancelled,
 similar to Fig. \ref{fig:PedagogAlt2}a, b: the annual positive population effects of $\epsilon^{(1)}$
@@
 In the antagonistic scenario ($\alpha 2$), lags are misaligned, i.e., 
-$l_n+l_{e1}=l_{e2}+\sigma/2$, where $\sigma=20$ is the period, so that peak positive 
+$l_n+l_{e1}=l_{e2}-\sigma/2$, where $\sigma=20$ is the period, so that peak positive 
 effects of $\epsilon_i^{(1)}$ coincide with maximally negative effects of $\epsilon_i^{(2)}$, and  maximally 
@@
 on populations. We here assume that larger values of $\epsilon_i^{(1)}$ have more 
-positive population influences (as in the $\alpha$ scenarios of Fig. \ref{fig:PedagogAlt}), 
+positive population influences (as in Fig. \ref{fig:PedagogAlt}), 
 but that larger values of $\epsilon_i^{(2)}$ have more negative population influences (contrasting
-with the Fig. \ref{fig:PedagogAlt} scenarios). In scenario $\beta 1$, 
+with Fig. \ref{fig:PedagogAlt}). In scenario $\beta 1$, 
 $l_n+l_{e1}=l_{e2}$, as in $\alpha 1$; but Moran interactions are nonetheless antagonistic because the 
@@
 variable correspond to the peak negative effects of $\epsilon_i^{(1)}$. In scenario $\beta 2$, as in $\alpha 2$,
-$l_n+l_{e1}=l_{e2}+\sigma/2$ for $\sigma=20$ the period, but Moran interactions are synergistic because the 
+$l_n+l_{e1}=l_{e2}-\sigma/2$ for $\sigma=20$ the period, but Moran interactions are synergistic because the 
 negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
@@
 
+\clearpage
+\newpage
+
+\begin{table}[!h]
+\caption{Summary of notation and abbreviations.}\label{tab:notation}
+\begin{center}
+\begin{tabular}{l|l}
+Notation & Meaning \\
+ \hline
+$i=1,\ldots,N$ & Locations of population sampling \\
+$t=1,\ldots,T$ & Times of population sampling \\
+$w_i(t)$ & A population index at location $i$ and time $t$ \\
+$\epsilon_i^{(a)}(t)$ & Environmental variable $a$ measured at location $i$ and time $t$ \\
+$l_{ek}$ & Lag of the effects of environmental variable $k$ on populations \\
+$l_n$ & A lag between two environmental variables \\
+$\sigma$ & A timescale \\
+$f=1/\sigma$ & Frequency \\
+$S_{w_i w_i}$ & The power spectrum of the process $w_i$ \\
+$S_{w_i w_j}$ & The cross spectrum of the processes $w_i$ and $w_j$ \\
+$S_{ww}$ & The spectral matrix of the $N$-dimensional process $w$ \\
+$\rho_{ww}$ & Synchrony of the process $w$ \\
+$\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ & Cross synchrony of the processes $\epsilon^{(1)}$ and $\epsilon^{(2)}$ \\
+$b_1,\ldots,b_n$ & Autoregressive coefficients in the population model \\
+$p_0^{(k)},\ldots,p_{m_k}^{(k)}$ & Coefficients for the lagged effects of $\epsilon^{(k)}$ on populations \\
+$\delta$ & Effects of unmeasured environmental variables \\
+$\mu$ & $\exp(-2 \pi \iota f)$ for $\iota$ the imaginary unit \\
+$f_{\mcB}$ & $1-b_1 \mu - b_2 \mu^2 - \cdots - b_n \mu^n$ \\
+$f_{\mcP^{(k)}}$ & $p_0^{(k)} + p_1^{(k)} \mu + \cdots + p_{m_k}^{(k)} \mu^{m_k}$ \\
+$\frac{|f_{\mcP^{(k)}}|^2}{|f_\mcB|^2}\rho_{\epsilon^{(k)}\epsilon^{(k)}}$ & Direct Moran effects of variable $k$ in the theory \\
+$2 \Rp(f_{\mcP^{(1)}}\conj{f_{\mcP^{(2)}}} \epsilon^{(1)} \epsilon^{(2)})$ & Moran interactions \\
+CaseA, CaseB, CaseC1, CaseC2 & Names of theoretical case studies \\
+CCA1 & Region 1 of central California (Fig. \ref{fig:map}) \\
+CCA2 & Region 2 of central California (Fig. \ref{fig:map}) \\
+SB & Region around Santa Barbara (Fig. \ref{fig:map}) \\
+\end{tabular}
+\end{center}
+\end{table}
+
 \end{document}

### 2022-12-09 Paper/Paper.Rnw
dropped a bunch of complexity connecting to specific portions of the theory which I had been planning to put in sup mat but decided just to drop

@@
 the intrinsic nature of population dynamics modulates Moran influences.
-%***DAN: I need to add some stuff to the SI section mentioned to support the claim that 
-%"details supporting the above assertions are in SI section X"
-%***DAN: I am currently (2022 11 17) unsure of how much of this belongs in the main text
-%versus the sup mat, and how to structure new material I will add to the sup mat, which will 
-%influence what section I site above (currently not the right section). 
-%***DAN: One way to go may be to put the 2 paras immediately above at the end of the theory
-%derivation section in sup mat, and just state here that interpretations of theory 
-%consituents, as well as the derivation itself, are in that section. The you could add 
-%a detailed elaboration of the theoretical constituents for the theory case studies to S6,
-%including a reinstatement of the components of Fig. 5 that were cut. And then I guess you could
-%add a new SI section S7 which would have a detailed exploration of the theory constituents
-%for kelp.
-%***DAN: Decision, and also started implementing a bit: Keep the above paras, no extension of
-%them in sup mat, but do add 
-%a detailed elaboration of the theoretical constituents for the theory case studies to S6,
-%including a reinstatement of the components of Fig. 5 that were cut. And then 
-%add a new SI section S7 which would have a detailed exploration of the theory constituents
-%for kelp.
 
@@
 quantity, not altering their relative importance or sign [see (\ref{rhoww_T1to3})].
-%See SI section \ref{SM-sect:breakdown_case_studies} for a detailed technical breakdown of how the terms in 
-%our theory capture the complexities described above.
 
@@
 were, respectively, positive and negative for these timescales (Fig. \ref{fig:TheoryCaseA}c).
-%See SI section \ref{SM-sect:breakdown_case_studies} for a detailed technical breakdown of how the terms in 
-%our theory capture the complexities described above.
-%***DAN: May again want to cite a new section of sup mat, in a similar way suggested in a comment in the previous paragraph
-%Theory captures these complexities because 
-%the size of the $3$q sum of lags for this scenario, relative to $\sigma$,
-%is again captured by the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$; the phase of that
-%quantity, represented by colors on Fig. \ref{fig:TheoryCaseA}f,g, fluctuates analogously to the
-%interaction effects on Fig. \ref{fig:TheoryCaseA}a.
 
@@
 (Fig. \ref{fig:TheoryCaseA}g,h). 
-%***DAN: Maybe cite a new sup mat section here, once written
-%Our theory captures these complexities
-%because $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$
-%quantifies the degree of alignment between the effects of $\epsilon^{(1)}$ and 
-%$\epsilon^{(2)}$ compared to a given timescale, while 
-%$\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ captures the degree of alignment between 
-%oscillations in the variables themselves, compared to the timescale (Fig. \ref{fig:TheoryCaseA}j,k,n,o).
-%The phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$
-%therefore captures whether delays in effects and lags between 
-%variables align or misalign, with alignment
-%reflected in the real part of this quantity, which is half the interaction term in our theory (\ref{rhoww_T1to3}).
 On long timescales, interactions were the
@@
 positive and those of $\epsilon^{(2)}$ were negative.
-%See SI section \ref{SM-sect:breakdown_case_studies} for a detailed technical breakdown of how the terms in 
-%our theory capture the complexities described above.
 
@@
 were positive and delayed by $1$q and $\epsilon^{(2)}$ effects were negative and immediate.
-%Delayed effects are reflected, for CCA1, in Fig. \ref{fig:indirectMoran}a, and for 
-%CaseC1, in Fig. \ref{fig:TheoryCaseA}j, both of which show 
-%that $f_{\mcP^{(1)}}$ has phase about $-\pi/2$ at the annual timescale. 
 The delayed effects of nitrates in CCA1 are reflected in Table \ref{SM-tab:regcoefs}, where 
 regression results for $p_0^{(1)}$ are close to $0$ and those for $p_1^{(1)}$ are positive.
-%Immediate negative effects of waves in CCA1 are reflected in Fig. \ref{fig:indirectMoran}b,g,l
-%and for CaseC1 in Fig. \ref{fig:TheoryCaseA}j, both of which show 
-%that $f_{\mcP^{(2)}}$ has phase $-\pi$ at all timescales. 
 Immediate negative effects of waves in CCA1 are also reflected in Table \ref{SM-tab:regcoefs},
@@
 peaks by $1$q in CaseC1.
-%***DAN: We would keep the portion of the following sentence that is about CCA1,
-%and we would keep a version of Fig 8d, which shows the cross synchrony between 
-%these drivers corresponds to a lag of 1 quarter. The reference to 4k will just 
-%be dropped, since anyway we built that theoretical case study to have that lag 
-%so don't need a confusing spectral reflection of the lag. In the sup mat section 
-%where we go through the constituents of the theory in detail can have this kinda 
-%stuff.
 For CCA1, this is reflected in Fig. \ref{fig:noise}c, which shows that 
@@
 producing reinforcing interactions between Moran effects. 
-%Likewise, the $\pi/2$ phase 
-%of $f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}$ on annual timescales 
-%(Fig. \ref{fig:indirectMoran}c for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1)
-%approximately cancels the $-\pi/2$ phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ on annual timescales
-%(Fig. \ref{fig:indirectMoran}d for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1), so that  
-%$f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}}$ has phase 
-%approximately $0$, making the interaction-effect term 
-%$\Rp(f_{\mcP^{(1)}}\overline{f_{\mcP^{(2)}}}\rho_{\epsilon^{(1)}\epsilon^{(2)}})$ strongly positive 
-%on annual timescales for both the theoretical and emprical cases 
-%(Fig. \ref{fig:indirectMoran}e for CCA1, Fig. \ref{fig:TheoryCaseA}k for CaseC1). 
 Interactions between Moran effects on long-timescales ($>8$y) were also similar for both CCA1 and
@@
 were observable within the same quarter: 
-%(Fig. \ref{fig:indirectMoran}a v. p; see also 
 see Table \ref{SM-tab:regcoefs}, where $p_0^{(1)}$ terms were close to zero for CCA1 and positive for SB,
@@
 immediate in both regions, i.e., $p_0^{(2)}<0$ in Table \ref{SM-tab:regcoefs}. 
-%(Fig. \ref{fig:indirectMoran}b,g,l v. q,v,aa; see also 
-%Table \ref{SM-tab:regcoefs}). 
 Annual peaks in nitrates still tended to lag annual peaks in waves by $1$q in SB,

### 2022-12-15 Paper/Paper.Rnw
continued editing to the paper

@@
 are environmental random variables measured in locations $i=1,2$ at times $t$, and assume these are independent
-through time and standard normally distributed for all $i$ and $t$. If a population index $w_i(t)$ 
+through time and standard normally distributed for all $i$ and $t$. 
+If a population index $w_i(t)$ 
 follows the autoregressive process $w_i(t)=b w_i(t-1)+\epsilon_i^{(1)}(t)$ for $i=1,2$, then the classic Moran theorem 
@@
 populations, the Moran theorem then implies that population synchrony is $\cor(w_1,w_2)=
-\cor(\epsilon_1^{(1)}+\epsilon_1^{(2)},\epsilon_2^{(1)}+\epsilon_2^{(2)})=
-[\cov(\epsilon_1^{(1)},\epsilon_2^{(1)})+\cov(\epsilon_1^{(1)},\epsilon_2^{(2)})+
-\cov(\epsilon_1^{(2)},\epsilon_2^{(1)})+\cov(\epsilon_1^{(2)},\epsilon_2^{(2)})]/
-[\sqrt{2+2\cov(\epsilon_1^{(1)},\epsilon_1^{(2)})}\sqrt{2+2\cov(\epsilon_2^{(1)},\epsilon_2^{(2)})}]$,
-a quantity which depends not only on the standard environmental synchrony 
+\cor(\epsilon_1^{(1)}+\epsilon_1^{(2)},\epsilon_2^{(1)}+\epsilon_2^{(2)})$.
+As we show in detail in SI section \ref{SM-Sect:IntroExample}, it is straightforward to show that 
+this quantity depends not only on the standard environmental synchrony 
 measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(a)})$, for 
 $a=1,2$, but also on the ``cross synchrony'' measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(b)})$
-for $a \neq b$. Cross synchrony measures represent synchrony between an environmental variable in one location and 
-a different variable in another location. Fig. \ref{fig:IntroTheory} illustrates how population synchrony can 
+for $a \neq b$. 
+Cross synchrony measures represent synchrony between an environmental variable in one location and 
+a different variable in another location. Fig. \ref{fig:IntroTheory} illustrates, using this model, 
+how population synchrony can 
 therefore differ from what synchrony would be if the environmental processes $\epsilon^{(1)}$ and 
@@
 Consider $N$ children, each riding on their own playground swing and each
-being pushed both by their own mother and father. If the fathers from seperate families
+being pushed both by their own mother and father. 
+Here, the children correspond to ecological populations and the parental pushing to environmental
+influences on the populations. 
+If the fathers from seperate families
 were to synchronize their pushes, they would act as a Moran-like influence, tending to synchronize
@@
 were not appropriately coordinated, the children 
-would swing more or less randomly, with small amplitude, and would become asynchronous with each other. 
+would swing more or less randomly, with smaller amplitude, and would become asynchronous with each other. 
 This second case demonstrates
 an antagonistic interaction between synchronizing effects, i.e., a tendency for the two synchronizing
-effects to cancel. As oscillators, children on swings are distinct in many ways from ecological
+effects to cancel. As oscillators, children on swings are distinct in many ways from 
 populations, not least because swings oscillate on one frequency/timescale, whereas populations
-typically oscillate simultaneously on many timescales, having a broadband spectrum. 
-We will see that a key insight of our theory is the use of a timescale-specific approach: 
-environmental signals are decomposed, using Fourier analysis, into components that occur 
-on different timescales,
-and the nature of interacting Moran effects can be understood on a timescale-by-timescale basis.
+typically oscillate simultaneously on many timescales. 
+We will see that 
+a key insight is the use of a timescale-specific approach. By timescale-specific, we mean that 
+fluctuations having different characteristic periods, which are superimposed in actual population time series data, 
+can be understood separately. 
+Timescale-specific approaches have been shown to be important to understanding synchrony and other 
+phenomena (e.g., \citealp{Sheppard2016,Anderson2021,Zhao2020}). 
+In the approach used in this paper, environmental and population signals are decomposed using Fourier analysis.
 The simple swing analogy turns out to supply, in much idealized form, the basic intuition our formal
@@
 effects on populations, by the amounts $l_{e1}$ and $l_{e2}$, respectively, i.e., peaks in the environmental signals
-manifest as maximum positive influence on populations after delays of $l_{e1}$ and $l_{e2}$, respectively. In this example,
+manifest as maximum positive influence on populations after delays of $l_{e1}$ and $l_{e2}$, respectively. 
+Such delays can be due to a variety of biological mechanisms associated with the life history of the organisms 
+which comprise the populations. In this example,
 we assume for simplicity that larger values of both environmental variables are beneficial to 
@@
 theory described below does not require that assumption, and see also the next example which makes a
-different assumption. If, as in Fig. \ref{fig:PedagogAlt}a, b, $l_n+l_{e1}=l_{e2}$ (or $l_n+l_{e1}-l_{e2}$ is any integer
+different assumption. If, as in Fig. \ref{fig:PedagogAlt}a, b, $l_{e1}-l_{e2}+l_n=0$ (or $l_{e1}-l_{e2}+l_n$ is any integer
 multiple of $\sigma=20$, the period), then the 
@@
 tend to coincide with positive influences of variable 2 in location $j$. In contrast, Fig. \ref{fig:PedagogAlt}c, d
-shows the opposite scenario, for which $l_n+l_{e1}=l_{e2}-\sigma/2$. Thus
+shows the opposite scenario, for which $l_{e1}-l_{e2}+l_n=-\sigma/2$. Thus
 the periodic maximal positive influence of environmental variable 1 coincides with the periodic maximal negative
 influence of environmental variable 2, reducing synchrony relative to the case of unrelated environmental variables. 
-(The same outcome would occur if $l_n+l_{e1}-l_{e2}$ were any integer multiple of $\sigma$ plus or minus $\sigma/2$.)
+(The same outcome would occur if $l_{e1}-l_{e2}+l_n$ were any integer multiple of $\sigma$ plus or minus $\sigma/2$.)
 Intermediate scenarios between the two scenarios of Fig. \ref{fig:PedagogAlt} are also possible, as will be revealed by 
-the theory. If we replace the assumption that
+the theory. We henceforth refer to the quantity $l_{e1}-l_{e2}+l_n$ as the \emph{environmental effect alignment measure},
+because it measures the extent to which the timing of the population influences of the two environmental variables
+are aligned. If we replace the assumption that
 larger values of both environmental variables are beneficial to populations with an assuption that larger 
@@
 %Description of data after cleaning
-The core of this paper is the theory, but to help illustrate that theory, for goal G3 of the 
-Introduction, we also apply
+To help illustrate our theory, for goal G3 of the Introduction, we also apply
 the theory to an exceptional dataset on giant kelp (\emph{Macrocyctis pyrifera}) dynamics off the CA coast, and 
@@
 last for multiple years, so kelp lag effects were considered:
-we used $n=4, 8, 12$ in separate analysis, with this choice making no substantive difference to results (see Results).
+we used $n=4, 8, 12$q in separate analyses, with this choice making no substantive difference to results (see Results).
 Fitted regression coefficients determined the quantities $b_1,\ldots,b_n$, $p_0^{(1)}$, $p_1^{(1)}$, and $p_0^{(2)}$
@@
 
-%***DAN: Another potentially good Discussion para: make connections to the work (or Wang and Loreau and
-%others) that is fusing spatial and community ecology. What are the implications of this work for that work?
-%Making such additions may help it appeal to a broader audience, which is something we need to work on as 
-%much as we can for Ecol Lett.
-
 %*DONE Summary
@@
 %*DONE Revist intuition
-%***DAN: Do we still need this if we're cutting? Are we cutting?
 We again revisit the intuition behind interacting Moran effects, using the central CA kelp
@@
 %*DONE Comparison to Castorani et al paper, and maybe also a bit on comparison to the Sheppard et al 2019 paper
-%***DAN: Could we put this in sup mat if we're cutting? Are we cutting?
 Our results about kelp were consistent with those of \cite{CastoraniSheppardKelp}, though
@@
 %*DONE relationship to earlier papers comparing synchrony of drivers and pops
-%***DAN: I guess you could move this to sup mat if we're cutting
 It has been a frequent topic of research why populations are often less synchronous, or sychronous over a smaller
@@
 % %*DONE Revisit the kendall paper, briefly. Talk about possible three-way interactions.
-% %***DAN: Could def put this in sup mat if we're cutting
+% %***DAN: Could def put this in sup mat if we're cutting, or drop it because it's a bit lame
 % \cite{Kendall2000} considered interactions between mechanisms of synchrony. Their research 
@@
 % %has been made many other places. Though I DO like the ending of this para, the ending of the prior
-% %para is also a good place to stop, maybe better. 
+% %para is also a good place to stop, maybe better. Anyway, just drop it.
 % An ever growing body of work shows that synchrony is, and should be analyzed as, a timescale-specific 
@@
 the environmental variables ($l_n$) and the lags of the population effects
-of the variables ($l_{e1}$ and $l_{e2}$) are aligned, i.e., $l_n+l_{e1}=l_{e2}$, so that peak positive 
+of the variables ($l_{e1}$ and $l_{e2}$) are aligned, i.e., the \emph{environmental effect alignment measure},
+$l_n+l_{e1}-l_{e2}$ (see main text), equals $0$. So peak positive 
 effects of $\epsilon_i^{(1)}$ coincide with peak positive effects of $\epsilon_i^{(2)}$ (the pink rectangles are
@@
 In the antagonistic scenario ($\alpha 2$), lags are misaligned, i.e., 
-$l_n+l_{e1}=l_{e2}-\sigma/2$, where $\sigma=20$ is the period, so that peak positive 
+$l_n+l_{e1}-l_{e2}=-\sigma/2$, where $\sigma=20$ is the period. So peak positive 
 effects of $\epsilon_i^{(1)}$ coincide with maximally negative effects of $\epsilon_i^{(2)}$, and  maximally 
@@
 variable correspond to the peak negative effects of $\epsilon_i^{(1)}$. In scenario $\beta 2$, as in $\alpha 2$,
-$l_n+l_{e1}=l_{e2}-\sigma/2$ for $\sigma=20$ the period, but Moran interactions are synergistic because the 
+the environmental effect alignment measure, $l_n+l_{e1}-l_{e2}$, equals $-\sigma/2$ for $\sigma=20$ the period. 
+But Moran interactions are synergistic because the 
 negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that

### 2022-12-15 Paper/SupMat.Rnw
edits to paper
continued editing to the paper

@@
 \listoffigures
+
+\section{Mathematical details of the example in the Introduction}\label{Sect:IntroExample}
+In the case of two environmental variables,
+the Moran theorem implies that population synchrony is 
+\begin{align}
+\cor(w_1,w_2) &= \cor(\epsilon_1^{(1)}+\epsilon_1^{(2)},\epsilon_2^{(1)}+\epsilon_2^{(2)}) \\
+&= \frac{\cov(\epsilon_1^{(1)},\epsilon_2^{(1)})+\cov(\epsilon_1^{(1)},\epsilon_2^{(2)})+
+\cov(\epsilon_1^{(2)},\epsilon_2^{(1)})+\cov(\epsilon_1^{(2)},\epsilon_2^{(2)})}{
+\sqrt{2+2\cov(\epsilon_1^{(1)},\epsilon_1^{(2)})}\sqrt{2+2\cov(\epsilon_2^{(1)},\epsilon_2^{(2)})}}, \\
+\end{align}
+which depends not only on the 
+standard environmental synchrony 
+measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(a)})$, for 
+$a=1,2$, but also on the ``cross synchrony'' measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(b)})$
+for $a \neq b$.
 
@@
 
-% \section{Detailed breakdown of terms in the general theory for the case studies}\label{sect:breakdown_case_studies}
-% 
-% We here provide a detailed technical breakdown how the terms in our main theoretical equation, 
-% (\ref{MT-rhoww_T1to3}), capture the complexities of CaseA described in the main text.
-% The size of the $1$q lag of the effects of 
-% 
-% 
-% Theory captures these complexities because
-% the size of the $1$q lag of this scenario, relative to a timescale, $\sigma$, is captured by
-% the phase of $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, which equals the phase of
-% $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}$ because $l_n=0$ and $\rho_{\epsilon^{(1)} \epsilon^{(2)}}$ is a real number
-% (SI section \ref{SM-sect:examples_cross_sync}). But $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}}=
-% p_1^{(1)} p_0^{(2)} \mu = p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota f) =
-% p_1^{(1)} p_0^{(2)} \exp(-2 \pi \iota /\sigma)$ (SI section \ref{SM-sect:details_case_studies}),
-% which has phase $-\pi$ at short timescales
-% and phase almost $0$ at long timescales (Fig. \ref{fig:TheoryCaseA}b,c).
-% The phase controls the sign of $2\Rp(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}})$, the
-% interaction term in (\ref{rhoww_T1to3}), as described in General Theory.
-% 
-% 
-% \section{Detailed breakdown of terms in the general theory for kelp}\label{sect:breakdown_kelp}
-
 \clearpage

### 2022-12-15 Paper/Paper.Rnw
done addressing Jon's comments from Feb 2022

@@
 through the phase of $\rho_{\epsilon^{(1)}\epsilon^{(2)}}$. 
-The expression $l_{e1}-l_{e2}+l_{n}$, which quantifies the degree of mismatch among the lags,
-corresponds to the phase of the expression 
+The environmental effect alignment measure, $l_{e1}-l_{e2}+l_{n}$, corresponds to the phase of the expression 
 $f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}$, because the phase of
@@
 taken to be a white noise process (see SI section \ref{SM-sect:examples_cross_sync} for details).
-Thus, what determines the sign of interactions between Moran effects for this example is how 
-$l_{e1}-l_{e2}+l_n=1$q compares to the timescale/period, $\sigma$, being considered. 
+Thus, the environmental effect alignment measure, $l_{e1}-l_{e2}+l_n$, equals $1$q.
+What determines the sign of interactions between Moran effects for this example is how 
+this measure compares to the timescale/period, $\sigma$, being considered. 
 On the shortest timescales ($\sigma=0.5 \text{y} = 2 \text{q}$; Fig. \ref{fig:TheoryCaseA}a, left side of panel), $1$q
 is half the period, so interaction effects are negative. On long timescales (e.g., $\sigma>8$y), $1$q is a negligible portion of the
-period, so $l_{e1}-l_{e2}+l_n=1$q is close to zero, relative to that period, and interaction effects
+period, so $l_{e1}-l_{e2}+l_n$ is close to zero, relative to that period, and interaction effects
 are positive: compared to such long timescales, $\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects happen close to simultaneously, so
-the two noise variables reinforce each other.
-Comparing
+the two noise variables reinforce each other. Comparing
 Fig. \ref{fig:TheoryCaseA}a and b shows the additional influence of autoregressive population effects, which
@@
 As for CaseA, because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
-a white noise process for CaseB, $l_n=0$.
+a white noise process for CaseB, $l_n=0$, so the environmental effect alignment measure is $3$q.
 Interaction effects were again negative on short timescales 
@@
 $1$q (Methods, SI section \ref{SM-sect:examples_cross_sync}), so that $l_n=1$. 
-Thus $l_{e1}-l_{e2}+l_n=2 \text{q}=0.5 \text{y}$, and, on annual timescales, lags compounded, similar to 
+Thus the environmental effect alignment measure, $l_{e1}-l_{e2}+l_n$, equals $2 \text{q}=0.5 \text{y}$, 
+and, on annual timescales, lags compounded, similar to 
 Fig. \ref{fig:PedagogAlt2}c, d: the 
@@
 has a phase of about $-\pi/2$ at the annual timescale.
-Thus lags compound on the annual timescale, in the CCA1 case as in CaseC1: the annual positive 
+Thus $l_{e1}=1$q, $l_{e2}=0$q, $l_n=1$q, and so the environmental effect alignment measures is 
+$l_{e1}-l_{e2}+l_n=2$q. Lags compound on the annual timescale, in the CCA1 case as in CaseC1: the annual positive 
 effects of nitrates on kelp tend to come in summer, and the annual negative effects of waves come in winter,
@@
 as in CCA1, though the seasonal periodicity of both variables was reduced in SB compared to 
-CCA1 (Fig. \ref{fig:noise}l v. c). 
+CCA1 (Fig. \ref{fig:noise}l v. c). Thus, in SB, the environment effect alignment measure was 
+$1$q, contrasting with its value of $2$q in CCA1.
 Whereas in CCA1, the tendency of nitrate annual peaks to follow wave annual peaks by $1$q, together with to the tendency of 
@@
 between Moran effects were detected on both of the two occassions they have been tested 
-for \citep{Sheppard2019,CastoraniSheppardKelp}, interactions may be common. 
+for \citep{Sheppard2019,CastoraniSheppardKelp}, interactions may be common (also see below). 
 Moran interactions are therefore
 a newly recognized and potentially widespread aspect of a fundamental means (Moran effects) by which 
-environmental factors influence populations.
+environmental factors influence populations and diverse, synchrony-related phenomena such as ecosystem stability 
+\citep{Wilcox2017}, population cycling \citep{Anderson2021}, and extinction \citep{Ghosh2020}.
 Climate change is altering many aspects of environmental variables, including their means, variances and
@@
 %*DONE Revist intuition
-We again revisit the intuition behind interacting Moran effects, using the central CA kelp
+We again revisit the intuition behind interacting Moran effects using the central CA kelp
 example as a vehicle. Large winter waves 
@@
 drivers must always be compared to the timescale of interest to determine interaction effects,
-as in Fig. \ref{fig:PedagogAlt}, and as captured formally in our theory.
+as in Figs \ref{fig:PedagogAlt} and \ref{fig:PedagogAlt2}, and as captured formally in our theory.
 
@@
 and southern CA on long timescales. The Fourier approach of our study was designed to facilitate 
-mathematical examination of interactions between Moran effect as a general mechanism and development
-of general theory, whereas the 
+mathematical examination of interactions between Moran effect as a general mechanism, and development
+of general theory; whereas the 
 wavelet approach of \cite{CastoraniSheppardKelp} was instead optimized for detecting interactions 
@@
 other complicating features which are present in many ecological datasets. The study of
-\cite{Sheppard2019} developed the wavelet methods applied by \cite{CastoraniSheppardKelp}.
+\cite{Sheppard2019} developed the wavelet methods applied by \cite{CastoraniSheppardKelp}; an open-source implementation of 
+these methods \citep{wsyn} can facilitate future work.
 The modelling approach of this study relates indirectly to the approach of \cite{Anderson2021}, though that
@@
 synchrony of environmental drivers \citep{Herfindal2022}.
-Antagonistic interactions between Moran drivers may be one common and previously
-unrecognized mechanism contributing to this discrepancy. On the other hand,
+Mechanisms include demographic stochasticity, measurement error, and influences of unmeasured environmental variables. 
+Antagonistic interactions between Moran drivers may be a common and previously
+unrecognized mechanism also contributing to this discrepancy. On the other hand,
 two recent papers descibed ``enhanced Moran effects'' by which specific 
@@
 than in drivers \citep{Massie2015,Desharnais2018}.
-Synergistic interaction effects is another mechanism by which populations can
+Synergistic interaction effects are another mechanism by which populations can
 be more synchronous than expected.
@@
 $l_n$ & A lag between two environmental variables \\
+$l_{e1}-l_{e2}+l_n$ & The \emph{environmental effect alignment measure} \\
 $\sigma$ & A timescale \\

### 2022-12-15 Paper/Paper.Rnw
finished editing for now and sent to coauthors

@@
 \author[a,*]{Daniel C. Reuman}
-\author[b]{Max Castorani}
-\author[c]{Kyle Cavanaugh}
+\author[b]{Max C.N. Castorani}
+\author[c]{Kyle C. Cavanaugh}
 \author[d]{Lawrence W. Sheppard}
-\author[b]{Jonathan Walter}
-\author[e,f]{Tom Bell}
+\author[e]{Jonathan A. Walter}
+\author[f,g]{Tom W. Bell}
 
@@
 \affil[d]{Marine Biological Association of the United Kingdom}
-\affil[e]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
-\affil[f]{Earth Research Institute, University of California, Santa Barbara}
+\affil[e]{Center for Watershed Sciences, University of California, Davis}
+\affil[f]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
+\affil[g]{Earth Research Institute, University of California, Santa Barbara}
 \affil[*]{Corresponding author, reuman@ku.edu}
@@
 
-\keywords{Fourier, interactions, kelp, \emph{Macrocyctis pyrifera}, Moran effects, nutrients, synchrony, waves}
+\keywords{Fourier, interactions, kelp, \emph{Macrocystis pyrifera}, Moran effects, nutrients, synchrony, waves}
 
@@
 periodic environmental fluctuations sampled quarterly,
-e.g., kelp wave action in central CA peaks annually in the winter, whereas surface-water
+e.g., wave action in central CA peaks annually in the winter, whereas surface-water
 nitrate concentrations also fluctuate with period one year, but peak in the spring, a delay of $1$q
@@
 To help illustrate our theory, for goal G3 of the Introduction, we also apply
-the theory to an exceptional dataset on giant kelp (\emph{Macrocyctis pyrifera}) dynamics off the CA coast, and 
+the theory to an exceptional dataset on giant kelp (\emph{Macrocystis pyrifera}) dynamics off the CA coast, and 
 associated environmental measurements. 
@@
 (called SB; see Fig. \ref{fig:map}). Each kelp measurement is an estimate of 
-mean quarterly kelp biomass (kg wet) per unit 
+mean quarterly kelp canopy biomass (kg wet) per unit 
 useable habitat ($\text{m}^2$) along a 500m stretch of coastline. We used coastline segments where kelp was persistent
@@
 effects appear delayed by about $1$q because our kelp data quantify canopy (surface) biomass, and it can
-take time for damaged kelp to grow back to the surface 
+take time for subsurface kelp to grow back to the surface 
 \citep{Cavanaugh2011,Bell2015,Schiel2015}. 
@@
 for CCA2), possibly due to the North Pacific Gyre Oscillation 
-\citep{CastoraniSheppardKelp}. The long-timescale synchrony
+\citep{CastoraniSheppardKelp, DiLorenzo2008}. The long-timescale synchrony
 in nitrates and waves produced some of the long-timescale
@@
 and the values $\cov(\epsilon_i^{(1)},\epsilon_j^{(2)})$, for $i,j=1,2$, were all set equal to each
-other and to a parameter, $\phi$. 
+other and to a parameter, $\varphi$. 
 See the online version for color
@@
 central California 1 (CCA1, blue points), central california 2 (CCA2, green points) and the region
-around Santa Barbara (SB, red points). Kelp density is shown with color intensity in (b)-(d),
+around Santa Barbara (SB, red points). Kelp density in $500$m coastline segments 
+is shown with color intensity in (b)-(d),
 and those panels correspond to the regions. One example time series from each region is shown in
@@
 and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.
-Synch. contrib. = Synchrony contribution; Rel. synch. contrib. = Relative synchrony contribution. 
+Synch. contrib. = Synchrony contribution refers to contributions to synchrony of the individual 
+terms in our theory; Rel. synch. contrib. = Relative synchrony contribution refers to contributions 
+expressed without accounting for the influence of autoregressive population effects. 
 See the online version for color
@@
 \begin{figure}
+\hspace{2cm} {\small Central CA region 1 (CCA1)} \hspace{3cm} {\small Santa Barbara region (SB)}  \\ 
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_MainPlot_ForMS.pdf}
@@
 \begin{figure}
+\hspace{2cm} {\small Central CA region 1 (CCA1)} \hspace{3cm} {\small Santa Barbara region (SB)}  \\ 
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal1_KelpLag4_ExplainDirectMoran_ForMS_short.pdf}

### 2023-01-10 Paper/Paper.Rnw
implemented editorial comments of Max and Jon

@@
 \author[d]{Lawrence W. Sheppard}
-\author[e]{Jonathan A. Walter}
+\author[b,e]{Jonathan A. Walter}
 \author[f,g]{Tom W. Bell}
@@
 \emph{synchrony}) has been studied for decades
-in a wide variety of species ranging from viruses to plants to mammals, and at 
+in a wide variety of species ranging from viruses and plants to mammals, and at 
 spatial scales ranging from centimeters to over 1000 km \citep{Liebhold2004},
@@
 that synchrony is changing as a newly recognized component of climate 
-change \citep{Defriez2016,Sheppard2016,Hansen2020}. Spatial and community 
-viewpoints on synchrony are becoming integrated, leading to a more wholistic understanding 
+change \citep{Defriez2016,Sheppard2016,Hansen2020}. 
+Distinct viewpoints on synchrony from population and community ecology are 
+becoming integrated, leading to a more wholistic understanding 
 of the importance of synchrony for ecosystem stability \citep{Wang2014}. These new developments 
@@
 environmental drivers can, in fact, interact, either synergistically or antagonistically,
-so that total population synchrony can be either greater or less than if the distinct Moran 
+so that total population synchrony can be either greater than or less than what synchrony would be if the distinct Moran 
 drivers operated independently \citep{Sheppard2019,CastoraniSheppardKelp}. 
@@
 Having demonstrated that interactions between Moran effects can occur, we now 
-use a simple analogy from everyday experience to begin to provide intuition. 
+use a simple analogy from common experience to begin to provide intuition. 
 Consider $N$ children, each riding on their own playground swing and each
-being pushed both by their own mother and father. 
+being pushed both by their own mother and their own father. 
 Here, the children correspond to ecological populations and the parental pushing to environmental
@@
 synchronize their pushes, it would tend to have a separate synchronizing effect. If the pushes
-of the fathers were appropriately coordinated with the pushes of the mothers, the children would swing higher, and
+of the fathers were appropriately coordinated with the pushes of the mothers (either happening at the same
+time, if the mothers and fathers are standing on the same sides of the swining children, or at
+opposite times in the swing period if the parents are on opposite sides), the children would swing higher, and
 would also swing more synchronously, demonstrating a synergistic interaction between synchronizing
@@
 period, so $l_{e1}-l_{e2}+l_n$ is close to zero, relative to that period, and interaction effects
-are positive: compared to such long timescales, $\epsilon^{(1)}$ and $\epsilon^{(2)}$ effects happen close to simultaneously, so
+are positive: relative to long timescales, $\epsilon^{(1)}$ 
+and $\epsilon^{(2)}$ effects happen close to simultaneously, so
 the two noise variables reinforce each other. Comparing
@@
 has a phase of about $-\pi/2$ at the annual timescale.
-Thus $l_{e1}=1$q, $l_{e2}=0$q, $l_n=1$q, and so the environmental effect alignment measures is 
+Thus $l_{e1}=1$q, $l_{e2}=0$q, $l_n=1$q, and so the environmental effect alignment measure is 
 $l_{e1}-l_{e2}+l_n=2$q. Lags compound on the annual timescale, in the CCA1 case as in CaseC1: the annual positive 
@@
 $1$q, contrasting with its value of $2$q in CCA1.
-Whereas in CCA1, the tendency of nitrate annual peaks to follow wave annual peaks by $1$q, together with to the tendency of 
+Whereas in CCA1, the tendency of nitrate annual peaks to follow wave annual peaks by $1$q, together with the tendency of 
 nitrate effects on kelp to be delayed by $1$q, resulted in summer positive nitrate effects and winter 
@@
 or when distinct variables are measured in the same part of the 
-year (e.g., spring temperature and precipitation). Future work should systematically 
+year (e.g., spring temperature and precipitation). Such scenarios involving seasonality of effects, 
+which were recently explored by \cite{WalterInPress}, can produce specific manifestations of the general mechanisms
+explored here. \cite{WalterInPress} found interactions and cross synchrony to be important.
+Future work should systematically 
 investigate cross synchrony ($\rho_{\epsilon^{(1)} \epsilon^{(2)}}$) of environmental 
@@
 and it takes time for kelp to grow to the surface and become detectable by satellite, only
-then revealing the effects of elevated nitrates. In southern CA, kelp are less likely to be
-completely removed by winter waves. Effects of elevated nitrates on the growth of kelp stands already reaching
-the surface can be immediately observed. Climate change may similarly alter 
+then revealing the effects of elevated nitrates \citep{Cavanaugh2011,Bell2015,Schiel2015}. 
+In southern CA, kelp are less likely to be completely removed by winter waves. 
+Effects of elevated nitrates on the growth of kelp stands already reaching
+the surface can be immediately observed \citep{Cavanaugh2011,Bell2015,Schiel2015}. 
+Climate change may similarly alter 
 lags and delays associated with environmental effects on populations in at least two ways:

### 2023-02-06 Paper/Paper.Rnw
impelemented comments of coauthors

@@
 
-%***Tentative author order, Bell last for his data contributions
+%authors
 \author[a,*]{Daniel C. Reuman}
@@
 A common mechanism of spatial synchrony is the Moran effect, whereby spatially synchronized environmental signals 
-drive population dynamics and hence induce population synchrony. We here elaborate a general theory of how 
+drive population dynamics and hence induce population synchrony. 
+%***DAN: the below line is the new line added to fit into Ecography, delete if you don't go that way
+After reviewing recent progress in understanding Moran effects, 
+%***DAN: the above line is the new line added to fit into Ecography, delete if you don't go that way
+we here elaborate a general theory of how 
 Moran effects of different environmental drivers acting on the same populations can interact, either 
@@
 of the fathers were appropriately coordinated with the pushes of the mothers (either happening at the same
-time, if the mothers and fathers are standing on the same sides of the swining children, or at
+time, if the mothers and fathers are standing on the same sides of the swinging children, or at
 opposite times in the swing period if the parents are on opposite sides), the children would swing higher, and
@@
 \section{General Theory}
+
+\subsection{Building intuition for Moran interactions: a single timescale}
 
@@
 
+\subsection{Formal theory}
+
 Our formal theory requires a conceptual understanding of the \emph{spectrum} of an environmental 
@@
 
-%***DAN: Potential reviewers or editors:
-%Maybe Holt as editor if we go to Ecography
-%Gonzalez
-%Loreau?
-%Bjornstad
-%Holt's postdoc, name slipping my mind
-%Barfield? 
-%Holt as reviewer? Will he even do it?
-%Cohen? 
-%Arnoldi? Bart Haegeman?
-%Montoya?
-
 Our population model is 
@@
 $|f_{\mcP^{(1)}}|^2$, and modulated
-by the autoregressive nature of kelp dynamics,
+by the autoregressive nature of population dynamics,
 $1/|f_{\mcB}|^2$. The term for the direct Moran effects of $\epsilon^{(2)}$
@@
 \section{Methods}
+
+\subsection{Theoretical case studies}
 
@@
 the autoregressive order of population dynamics is assumed, for simplicity, to be $1$, i.e, $b_1 \neq 0$ 
-but $b_i=0$ for $i>1$.
+but $b_i=0$ for $i>1$. Case studies do not cover the full range of possible scenarios which can be illuminated by 
+our general theory; they were selected for the intuition they provide, and for the correspondence of some of the cases to the 
+kelp examples we later discuss.
+
+\subsection{Kelp examples}
 
@@
 
+\subsection{Illustrating properties of Moran interactions: case studies}
+
 %Para basically confirming earlier work, as a starter
@@
 straightforwardly, and those papers are case studies, whereas our results provide general theory.
+
+\subsection{Building intuition for Moran interactions: case studies}
 
@@
 
+\subsection{Illustrating properties of Moran interactions: kelp examples}
+
 %Paralels the first para, but for kelp
@@
 negative on timescales $>8$y (Fig. \ref{fig:SummaryKelpFig}c).
+
+\subsection{Direct Moran effects in the kelp examples}
 
@@
 
+\subsection{Building intuition for Moran interactions: kelp examples}
+
 %Interactions between Moran effects, CCA1
@@
 where entries for $p_0^{(2)}$ are negative.
+Nitrate effects on kelp 
+are probably delayed by a quarter in central CA because winter waves commonly remove kelp,
+and it takes time for kelp to grow to the surface and become detectable by satellite, only
+then revealing the effects of elevated nitrates \citep{Cavanaugh2011,Bell2015,Schiel2015}.
 In CCA1, annual peak nitrates tend to come in spring, whereas annual peak waves 
@@
 see Table \ref{SM-tab:regcoefs}, where $p_0^{(1)}$ terms were close to zero for CCA1 and positive for SB,
-whereas $p_1^{(1)}$ terms were positive for CCA1 and close to zero for SB. Wave effects were negative and
+whereas $p_1^{(1)}$ terms were positive for CCA1 and close to zero for SB. 
+Wave effects were negative and
 immediate in both regions, i.e., $p_0^{(2)}<0$ in Table \ref{SM-tab:regcoefs}. 
+In southern CA, kelp are less likely to be completely removed by winter waves \citep{Reed2011}. 
+Effects of elevated nitrates on the growth of kelp stands already reaching
+the surface can be observed within the quarter \citep{Cavanaugh2011,Bell2015,Schiel2015}. 
 Annual peaks in nitrates still tended to lag annual peaks in waves by $1$q in SB,
@@
 nitrate effects on kelp to be delayed by $1$q, resulted in summer positive nitrate effects and winter 
-negative wave effects which reinforced sycnhrony; in SB nitrate instead both peaked and had its positive effects on kelp in spring,
+negative wave effects which reinforced synchrony; in SB nitrate instead both peaked and had its positive effects on kelp in spring,
 while waves still had negative effects on kelp in winter. Thus the effects of nitrates and waves in SB
@@
 and are also often related to each other because of their common origin in underlying climatic 
-phenomena such as the North Pacific Gyre Oscillation or El Ni\~{n}o Southern Oscillation (ENSO).
+phenomena such as the North Pacific Gyre Oscillation (NPGO) or El Ni\~{n}o Southern Oscillation (ENSO).
 Driver pairs which may commonly produce interacting Moran effects include instances where the same quantity 
@@
 as in Figs \ref{fig:PedagogAlt} and \ref{fig:PedagogAlt2}, and as captured formally in our theory.
+Thus interactions between Moran effects provide yet another reason, among reasons our previous work
+has already documented (e.g., \citealp{Sheppard2016,Walter2017,Anderson2021,Zhao2020}), that patterns 
+of synchrony mush be considered from a timescale-specific 
+viewpoint to achieve fuller understanding.
 
@@
 spatial extent, than might be expected given the strength and extent of 
-synchrony of environmental drivers \citep{Herfindal2022}.
-Mechanisms include demographic stochasticity, measurement error, and influences of unmeasured environmental variables. 
+synchrony of an environmental driver \citep{Herfindal2022}. Our new mechanism
+of interacting Moran effects provides both a new means by which populations may be less synchronous than 
+population drivers; and also a new means by which populations can be more 
+synchronous than environmental drivers. 
+Previously known mechanisms by which populations can be less synchronous than environmental drivers 
+include demographic stochasticity and measurement error. 
 Antagonistic interactions between Moran drivers may be a common and previously
-unrecognized mechanism also contributing to this discrepancy. On the other hand,
+unrecognized additional mechanism contributing to this discrepancy. On the other hand,
 two recent papers descibed ``enhanced Moran effects'' by which specific 
@@
 than in drivers \citep{Massie2015,Desharnais2018}.
-Synergistic interaction effects are another mechanism by which populations can
-be more synchronous than expected.
+Synergistic interaction between Moran drivers are another mechanism by which populations can
+be more synchronous than expected. Although we considered only two drivers in 
+our theory and examples, populations may, in principle, depend on many environmental
+factors, all related to a large-scale climatic phenomenon such as 
+ENSO and hence related to each other. Future work should consider whether and when such scenarios can lead to elevated synchrony.
 
@@
 If climate change modifies environmental effects on populations in a related way it would 
-be expected to produce similarly large changes in Moran interactions. Nitrate effects on kelp 
-are probably delayed by a quarter in central CA because winter waves commonly remove kelp,
-and it takes time for kelp to grow to the surface and become detectable by satellite, only
-then revealing the effects of elevated nitrates \citep{Cavanaugh2011,Bell2015,Schiel2015}. 
-In southern CA, kelp are less likely to be completely removed by winter waves. 
-Effects of elevated nitrates on the growth of kelp stands already reaching
-the surface can be immediately observed \citep{Cavanaugh2011,Bell2015,Schiel2015}. 
-Climate change may similarly alter 
+be expected to produce similarly large changes in Moran interactions. 
+Climate change may alter 
 lags and delays associated with environmental effects on populations in at least two ways:
@@
 decreasing or increasing delays. Though it is too early to conclude that
-effects on Moran interactions are among the important impacts of climate change, we feel the mechanisms
+effects on Moran interactions are among the most important impacts of climate change, we feel the mechanisms
 outlined above are sufficient to warrant further investigation.
@@
 Maowei Liang, Kyle Emery, Nat Coombs, Adeola Adeboje and Ethan Kadiyala for helpful discussions.
-%***DAN: If you end up adding any of these people to the author list, take them off here. Also in the acknowledgments for the README.
 
@@
 and the values $\cov(\epsilon_i^{(1)},\epsilon_j^{(2)})$, for $i,j=1,2$, were all set equal to each
-other and to a parameter, $\varphi$. 
+other and to a parameter, $\varphi$, which appears on the horizontal axis and which characterizes the 
+strength of cross synchrony of the environmental variables. 
 See the online version for color
@@
 \caption{%\doublespacing 
-Theoretical case studies A (panels a-b), B (c-d), C1 (e-h) and C2 (g-h). 
+Theoretical case studies A (panels a-b), B (c-d), C1 (e-f) and C2 (g-h). 
 Left panels (a, c, e, g) show the terms on the right side

### 2023-02-06 Paper/SupMat.Rnw
implemented editorial comments of Max and Jon
impelemented comments of coauthors

@@
 \tableofcontents
-%\listoftables
+\listoftables
 \listoffigures
@@
 $\epsilon^{(2)}=(\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)})$ and $\delta=(\delta_1,\ldots,\delta_N)$ are
-environmental processes influencing the populations. Note that this model assumes from the outset 
-a homogeneous influence of the environmental variables on the populations in the different
-sampling locations, i.e., the parameters $p$ are spatially homogeneous. 
-%We also assume spatial homogeneity of the AR coefficients, need to add a mention of that
-We also assume that the 
-combined process $(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ is a second-order stationary stochastic process \citep{Brillinger2001}
-with the expected values of its component processes equal to $0$. Finally, we assume the complex roots of the
+environmental processes influencing the populations. Note that this model makes some assumptions from the outset. 
+First, it assumes a homogeneous influence of the environmental variables on the populations in the different
+sampling locations, i.e., the parameters $p$ are spatially homogeneous. Second, autoregressive population effects
+are also assumed spatially homogeneous. Third, we assume that the 
+combined process $(\epsilon_1^{(1)},\ldots,\epsilon_N^{(1)},\epsilon_1^{(2)},\ldots,\epsilon_N^{(2)},\delta_1,\ldots,\delta_N)$ 
+is a second-order stationary stochastic process \citep{Brillinger2001}
+with the expected values of its component processes equal to $0$. We do \emph{not} assume that the unmeasured 
+noise terms $(\delta_1,\ldots,\delta_N)$ are spatially asynchronous or independent through time. 
+Finally, we assume the complex roots of the
 polynomial $1-b_1 z - \cdots - b_n z^n = 0$ have modulus greater than $1$, an assumption needed for the 

### 2023-02-06 Paper/Paper.Rnw
started work on additions of review materials for Ecography

@@
 
-%Intro to Moran effects and interactions
-One of the classically theorized \citep{Moran1953} and now widely empirically validated mechanisms of 
-synchrony, called the Moran effect, operates when spatially synchronized environmental signals, such
-as temperature or precipitation, help drive population dynamics and thereby induce synchrony in
-populations. Two recent papers have documented that Moran effects of distinct 
+%Intro to Moran effects and history of the early years where people had trouble identifying them
+Spatial autocorrelations in fluctuating environmental drivers of populations are now known to be one of the most important
+causes of synchrony, through a process called the Moran effect. But this effect was originally proposed 
+theoretically \citep{Moran1953}, and decades passed during which it was 
+considered difficult to distinguish this cause of synchrony from others, except under special circumstances.
+Synchrony has long been thought to have three causes: dispersal, and trophic interactions with a synchronous or
+mobile species, in addition to the Moran effect \citep{Liebhold2004}. But using common past approaches
+examining declines in population correlations with distance, statistical patterns of synchrony produced
+by each of these mechanisms are quite similar \citep{Ranta1999,Abbott2007,Liebhold2004,Walter2017}. So examination
+of such patterns provides little or no traction for performing inferences about causes of synchrony. 
+Early papers demonstrating the importance of Moran effects resorted to special cases where dispersal was 
+impossible and predators were absent \citep{Grenfell1998,Tedesco2004}. Though controlled experiments revealed that all
+three mechanisms could be involved in synchrony [e.g., \citealp{Vasseur2009}], the problem of determining
+the relative contribution of each mechanism to synchrony observed in field systems was considered a challenge 
+for decades.
+
+%Recent statistical explorations of Moran effects
+
+%Intro to interactions
+Two recent papers have documented that Moran effects of distinct 
 environmental drivers can, in fact, interact, either synergistically or antagonistically,
@@
 as in Figs \ref{fig:PedagogAlt} and \ref{fig:PedagogAlt2}, and as captured formally in our theory.
-Thus interactions between Moran effects provide yet another reason, among reasons our previous work
-has already documented (e.g., \citealp{Sheppard2016,Walter2017,Anderson2021,Zhao2020}), that patterns 
-of synchrony mush be considered from a timescale-specific 
-viewpoint to achieve fuller understanding.
+Thus interactions between Moran effects provide yet another reason, among many reasons previous work
+has already documented (e.g., \citealp{Vasseur2007,Keitt2008,Vasseur2014,Sheppard2016,
+Defriez2016,Defriez2017,Defriez2017b,Desharnais2018,Walter2017,
+Anderson2021,Zhao2020}), that patterns 
+of synchrony must be considered from a timescale-specific 
+viewpoint for full understanding.
 
@@
 ENSO and hence related to each other. Future work should consider whether and when such scenarios can lead to elevated synchrony.
-
-% %*DONE Revisit the kendall paper, briefly. Talk about possible three-way interactions.
-% %***DAN: Could def put this in sup mat if we're cutting, or drop it because it's a bit lame
-% \cite{Kendall2000} considered interactions between mechanisms of synchrony. Their research 
-% questions were distinct from ours, but combining their viewpoints and ours may lead to valuable
-% future work. \cite{Kendall2000} considered interactions between dispersal and Moran causes of 
-% synchrony, whereas we analyzed interactions between two Moran drivers, a distinct context.
-% But it seems likely that numuerous ecological systems involve important dispersal effects on 
-% synchrony as well as influences of at least two Moran effects. It may be necessary in future
-% work to consider three-way interactions between dispersal and two Moran drivers and their 
-% influence on synchrony. Dispersal can readily be added to our modelling framework.
-% \cite{Desharnais2018} performed a spectral analysis on a model similar to ours
-% which included dispersal but which only considered two habitat patches and one environmental driver.
 
@@
 % long-term ecosystem monitoring.
+
+% %*DONE Revisit the kendall paper, briefly. Talk about possible three-way interactions.
+% %***DAN: Could def put this in sup mat if we're cutting, or drop it because it's a bit lame
+% \cite{Kendall2000} considered interactions between mechanisms of synchrony. Their research 
+% questions were distinct from ours, but combining their viewpoints and ours may lead to valuable
+% future work. \cite{Kendall2000} considered interactions between dispersal and Moran causes of 
+% synchrony, whereas we analyzed interactions between two Moran drivers, a distinct context.
+% But it seems likely that numuerous ecological systems involve important dispersal effects on 
+% synchrony as well as influences of at least two Moran effects. It may be necessary in future
+% work to consider three-way interactions between dispersal and two Moran drivers and their 
+% influence on synchrony. Dispersal can readily be added to our modelling framework.
+% \cite{Desharnais2018} performed a spectral analysis on a model similar to ours
+% which included dispersal but which only considered two habitat patches and one environmental driver.
 

### 2023-02-07 Paper/Paper.Rnw
almost final version, sent to coauthors for final look

@@
 %Intro to Moran effects and history of the early years where people had trouble identifying them
-Spatial autocorrelations in fluctuating environmental drivers of populations are now known to be one of the most important
-causes of synchrony, through a process called the Moran effect. But this effect was originally proposed 
+Correlations between weather time series measured in different locations can 
+induce synchrony between populations in those locations if the weather variables influence
+population processes. This mechanism, called the Moran effect, is now known to be one of the most
+important causes of synchrony. But the mechanism was originally proposed 
 theoretically \citep{Moran1953}, and decades passed during which it was 
-considered difficult to distinguish this cause of synchrony from others, except under special circumstances.
+considered difficult to distinguish this potential cause of synchrony from others in field systems.
 Synchrony has long been thought to have three causes: dispersal, and trophic interactions with a synchronous or
-mobile species, in addition to the Moran effect \citep{Liebhold2004}. But using common past approaches
-examining declines in population correlations with distance, statistical patterns of synchrony produced
-by each of these mechanisms are quite similar \citep{Ranta1999,Abbott2007,Liebhold2004,Walter2017}. So examination
-of such patterns provides little or no traction for performing inferences about causes of synchrony. 
-Early papers demonstrating the importance of Moran effects resorted to special cases where dispersal was 
-impossible and predators were absent \citep{Grenfell1998,Tedesco2004}. Though controlled experiments revealed that all
-three mechanisms could be involved in synchrony [e.g., \citealp{Vasseur2009}], the problem of determining
-the relative contribution of each mechanism to synchrony observed in field systems was considered a challenge 
-for decades.
+mobile species, in addition to the Moran effect \citep{Liebhold2004}. But using common past statistical approaches
+which focussed on declines in population correlations with distance, patterns of synchrony produced
+by each of these mechanisms are quite similar \citep{Ranta1999,Abbott2007,Liebhold2004,Walter2017}; so examination
+of such patterns provides little or no traction for inferring the causes of synchrony. 
+Early papers demonstrating the Moran effect mechanism resorted to special cases where dispersal was 
+impossible and predators were absent \citep{Grenfell1998,Tedesco2004}. Controlled experiments also confirmed that all
+three theorized mechanisms could be involved in synchrony [e.g., \citealp{Vasseur2009}]. But the problem of inferring
+the relative contribution of the three mechanisms in field systems was considered a challenge 
+until recently.
 
 %Recent statistical explorations of Moran effects
+Recent research has provided new statistical viewpoints which have, when sufficient data are available, 
+essentially solved the problem of inference
+of the causes of synchrony, and the research has revealed the broad importance of Moran effects. 
+%geographic approaches
+Approaches based on formal statistical comparisons between population synchrony maps and maps of 
+environmental correlations produced inferences that precipitation and temperature Moran effects are important
+causes of synchrony of terrestrial \citep{Defriez2017b} and marine \citep{Defriez2017} primary productivity. 
+Another geographic approach, based on multiple regression of distance matrices (MRM), was used in an early paper
+to infer that a precipitation Moran effect is a cause of synchrony for the spongy moth \citep{Haynes2013},
+an invasive forest pest whose name was recently changed from ``gypsy moth.'' 
+Geographic approaches to identifying causes of synchrony, often MRM approaches 
+specifically \citep{Koenig2017,Walter2020,Bogdziewicz2021}, 
+have become widespread, and the geography of synchrony has become a mainstream area of study \citep{Walter2017}.
+The approach of \cite{Gouveia2016}, called ``noncentered local indicators of spatial association'' (ncLISA), is another 
+geographic approach; as is the ``fluvial synchrogram'' of \cite{Larsen2021}, a tool for studying synchrony in 
+river networks.
+MRM approaches have also been used to identify dispersal as a cause of synchrony \citep{Anderson2018},
+sometimes combined with genetic methods \cite{Haynes2022}.
+%Time and timescale methods
+Another class of methods exploit the time and timescale structure of synchrony.
+Using a method that examined the match between temporal variations in population and environmental synchrony, 
+\cite{Allstadt2015} confirmed the importance of a precipitation Moran effect for synchrony of spongy moth.
+A suite of wavelet and related Fourier techniques have been developed that can comprehensively describe the time and timescale
+structure of synchrony, identify Moran drivers of synchrony opperating in distinct timescale bands, and apportion fractions
+of observed synchrony to respective Moran drivers \citep{Sheppard2016,Sheppard2017,Desharnais2018,Sheppard2019,wsyn,Anderson2021}.
+The methods have been used, for instance, to provide one of the first clear demonstrations that changes in 
+Moran effects are another of the impacts of climatic changes \citep{Sheppard2016,Sheppard2019}; and to illustrate the importance of the
+timescale structure of environmental covariation for altering Moran effects \citep{Desharnais2018}.
+These methods have since been used to identify Moran effects in several additional systems
+(e.g., \citealp{Anderson2019,Walter2019,Walter2019,GarciaCarreras2022,CastoraniSheppardKelp}).
 
 %Intro to interactions
-Two recent papers have documented that Moran effects of distinct 
-environmental drivers can, in fact, interact, either synergistically or antagonistically,
+The techniques reviewed above have made it possible for several recent papers to identify cases in which two or more
+distinct Moran drivers opperate simultaneously on the same populations 
+(e.g., \citealp{Defriez2017b,Haynes2019,Walter2019,Walter2020,Anderson2021}).
+In fact, two recent papers have documented that Moran effects of distinct 
+environmental drivers can interact, either synergistically or antagonistically,
 so that total population synchrony can be either greater than or less than what synchrony would be if the distinct Moran 
@@
 which has $ij^{\text{th}}$ entry $S_{w_i w_j}(f)$. Spectral methods are standard  
-[e.g., \cite{Vasseur2007,Reuman2008,Defriez2017}], and many background references are available 
+[e.g., \cite{Vasseur2007,Defriez2017}], and many background references are available 
 [e.g., \cite{Brillinger2001}].
@@
 theoretical progress in ecology, and it has been demonstrated that results based on a
-weak noise assumption often hold for noise which is fairly strong \citep{Nisbet1977,Reuman2006,Desharnais2018}.
+weak noise assumption often hold for noise which is fairly strong \citep{Nisbet1977,Desharnais2018}.
 See \cite{Brillinger2001} for background on stochastic processes.
@@
 Synergistic interaction between Moran drivers are another mechanism by which populations can
-be more synchronous than expected. Although we considered only two drivers in 
-our theory and examples, populations may, in principle, depend on many environmental
-factors, all related to a large-scale climatic phenomenon such as 
-ENSO and hence related to each other. Future work should consider whether and when such scenarios can lead to elevated synchrony.
+be more synchronous than expected.
 
@@
 
-% %*DONE Revisit the kendall paper, briefly. Talk about possible three-way interactions.
-% %***DAN: Could def put this in sup mat if we're cutting, or drop it because it's a bit lame
-% \cite{Kendall2000} considered interactions between mechanisms of synchrony. Their research 
-% questions were distinct from ours, but combining their viewpoints and ours may lead to valuable
-% future work. \cite{Kendall2000} considered interactions between dispersal and Moran causes of 
-% synchrony, whereas we analyzed interactions between two Moran drivers, a distinct context.
-% But it seems likely that numuerous ecological systems involve important dispersal effects on 
-% synchrony as well as influences of at least two Moran effects. It may be necessary in future
-% work to consider three-way interactions between dispersal and two Moran drivers and their 
-% influence on synchrony. Dispersal can readily be added to our modelling framework.
-% \cite{Desharnais2018} performed a spectral analysis on a model similar to ours
-% which included dispersal but which only considered two habitat patches and one environmental driver.
+%*Revisit the kendall paper, briefly. Talk about possible three-way interactions.
+We have focussed on interactions between two Moran drivers, but synchrony in most systems
+may be a phenomenon with multiple (more than two) interacting causes.
+\cite{Kendall2000} considered interactions between dispersal and a Moran mechanism of synchrony. Their research
+questions were therefore distinct from ours, but combining their viewpoints and ours may lead to 
+future work about interactions between dispersal and more than one Moran driver of synchrony. 
+Although we considered only two Moran drivers in 
+our theory and examples, essentially all population systems are influenced by multiple environmental drivers,
+and environmental drivers very commonly are associated with large scale climatic phenomena
+such as ENSO, and hence are associated with each other. Thus it may be quite common for synchrony to simultaneously be caused by dispersal and
+multiple distinct Moran effects, and these influences may interact in multifarious combinations. 
+It may be necessary in future
+work to consider interactions between dispersal and two Moran drivers. 
+It may be useful to consider cases for which multiple related Moran drivers all interact.
+Dispersal can readily be added to our modelling framework:
+\cite{Desharnais2018} performed a spectral analysis on a model similar to ours
+which included dispersal.
+Future work should consider whether and when scenarios of multi-driver interactions between causes of synchrony 
+can lead to synchrony patterns which differ fundamentally from what one would expect from one
+or two mechanisms. Potential interactions increase as the square of the number of 
+drivers, so interactions seem likely to become more important as our viewpoint of synchrony expands.
 

### 2023-02-08 Paper/Paper.Rnw
minor changes suggested by coauthors, planning to submit this version to Ecography

@@
 
+%length stats
+%Intro through Discussion: 7719 as of 2023 02 08 (limit is 7500)
+%7 figures, 1 table (limit is 10)
+%so I think I'm OK - slightly over on the words, but with 2 fewer displays
+
 %header material for paper
@@
 \affil[e]{Center for Watershed Sciences, University of California, Davis}
-\affil[f]{Applied Ocean Physics and Engineering Department, Woods Hole Oceanographic Institution}
+\affil[f]{Department of Applied Ocean Physics and Engineering, Woods Hole Oceanographic Institution}
 \affil[g]{Earth Research Institute, University of California, Santa Barbara}
@@
 %Intro to Moran effects and history of the early years where people had trouble identifying them
+%***DAN: new review para for ecography
 Correlations between weather time series measured in different locations can 
@@
 Synchrony has long been thought to have three causes: dispersal, and trophic interactions with a synchronous or
-mobile species, in addition to the Moran effect \citep{Liebhold2004}. But using common past statistical approaches
+mobile species, in addition to the Moran effect \citep{Liebhold2004}. However, using common past statistical approaches
 which focussed on declines in population correlations with distance, patterns of synchrony produced
@@
 impossible and predators were absent \citep{Grenfell1998,Tedesco2004}. Controlled experiments also confirmed that all
-three theorized mechanisms could be involved in synchrony [e.g., \citealp{Vasseur2009}]. But the problem of inferring
-the relative contribution of the three mechanisms in field systems was considered a challenge 
+three theorized mechanisms could be involved in synchrony [e.g., \citealp{Vasseur2009}]. Nevertheless, the problem of inferring
+specific mechanisms of synchrony in field systems was considered a challenge 
 until recently.
@@
 %Recent statistical explorations of Moran effects
+%***DAN: new review para for ecography
 Recent research has provided new statistical viewpoints which have, when sufficient data are available, 
@@
 MRM approaches have also been used to identify dispersal as a cause of synchrony \citep{Anderson2018},
-sometimes combined with genetic methods \cite{Haynes2022}.
+sometimes combined with genetic methods \citep{Haynes2022}.
 %Time and timescale methods
@@
 These methods have since been used to identify Moran effects in several additional systems
-(e.g., \citealp{Anderson2019,Walter2019,Walter2019,GarciaCarreras2022,CastoraniSheppardKelp}).
+(e.g., \citealp{Anderson2019,Walter2019,Walter2020,GarciaCarreras2022,CastoraniSheppardKelp}).
 
@@
 the timescale of analysis.
+
+%main purpose
 The main purpose of this study is to establish a general theory of the mechanism
@@
 and $f_{\mcB}$ are in Table \ref{SM-tab:regcoefs}.
-%***DAN: Later you will have to make sure all rules are followed for citing all sup mat figures and 
-%tables and sections from the main text, in order. I think currently they are not all cited, in part
-%because I have been shifting around what information is where in the whole paper.
 
@@
 could also be assessed, by using either time-windowed versions or wavelet adaptations 
-of this statistic, applied to either long-term climate records or future climate scanarios. 
+of this statistic, applied to either long-term climate records or future climate scenarios. 
 Differences in $f_{\mcP^{(1)}}$ were responsible for differences
@@
 %*Revisit the kendall paper, briefly. Talk about possible three-way interactions.
+%***DAN: new review para for ecography
 We have focussed on interactions between two Moran drivers, but synchrony in most systems

### 2023-02-08 Paper/SupMat.Rnw
minor changes suggested by coauthors, planning to submit this version to Ecography

@@
 \begin{table}[!h]
-\caption{Regression coefficients used to parameters (\ref{MT-main_model_1})-(\ref{MT-main_model_4}) of the main text for each 
+\caption{Regression coefficients used to parameterize (\ref{MT-main_model_1})-(\ref{MT-main_model_4}) of the main text for each 
 of our three regions. All values rounded to two decimal places.}\label{tab:regcoefs}

### 2023-05-30 Paper/Paper.Rnw
version actually submitted to Ecography
changes in response to referees, minor revisions

@@
 
-%external documents
-\externaldocument[SM-]{SupMat}
-
 %Get line numbers to work, also in figure captions
@@
 
-%\doublespacing
+\doublespacing
 
@@
 
+\newpage
+
+\large{\textbf{How environmental drivers of spatial synchrony interact}}
+
 \begin{abstract}
@@
 drive population dynamics and hence induce population synchrony. 
-%***DAN: the below line is the new line added to fit into Ecography, delete if you don't go that way
 After reviewing recent progress in understanding Moran effects, 
-%***DAN: the above line is the new line added to fit into Ecography, delete if you don't go that way
 we here elaborate a general theory of how 
@@
 
-\keywords{Fourier, interactions, kelp, \emph{Macrocystis pyrifera}, Moran effects, nutrients, synchrony, waves}
+\keywords{Fourier, interactions, kelp, \emph{Macrocystis pyrifera}, Moran effects, synchrony}
 
@@
 synchrony is now known to have a complex and pronounced geography \citep{Defriez2017,Defriez2017b,Walter2017,Walter2020}, 
-and patterns of synchrony are now known to be asymmetric in distribution tails in a way that
-is important for system stability \citep{Ghosh2020b,Ghosh2020c,Walter2021}. Our ability to infer the causes of 
+and patterns of synchrony are now known to be ``asymmetric in distribution tails'' in a way that
+is important for system stability \citep{Ghosh2020b,Ghosh2020c,Walter2021}. 
+Synchrony is asymmetric in distribution tails when two positively associated (synchronous) variables are 
+more (respectively, less) synchronous in the lower parts of their distributions than in the 
+upper parts. Our ability to infer the causes of 
 synchrony is much improved in recent years \citep{Sheppard2016,Defriez2017b,Walter2017,Anderson2018,Sheppard2019}, 
@@
 %Intro to Moran effects and history of the early years where people had trouble identifying them
-%***DAN: new review para for ecography
 Correlations between weather time series measured in different locations can 
@@
 theoretically \citep{Moran1953}, and decades passed during which it was 
-considered difficult to distinguish this potential cause of synchrony from others in field systems.
+considered difficult to distinguish this potential cause of synchrony from others (e.g., dispersal) in field systems.
 Synchrony has long been thought to have three causes: dispersal, and trophic interactions with a synchronous or
@@
 of such patterns provides little or no traction for inferring the causes of synchrony. 
-Early papers demonstrating the Moran effect mechanism resorted to special cases where dispersal was 
+Early empirical papers demonstrating the Moran effect mechanism resorted to special cases where dispersal was 
 impossible and predators were absent \citep{Grenfell1998,Tedesco2004}. Controlled experiments also confirmed that all
@@
 %Recent statistical explorations of Moran effects
-%***DAN: new review para for ecography
 Recent research has provided new statistical viewpoints which have, when sufficient data are available, 
 essentially solved the problem of inference
-of the causes of synchrony, and the research has revealed the broad importance of Moran effects. 
+of the causes of synchrony (including identifying specific Moran drivers), and the research has 
+revealed the broad importance of Moran effects.
 %geographic approaches
-Approaches based on formal statistical comparisons between population synchrony maps and maps of 
-environmental correlations produced inferences that precipitation and temperature Moran effects are important
+Approaches based on spatial statistics produced inferences that precipitation and temperature Moran effects are important
 causes of synchrony of terrestrial \citep{Defriez2017b} and marine \citep{Defriez2017} primary productivity. 
-Another geographic approach, based on multiple regression of distance matrices (MRM), was used in an early paper
-to infer that a precipitation Moran effect is a cause of synchrony for the spongy moth \citep{Haynes2013},
-an invasive forest pest whose name was recently changed from ``gypsy moth.'' 
-Geographic approaches to identifying causes of synchrony, often MRM approaches 
-specifically \citep{Koenig2017,Walter2020,Bogdziewicz2021}, 
-have become widespread, and the geography of synchrony has become a mainstream area of study \citep{Walter2017}.
-The approach of \cite{Gouveia2016}, called ``noncentered local indicators of spatial association'' (ncLISA), is another 
-geographic approach; as is the ``fluvial synchrogram'' of \cite{Larsen2021}, a tool for studying synchrony in 
-river networks.
+Another geographic approach, based on multiple regression of distance matrices (MRM), was used 
+to infer that a precipitation Moran effect is a cause of synchrony for the spongy moth \citep{Haynes2013}. 
+Geographic approaches to identifying causes of synchrony \citep{Koenig2017,Walter2020,Bogdziewicz2021} 
+have become widespread \citep{Gouveia2016,Walter2017,Larsen2021}.
 MRM approaches have also been used to identify dispersal as a cause of synchrony \citep{Anderson2018},
@@
 Another class of methods exploit the time and timescale structure of synchrony.
-Using a method that examined the match between temporal variations in population and environmental synchrony, 
-\cite{Allstadt2015} confirmed the importance of a precipitation Moran effect for synchrony of spongy moth.
-A suite of wavelet and related Fourier techniques have been developed that can comprehensively describe the time and timescale
-structure of synchrony, identify Moran drivers of synchrony opperating in distinct timescale bands, and apportion fractions
-of observed synchrony to respective Moran drivers \citep{Sheppard2016,Sheppard2017,Desharnais2018,Sheppard2019,wsyn,Anderson2021}.
-The methods have been used, for instance, to provide one of the first clear demonstrations that changes in 
-Moran effects are another of the impacts of climatic changes \citep{Sheppard2016,Sheppard2019}; and to illustrate the importance of the
-timescale structure of environmental covariation for altering Moran effects \citep{Desharnais2018}.
-These methods have since been used to identify Moran effects in several additional systems
-(e.g., \citealp{Anderson2019,Walter2019,Walter2020,GarciaCarreras2022,CastoraniSheppardKelp}).
+E.g., \cite{Allstadt2015} confirmed the importance of a precipitation Moran effect in spongy moth.
+A suite of wavelet and related Fourier techniques have been developed that can comprehensively 
+describe the time and timescale
+structure of synchrony, identify Moran drivers of synchrony operating in distinct timescale bands, and apportion fractions
+of observed synchrony to respective Moran 
+drivers \citep{Sheppard2016,Sheppard2017,Desharnais2018,Sheppard2019,wsyn,Anderson2021}.
 
@@
 \citep{Moran1953} implies that temporal population correlation equals temporal environmental correlation, i.e., 
-$\cor(w_1,w_2)=\cor(\epsilon_1^{(1)},\epsilon_2^{(1)})$. If, instead, 
+$\cor(\epsilon_1^{(1)},\epsilon_2^{(1)})$. If, instead, 
 $w_i(t)=b w_i(t-1)+\epsilon_i^{(1)}(t)+\epsilon_i^{(2)}(t)$, so that both environmental variables influence 
-populations, the Moran theorem then implies that population synchrony is $\cor(w_1,w_2)=
-\cor(\epsilon_1^{(1)}+\epsilon_1^{(2)},\epsilon_2^{(1)}+\epsilon_2^{(2)})$.
-As we show in detail in SI section \ref{SM-Sect:IntroExample}, it is straightforward to show that 
+populations, the Moran theorem then implies that population synchrony is 
+$\cor(\epsilon_1^{(1)}+\epsilon_1^{(2)},\epsilon_2^{(1)}+\epsilon_2^{(2)})$.
+As we show in detail Supporting information, it is straightforward to show that 
 this quantity depends not only on the standard environmental synchrony 
-measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(a)})$, for 
-$a=1,2$, but also on the ``cross synchrony'' measures $\cov(\epsilon_1^{(a)},\epsilon_2^{(b)})$
-for $a \neq b$. 
+measures $\cov(\epsilon_1^{(1)},\epsilon_2^{(1)})$ and $\cov(\epsilon_1^{(2)},\epsilon_2^{(2)})$, 
+but also on ``cross synchrony'' measures such as $\cov(\epsilon_1^{(1)},\epsilon_2^{(2)})$
+and $\cov(\epsilon_1^{(2)},\epsilon_2^{(1)})$. 
 Cross synchrony measures represent synchrony between an environmental variable in one location and 
@@
 therefore differ from what synchrony would be if the environmental processes $\epsilon^{(1)}$ and 
-$\epsilon^{(2)}$ were unrelated, so that covariances between components of those processes are zero. 
-The difference is due to interactions
+$\epsilon^{(2)}$ were unrelated. The difference is due to interactions
 between the two Moran effects, and can be substantially positive or negative. This example was adapted 
@@
 already carried out a detailed analysis of interacting Moran effects in CA kelp populations and their importance 
-for kelp ecology, we instead use a simplified subset of the available kelp data. Our kelp analysis is intended 
-to illuminate the inner workings of the general mechanism we describe rather than being an exploration
-of kelp dynamics, specifically. Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
+for kelp ecology, we instead use a simplified subset of the available kelp data. 
+Our kelp results are not intended to extend what is known about kelp ecology (beyond what was discovered 
+in our earlier paper, \citealp{CastoraniSheppardKelp}), but rather serve as a
+vehicle for understanding the inner workings of our general theory of interacting Moran effects.
+Overall, this study introduces a general theory of Moran interactions and uses it to conceptually
 illuminate this newly observed but potentially quite important mechanism of spatiotemporal population dynamics.
@@
 Intermediate scenarios between the two scenarios of Fig. \ref{fig:PedagogAlt} are also possible, as will be revealed by 
-the theory. We henceforth refer to the quantity $l_{e1}-l_{e2}+l_n$ as the \emph{environmental effect alignment measure},
+our theory. We henceforth refer to the quantity $l_{e1}-l_{e2}+l_n$ as the \emph{environmental effect alignment measure},
 because it measures the extent to which the timing of the population influences of the two environmental variables
@@
 antagonistic Moran interactions are reversed, but with the same general concepts still operating 
-(Fig. \ref{fig:PedagogAlt2}).
+(Supporting information analogue of Fig. \ref{fig:PedagogAlt}).
 
@@
 the timescale, or period, $\sigma$, of the oscillation. 
-The spectrum $S_{w_i w_i}(f)$ is larger for frequencies at which $w_i(t)$ oscillates more. So, for example, a 
+The spectrum $S_{w_i w_i}(f)$ is larger for frequencies at which $w_i(t)$ oscillates with 
+larger amplitude. So, for example, a 
 population $w_i(t)$ that shows strong oscillatory dynamics of $5$-year period will have a large value of 
@@
 that relationship. The notation $S_{ww}(f)$ refers to the \emph{spectral matrix}, 
-which has $ij^{\text{th}}$ entry $S_{w_i w_j}(f)$. Spectral methods are standard  
+which has $ij^{\text{th}}$ entry $S_{w_i w_j}(f)$. The notation $S_{w_i w_i}$ for the spectrum makes sense
+and is consistent with the notation $S_{w_i w_j}$ for the cross spectrum because the spectrum of $w_i$ is the 
+cross spectrum of $w_i$ with itself. Spectral methods are standard  
 [e.g., \cite{Vasseur2007,Defriez2017}], and many background references are available 
@@
 is the classic, non-frequency-specific synchrony 
-measure $\left( \sum_{i \neq j} \cov(w_i,w_j) \right)/(N^2-N)$ (see SI section
-\ref{SM-Sect:measuring_sync} for details, here). If two time series $\epsilon_i^{(1)}(t)$ and $\epsilon_i^{(2)}(t)$
-($t=1,\ldots,T$ and $i=1,\ldots,N$) were measured at each sampling location (e.g., two environmental variables),
+measure $\left( \sum_{i \neq j} \cov(w_i,w_j) \right)/(N^2-N)$ (see Supporting information 
+for details, here, and additional comparisons to previous measures of synchrony). 
+Because $\rho_{ww}$ is a real-valued function of frequency, it equals $\rho_{ww}=\Rp(\rho_{ww})=
+\left( \sum_{i \neq j} \Rp(S_{w_i w_j}) \right)/(N^2-N)$, which only involves the cospectrum, $\Rp(S_{w_i w_j})$,
+not the full cross-spectrum, $S_{w_i w_j}$.
+If two time series $\epsilon_i^{(1)}(t)$ and $\epsilon_i^{(2)}(t)$ (e.g., two environmental variables)
+were measured at each sampling location and time,
 cross-variable synchrony (or, simply, \emph{cross synchrony}) between the variables
@@
 This is interpretable as pertaining to spatial synchrony because it makes comparisons across 
-distinct locations, i.e., $i \neq j$. 
+distinct locations. 
 It is interpretable as cross-variable synchrony because comparisons are between the two variables, i.e., cross spectra 
@@
 phase at that frequency equal to $\pi/2$, reflecting the one-year lag.
-See SI section \ref{SM-sect:examples_cross_sync} for detailed examples.
+See Supporting information for detailed examples of cross synchrony.
 
@@
 with expected values of its components equal to $0$. We make additional mild
-regularity assumptions for model stability 
-(SI section \ref{SM-sect:model_spec}). This model can be seen as 
+regularity assumptions for model stability, detailed in the Supporting information. 
+Our model can be seen as 
 a linearization of a very general dynamical model, influenced by ``weak noise'' [see, e.g.,
@@
 weak noise assumption often hold for noise which is fairly strong \citep{Nisbet1977,Desharnais2018}.
+The topic of nonlinearity is revisited in the Discussion.
 See \cite{Brillinger2001} for background on stochastic processes.
@@
 $\mu=\exp(-2 \pi \iota f)$, and where $\iota$, the Greek letter iota, is the imaginary unit.
-The derivation of the theory is in SI section \ref{SM-sect:derivation}.
+The derivation of the theory is in Supporting information, 
+where additional perspectives on our theory can also be found.
 
@@
 $\epsilon^{(1)}$ is assumed to have a simple positive effect on populations, but 
-lagged by one model time step (1q), i.e.,
-$p_1^{(1)}>0$, and $p_i^{(1)}=0$ for $i \neq 1$. 
+lagged by one model time step (1q). 
 For case study B (henceforth CaseB), $\epsilon^{(1)}$ is again assumed to have a simple 
 positive effect on populations,
-but now lagged by $3$q, so that $p_3^{(1)}>0$ while $p_i^{(1)}=0$ for $i \neq 3$.
+but now lagged by $3$q.
 For both CaseA and CaseB, the effects of $\epsilon^{(2)}$
-are assumed to be un-lagged and positive, i.e., $p_0^{(2)} > 0$ and $p_i^{(2)}=0$ for $i>0$.
+are assumed to be un-lagged and positive.
 For case study C (CaseC), the effects of $\epsilon^{(1)}$ are positive and lagged by $1$q and the 
-effects of $\epsilon^{(2)}$ are unlagged and negative, i.e.,
-$p_{1}^{(1)} > 0$ while $p_i^{(1)}=0$ for $i \neq 1$, and 
-$p_{0}^{(2)} < 0$ while $p_i^{(2)}=0$ for $i \neq 0$.
+effects of $\epsilon^{(2)}$ are unlagged and negative.
 The noise process
@@
 nitrate concentrations also fluctuate with period one year, but peak in the spring, a delay of $1$q
-compared to the wave peak. 
-Details of the noise processes are in SI section \ref{SM-sect:examples_cross_sync}.
-For all case studies,
-the autoregressive order of population dynamics is assumed, for simplicity, to be $1$, i.e, $b_1 \neq 0$ 
-but $b_i=0$ for $i>1$. Case studies do not cover the full range of possible scenarios which can be illuminated by 
-our general theory; they were selected for the intuition they provide, and for the correspondence of some of the cases to the 
-kelp examples we later discuss.
+compared to the wave peak. For all case studies,
+the autoregressive order of population dynamics is assumed, for simplicity, to be $1$.
+Details of the noise processes are Supporting information in the same section that
+has cross synchrony examples. 
+Model coefficients and other setup details for the theoretical case studies are also 
+specified in the Supporting information.
+Our case studies do not cover the full range of possible scenarios which can be illuminated by 
+our general theory; they were selected for the intuition they provide, and for the 
+correspondence of some of the cases to the kelp examples we later discuss.
 
@@
 The data are based on a subset of those used by \cite{CastoraniSheppardKelp} and
-\cite{Walter2021}, and data details are given in those papers and in SI section \ref{SM-sect:data};
+\cite{Walter2021}, and data details are given in those papers and in Supporting information;
 \cite{CastoraniSheppardKelp} and \cite{Walter2021} also provide an introduction to 
@@
 through essentially all of the $\Sexpr{quarters[1,2]}$-$\Sexpr{quarters[dim(quarters)[1],2]}$ 
-period (SI section \ref{SM-sect:data}). We also had estimates
+period (Supporting information). We also had estimates
 of maximum wave height and mean nitrate concentration for each quarter and location. Both waves
@@
 All computations were done in R version 3.6.3 on a laptop running Ubuntu Linux
-16.04. Complete codes for the project workflow are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.
+16.04. Complete codes for the project workflow will be published online, simultaneously with the final version of the paper.
+%are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.
 
@@
 (Fig. \ref{fig:TheoryCaseA}g).
-Thus interaction effects can either augment or reduce synchrony. Finally, interaction effects depended strongly on timescale 
+Thus interaction effects can either augment or reduce synchrony. Finally, 
+interaction effects depended strongly on timescale 
 for all case studies. This result complements earlier studies that showed direct Moran effects can 
@@
 straightforwardly, and those papers are case studies, whereas our results provide general theory.
+See Supporting information for mathematical details
+of the theoretical case studies.
 
@@
 The between-noise lag, $l_n$, of Fig. \ref{fig:PedagogAlt} is $0$q because $(\epsilon^{(1)},\epsilon^{(2)})$ was 
-taken to be a white noise process (see SI section \ref{SM-sect:examples_cross_sync} for details).
+taken to be a white noise process (Methods).
 Thus, the environmental effect alignment measure, $l_{e1}-l_{e2}+l_n$, equals $1$q.
@@
 But CaseC illustrates what can happen when noise processes have lagged associations, e.g., 
-both processes oscillate with annual periodicity and distinct phenology, an important scenario because seasonality is common.
-Recall that, for CaseC, $\epsilon^{(1)}$ effects on populations were positive and lagged by $1$q, i.e., $l_{e1}=1$ in the 
-language of Fig. \ref{fig:PedagogAlt2}; whereas 
+both processes oscillate with annual periodicity and distinct phenology, an important 
+scenario because seasonality is common.
+Recall that, for CaseC, $\epsilon^{(1)}$ effects on populations were positive 
+and lagged by $1$q, i.e., $l_{e1}=1$ in the 
+language of the Supporting information analogue of Fig. \ref{fig:PedagogAlt}; whereas 
 $\epsilon^{(2)}$ effects were unlagged and negative, i.e., $l_{e2}=0$ (Methods). 
-We compare CaseC to the Fig. \ref{fig:PedagogAlt2} $\beta$ scenarios instead of the Fig. \ref{fig:PedagogAlt} 
-$\alpha$ scenarios because
+We compare CaseC to the $\beta$ scenarios in the Supporting information analogue of Fig. \ref{fig:PedagogAlt}
+instead of the $\alpha$ scenarios of Fig. \ref{fig:PedagogAlt} itself because
 $\epsilon^{(2)}$ effects were negative for CaseC, the situation considered by the $\beta$ scenarios.
@@
 noise process $\epsilon^{(1)}$ were set up to lag peaks in the periodic process $\epsilon^{(2)}$ by
-$1$q (Methods, SI section \ref{SM-sect:examples_cross_sync}), so that $l_n=1$. 
+$1$q (Methods, Fig. \ref{fig:TheoryCaseA}), so that $l_n=1$. 
 Thus the environmental effect alignment measure, $l_{e1}-l_{e2}+l_n$, equals $2 \text{q}=0.5 \text{y}$, 
 and, on annual timescales, lags compounded, similar to 
-Fig. \ref{fig:PedagogAlt2}c, d: the 
+panels c, d of the Supporting information analogue of 
+Fig. \ref{fig:PedagogAlt}: the 
 annual positive population effects of $\epsilon^{(1)}$ 
@@
 $1$q, so that $l_n=-1$. Thus $l_{e1}-l_{e2}+l_n=0$ and, on annual timescales, lags cancelled,
-similar to Fig. \ref{fig:PedagogAlt2}a, b: the annual positive population effects of $\epsilon^{(1)}$
+similar to panels a, b of the Supporting information analogue of Fig. \ref{fig:PedagogAlt}: 
+the annual positive population effects of $\epsilon^{(1)}$
 coincided with the negative effects of $\epsilon^{(2)}$ every year, because $\epsilon^{(1)}$
@@
 effects was comparable to that of direct Moran effects (Fig. \ref{fig:SummaryKelpFig}
-for the CCA1 and SB regions and Fig. \ref{SM-fig:SummaryKelpFig_SM_lag4} for the CCA2 region).
+for the CCA1 and SB regions and Supporting information for the CCA2 region).
 Second, interaction effects could be positive (e.g., annual timescales, CCA1 region on Fig. 
@@
 synchrony had a strong annual component (Fig. \ref{fig:noise}a,b,j,k for the CCA1 and SB regions,
-Fig. \ref{SM-fig:directMoran_SM_lag4} for CCA2), which produced 
+Supporting information for CCA2), which produced 
 some of the annual synchrony observed in kelp 
@@
 Nitrates and waves are also synchronous on long timescales $>8$y  
-(Fig. \ref{fig:noise}g,h,p,q for the CCA1 and SB regions, Fig. \ref{SM-fig:directMoran_SM_lag4} 
+(Fig. \ref{fig:noise}g,h,p,q for the CCA1 and SB regions, Supporting information 
 for CCA2), possibly due to the North Pacific Gyre Oscillation 
@@
 regression coefficients determining $f_{\mcP^{(1)}}$, $f_{\mcP^{(2)}}$
-and $f_{\mcB}$ are in Table \ref{SM-tab:regcoefs}.
+and $f_{\mcB}$ are Supporting information.
 
@@
 were positive and delayed by $1$q and $\epsilon^{(2)}$ effects were negative and immediate.
-The delayed effects of nitrates in CCA1 are reflected in Table \ref{SM-tab:regcoefs}, where 
-regression results for $p_0^{(1)}$ are close to $0$ and those for $p_1^{(1)}$ are positive.
-Immediate negative effects of waves in CCA1 are also reflected in Table \ref{SM-tab:regcoefs},
+The delayed effects of nitrates in CCA1 are reflected in the tabulated regression results of 
+Supporting information, where 
+results for $p_0^{(1)}$ are close to $0$ and those for $p_1^{(1)}$ are positive.
+Immediate negative effects of waves in CCA1 are also reflected in the same table,
 where entries for $p_0^{(2)}$ are negative.
@@
 effects of waves/$\epsilon^{(2)}$ therefore lead to negative interactions.
-Interactions between Moran effects in CCA2 operated similarly to CCA1 (Fig. \ref{SM-fig:indirectMoran_SM_lag4}).
+Interactions between Moran effects in CCA2 operated similarly to CCA1 (Supporting information).
 
@@
 were observable within the same quarter: 
-see Table \ref{SM-tab:regcoefs}, where $p_0^{(1)}$ terms were close to zero for CCA1 and positive for SB,
+see the tabulated regression results of Supporting information, 
+where $p_0^{(1)}$ terms were close to zero for CCA1 and positive for SB,
 whereas $p_1^{(1)}$ terms were positive for CCA1 and close to zero for SB. 
 Wave effects were negative and
-immediate in both regions, i.e., $p_0^{(2)}<0$ in Table \ref{SM-tab:regcoefs}. 
+immediate in both regions, i.e., $p_0^{(2)}<0$ in the same Supporting information table. 
 In southern CA, kelp are less likely to be completely removed by winter waves \citep{Reed2011}. 
@@
 
-%*DONE Summary
+%Summary
 We provided a general mathematical theory of the new mechanism of 
@@
 substantially benefit both basic \citep{Liebhold2004} 
-and applied \citep{Hansen2020,Larsen2021,Herfindal2022} ecological research.
-
-%*DONE arguments about likelihood that interactions will be common, extending earlier arguments
+and applied [\cite{Hansen2020,Larsen2021,Herfindal2022}, see also below] ecological research.
+
+%arguments about likelihood that interactions will be common, extending earlier arguments
 We argue that it is likely that interactions between Moran effect are common. Most species
@@
 
-%*DONE Revist intuition
+%Revist intuition
 We again revisit the intuition behind interacting Moran effects using the central CA kelp
@@
 drivers must always be compared to the timescale of interest to determine interaction effects,
-as in Figs \ref{fig:PedagogAlt} and \ref{fig:PedagogAlt2}, and as captured formally in our theory.
+as in Figs \ref{fig:PedagogAlt} and its analogue in Supporting information, and as captured formally in our theory.
 Thus interactions between Moran effects provide yet another reason, among many reasons previous work
@@
 
-%*DONE Comparison to Castorani et al paper, and maybe also a bit on comparison to the Sheppard et al 2019 paper
+%Comparison to Castorani et al paper
 Our results about kelp were consistent with those of \cite{CastoraniSheppardKelp}, though
-that study uses distinct methods; our results complement the results of \cite{CastoraniSheppardKelp} 
-in important ways. In spite of 
+that study uses distinct methods. In spite of 
 numerous methodological choices which differed between the 
@@
 in central CA, but negative interactions in southern CA on annual timescales and in both central
-and southern CA on long timescales. The Fourier approach of our study was designed to facilitate 
+and southern CA on long timescales. As stated in the Introduction, our results about kelp 
+were not intended to reveal new aspects of kelp ecology -- we did that already in 
+\cite{CastoraniSheppardKelp} -- instead our purpose was to use kelp as a vehicle for helping explain 
+the mechanism of interacting Moran effects. Kelp is a good vehicle for this purpose
+because of the clarity of the lagged effects in the kelp system, and the differences in the 
+lags between Southern and Central CA.
+
+%Fourier versus wavelet
+Since both our Fourier approach and the wavelet methods of \cite{CastoraniSheppardKelp} provided
+similar results when applied to kelp, when should one approach be applied in favor of the other?
+The Fourier approach of our study was designed to facilitate 
 mathematical examination of interactions between Moran effect as a general mechanism, and development
-of general theory; whereas the 
-wavelet approach of \cite{CastoraniSheppardKelp} was instead optimized for detecting interactions 
+of general theory. The Fourier
+approach facilitates deriving analytic results from mathematical models such as (\ref{main_model_4}).
+The wavelet approach of \cite{CastoraniSheppardKelp} was instead optimized for detecting interactions 
 and identifying Moran mechanisms in data, in spite of non-stationarity and
-other complicating features which are present in many ecological datasets. The study of
+other complicating features which are present in many ecological datasets. 
+The study of
 \cite{Sheppard2019} developed the wavelet methods applied by \cite{CastoraniSheppardKelp}; an open-source implementation of 
@@
 
-%*DONE relationship to earlier papers comparing synchrony of drivers and pops
+%relationship to earlier papers comparing synchrony of drivers and pops
 It has been a frequent topic of research why populations are often less synchronous, or sychronous over a smaller
@@
 
-%*DONE Climate change
+%nonlinearity
+Our model is a linear population dynamical model, but most population dynamics are probably 
+nonlinear, at least to some degree. We speculate on what may be the outcome if future work
+explores interactions between two or more Moran effects using nonlinear population dynamical 
+models. 
+Some researchers have explored Moran effects of one environmental variable acting on nonlinear
+population models and found generally that synchronous environmental fluctuations tend to promote
+synchronous population fluctuations, in spite of nonlinearities, though often to a lesser 
+degree than for the linear case \citep{Royama2005,Engen2005}. 
+Our model is a linearization of a very general population dynamical model [SI section
+S1.2 of \cite{Desharnais2018} and SI section S1 of \cite{Walter2017}], and it is well known
+that analytic results for a linearized models, such as we have developed, also hold, approximately, for
+the original nonlinear model, as long as the noise impinging the model is sufficiently weak. 
+How weak is ``sufficiently weak'' is model dependent, but prior work examining that question
+provided evidence that weak noise approximations using linearizations of commonly used 
+ecological models are often actually quite accurate, even for fairly strong 
+noise \cite{Nisbet1977}. 
+Therefore we speculate that our main observations will hold even for nonlinear systems:
+1) Moran effects can substantially interact, either synergistically or destructively;
+2) interactions can depend on timescale; and 3) the nature of effects depend on relative,
+timescale-specific lags of environmental influences on populations and between 
+environmental variables. However, future work should nevertheless examine Moran interactions and 
+nonlinear models, since much has been learned by exploring deviations of ecological models 
+from predictions of theory based on linearization (e.g., \citealp{Reuman2006}).
+
+%Conservation
+Improved understanding of fundamental ecological phenomena often can be used to improve 
+conservation efforts. But how, specifically, might the improved understanding of Moran effects that 
+this study provides aid kelp conservation or restoration? Coauthor TWB is part of an effort
+%***add detail on what effort, official names, maybe funders, etc. - whatever is appropriate to identify the project officially
+to develop a stability model that informs the when, where and how of kelp restoration
+following the 20XX crash of kelp populations in large areas along the west coast from X down to Y.
+%***Tom to fill in the above missing info
+Specifically, the stability model informs kelp restoration efforts by... 
+%***Tom to fill in a very brief summary of how it works
+It is well known that stability of regional populations is closely related to synchrony,
+both generally \citep{Wilcox2017} and for kelp specifically \citep{Walter2021};
+and the first step in the decision tree that TWB and his collaborators have developed 
+involves computing the predicted stability of every subtidal reef across a large area.
+%***Tom, we should cite your report, or other appropriate reference, somewhere around here.
+Future iterations of TWB's and others' efforts to understand kelp stability and thereby 
+support kelp restoration initiatives should incorporate consideration of interacting Moran
+effects and their impacts on stability, and such consideration could likely be straightforwardly
+integrated since existing methods already compute stability metrics as a key part of their process.
+
+%Climate change
 Climate change has the potential to influence interactions between Moran effects in two specific ways
@@
 % The results of our present study extend this earlier work by showing that interactions between Moran effects 
-% are also timescale-specific, differing stongly in strength, and in sign, across timescales. 
+% are also timescale-specific, differing strongly in strength, and in sign, across timescales. 
 % A non-timescale-specific version of our theory could be obtained by appropriately integrating our main equation 
@@
 
-%*Revisit the kendall paper, briefly. Talk about possible three-way interactions.
-%***DAN: new review para for ecography
-We have focussed on interactions between two Moran drivers, but synchrony in most systems
+%Revisit the kendall paper, briefly. Talk about possible three-way interactions.
+We have focused on interactions between two Moran drivers, but synchrony in most systems
 may be a phenomenon with multiple (more than two) interacting causes.
@@
 
+\section{Author contributions}
+DCR conceptualized the paper, developed methodology, performed formal analysis and 
+investigation, carried out project administration and supervision, provided computing 
+resources, wrote and validated the software, produced most of the figures and wrote the 
+first draft. KCC and TWB performed data curation. MCNC helped with data visualization 
+and curation and project conceptualization. JAW and LWS helped with conceptualization, 
+methodology and formal analysis. MCNC and JAW helped with project administration. 
+All authors acquired the funding and performed substantial editing of manuscript drafts.
+
 \bibliographystyle{ecology_letters2}
@@
 and the values $\cov(\epsilon_i^{(1)},\epsilon_j^{(2)})$, for $i,j=1,2$, were all set equal to each
-other and to a parameter, $\varphi$, which appears on the horizontal axis and which characterizes the 
-strength of cross synchrony of the environmental variables. 
+other and the common value appears on the horizontal axis -- cross synchrony of the environmental variables. 
 See the online version for color
@@
 \begin{figure}
-\includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt_SM.pdf}
-\caption{%\doublespacing 
-Similar to Fig. \ref{fig:PedagogAlt}, but for slightly altered scenarios. For Fig. \ref{fig:PedagogAlt},
-for simplicity, we assumed that larger values of environmental variables had more positive influences
-on populations. We here assume that larger values of $\epsilon_i^{(1)}$ have more 
-positive population influences (as in Fig. \ref{fig:PedagogAlt}), 
-but that larger values of $\epsilon_i^{(2)}$ have more negative population influences (contrasting
-with Fig. \ref{fig:PedagogAlt}). In scenario $\beta 1$, 
-$l_n+l_{e1}=l_{e2}$, as in $\alpha 1$; but Moran interactions are nonetheless antagonistic because the 
-negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
-variable correspond to the peak negative effects of $\epsilon_i^{(1)}$. In scenario $\beta 2$, as in $\alpha 2$,
-the environmental effect alignment measure, $l_n+l_{e1}-l_{e2}$, equals $-\sigma/2$ for $\sigma=20$ the period. 
-But Moran interactions are synergistic because the 
-negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
-variable correspond to the peak positive effects of $\epsilon_i^{(1)}$. 
-See the online version for color renderings of all figures.}\label{fig:PedagogAlt2}
-\end{figure}
-
-\clearpage
-\newpage
-
-\begin{figure}
 \includegraphics[width=\textwidth]{./Fig2.pdf}
@@
 \caption{%\doublespacing 
-Theoretical case studies A (panels a-b), B (c-d), C1 (e-f) and C2 (g-h). 
+Results from theoretical case studies A (panels a-b), B (c-d), C1 (e-f) and C2 (g-h). 
 Left panels (a, c, e, g) show the terms on the right side
@@
 synchrony [see (\ref{rhoww_T1to3})].
-Model parameters were: $N=\Sexpr{N_A}$ and $b_1=\Sexpr{b_A}$ for all case studies; 
-$p_1^{(1)}=\Sexpr{p1_A[2]}$ and $p_0^{(2)}=\Sexpr{p2_A}$ for case study A; 
-$p_3^{(1)}=\Sexpr{p1_B[4]}$ and $p_0^{(2)}=\Sexpr{p2_B[1]}$ for case study B;
-$p_1^{(1)}=\Sexpr{p1_C1[2]}$, $p_0^{(2)}=\Sexpr{p2_C1[1]}$ 
-%$c_1=0$ and $c_2=-4/9$ 
-for case study C. For C1, peaks in the periodic noise process $\epsilon^{(2)}$
+For C1, peaks in the periodic noise process $\epsilon^{(2)}$
 lagged peaks in the periodic process $\epsilon^{(1)}$ by 1 quarter
@@
 amount (g-h).
-See SI section \ref{SM-sect:examples_cross_sync} and Fig. \ref{SM-fig:rhosForC} for 
-parameters associated with the noise for each case study,
-and see Methods and SI section \ref{SM-sect:details_case_studies} for additional details.
 Synch. contrib. = Synchrony contribution refers to contributions to synchrony of the individual 
@@
 study C scenario 1 (Fig. \ref{fig:TheoryCaseA}e; see text for details).
-See Fig. \ref{SM-fig:SummaryKelpFig_SM_lag4} for the central CA 2 (CCA2) region,
+See Supporting information for the central CA 2 (CCA2) region,
 for which results were substantially the same as for CCA1. This figure used 
 kelp lag $n=4$ [see (\ref{main_model_1})]; 
-see Figs \ref{SM-fig:SummaryKelpFig_SM_lag8} 
-and \ref{SM-fig:SummaryKelpFig_SM_lag12} for $n=8, 12$, respectively.
+see analogue figures in Supporting information for $n=8, 12$.
 See the online version for color renderings of all 
@@
 Cross synchrony is complex valued, with magnitude plotted on the vertical axis and phase displayed using color. 
-See Figs \ref{SM-fig:directMoran_SM_lag4} and \ref{SM-fig:indirectMoran_SM_lag4} for CCA2,
+See Supporting information for CCA2,
 for which results were similar to CCA1. This figure used 
-kelp lag $n=4$; see Figs \ref{SM-fig:directMoran_SM_lag8}, \ref{SM-fig:indirectMoran_SM_lag8}, 
-\ref{SM-fig:directMoran_SM_lag12} and \ref{SM-fig:indirectMoran_SM_lag12} for $n=8$ and $12$.
+kelp lag $n=4$; see analogue figures in Supporting information for $n=8$ and $12$.
 See the online version for color

### 2023-05-30 Paper/SupMat.Rnw
changes in response to referees, minor revisions

@@
 
+We briefly compare our definition of synchrony to other measures and terminology 
+which have been used. Perhaps the most commonly 
+used measure of synchrony is simple Pearson correlation between time series in different locations. When 
+time series from multiple locations are available, pairwise Pearson correlation values have sometimes
+been considered individually, often as a function of geographic distance between pairs of sites \citep{BjornstadFalck2001},
+or have simply been averaged:
+\begin{equation}
+\left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} \cor(w_i,w_j). \label{eq:classic_sync}
+\end{equation}
+As explained above, our metric, $\rho_{ww}$, is a timescale-specific generalization of 
+\begin{equation}
+\left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} \cov(w_i,w_j), \label{eq:nontssync}
+\end{equation}
+which is similar to (\ref{eq:classic_sync}), but weights time series of greater variance
+more heavily in the average. The latter definition, (\ref{eq:nontssync}), is better connected
+to concepts of regional stability because  
+\begin{equation}
+\var(\sum_i w_i) = \sum_{i} \var(w_i) + \sum_{i \neq j} \cov(w_i,w_j).
+\end{equation}
+The quantity $\var(\sum_i w_i)$ and related measures are often used as measures of 
+regional temporal instability.
+
+The concept of \emph{phase synchonization} has sometimes been used in studies
+of populations fluctuating with a single dominant periodicity [e.g., 
+\cite{Blasius1999,Massie2010}]. Phase synchronization refers to the tendency of population
+peaks in different places to occur simultaneously, and phase synchronization
+can occur independently of whether peak magnitudes are related across locations.
+Strong phase synchronization would tend to produce a large value of $\rho_{ww}$ at the
+common frequency of oscillation.
+
+The term ``asynchrony'' has been used to refer to either
+uncorrelated or independent dynamics in different locations. 
+If the processes $w_i$ and $w_j$ are independent, then not only are the classic synchrony quantities
+$\cor(w_i,w_j)$ and $\cov(w_i,w_j)$ equal to $0$, but so is the cospectrum $\Rp(w_i w_j)$.
+On the other hand, if $\cor(w_i,w_j)=\cov(w_i,w_j)=0$, then the cospectrum may nevertheless be non-zero 
+if it is positive for some 
+frequencies and negative for others in a manner which causes the opposing quantities to cancel.
+Research exploring ``asynchrony'' in a timescale-specific context would therefore have to begin 
+by clarifying the term.
+
+The term ``anti-synchrony'' has been used to refer to
+two time series which are negatively correlated. 
+Standard past uses of the term ``anti-synchrony'' may likewise need to be clarified 
+in a timescale-specific context.
+A negative correlation, $\cor(w_i,w_j)$, or covariance, $\cov(w_i,w_j)$, 
+may result from negative associations on some timescales outweighing positive 
+associations on other timescales, a pattern which would be clarified by 
+the cospectrum, $\Rp(w_i, w_j)$. Related ideas in a community context
+were explored by \cite{Zhao2020}.
+
+We now move on to defining cross-variable synchrony. 
 If $\epsilon^{(1)}=(\epsilon^{(1)}_1,\ldots,\epsilon^{(1)}_N)$ and 
@@
 
-\section{Model specification, full details}\label{sect:model_spec}
+\section{Main model specification, full details}\label{sect:model_spec}
 
@@
 &+ \frac{2}{|f_{\mcB}|^2} \Rp \left(f_{\mcP^{(1)}} \conj{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}} \right) \label{rhoww_Tint} \\
-&+ \text{other contributions}.
+&+ \text{other contributions}.\label{OCeq}
 \end{align}
@@
 
-\section{Details of data and data cleaning}\label{sect:data}
+\section{Additional perspectives on the theory}\label{sect:theorypersp}
+
+Our model (\ref{main_model_stochproc}) can be reconceptualized as 
+\begin{equation}
+\mcB w = \epsilon^{(T)} + \delta,\label{eq:model_simple}
+\end{equation}
+if one is willing to consider ``the environment'' to be 
+$\epsilon^{(T)} = \mcP^{(1)} \epsilon^{(1)} + \mcP^{(2)} \epsilon^{(2)}$. 
+This simplifies the model, producing some advantages, but requires an abstract
+definition of the environment which would be harder to measure and less 
+directly interpretable. Though this reconceptualization of our model is not
+directly useful for the main research goals of our paper, we nevertheless
+explore its benefits and drawbacks here, beginning with benefits.
+
+It is straightforward to derive the population spectral characteristics of (\ref{eq:model_simple})
+and a relationship between population synchrony and synchrony of the environment. 
+Taking the spectral matrix of both sides of (\ref{eq:model_simple}) gives
+\begin{equation}
+T(\mcB) S_{ww} T(\mcB)^* = S_{\epsilon^{(T)} \epsilon^{(T)}} + S_{\delta \delta} + S_{\epsilon^{(T)} \delta}
++S_{\delta \epsilon^{(T)}}
+\end{equation}
+and therefore
+\begin{equation}
+S_{ww} = \frac{1}{|f_{\mcB}|^2}[S_{\epsilon^{(T)} \epsilon^{(T)}} + S_{\delta \delta} + S_{\epsilon^{(T)} \delta}
++S_{\delta \epsilon^{(T)}}].
+\end{equation}
+Now averaging over all $i,j$ such that $i \neq j$, we have
+\begin{align}
+\left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{w_i w_j} &= \frac{1}{|f_{\mcB}|^2} \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{\epsilon^{(T)}_i \epsilon^{(T)}_j} \\
+&+ \frac{1}{|f_{\mcB}|^2} \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} S_{\delta_i \delta_j} \\
+&+ \frac{2}{|f_{\mcB}|^2} \left( \frac{1}{N^2-N}  \right) \sum_{i \neq j} \Rp (S_{\epsilon^{(T)}_i \delta_j}),
+\end{align}
+which gives
+\begin{align}
+\rho_{ww} &= \frac{1}{|f_{\mcB}|^2} \rho_{\epsilon^{(T)} \epsilon^{(T)}} \label{eq:main_result_simple_1}\\
+&+ \text{other contributions}. \label{eq:main_result_simple_2}
+\end{align}
+This is a simpler version of our main result (equations (\ref{MT-rhoww_T1to3})-(\ref{MT-rhoww_OC})
+of the main text and (\ref{rhoww_T1})-(\ref{OCeq}) of SI section \ref{sect:derivation}),
+and ``other contributions'' has the same meaning as was explained below (\ref{OCeq}).
+
+We can also use (\ref{eq:model_simple}) to recover the classic Moran theorem, as follows.
+If we impose the same assumptions on our model which were used to derive the original
+Moran theorem, i.e., $\epsilon^{(T)}$ is iid through time and $\delta$ is $0$, then 
+(\ref{eq:model_simple}) gives
+\begin{equation}
+S_{ww} = \frac{1}{|f_{\mcB}|^2} S_{\epsilon^{(T)} \epsilon^{(T)}},
+\end{equation}
+and $S_{\epsilon^{(T)} \epsilon^{(T)}}$ does not depend on frequency, $f$.
+For a pair of locations, $i$ and $j$, with $i \neq j$, we get
+\begin{align}
+S_{w_i w_j} &= \frac{1}{|f_{\mcB}|^2} S_{\epsilon^{(T)}_i \epsilon^{(T)}_j}\\
+S_{w_i w_i} &= \frac{1}{|f_{\mcB}|^2} S_{\epsilon^{(T)}_i \epsilon^{(T)}_i}\\
+S_{w_j w_j} &= \frac{1}{|f_{\mcB}|^2} S_{\epsilon^{(T)}_j \epsilon^{(T)}_j},
+\end{align}
+so
+\begin{align}
+\cor(w_i, w_j) &= \frac{\cov(w_i,w_j)}{\sqrt{\var(w_i)\var(w_j)}} \\
+&= \frac{\int S_{w_i w_j} df}{\sqrt{\int S_{w_i w_i} df \int S_{w_j w_j} df}} \\
+&= \frac{\frac{1}{|f_{\mcB}|^2} \int S_{\epsilon^{(T)}_i \epsilon^{(T)}_j} df}{\sqrt{\frac{1}{|f_{\mcB}|^2} \int S_{\epsilon^{(T)}_i \epsilon^{(T)}_i} df \frac{1}{|f_{\mcB}|^2} \int S_{\epsilon^{(T)}_j \epsilon^{(T)}_j} df}} \\
+&= \frac{\int S_{\epsilon^{(T)}_i \epsilon^{(T)}_j} df}{\sqrt{\int S_{\epsilon^{(T)}_i \epsilon^{(T)}_i} df 
+\int S_{\epsilon^{(T)}_j \epsilon^{(T)}_j} df}} \\
+&= \frac{\cov(\epsilon^{(T)}_i,\epsilon^{(T)}_j)}{\sqrt{\var(\epsilon^{(T)}_i)\var(\epsilon^{(T)}_j)}} \\
+&= \cor(\epsilon^{(T)}_i,\epsilon^{(T)}_j),
+\end{align}
+which proves the original Moran theorem.
+
+Although (\ref{eq:main_result_simple_1})-(\ref{eq:main_result_simple_2}) is simpler than 
+our main result in the main text, and relating our results back to the original Moran 
+theorem is a benefit, (\ref{eq:main_result_simple_1})-(\ref{eq:main_result_simple_2})
+is of limited use for understanding Moran effects because of the abstract notion of 
+the environment that it employs. We explain by way of example. Suppose $\epsilon^{(1)}$ 
+represents temperature and $\epsilon^{(2)}$ represents precipitation. Then $\mcP^{(1)}$
+encodes the nature of the temperature effect on our populations, and $\mcP^{(2)}$
+encodes the precipitation effect. Thus $\mcP^{(1)} \epsilon^{(1)} + \mcP^{(2)} \epsilon^{(2)}$
+is an abstract environmental quantity, incorporating two distinct traditional environmental
+variables and multiple lags. It would not necessarily be straightforward to interpret this
+quantity in a more concrete manner than simply ``the aggregate effect of temperature and 
+precipitation'' on the populations. Ecologists are more directly interested in the effects 
+of directly measurable environmental 
+variables such as temperature and precipitation than they are in the effects of highly derived 
+variables like $\mcP^{(1)} \epsilon^{(1)} + \mcP^{(2)} \epsilon^{(2)}$. Furthermore,
+interactions between Moran drivers cannot be effectively studied if those drivers 
+have been combined into a single unified influence.
+
+Although the simpler formulation (\ref{eq:model_simple}) has its strengths and weaknesses,
+it is mathematically equivalent to our original model. Because 
+\begin{align}
+S_{\epsilon^{(T)} \epsilon^{(T)}} &= T(\mcP^{(1)}) S_{\epsilon^{(1)} \epsilon^{(1)}} T(\mcP^{(1)})^* \\
+&+ T(\mcP^{(2)}) S_{\epsilon^{(2)} \epsilon^{(2)}} T(\mcP^{(2)})^* \\
+&+ T(\mcP^{(1)}) S_{\epsilon^{(1)} \epsilon^{(2)}} T(\mcP^{(2)})^* \\
+&+ T(\mcP^{(2)}) S_{\epsilon^{(2)} \epsilon^{(1)}} T(\mcP^{(1)})^*,
+\end{align}
+which gives 
+\begin{align}
+S_{\epsilon^{(T)} \epsilon^{(T)}} &= |f_{\mcP^{(1)}}|^2 S_{\epsilon^{(1)} \epsilon^{(1)}}\\
+&= |f_{\mcP^{(2)}}|^2 S_{\epsilon^{(2)} \epsilon^{(2)}}\\
+&= f_{\mcP^{(1)}} \overline{f_{\mcP^{(2)}}} S_{\epsilon^{(1)} \epsilon^{(2)}}\\
+&= f_{\mcP^{(2)}} \overline{f_{\mcP^{(1)}}} S_{\epsilon^{(2)} \epsilon^{(1)}},
+\end{align}
+we can average over $i \neq j$ to see that
+\begin{align}
+\rho_{\epsilon^{(T)} \epsilon^{(T)}} &= |f_{\mcP^{(1)}}|^2 \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon_i^{(1)} \epsilon_j^{(1)}} \\
+&+ |f_{\mcP^{(2)}}|^2 \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon_i^{(2)} \epsilon_j^{(2)}}\\
+&+ f_{\mcP^{(1)}} \overline{f_{\mcP^{(2)}}} \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon_i^{(1)} \epsilon_j^{(2)}}\\
+&+ f_{\mcP^{(2)}} \overline{f_{\mcP^{(1)}}} \left( \frac{1}{N^2-N} \right) \sum_{i \neq j} S_{\epsilon_i^{(2)} \epsilon_j^{(1)}}.
+\end{align}
+Therefore, 
+\begin{align}
+\rho_{\epsilon^{(T)} \epsilon^{(T)}} &= |f_{\mcP^{(1)}}|^2 \rho_{\epsilon^{(1)} \epsilon^{(1)}}\\
+&+ |f_{\mcP^{(2)}}|^2 \rho_{\epsilon^{(2)} \epsilon^{(2)}}\\
+&+ 2\Rp(f_{\mcP^{(1)}} \overline{f_{\mcP^{(2)}}} \rho_{\epsilon^{(1)} \epsilon^{(2)}}).
+\end{align}
+Plugging this into (\ref{eq:main_result_simple_1}) recovers our main result.
+
+\section{Details of the setup of the theoretical case studies}\label{sect:case_studies_setup}
+
+For CaseA, we assumed $p_1^{(1)}=\Sexpr{p1_A[2]}$, and $p_i^{(1)}=0$ for $i \neq 1$ and 
+$p_0^{(2)}=\Sexpr{p2_A}$ and $p_i^{(2)}=0$ for $i>0$.
+For CaseB, we assumed $p_3^{(1)}=\Sexpr{p1_B[4]}$ while $p_i^{(1)}=0$ for $i \neq 3$ and 
+$p_0^{(2)} = \Sexpr{p2_B[1]}$ and $p_i^{(2)}=0$ for $i>0$.
+For CaseC, we assumed $p_{1}^{(1)} = \Sexpr{p1_C1[2]}$ while $p_i^{(1)}=0$ for $i \neq 1$, and 
+$p_{0}^{(2)} = \Sexpr{p2_C1[1]}$ while $p_i^{(2)}=0$ for $i \neq 0$.
+For all theoretical case studies, we used $N=\Sexpr{N_A}$, $b_1=\Sexpr{b_A}$,
+and $b_i=0$ for $i>1$.
+See SI section \ref{sect:examples_cross_sync} and Fig. \ref{fig:rhosForC} for 
+information and parameters associated with the noise for each case study.
+
+\section{Details of kelp and related data, and data cleaning}\label{sect:data}
 
@@
 
-\section{Details of the theoretical case studies}\label{sect:details_case_studies}
+\section{Mathematical details of results from the theoretical case studies}\label{sect:details_case_studies}
 
@@
 $\rho_{\epsilon^{(1)}\epsilon^{(2)}} = \left| \frac{1}{1-c_1 \mu - c_2 \mu^2} \right|^2 w_{12} \exp(-2 \pi \iota f (l_1-l_2))$.
+
+\clearpage
+\newpage
+
+\begin{figure}
+\includegraphics[width=\textwidth]{../Results/Theory/PedagogFig_Alt_SM.pdf}
+\caption[Like main text Fig. \ref{MT-fig:PedagogAlt}, but for slightly altered scenarios]{%\doublespacing 
+Similar to Fig. \ref{MT-fig:PedagogAlt}, but for slightly altered scenarios. For Fig. \ref{MT-fig:PedagogAlt},
+for simplicity, we assumed that larger values of environmental variables had more positive influences
+on populations. We here assume that larger values of $\epsilon_i^{(1)}$ have more 
+positive population influences (as in Fig. \ref{MT-fig:PedagogAlt}), 
+but that larger values of $\epsilon_i^{(2)}$ have more negative population influences (contrasting
+with Fig. \ref{MT-fig:PedagogAlt}). In scenario $\beta 1$, 
+$l_n+l_{e1}=l_{e2}$, as in $\alpha 1$; but Moran interactions are nonetheless antagonistic because the 
+negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
+variable correspond to the peak negative effects of $\epsilon_i^{(1)}$. In scenario $\beta 2$, as in $\alpha 2$,
+the environmental effect alignment measure, $l_n+l_{e1}-l_{e2}$, equals $-\sigma/2$ for $\sigma=20$ the period. 
+But Moran interactions are synergistic because the 
+negative effects of $\epsilon_i^{(2)}$ means that the peak positive effects of that
+variable correspond to the peak positive effects of $\epsilon_i^{(1)}$. 
+See the online version for color renderings of all figures.}\label{fig:PedagogAlt2}
+\end{figure}
 
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainDirectMoran_ForMS.pdf}
-\caption[All theory constituents for kelp for direct Moran effects terms, kelp lags up to 4 quarters]{Plots of all constituents of 
+\caption[Extension of main text Fig. \ref{MT-fig:noise} to CCA2, with additional theoretical components pertaining to direct Moran effects]{Plots 
+of all constituents of 
 the terms in theory that pertain to direct Moran effects, for kelp, for the CCA1 region (a-o), the SB region
@@
 \includegraphics[width=.5\textwidth]{../Results/Kelp_SpectralDecompResults/RegionalAnalysis_CentCal2_KelpLag4_ExplainInteractingMoran_ForMS.pdf} 
-\caption[All theory constituents for kelp for Moran interaction terms, kelp lags up to 4 quarters]{Plots of 
+\caption[Extension of main text Fig. \ref{MT-fig:noise} to CCA2, with additional theoretical components pertaining to Moran interactions]{Plots of 
 all constituents of the terms in theory that pertain to interactions between Moran effects, for kelp, for CCA1 region
@@
 \begin{table}[!h]
-\caption{Regression coefficients used to parameterize (\ref{MT-main_model_1})-(\ref{MT-main_model_4}) of the main text for each 
+\caption[Regression coefficients used to parameterize (\ref{MT-main_model_1})-(\ref{MT-main_model_4}) of the main text for three kelp regions]{Regression coefficients used to parameterize (\ref{MT-main_model_1})-(\ref{MT-main_model_4}) of the main text for each
 of our three regions. All values rounded to two decimal places.}\label{tab:regcoefs}

### 2023-05-30 Code/Theory_IntroExample.R

@@ -83,7 +83,7 @@ par(fig=c((yaxwd)/totwd,
 plot(res$od,res$sync,type="l",xlab="Parameter",ylab="Population correlation")
 lines(res$od,res$sync.in,type="l",lty="dashed")
 
-mtext(latex2exp::TeX("Parameter, $\\phi$"),1,1.2)
+mtext("Environmental cross sycnhrony",1,1.2)
 mtext("Population synchrony (correlation)",2,1.2)
 
 inds<-which(res$sync<=res$sync.in)

### 2023-05-30 Code/Theory_PedagogFig_Alt.R

@@ -94,12 +94,12 @@ par(fig=c((yaxwd)/totwd,
     mai=c(0,0,0,0),mgp=c(3,.15,0),tcl=-.25)
 
 #do the start of the plotting
+xlimits<-c(0,60) #range(tim)
 ylimits<-range(envvar_1)
 ylimits[2]<-ylimits[2]+.2*diff(ylimits)
-plot(tim,envvar_1[1,],type="n",ylim=ylimits,xaxs="i",xaxt="n",yaxt="n")
+plot(tim,envvar_1[1,],type="n",ylim=ylimits,xlim=xlimits,xaxs="i",xaxt="n",yaxt="n")
 axis(side=1,label=FALSE)
 #put the population rectangles, only done here so other stuff plots over it
-xlimits<-range(tim)
 rect(xlimits[1],vertoffset/2-1,xlimits[2],vertoffset/2+1,col="lightgrey",border=NA)
 rect(xlimits[1],3*vertoffset/2-1,xlimits[2],3*vertoffset/2+1,col="lightgrey",border=NA)
 rect(xlimits[1],5*vertoffset/2-1,xlimits[2],5*vertoffset/2+1,col="lightgrey",border=NA)
@@ -116,9 +116,9 @@ text(10,3*vertoffset/2,latex2exp::TeX("population, $w_2(t)"),adj=c(0,.5),cex=0.6
 text(10,5*vertoffset/2,latex2exp::TeX("population, $w_1(t)"),adj=c(0,.5),cex=0.6)
 
 #add arrows showing max positive and negative effects on the pops
-for (csw in 1:dim(sinpks_1_x)[1])
+for (csw in 1:(dim(sinpks_1_x)[1]))
 {
-  for (cpk in 1:dim(sinpks_1_x)[2])
+  for (cpk in 1:(dim(sinpks_1_x)[2]))
   {
     #plot the max positive effects on the pops as black arrows
     sx<-sinpks_1_x[csw,cpk]
@@ -164,8 +164,8 @@ for (csw in 1:dim(sinpks_1_x)[1])
 
 #label the arrows to show the lags
 text(sinpks_1_x[3,2]+le1/2,2*vertoffset+1+.5,latex2exp::TeX("$l_{e1}$"),cex=0.7,adj=c(0.5,-.1))
[…]

### 2023-05-31 Paper/Paper.Rnw
continued edits for referees

@@
 S1.2 of \cite{Desharnais2018} and SI section S1 of \cite{Walter2017}], and it is well known
-that analytic results for a linearized models, such as we have developed, also hold, approximately, for
+that analytic results for linearized models, such as we have developed, also hold, approximately, for
 the original nonlinear model, as long as the noise impinging the model is sufficiently weak. 
@@
 
-%Conservation
-Improved understanding of fundamental ecological phenomena often can be used to improve 
-conservation efforts. But how, specifically, might the improved understanding of Moran effects that 
-this study provides aid kelp conservation or restoration? Coauthor TWB is part of an effort
-%***add detail on what effort, official names, maybe funders, etc. - whatever is appropriate to identify the project officially
-to develop a stability model that informs the when, where and how of kelp restoration
-following the 20XX crash of kelp populations in large areas along the west coast from X down to Y.
-%***Tom to fill in the above missing info
-Specifically, the stability model informs kelp restoration efforts by... 
-%***Tom to fill in a very brief summary of how it works
-It is well known that stability of regional populations is closely related to synchrony,
-both generally \citep{Wilcox2017} and for kelp specifically \citep{Walter2021};
-and the first step in the decision tree that TWB and his collaborators have developed 
-involves computing the predicted stability of every subtidal reef across a large area.
-%***Tom, we should cite your report, or other appropriate reference, somewhere around here.
-Future iterations of TWB's and others' efforts to understand kelp stability and thereby 
-support kelp restoration initiatives should incorporate consideration of interacting Moran
-effects and their impacts on stability, and such consideration could likely be straightforwardly
-integrated since existing methods already compute stability metrics as a key part of their process.
+%Conservation - new
+A better understanding of fundamental ecological phenomena can be used to improve 
+conservation efforts. But how, specifically, might improved understanding of Moran effects 
+from this study aid kelp conservation or restoration? Scientists and managers are 
+currently developing a stability model that informs when, where, and how to restore kelp 
+forests following recent large-scale declines \citep{Bell2023}.
+Specifically, these efforts inform kelp restoration by modeling the 
+dynamics of kelp abundance (measured by diver surveys) as a function of environmental and 
+bathymetric variables. Resulting maps of kelp stability are then compared to maps of recent  
+canopy declines to identify sites where restoration 
+activities are needed and would lead to stable future forests \citep{GiraldoOspinoInPrep}.
+It is well known that stability of regional populations is influenced by synchrony, both 
+generally \citep{Wilcox2017} and for kelp specifically \citep{Walter2021}; and the first step in 
+the emerging kelp restoration decision tree of \cite{Bell2023} and \cite{GiraldoOspinoInPrep} 
+involves predicting kelp stability across all 
+suitable habitat in California. Future iterations of this process may consider 
+interacting Moran effects to better predict stability and thereby better support kelp restoration.
 

### 2023-05-31 Code/Theory_IntroExample.R

@@ -83,7 +83,7 @@ par(fig=c((yaxwd)/totwd,
 plot(res$od,res$sync,type="l",xlab="Parameter",ylab="Population correlation")
 lines(res$od,res$sync.in,type="l",lty="dashed")
 
-mtext("Environmental cross sycnhrony",1,1.2)
+mtext("Environmental cross synchrony",1,1.2)
 mtext("Population synchrony (correlation)",2,1.2)
 
 inds<-which(res$sync<=res$sync.in)

### 2023-06-27 Paper/Paper.Rnw
various changes

@@
 All computations were done in R version 3.6.3 on a laptop running Ubuntu Linux
-16.04. Complete codes for the project workflow will be published online, simultaneously with the final version of the paper.
-%are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.
+16.04. Complete codes for the project workflow are at \url{https://github.com/reumandc/InteractingMoranEffects.git}.
 
@@
 The Fourier approach of our study was designed to facilitate 
-mathematical examination of interactions between Moran effect as a general mechanism, and development
+mathematical examination of interactions between Moran effects as a general mechanism, and development
 of general theory. The Fourier
@@
 ecological models are often actually quite accurate, even for fairly strong 
-noise \cite{Nisbet1977}. 
+noise \citep{Nisbet1977}. 
 Therefore we speculate that our main observations will hold even for nonlinear systems:
@@
 \section{Author contributions}
-DCR conceptualized the paper, developed methodology, performed formal analysis and 
-investigation, carried out project administration and supervision, provided computing 
-resources, wrote and validated the software, produced most of the figures and wrote the 
-first draft. KCC and TWB performed data curation. MCNC helped with data visualization 
-and curation and project conceptualization. JAW and LWS helped with conceptualization, 
-methodology and formal analysis. MCNC and JAW helped with project administration. 
-All authors acquired the funding and performed substantial editing of manuscript drafts.
+\textbf{Reuman:} Conceptualization (lead), formal analysis (lead), funding acquisition (equal), investigation (lead), methodology (lead), project administration (equal), resources (lead), software (lead), validation (lead), visualization (lead), writing - original draft (lead), writing - review and editing (lead).
+\textbf{Castorani:} Conceptualization, data curation, funding acquisition (equal), project administration (equal), visualization, writing - original draft, writing - review and editing.
+\textbf{Cavanaugh:} Data curation, funding acquisition (equal), writing - review and editing. 
+\textbf{Sheppard:} Conceptualization, formal analysis, funding acquisition (equal), methodology, writing - review and editing.
+\textbf{Walter:} Conceptualization, formal analysis, funding acquisition (equal), methodology, project administration (equal), writing - review and editing.
+\textbf{Bell:} Data curation (lead), funding acquisition (equal), writing - review and editing.
 

### 2023-06-27 Paper/Paper.Rnw
small

@@
 %Get line numbers to work, also in figure captions
-\linenumbers
-\DeclareCaptionFont{linenumbers}{\internallinenumbers}
-\captionsetup[figure]{font+=linenumbers} % use \internallinenumbers inside figure caption
-
-\doublespacing
+%\linenumbers
+%\DeclareCaptionFont{linenumbers}{\internallinenumbers}
+%\captionsetup[figure]{font+=linenumbers} % use \internallinenumbers inside figure caption
+
+%\doublespacing
 
@@
 %header material for paper
-\title{How environmental drivers of spatial synchrony interact}
+\title{Latex-rendered version of the main text of ``How environmental drivers of spatial synchrony interact''}
 \date{}
@@
 \maketitle
-
-\newpage
-
-\large{\textbf{How environmental drivers of spatial synchrony interact}}
 

### 2023-06-27 Paper/Paper.Rnw
making sure it runs from a fresh clone

@@
 
-resloc<-"../Results/Kelp_SpectralDecompResults/"
+resloc<-"../Results/Kelp_MapFig/"
 SBlocstouse<-readRDS(paste0(resloc,"SBlocstouse.Rds"))

### 2023-06-27 Code/main.R

@@ -10,13 +10,12 @@ graphics.off()
 #checkpoint, to help ensure it continues to work into the future
 #***
 
-print("Running checkpoint, this may take a while the first time...")
-if (!dir.exists("./.checkpoint/")){
-  dir.create("./.checkpoint/")
-}
-checkpoint::checkpoint(snapshot_date="2022-01-01",r_version=getRversion(),checkpoint_location=getwd(),
-                       scan_now = TRUE) #mine is R-3.6.3
-#The version of the checkpoint package I used was installed from CRAN on 2022 01 31
+#print("Running checkpoint, this may take a while the first time...")
+#if (!dir.exists("./.checkpoint/")){
+#  dir.create("./.checkpoint/")
+#}
+#checkpoint::checkpoint(snapshot_date="2023-06-01",r_version=getRversion(),checkpoint_location=getwd(),
+#                       scan_now = TRUE) 
 
 #***
 #Develop the theoretical case studies

### 2023-06-27 Code/main.R
trivial

@@ -55,7 +55,7 @@ source("Kelp_AnalysisStep_CleanData_Final.R")
 
 print("Mapping the system...")
 source("Kelp_AnalysisStep_MakeMapFigDat.R") #preps the data for Max Castorani code,
-#run separately on another machine to produce Fig. 2, which is purely descriptive
+#run separately on another machine to produce Fig. 3, which is purely descriptive
 
 #***
 #Do the spectra decomposition analysis for kelp

### 2023-06-27 Code/Kelp_AnalysisStep_CleanData_Final.R
getting ready for complete near-term reproducibility

@@ -8,7 +8,7 @@ rm(list=ls())
 #External codes needed
 #***
 
-#packages needed (invoked with "::"): 
+#packages needed (invoked with "::"): wsyn
 
 #***
 #Locations for storing results, and other prep

### 2023-06-27 Code/Kelp_AnalysisStep_MakeMapFigDat.R

@@ -6,7 +6,7 @@ rm(list=ls())
 #External codes needed
 #***
 
-#packages needed (invoked with "::"): 
+#packages needed (invoked with "::"): ncf
 
 #***
 #Location for storing results and other prep

### 2023-06-27 Code/main.R

@@ -7,17 +7,6 @@ rm(list=ls())
 graphics.off()
 
 #***
-#checkpoint, to help ensure it continues to work into the future
-#***
-
-#print("Running checkpoint, this may take a while the first time...")
-#if (!dir.exists("./.checkpoint/")){
-#  dir.create("./.checkpoint/")
-#}
-#checkpoint::checkpoint(snapshot_date="2023-06-01",r_version=getRversion(),checkpoint_location=getwd(),
-#                       scan_now = TRUE) 
-
-#***
 #Develop the theoretical case studies
 #***
 
@@ -65,4 +54,3 @@ print("Spectral analysis...")
 source("Kelp_AnalysisStep_SpectralDecomp.R")
 print("Done!")
 
-checkpoint::uncheckpoint()
